### 2021-07-31 paper.tex
Upload files of July 30 version

\documentclass[12pt, a4paper]{article}

\usepackage{mathptmx} % this doesn't have \jmath
%\usepackage{newtxtext, newtxmath} % I can't make this work
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue,linktocpage=true]{hyperref}
%\RequirePackage{hypernat}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[tight]{subfigure}
\usepackage{authblk}

\addtolength{\textheight}{28mm}
\addtolength{\voffset}{-17mm}
\addtolength{\textwidth}{-3mm}
\addtolength{\hoffset}{1.5mm}

%\arxiv{arXiv: ***}

\newcommand{\mr}{\mathrm}
\newcommand{\st}{\mid}%{\,|\,}
\newcommand{\funt}{\tau} % function that gives maximum number of touched tiles for real-valued length
\newcommand{\funl}{\lambda} % function that gices infimum real-valued length for a given number of touched tiles
\newcommand{\funti}{T} % function that gives maximum number of touched tiles for integer-valued length
\newcommand{\funli}{\Lambda} % function that gives minimum integer-valued length for a given number of touched tiles
\newcommand{\len}{\ell} % length variable, real-valued
\newcommand{\leni}{\ell} % length variable, integer-valued. Same notation, but I keep the different command just in case
\newcommand{\tiles}{t} % number of tiles (integer-valued)
\let\Re\relax % remove the command...
\DeclareMathOperator{\Re}{Re} % ...to define it differently
\let\Im\relax
\DeclareMathOperator{\Im}{Im}
\newcommand{\isolr}{i^+}
\newcommand{\jsolr}{j^+}
\newcommand{\isoli}{i^\ast}
\newcommand{\jsoli}{j^\ast}
\newcommand{\mss}{M}

\newtheorem{proposition}{Proposition}%[section]
\newtheorem{theorem}{Theorem}%[section]
\newtheorem{corollary}{Corollary}%[section]
%\newtheorem{lemma}{Lemma}%[section]

\makeatletter
\renewcommand\theenumi{(\@roman\c@enumi)} % this affects \ref's too
\renewcommand\labelenumi{\theenumi} % this doesn't affect \ref's
\makeatother

\graphicspath{{figures/}}

% !TeX spellcheck = en_GB
% This is so that TexStudio sets language automatically

\begin{document}

\title{
On the maximum number of tiles touched\\
by a line segment on a rectangular grid
}

% Authors with different affiliations: from https://tex.stackexchange.com/a/370813/90222
\author[1]{Luis Mendo}
\author[2]{**Name**}

\affil[1]{\small Universidad Polit\'ecnica de Madrid. \texttt{luis.mendo@upm.es}}
%Information Processing and Telecommunications Center, Universidad Polit\'ecnica de Madrid\\
%Avenida Complutense, 30. 28040 Madrid, Spain.\\
\affil[2]{\small **Affiliation, e-mail**}

%ORCID: Luis Mendo: 0000-0001-5691-714X

\maketitle

\begin{abstract}
***

\emph{Keywords:} ***.

%\emph{MSC2010:} 65C10, 65C50.
\end{abstract}

\section{Introduction}
\label{part: intro}

Given $a, b \in \mathbb R^+$, consider a grid on $\mathbb R^2$ formed by rectangular \emph{tiles} of width $a$ and height $b$. A line segment of length $\len \in \mathbb R^+$ is located on the plane with arbitrary position and orientation. The segment is said to \emph{touch} a tile if it intersects its interior.\footnote{
The definition uses the interior of the tile, excluding the border, to avoid uninteresting results such as a ``zero-length'' segment touching (a vertex of) $4$ tiles.}
What is the maximum number of tiles that the segment can touch?

As an example, consider $a=1.35$, $b=1$. A segment of unit length can be placed as shown in Figure~\ref{fig: examples} (left) to touch $3$ tiles. In fact, the solution for $\len=1$ is $3$ tiles. The figure also illustrates that the solution for length $2.4$ is $5$ (center), and for $4.7$ it is $8$ (right).

\begin{figure}
\centering%
\includegraphics[width=.85\textwidth]{examples_1p35}%
\caption{Examples for $a=1.35, b=1$; $\len=1$, $\len=2.4$ and $\len=4.7$
}%
\label{fig: examples}%
\end{figure}%

An equivalent formulation of the problem is obtained allowing segments of length $\len$ \emph{or smaller}. The equivalence is clear from the fact that reducing the length cannot increase the number of touched tiles. Either of these formulations will be referred to as the \emph{direct} problem.

The \emph{inverse} problem is, given $\tiles \in \mathbb N$, to determine the infimum length of all segments that touch at least $\tiles$ tiles. For real-valued lengths this infimum is not a minimum, because given a segment that touches $\tiles$ tiles, its length can always be reduced by some non-zero amount without changing the number of touched tiles. This is a consequence of the interior of the each tile being an open set.

% ***Do we need to discuss potential applications of this (if we can find any)? This probably depends on the journal
% ***Do we need a section with conclusions? Probably not; depends on the journal

These problems are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem. Some references: ``Buffon's Noodle Problem'', J.F.~Ramaley, 1969. ``On Laplace's Extension of the Buffon Needle Problem'', B.J.~Arnow, 1994. \\
\url{https://www.jstor.org/stable/2317945} \\
\url{https://www.jstor.org/stable/2687085}

This paper analyzes both problems for real-valued lengths, and for integer-valued lengths on a unit square grid. Section~\ref{part: fund results} presents fundamental results, from which the characterization is obtained in Section~\ref{part: arbitrary grid, real lengths} for an arbitrary grid with real-valued lengths; in Section~\ref{part: square grid, real lengths} for a square grid with real-valued lengths; and in Section~\ref{part: unit square grid, integer lengths} for a unit square grid with integer-valued lengths.

% The rounding down (floor) and rounding up (ceiling) operations will be denoted as $\lfloor \cdot \rfloor$ and $\lceil \cdot \rceil$. The notation $[\mathsf S]$ will be used for the Iverson function, which equals $1$ if statement $\mathsf S$ is true and $0$ otherwise.

\section{Fundamentals}
\label{part: fund results}

For a grid with horizontal spacing $a$ and vertical spacing $b$, lines of the form $x = ka$ or $y = kb$ with $k \in \mathbb Z$ will be called \emph{grid lines}. A \emph{tile} is delimited by two pairs of consecutive horizontal an vertical grid lines. The intersection points of horizontal and vertical grid lines will be called \emph{grid points}. These correspond to vertices of the tiles.

Every segment has an associated \emph{canonical rectangle}, which is the minimum-size rectangle that is formed by grid lines and contains the segment. More specifically, if the segment has endpoints $(x_1,y_1)$, $(x_2,y_2)$, where $x_1, x_2, y_1, y_2 \in \mathbb R$, its canonical rectangle has lower-left corner and upper-right corner given as
\begin{align*}
& (\lfloor\min\{x_1, x_2\}/a\rfloor a, \lfloor\min\{y_1,y_2\}/b\rfloor b), \\
& (\lceil\max\{x_1, x_2\}/a \rceil a, \lceil\max\{y_1,y_2\}/b \rceil b).
\end{align*}
The dimensions of the canonical rectangle, normalized to the tile width and height respectively, are two integer numbers $i$, $j$. Two examples are illustrated in Figure~\ref{fig: canonical rectangle and touched tiles}, both with $i=5$, $j=4$. Clearly, all tiles touched by the segment are contained in the canonical rectangle. Note also that the canonical rectangle can have $i=0$ or $j=0$ if the segment coincides with part of a grid line.

\begin{figure}
\centering%
\subfigure[The segment does not pass through any interior grid points]{%
\label{fig: S_nogridpoints}%
\includegraphics[width=.49\textwidth]{S_nogridpoints_1p35}%
}\hfill%
\subfigure[The segment passes through some interior grid points]{%
\label{fig: S_gridpoints}%
\includegraphics[width=.49\textwidth]{S_gridpoints_1p35}%
}%
\caption{Canonical rectangle and touched tiles
}%
\label{fig: canonical rectangle and touched tiles}%
\end{figure}%

\begin{proposition}
\label{prop: i+j-1}
Consider an arbitrary segment, and let $i, j$ respectively denote the normalized width and height of its canonical rectangle. If $i, j \geq 1$, the number of tiles touched by the segment is at most $i+j-1$. This bound is attained if and only if the segment does not pass through any grid point in the interior of the rectangle.
\end{proposition}

\begin{proof}
The segment touches, by definition, two tiles in opposite corners of the canonical rectangle. It can be assumed, without loss of generality, that those tiles are in the lower-left and upper-right corners of the rectangle, as in Figure~\ref{fig: canonical rectangle and touched tiles}. The touched tiles can be thought of as following a path within the canonical rectangle. Starting at the lower-left tile, the next tile can be the one to the left, the one above, or the one above and to the left. The latter case occurs if and only if the segment passes through the grid point between those two tiles.

Since the segment follows a straight line, once it ``leaves'' a row of tiles in its path from the lower-left to the upper-right corner, it can never touch again a tile from that row. The same observation applies for the columns.

This implies that the maximum number of touched tiles is $i+j-1$, which is attained if and only if the segment avoids all grid points in the interior of the canonical rectangle, as in Figure~\ref{fig: S_nogridpoints}. Note that grid points at the corners of the rectangle do not count for this; and that the segment cannot pass through any other grid points on the rectangle border, because that would imply $i=0$ or $j=0$. Figure~\ref{fig: S_gridpoints} illustrates a case where the maximum is not attained.
\end{proof}

\begin{proposition}
\label{prop: len ineq i j}
Consider $a, b, \len \in \mathbb R^+$ and $i, j \in \mathbb N$, $i, j \geq 2$ arbitrary.
\begin{enumerate}
\item
\label{prop: len ineq i j: ineqs}
The following inequalities hold for any segment with length $\len$ whose canonical rectangle has normalized dimensions $i, j$:
\begin{align}
\label{eq: len > sqrt}
\len &> \sqrt{(i-2)^2 a^2 + (j-2)^2 b^2}, \\
\label{eq: len leq sqrt}
\len &\leq \sqrt{i^2 a^2 + j^2 b^2}.
\end{align}
\item
\label{prop: len ineq i j: exist}
Conversely, if $\len$, $i$, $j$ satisfy \eqref{eq: len > sqrt} and \eqref{eq: len leq sqrt} there exists a segment of length $\len$ whose canonical rectangle has normalized dimensions $i$ and $j$.
\item
\label{prop: len ineq i j: ineq, exist}
There is a segment of length not exceeding $\len$ that has a canonical rectangle with normalized dimensions $i, j$ if and only if \eqref{eq: len > sqrt} holds.
\end{enumerate}
\end{proposition}

\begin{proof}
\ref*{prop: len ineq i j: ineqs} The inequalities follow from the fact that the segment endpoints lie in the interiors or on the outer edges of two tiles in opposite corners of the canonical rectangle. This is illustrated in Figures~\ref{fig: len ineq 4 3} and \ref{fig: len ineq 4 2} for two specific $(i,j)$ pairs respectively. For each $(i,j)$, segments are shown with lengths close to either of the two bounds. Note that \eqref{eq: len > sqrt} is valid even for $i=2$, $j=2$, in which case it reduces to $\len>0$.

\ref*{prop: len ineq i j: exist} For $a$, $b$, $\len$, $i$, $j$ satisfying the two inequalities, a segment of length $\len$ can be found that has its endpoints in the interiors or on the outer edges of the two shaded tiles of a rectangle of normalized dimensions $i$ and $j$, as in Figure~\ref{fig: len ineq i j}. Therefore this segment has the given rectangle as canonical.
\begin{figure}
\centering%
\subfigure[$i=4$, $j=3$]{%
\label{fig: len ineq 4 3}%
\includegraphics[width=.49\textwidth]{L_ineq_4_3_1p35}%
}\hfill%
\subfigure[$i=4$, $j=2$]{%
\label{fig: len ineq 4 2}%
\includegraphics[width=.49\textwidth]{L_ineq_4_2_1p35}%
}%
\caption{Relationship between segment length $\len$ and dimensions $i$, $j$ of the canonical rectangle
}%
\label{fig: len ineq i j}
\end{figure}%

\ref*{prop: len ineq i j: ineq, exist} ``\eqref{eq: len > sqrt} $\Rightarrow$ there is a segment\ldots''. Assume that \eqref{eq: len > sqrt} holds. It is always possible to choose a length equal to or smaller than $\len$ such that both \eqref{eq: len > sqrt} and \eqref{eq: len leq sqrt} hold. The result follows, for that length, from part~\ref{prop: len ineq i j: exist}.

``There is a segment\ldots $\Rightarrow$ \eqref{eq: len > sqrt} ''. Assume that a segment exists with length not exceeding $\len$ and with a canonical rectangle with normalized dimensions $i$, $j$. From part~\ref{prop: len ineq i j: ineqs}, \eqref{eq: len > sqrt} holds for that segment length, and thus for $\len$.
\end{proof}

According to Proposition~\ref{prop: i+j-1}, in order to maximize the number of touched tiles for a given length, the position and orientation of the segment should be chosen to obtain $i+j-1$ as large as possible, where $i$ and $j$ are the normalized dimensions of its canonical rectangle. On the other hand, Proposition~\ref{prop: len ineq i j} restricts the $i, j$ values that can be achieved with a given length. A relevant question is: are there any $(i,j)$ pairs that can be disregarded irrespective of the length $\len$? In other words, what is the ``smallest'' subset of $\mathbb N^2$ such that the $(i,j)$ pair that maximizes the number of tiles for any length can always be found within that subset? A subset that contains a maximizing $(i,j)$ for any length will be called a \emph{sufficient} set. Clearly, this set must contain at least one such pair for each possible value of $i+j-1$, so that the set can produce that value as solution (for the corresponding lengths). A sufficient set that contains exactly one pair $(i,j)$ for each value of $i+j-1$ will be termed a \emph{minimal sufficient} set.

For instance, it is intuitively clear from Figure~\ref{fig: examples} that segment orientations near the vertical or horizontal directions (resulting in $i=1$ with large $j$, or $j=1$ with large $i$) will not maximize the number of touched tiles.
% This is true for any a, b: 2 can achieved instead of 1 without compromising the other dimension.
The corresponding $(i,j)$ pairs can be left out of the sufficient set. In general, the pairs included in a minimal sufficient set will be different depending on $a$ and $b$. Two specific examples of minimal sufficient sets will be presented later in this section.

Knowing a minimal sufficient set of $(i,j)$ pairs facilitates the solution of both the direct an inverse problems, because it reduces the number of pairs that need to be tried. In order to derive a general method to build a minimal sufficient set, it is insightful to consider two specific cases first, as will be presented in the following. It is also helpful to use the equivalent formulation of the direct problem referred to in Section~\ref{part: intro}, i.e.\ finding the maximum number of tiles touched by segments of length $\len$ \emph{or less}.

Thus each choice of $(i, j)$ implies a maximum of $i+j-1$ touched tiles on one hand; and sets a lower bound on the required length $\len$ on the other hand.

Consider first the case $a=b=1$. This is illustrated in Figure~\ref{fig: ijLS_1}. Note that in this and in the next figures the axes represent $ia$ and $jb$ (not $i$ and $j$). Each dashed diagonal line joins $(i, j)$ pairs with the same $i+j-1$. The radius of each arc represents the lower bound on $\len$ given by \eqref{eq: len > sqrt}, for certain $(i,j)$ pairs, which are marked with filled circles (these pairs form a minimal sufficient set, as will be seen).

\begin{figure}
\centering%
\includegraphics[width=.7\textwidth]{ijLS_1}%
\caption{Relationship between segment length and number of touched tiles with the width and height of the canonical rectangle, for $a=b$
}%
\label{fig: ijLS_1}%
\end{figure}%

For a given $\len \in \mathbb R^+$, the $(i,j)$ pairs that can be achieved with segments of length not exceeding $\len$ are, by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, those whose distance from $(2a,2b)$ is less than $\len$, i.e.~those delimited by the corresponding arc. It is clear from the figure that for $i+j-1$ odd, if some point $(i_1,j_1)$ with $|i_1-j_1|>1$ is achievable with the length bound $\len$, then necessarily there is a point $(i_0,i_0)$ with the same number of touched tiles $i+j-1$ (same diagonal line) that is also achievable with the same length restriction (because it has smaller distance to $(2a,2b)$). Similarly, for $i+j-1$ even, if a point $(i_1,j_1)$ with $|i_1-j_1|>1$ is achievable, then there is a point $(i_0,i_0+1)$ in the same diagonal that is also achievable with the same length restriction.

Thus for $a=b=1$, a minimal sufficient set is formed by the pairs $(i,j)$ with $j=i$ or $j=i+1$. Observe that due to symmetry, any point $(i_0,i_0+1)$ could be replaced by $(i_0+1,i_0)$. This illustrates that the minimal sufficient set is not unique in general.

The pairs in the minimal sufficient set are marked with filled circles in Figure~\ref{fig: ijLS_1}. Given a length, the maximum number of touched tiles will be achieved with one of those pairs, namely that in the diagonal line furthest from $(2a,2b)$ that is within the circle with radius $\len$ centered at that point.

As a second example, consider $a=1.35$, $b=1$. This is depicted in Figure~\ref{fig: ijLS_1p35}. In this case the minimal sufficient set does not follow a rule as simple as in the previous example.

\begin{figure}
\centering%
\includegraphics[width=.7\textwidth]{ijLS_1p35}%
\caption{Relationship between segment length and number of touched tiles with the width and height of the canonical rectangle, for $a=1.35b$
}%
\label{fig: ijLS_1p35}%
\end{figure}%

In general, a systematic way to obtain a minimal sufficient set is to gradually increase $\len$ starting from $0$, which gives an arc of increasing radius in the figure; and to take note of the first $(i,j)$ pair that is reached for each line of constant $i+j-1$. If two such pairs are reached at the same time, any of them can be used. In Figure~\ref{fig: ijLS_1p35} this rule produces the following pairs, in order: $(2,2)$, $(2,3)$, $(3,3)$, $(3,4)$, $(3,5)$, $(4,5)$, \ldots, as indicated by the filled circles.

Based on the above, an explicit procedure that produces a minimal sufficient set can be given. This set will be denoted as $\mss = \{(i_3,j_3), (i_4.j_4), \ldots\}$.

\begin{proposition}
\label{prop: min suff set}
Given grid parameters $a, b \in \mathbb R^+$, a minimal sufficient set $\mss$ can be obtained as follows. The chosen pair for $i+j-1=3$ is $(2,2)$. For each $\tiles \in \mathbb N, \tiles \geq 3$, let $(i_\tiles, j_\tiles)$ denote the pair chosen for $i+j-1=\tiles$. Then $(i_{\tiles+1}, j_{\tiles+1})$ is either $(i_\tiles+1, j_\tiles)$ or $(i_\tiles, j_\tiles+1)$, whichever minimizes $(i_{\tiles+1}-2)^2 a^2 + (j_{\tiles+1}-2)^2 b^2$; if both are minimizers $(i_\tiles+1, j_\tiles)$ is taken. Equivalently, $(i_{\tiles+1}, j_{\tiles+1})$ is $(i_\tiles, j_\tiles+1)$ if
\begin{equation}
\label{eq: cond for incrementing j}
j_\tiles < \frac{i_\tiles a^2}{b^2} - \frac{3a^2}{2b^2} + \frac 3 2,
\end{equation}
or $(i_\tiles+1, j_\tiles)$ otherwise.
\end{proposition}

\begin{proof}
For each $\tiles$, the pair $(i_\tiles,j_\tiles)$ of the minimal sufficient set should be chosen as that on the line $i+j-1=\tiles$ which minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$. This allows maximizing the sum $i+j-1$, and thus the number of touched tiles, for a given length restriction (direct problem); or minimizing the required lengths for a specified number of touched files (inverse problem).

The fact that $(i_t,j_t)$ minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$ among all pairs with $i+j-1=\tiles$ can be expressed as follows: for any $k \in \mathbb N$,
\begin{align}
\label{eq: chosen contiguous ineq g, 1}
(i_\tiles-2)^2 a^2 + (j_\tiles-2)^2 b^2 &\leq (i_\tiles-2-k)^2 a^2 + (j_\tiles-2+k)^2 b^2, \\
\label{eq: chosen contiguous ineq g, 2}
(i_\tiles-2)^2 a^2 + (j_\tiles-2)^2 b^2 &\leq (i_\tiles-2+k)^2 a^2 + (j_\tiles-2-k)^2 b^2.
\end{align}
To see that $(i_{\tiles+1},j_{\tiles+1})$ is either $(i_\tiles,j_\tiles+1)$ or $(i_\tiles+1,j_\tiles)$ it suffices to prove that, for any $k \in \mathbb N$,
\begin{align}
\label{eq: chosen contiguous ineq g+1, 1}
(i_{\tiles}-2)^2 a^2 + (j_{\tiles}-1)^2 b^2 &\leq (i_{\tiles}-2-k)^2 a^2 + (j_{\tiles}-1+k)^2 b^2, \\
\label{eq: chosen contiguous ineq g+1, 2}
(i_{\tiles}-1)^2 a^2 + (j_{\tiles}-2)^2 b^2 &\leq (i_{\tiles}-1+k)^2 a^2 + (j_{\tiles}-2-k)^2 b^2.
\end{align}
With reference to Figure~\ref{fig: ijLS_1p35}, consider for instance $(i_7,j_7) = (3,5)$. Graphically it can be seen that $(i_8,j_8)$ could only be $(3,6)$ or $(4,5)$, because for example $(2a,7b)$ is further from $(2a,2b)$ than $(3a,6b)$ is (this corresponds to \eqref{eq: chosen contiguous ineq g+1, 1} with $k=1$), and $(5a,4b)$ is also further from $(2a,2b)$ than $(4a,5b)$ is (this corresponds to \eqref{eq: chosen contiguous ineq g+1, 2} with $k=1$).

The inequality \eqref{eq: chosen contiguous ineq g+1, 1} can be rewritten as
\begin{equation}
\label{eq: chosen contiguous ineq g+1, 1b}
\begin{split}
& (i_{\tiles}-2)^2 a^2 + (j_{\tiles}-2)^2 b^2 + 2(j_{\tiles}-2)b^2 + b^2 \\
& \quad \leq (i_{\tiles}-2-k)^2 a^2 + (j_{\tiles}-2+k)^2 b^2 + 2(j_{\tiles}-2+k) b^2 + b^2.
\end{split}
\end{equation}
Clearly,
\begin{equation}
\label{eq: chosen contiguous ineq g+1, step}
2(j_{\tiles+1}-2)b^2 + b^2 < 2(j_{\tiles+1}-2+k) b^2 + b^2.
\end{equation}
Adding \eqref{eq: chosen contiguous ineq g, 1} and \eqref{eq: chosen contiguous ineq g+1, step} yields \eqref{eq: chosen contiguous ineq g+1, 1b}. Therefore \eqref{eq: chosen contiguous ineq g+1, 1} holds. A similar procedure can be used to establish \eqref{eq: chosen contiguous ineq g+1, 2}.

The condition for choosing $(i_t,j_t+1)$ over $(i_t+1,j_t)$ is that
\begin{equation}
\label{eq: chosen contiguous criterion 1}
(i_\tiles-2)^2 a^2 + (j_\tiles-1)^2 b^2 < (i_\tiles-1)^2 a^2 + (j_\tiles-2)^2 b^2.
\end{equation}
This can be expressed as
\begin{equation}
\label{eq: chosen contiguous criterion 2}
2(j_\tiles-2)b^2 + b^2 < 2(i_\tiles-2)a^2 + a^2.
\end{equation}
Rearranging terms, \eqref{eq: chosen contiguous criterion 2} is seen to be the same as \eqref{eq: cond for incrementing j}.
\end{proof}

The condition \eqref{eq: cond for incrementing j} for incrementing $j_\tiles$ in Proposition~\ref{prop: min suff set} can be expressed in terms of two bounding lines,
\begin{align}
\label{eq: upper bound, line}
j &= \frac{i a^2}{b^2} - \frac{3a^2}{2b^2} + \frac 5 2, \\
\label{eq: lower bound, line}
j &= \frac{i a^2}{b^2} - \frac{5a^2}{2b^2} + \frac 3 2.
\end{align}
All pairs $(i,j)$ in $\mss$ are strictly below the upper line \eqref{eq: upper bound, line} and above or on the lower line \eqref{eq: lower bound, line}. This is shown in Figure~\ref{fig: pairs_bounds}, using three different pairs of grid parameters $a$, $b$ as examples. Given $(i_\tiles,j_\tiles) \in \mss$, the next pair $(i_{\tiles+1},j_{\tiles+1})$ is obtained by incrementing $j$ if that results in a point below \eqref{eq: upper bound, line}. Else $j$ cannot be increased and $i$ is incremented instead, and the new pair is guaranteed to be above or on \eqref{eq: lower bound, line}.

\begin{figure}
\centering%
\subfigure[$a = 1$, $b = 1$]{%
\label{fig: pairs_bounds_1}%
\includegraphics[height=.24\textheight]{pairs_bounds_1}% *!*
}\hfill%
\subfigure[$a = 1.35$, $b = 1$]{%
\label{fig: pairs_bounds_1p35}%
\includegraphics[height=.24\textheight]{pairs_bounds_1p35}% *!*
}\hfill%
\subfigure[$a = \sqrt{2}$, $b = 1$]{%
\label{fig: pairs_bounds_sqrt2}%
\includegraphics[height=.24\textheight]{pairs_bounds_sqrt2}% *!*
}%
\caption{Minimal sufficient set $\mss$ and bounding lines
}%
\label{fig: pairs_bounds}%
\end{figure}%

The construction of $\mss$ in Proposition~\ref{prop: min suff set} computes the pairs $(i_\tiles,i_\tiles)$ \emph{sequentially}: $(i_3,i_3), \ldots, (i_{\tiles-1},i_{\tiles-1})$ need to be previously computed in order to obtain $(i_\tiles,i_\tiles)$. Proposition~\ref{prop: min suff set, form} avoids this by providing a direct formula. Its derivation is based on finding the intersection point of \eqref{eq: lower bound, line} with  $i+j-1=\tiles$ and appropriately rounding to integer coordinates in order to satisfy \eqref{eq: upper bound, line}.
% The proof method is briefly described here because in a later Theorem it is said (in the text, not in the proof) that the method is similar to this

\begin{proposition}
\label{prop: min suff set, form}
Given $a, b \in \mathbb R^+$ and $\tiles \in \mathbb N, \tiles \geq 3$, the pair $(i_\tiles,j_\tiles)$ from the set $\mss$ defined in Proposition~\ref{prop: min suff set} can be computed as
\begin{align}
\label{eq: min suff set, form, i}
i_\tiles &= \left\lfloor \frac {(2\tiles-1)b^2+5a^2}{2(a^2+b^2)} \right\rfloor, \\
\label{eq: min suff set, form, j}
j_\tiles &= \left\lceil \frac {2\tiles a^2 + 3(b^2-a^2)}{2(a^2+b^2)} \right\rceil. 
\end{align}
\end{proposition}

\begin{proof}
Let $(\isolr, \jsolr)$ denote the intersection of \eqref{eq: lower bound, line} and the line $i+j-1=\tiles$. This is easily seen to be
\begin{align}
\label{eq: min suff set, form, isolr}
\isolr &= \frac {(2\tiles-1)b^2+5a^2}{2(a^2+b^2)}, \\
\label{eq: min suff set, form, jsolr}
\jsolr &= \frac {2\tiles a^2 + 3(b^2-a^2)}{2(a^2+b^2)}. 
\end{align}
This point is shown with a square marker in Figure \ref{fig: ijLS_1p35_detail_inverse}, which also shows the elements of $\mss$ as filled circles, as well as the two bounding lines. The line \eqref{eq: upper bound, line} is \eqref{eq: lower bound, line} shifted one step left in $i$ and one step up in $j$, i.e.~it is shifted by a diagonal step. Since all points of $\mss$ are between these two lines or on the lower line, the only possibility for $(i_\tiles,j_\tiles)$, as illustrated in the figure, is to be the point with integer coordinates closest to $(\isolr,\jsolr)$ along the line $i+j-1=\tiles$ in the left, upwards direction. This gives \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}.

\begin{figure}%
\centering%
\includegraphics[width=.7\textwidth]{ijLS_1p35_detail_inverse}%
\caption{Obtaining $(\isoli,\jsoli)$ in Proposition~\ref{prop: min suff set, form}. Example with $a=1.35, b=1$.%
}%
\label{fig: ijLS_1p35_detail_inverse}%
\end{figure}%
\end{proof}

For $a^2/b^2$ arbitrary, the number of pairs with the same $i$ coordinate in the set $\mss$ is in general irregular, because the line \eqref{eq: lower bound, line} does not follow a ``natural'' direction of the grid. This happens for instance in Figure~\ref{fig: pairs_bounds_1p35}, where the number of pairs for the first $i$ values equals either $2$ or $3$ without a clear pattern.\footnote{Strictly, there is a periodic pattern whenever $a^2/b^2$ is rational, which is the case in Figure~\ref{fig: pairs_bounds_1p35}. However, this is not easily discernible unless $a^2/b^2$ is a ratio of small numbers.}
On the other hand, a simple pattern arises for $a^2/b^2 \in \mathbb N$, as seen in Figures~\ref{fig: pairs_bounds_1} and \ref{fig: pairs_bounds_sqrt2}.

A segment whose canonical rectangle has normalized width $i$ and height $j$ is oriented with approximate slope $jb/(ia)$ with respect to the $x$ axis (See Figure~\ref{fig: len ineq i j}); and the approximation becomes better for greater segment lengths. From \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} it can be seen that the positions of the pairs $(i,j) \in \mss$ have $j/i \approx a^2/b^2$ for large $i, j$. Therefore the optimal slope for long segments is approximately $a/b$. This formalizes the intuition that to maximize the number of touched tiles, the segment should follow a direction along which the perceived ``length'' of the tile is smaller.

\section{Arbitrary grid with real-valued lengths}
\label{part: arbitrary grid, real lengths}

For a grid with parameters $a, b \in \mathbb R^+$, the maximum number $\tiles$ of touched tiles for a given real-valued length $\len$ can be represented by a function $\funt: \mathbb R^+ \to \mathbb N$ such that $\tiles = \funt(\len)$. Similarly, for the inverse problem a function $\funl: \mathbb N \to \mathbb R^+$ can be defined such that $\funl(\tiles)$ gives the infimum length of all segments that touch at least $\tiles$ tiles.

The function $\funt$ can be obtained from $\funl$ by noting that the maximum number of tiles that can be touched by a segment of length $\len$ is the index of the largest term of the sequence $\funl(\tiles)$ that is less than $\len$:
\begin{equation}
\label{eq: funt funl}
\funt(\len) = \max \{\tiles \in \mathbb N \st \funl(\tiles)<\len\}.
\end{equation}
Conversely, $\funl$ can be obtained from $\funt$ as
\begin{equation}
\label{eq: funl funt}
\funl(\tiles) = \inf\{\len \in \mathbb R^+ \st \funt(\len) \geq \tiles\}.
\end{equation}

For arbitrary $a, b \in \mathbb R^+$, the functions $\funt$ and $\funl$ can be computed using an iterative procedure, which exploits the fact that the pairs $(i_3,j_3), (i_4,j_4), \ldots$ of $\mss$ are sorted by increasing $i+j-1$, and also by increasing $(i-2)^2 a^2 + (j-2)^2 b^2$. Namely, for $\funt$ the following procedure yields the solution: generate successive pairs to find the last one, $(i_\tiles,j_\tiles)$, that satisfies \eqref{eq: len > sqrt}; then $\funt(\len) = \tiles$. For $\funl$ the analogous method gives a direct formula. In addition, it is possible to obtain a direct formula also for $\funt$ using an approach similar to that used in Proposition~\ref{prop: min suff set, form}. These formulas are given in Theorems \ref{theo: funt, form} and \ref{theo: funl, form}.

% ***Should i-1, j-2 be used instead of i, j? I don't think so. The formulas would be simplified a little, but their interpretation would be more difficult
\begin{theorem}
\label{theo: funt, form}
For $a, b \in \mathbb R^+$, $a \geq b$ and $\len \in \mathbb R^+$,
\begin{equation}
\label{eq: theo: funt, form; funt}
\funt(\len) = \isoli+\jsoli-1
\end{equation}
with
\begin{align}
\label{eq: theo: funt, form; i}
\isoli &= \left\lceil \frac{a + b \Re \sqrt{4\len^2/(a^2+b^2)-1}}{2a} \right\rceil + 1, \\
\label{eq: theo: funt, form; j}
\jsoli &= \left\lceil \frac{\sqrt{\len^2-(\isoli-2)^2a^2}}{b} \right\rceil + 1.
\end{align}
The function $\funt$ is piecewise constant and left-continuous, with unit-height jumps. A jump occurs at $\len$ if and only if $\len = \funl(\tiles)$ for some $\tiles \in \mathbb N, \tiles \geq 4$; and then $\funt(\len) = \tiles-1$, $\lim_{h \rightarrow 0+} \funt(\len+h) = \tiles$.
\end{theorem}

\begin{proof}
The general idea is similar to that used in Proposition~\ref{prop: min suff set, form}: find the intersection $(\isolr, \jsolr)$ of the line \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ for $i, j \geq 2$, and from that obtain the actual pair $(\isoli, \jsoli)$ of integer numbers which yields the solution.

Specifically, assuming first that \eqref{eq: lower bound, line} holds, the largest $i$ that can be achieved with lengths not exceeding $\len$ is obtained, as given by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}. This value, $\isoli$, is computed by rounding down $\isolr$. Then, assuming that $i = \isoli$, the largest $j$ allowed by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, $\jsoli$, is obtained. As will be seen, in some cases the resulting $(\isoli,\jsoli)$ is in $\mss$, and in other cases it is not. However, in either case $(\isoli,\jsoli)$ has the largest $i+j-1$ sum that can be achieved with segments of length up to $\len$. Thus the desired result is $\isoli+\jsoli-1$.

The two possibilities are illustrated in Figure~\ref{fig: ijLS_1p35_detail} for $a=1.35, b=1$. In each case, the arc displayed in the graph is centered at $(2a,2b)$ and has radius $\len$. The inner region defined by the arc contains all $(i,j)$ pairs that are achievable according to Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}. As in previous figures, filled circles represent $(i,j)$ pairs that are in $\mss$, and empty circles are those that do not belong to $\mss$. The solid line is \eqref{eq: lower bound, line}. The intersection point $(\isolr, \jsolr)$ is displayed with a square marker.

\begin{figure}
\centering%
\subfigure[$\len = 3.1$: $(\isoli,\jsoli) \in \mss$]{%
\label{fig: ijLS_1p35_detail_in}%
\includegraphics[width=.7\textwidth]{ijLS_1p35_detail_in}%
}\\%\hfill%
\subfigure[$\len = 3.7$: $(\isoli,\jsoli) \notin \mss$]{%
\label{fig: ijLS_1p35_detail_notin}%
\includegraphics[width=.7\textwidth]{ijLS_1p35_detail_notin}%
}%
\caption{Obtaining $(\isoli,\jsoli)$ in Theorem~\ref{theo: funt, form}. Examples with $a=1.35, b=1$.
}%
\label{fig: ijLS_1p35_detail}%
\end{figure}%

The point $(\isolr, \jsolr)$ can be obtained as a solution of the equation system
\begin{align}
\label{eq: icont jcont 1}
(\isolr-2)^2 a^2 + (\jsolr-2)^2 b^2 &= \len^2, \\
\label{eq: icont jcont 2}
\jsolr = \frac{\isolr a^2}{b^2} - \frac{5a^2}{2b^2} + \frac 3 2.
\end{align}
Substituting \eqref{eq: icont jcont 2} into \eqref{eq: icont jcont 1} yields a quadratic equation for $\isolr$,
\begin{equation}
\label{eq: quadratic, i}
4a^2(\isolr-2)^2 - 4a^2(\isolr-2) + a^2+b^2-\frac{4\len^2 b^2}{a^2+b^2},
\end{equation}
which can have zero, one or two real-valued solutions (this respectively means that the line is exterior, tangent or secant to the circle \eqref{eq: icont jcont 1}). Only the solution with $\isolr, \jsolr \geq 2$, if any, is of interest. This solution can only exist as the one with largest $\isolr$ value when there are real-valued solutions, and it is obtained as
\begin{equation}
\label{eq: icont jcont 3}
\isolr = \frac{a + b \sqrt{4\len^2/(a^2+b^2)-1}}{2a} + 2.
\end{equation}

The case with no real-valued solutions to \eqref{eq: quadratic, i} corresponds to $\len < \sqrt{a^2+b^2}/2$. Since $a \geq b$, this implies that $\len < a$. Thus any achievable $(i,j)$ pair, if any, will have $i=2$. Therefore in this case $\isoli$ should be set to $2$; and then the maximum achievable $\jsoli$ will be obtained from \eqref{eq: icont jcont 2}.

If \eqref{eq: quadratic, i} has two real-valued solutions or one real-valued double solution, which is the case when $\len \geq \sqrt{a^2+b^2}/2$, the largest solution will have $\jsolr < 2$ if $\isolr < (b^2/a^2+5)/2$, as can be seen setting $\jsolr=2$ in \eqref{eq: icont jcont 2}; or equivalently if $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, as can be seen substituting $\jsolr < 2$ and $\isolr < (b^2/a^2+5)/2$ in \eqref{eq: icont jcont 1}. Since $\isolr < (b^2/a^2+5)/2 < 3$ for $a \geq b$, only pairs with $i=2$ are achievable in this case again, and thus the solution $\isoli$ should also be $2$.

Lastly, for $\len \geq (a^2+b^2) / (2a)$ the expressions \eqref{eq: icont jcont 2} and \eqref{eq: icont jcont 3} give $\isolr, \jsolr \geq 2$, and $\isoli$ should be taken as the greatest integer less than $\isolr$, i.e.~$\lceil \isolr \rceil-1$.

The three cases are unified by taking the real part of \eqref{eq: icont jcont 3} and computing $\isoli = \lceil \isolr \rceil-1$, as is easily checked. This gives \eqref{eq: theo: funt, form; i}. Once $\isoli$ is known, $\jsoli$ is computed, according to \eqref{eq: theo: funt, form; j}, as the greatest integer such that $(\isoli,\jsoli)$ is within the arc determined by $\len$. This ensures that $(\isoli,\jsoli)$ is achievable with lengths less than $\len$.

There are two possibilities for the point $(\isoli,\jsoli)$, as stated at the outset. These are illustrated in Figures~\ref{fig: ijLS_1p35_detail_in} and \ref{fig: ijLS_1p35_detail_notin} respectively. The first possibility is that $(\isoli,\jsoli) \in \mss$ (Figure~\ref{fig: ijLS_1p35_detail_in}). Then, by construction  $(\isoli,\jsoli)$ maximizes $i+j-1$ among all achievable pairs of $\mss$, and is therefore optimal.

The second possibility is that $(\isoli,\jsoli) \notin \mss$ (Figure~\ref{fig: ijLS_1p35_detail_notin}). This happens when the point from $\mss$ that has $i = \isoli$ ($(4,5)$ in the figure) is outside the arc, i.e.~it would require a length greater than $\len$. The selected point  $(\isoli,\jsoli)$ ($(4,4)$ in the figure), however, has the same $i+j-1$ sum as the point from $\mss$ that ``should'' be used, which is $(\isoli-1,\jsoli+1)$ ($(3,5)$ in the figure); and therefore gives the same result. This is always the case, because  $(\isoli,\jsoli+1) \in \mss$ (it is above or on the bounding line) and $(\isoli,\jsoli) \notin \mss$ (it is below the line), and due to how $\mss$ has been constructed, this implies that $(\isoli-1,\jsoli+1) \in \mss$ and $(\isoli-1,\jsoli+k) \notin \mss$ for $k =2, 3, \ldots$.  Therefore, $(\isoli,\jsoli)$ is achievable and maximizes $i+j-1$, which implies that $\isoli+\jsoli-1$ is the desired solution.

Therefore, whether $(\isoli,\jsoli) \in \mss$ or $(\isoli,\jsoli) \notin \mss$, equations \eqref{eq: theo: funt, form; i} and \eqref{eq: theo: funt, form; j} give $\isoli+\jsoli-1$ equal to $\funt(\len)$, as claimed.

It should be noted that for the specific case that $a^2/b^2 = 2k-1$, $k \in \mathbb N$ the lower bounding line \eqref{eq: lower bound, line} becomes $j = i(2k-1)-5k+4$, which gives integer $j$ for integer $i$. This means that the bound is actually achieved for all pairs in $\mss$ (see for example Figure~\ref{fig: pairs_bounds_1}), and the case $(\isoli,\jsoli) \notin \mss$ never occurs. Thus, when $a^2/b^2$ is an odd integer the obtained $(\isoli,\jsoli)$ is guaranteed to be in $\mss$, whereas for general $a^2/b^2$ either $(\isoli,\jsoli)$ or $(\isoli-1,\jsoli+1)$ are in in this set.
% ^ If this is needed outside of the proof, it can be included at the end of the theorem statement

As for the properties of $\funt$, it follows from \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j} that it is piecewise constant and left-continuous. From the procedure described in the previous paragraphs for obtaining $(\isoli, \jsoli)$ it is clear that $\isoli+\jsoli-1$ increases in steps of $1$ when $\len$ is increased continuously; that is, $\funt$ has jumps of unit value.

Consider an arbitrary $\len$ such that for some $\tiles \in \mathbb N, \tiles \geq 4$
\begin{equation}
\label{eq: theo funt, form: prop 1}
\funl(\tiles)=\len.
\end{equation}
To see that $\funt$ has a jump at $\len$, assume for the sake of contradiction that $\funt$ is continuous, therefore constant, at that point. Then there is $\epsilon > 0$ for which $\funt(\len+\epsilon) = \tiles = \funt(\len-\epsilon)$. This means that there exists a segment with length $\len-\epsilon$ that touches $\tiles$ tiles, and thus $\funl(\tiles) \leq \len-\epsilon < \len$, in contradiction with \eqref{eq: theo funt, form: prop 1}. Therefore $\funt$ is discontinuous, from the right, at $\len$. By definition of $\funl$, \eqref{eq: theo funt, form: prop 1} implies that $\funt(\len) < \tiles$, and that there exists $\epsilon > 0$ such that $\funt(\len+h) = \tiles$ for $0 < h < \epsilon$. This means that
\begin{align}
\label{eq: theo funt, form: prop 2}
\funt(\len) &< \tiles, \\
\label{eq: theo funt, form: prop 3}
\lim_{h \rightarrow 0+} \funt(\len+h) &= \tiles,
\end{align}
that is, $\funt$ has a jump at $\len$. In addition, since the jump has unit height, it follows from \eqref{eq: theo funt, form: prop 2} and \eqref{eq: theo funt, form: prop 3}  that $\funt(\len) = \tiles-1$.

Conversely, assume that \eqref{eq: theo funt, form: prop 2} and \eqref{eq: theo funt, form: prop 3} hold for some $\len \in \mathbb R^+$, $\tiles \in \mathbb N$. From \eqref{eq: theo funt, form: prop 2} it follows that  $\funl(\tiles) \geq \len$. On the other hand, \eqref{eq: theo funt, form: prop 3} implies that $\funl(\tiles) \leq \len$. Thus $\funl(\tiles)=\len$.
\end{proof}

Although Theorem~\ref{theo: funt, form} is valid for $a \geq b$, the result could also be applied for $a < b$ by simply swapping the values of $a$ and $b$. In other words, \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j}
% *!*
can be used for any grid if $a$, $b$ are interpreted as the largest and smallest sides of a tile, respectively.

\begin{theorem}
\label{theo: funl, form}
For $a, b \in \mathbb R^+$ and $\tiles \in \mathbb N$,
\begin{equation}
\label{eq: theo: funl, form; funl}
\funl(\tiles) = \sqrt{(\isoli-2)^2 a^2 + (\jsoli-2)^2 b^2}
\end{equation}
with
\begin{align}
\label{eq: theo: funl, form; i}
\isoli &= \max\left\{ \left\lfloor \frac {(2\tiles-1)b^2+5a^2}{2(a^2+b^2)} \right\rfloor, 2 \right\}, \\
\label{eq: theo: funl, form; j}
\jsoli &= \max\left\{ \left\lceil \frac {2\tiles a^2 + 3(b^2-a^2)}{2(a^2+b^2)} \right\rceil, 2 \right\}.
\end{align}
This function is monotone increasing for $\tiles \geq 3$.
\end{theorem}

\begin{proof}
The $(i_\tiles,j_\tiles)$ pair in the minimal sufficient set $\mss$ corresponds to a maximum of $\tiles$ touched tiles. By construction of this set, any segment that touches $\tiles$ tiles must have length greater than $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$.

For $\tiles \geq 3$ the expressions \eqref{eq: theo: funl, form; i} and \eqref{eq: theo: funl, form; j} coincide with \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}. Therefore \eqref{eq: theo: funl, form; funl} gives the desired result.

For $\tiles \in \{1, 2\}$ both \eqref{eq: theo: funl, form; i} and \eqref{eq: theo: funl, form; j} equal $2$, and \eqref{eq: theo: funl, form; funl} gives $0$, which is the correct result.

By Theorem~\ref{theo: funt, form}, $\funt$ has unit-height jumps at the values $\funl(\tiles)$, $\tiles \in \mathbb N, \tiles \geq 4$. This implies that $\funl$ is monotone increasing for $\tiles \geq 3$.
\end{proof}

Theorems~\ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the two problems stated in Section~\ref{part: intro}; they also provide a way to actually position a segment of length slightly greater than $\len$ or $\funl(\tiles)$ so that it touches $\tiles$ or $\funt(\len)$ tiles. Namely, the segment should have its endpoints respectively in the interior of two tiles separated $\isoli-1$ steps horizontally and $\jsoli-1$ steps vertically, with its exact position and orientation  adjusted to avoid any grid points.

It is interesting to consider the following particular cases: $\len \gg a,b$; $a \gg b$; and $a=b$. Regarding the first, from \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j}
% *!*
and from \eqref{eq: theo: funl, form; funl}--\eqref{eq: theo: funl, form; j}
% *!*
it is seen that for long segments the number of touched tiles and the segment length are approximately proportional, with
\begin{equation}
\label{eq: asympt}
\lim_{\len \rightarrow \infty} \frac{\funt(\len)}{\len}
= \lim_{\tiles \rightarrow \infty} \frac{\tiles}{\funl(\tiles)}
= \sqrt{1/a^2 + 1/b^2}.
\end{equation}

As for $a \gg b$, in this case the optimal canonical rectangle has $\isoli = 2$, and $\jsoli$ as large as allowed by $\len$ (direct problem) or as required by $\tiles$ (inverse problem), corresponding to an almost vertical segment. Obviously, for $a \gg b$ the length of the segment is best invested in increasing the number of tiles traversed vertically (but the segment should be slightly tilted to cross a vertical edge), and the asymptotic value of \eqref{eq: asympt} is approximately $1/b$.

For $a=b$, either from symmetry considerations or particularizing the formulas in this section it stems that the optimal orientation of the segment is close to $45^\circ$. This case will be dealt with in Section~\ref{part: square grid, real lengths}, as it lends itself to somewhat simplified formulas.

Figure~\ref{fig: funt funl examples} shows $\funt$ and $\funl$ for several pairs of grid parameters $a$, $b$. The graphs illustrate some of the observations of the previous paragraphs. Indeed, the asymptotic slope in Figure~\ref{fig: funt examples}, or the inverse of the asymptotic slope in Figure~\ref{fig: funl examples}, is approximately $\sqrt{2}$ for $a, b=1$, and $1/b$ for the case $a=5,b=1$, and even for $a=5,b=1.5$ and $a=10,b=3$. Comparing the latter two cases it is also seen that scaling $a$, $b$ and $\len$ by the same factor does not alter $\funt(\len)$; or scaling $a$, $b$ results in $\funl(\tiles)$ being scaled by the same factor.

\begin{figure}
\centering%
\subfigure[Function $\funt$]{%
\label{fig: funt examples}%
\includegraphics[width=.49\textwidth]{funt_examples}%
}\hfill%
\subfigure[Function $\funl$]{%
\label{fig: funl examples}%
\includegraphics[width=.49\textwidth]{funl_examples}%
}%
\caption{Functions $\funt$ and $\funl$ for several pairs $a$, $b$
}%
\label{fig: funt funl examples}%
\end{figure}%

\section{Square grid with real-valued lengths}
\label{part: square grid, real lengths}

Particularizing the results in Section~\ref{part: arbitrary grid, real lengths} to a square grid, $a=b$, obviously yields simpler formulas.

\begin{corollary}
\label{cor: funt, sq, form}
For a square grid with $a \in \mathbb R^+$, and for $\len \in \mathbb R^+$,
\begin{equation}
\label{eq: cor: funt, sq, form}
\funt(\len) = \isoli+\jsoli-1
\end{equation}
with
\begin{align}
\label{eq: cor: funt, sq, form; i}
\isoli &= \left\lceil \frac{1 + \Re \sqrt{2\len^2/a^2-1}}{2} \right\rceil + 1, \\
\label{eq: cor: funt, sq, form; j}
\jsoli &= \left\lceil \sqrt{\len^2/a^2-(\isoli-2)^2} \right\rceil + 1,
\end{align}
\end{corollary}

\begin{corollary}
\label{theo: funl, sq, form}
For a square grid with $a \in \mathbb R^+$, and for $\tiles \in \mathbb N$,
\begin{equation}
\label{eq: theo: funt, sq, form; funl}\\
\frac{\funl(\tiles)}{a} = \begin{cases}
\displaystyle
0 & \text{for } \tiles =1, 2 \\[1.4mm]
\displaystyle
\frac{\tiles-3}{\sqrt{2}} & \text{for } \tiles \text{ odd, } \tiles \geq 3 \\[4.5mm]
\displaystyle
\sqrt{\frac{(\tiles-3)^2+1} {2}} & \text{for } \tiles \text{ even, } \tiles \geq 3.
\end{cases}
\end{equation}
\end{corollary}

Furthermore, an even simpler formula can be obtained for $\funt$. Its derivation uses a variation of the minimal sufficient set $\mss$ which is  more suited to this case.

\begin{theorem}
\label{theo: funt, sq, form}
For a square grid with $a \in \mathbb R^+$, and for $\len \in \mathbb R^+$,
\begin{equation}
\label{eq: theo: funt, sq, form; funt}
\funt(\len) = \isoli+\jsoli-1
\end{equation}
with
\begin{align}
\label{eq: theo: funt, sq, form; i}
\isoli &= \left\lceil \frac{\len}{a \sqrt{2}} \right\rceil + 1, \\
\label{eq: theo: funt, sq, form; j}
\jsoli &= \left\lceil \sqrt{\len^2/a^2-(\isoli-2)^2} \right\rceil + 1.
\end{align}
\end{theorem}

\begin{proof}
For $a=b$ the set $\mss$ from Proposition~\ref{prop: min suff set} consists of points of the form $(i,i)$ and $(i+1,i)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}. By symmetry, replacing each point $(i+1,i)$ by $(i,i+1)$ gives a set $\mss'$ which is also minimal sufficient. For this new set, the lower bounding line \eqref{eq: lower bound, line} can be replaced by the simpler $j=i$. The same approach used in the proof of Theorem~\ref{theo: funt, form} can be used, but using this line. Thus $(\isolr, \jsolr)$ is obtained from
\begin{align}
(\isolr-2)^2 a^2 + (\jsolr-2)^2 a^2 &= \len^2, \\
\jsolr &= \isoli,
\end{align}
which gives
\begin{equation}
\label{eq: theo: funt, sq, form; proof 1}
\isolr = \frac{\len}{a \sqrt{2}} + 2.
\end{equation}
The pair $(\isoli,\jsoli)$ resulting from \eqref{eq: theo: funt, sq, form; proof 1} is in $\mss'$, and is given in \eqref{eq: theo: funt, sq, form; i} and \eqref{eq: theo: funt, sq, form; j}.
\end{proof}

\section{Unit square grid with integer lengths}
\label{part: unit square grid, integer lengths}

A natural variation of the two problems introduced in Section~\ref{part: intro} is to consider $a=b=1$, with the additional restriction that the segment length can only be a positive integer (equivalently, a square grid is considered with step $a$ and the segment length can only be an integer multiple of $a$).

The \emph{direct problem} in this setting corresponds to the restriction of the function $\funt$ to $\mathbb N$. This will be denoted as a function $\funti: \mathbb N \to \mathbb N$ for greater clarity, although obviously $\funti(\leni) = \funt(\leni)$ for all $\leni \in \mathbb N$. The sequence $\funti(\leni)$, $\leni \in \mathbb N$ takes values $3, 5, 7, 8, 9, \ldots$, and is depicted in Figure~\ref{fig: funti}. This is A346232 in the On-Line Encyclopedia of Integer Sequences \cite{OEIS_unitsq_int_dir}. For this sequence, a simpler expression can be obtained than those resulting from Section~\ref{part: square grid, real lengths} with $a=1$, and the following properties hold.

\begin{figure}%
\centering%
\subfigure[Sequence $\funti$]{%
\label{fig: funti}%
\includegraphics[width=.49\textwidth]{funti}%
}\hfill%
\subfigure[Sequence $\funli$]{%
\label{fig: funli}%
\includegraphics[width=.49\textwidth]{funli}%
}%
\caption{Sequences $\funti$ and $\funli$
}%
\label{fig: funti funti}%
\end{figure}%

\begin{theorem}
\label{theo: funti}
For $\leni \in \mathbb N$,
% $\funti(\leni)$ can be computed as
\begin{equation}
\label{eq: funti leni}
\funti(\leni) = \left\lfloor \sqrt{2\leni^2-2} \right\rfloor + 3.
\end{equation}
In addition,
\begin{enumerate}
\item
This sequence is increasing, with $\funti(\leni+1)-\funti(\leni) \in \{1, 2\}$.
\item
There can be no more than $2$ consecutive increments equal to $1$.
\item
Increments equal to $2$ always appear isolated, except at the initial sequence terms $3, 5, 7$.
\end{enumerate}
\end{theorem}

\begin{proof}
***The derivation of \eqref{eq: funti leni} in CGCC can probably be adapted. $i$, $j$ there are $i-2$, $j-2$ here. (It cannot be expressed with $\lceil \cdot \rceil$ instead of $\lfloor \cdot \rfloor$, which would be more ``aesthetically aligned'' with previous expressions, because for example $\len=3$ makes $2\len^2-2$ a square).

%***Otherwise, perhaps we can prove the formula for the inverse problem first (starting from the one for real lenghts), and derive this from that

In order to prove that $\funti(\leni+1)-\funti(\leni) \in \{1,2\}$, consider the function $f(x) = \sqrt{2x^2-2}$ for $x \in \mathbb R$, $x>1$. Its first derivative is
\begin{equation}
f'(x) = \frac {\sqrt{2} \, x} {\sqrt{x^2-1}},
\end{equation}
and its second derivative is easily seen to be negative. Therefore $f'(x)$ can be bounded for $x \geq 3$ as
\begin{equation}
\label{eq: der bounds}
\lim_{x \rightarrow \infty} f'(x) = \sqrt{2} < f'(x) < f'(3) = 3/2.
\end{equation}
For $\leni \geq 2$, by the mean value theorem \cite[section~5.3]{Abbott15}, when $\leni$ is increased to $\leni+1$ the term $\sqrt{2\leni^2-2}$ in \eqref{eq: funti leni} has an increment that equals $f'(x)$ for some $\leni < x < \leni+1$. Therefore
\begin{equation}
\label{eq: der bounds 2}
\sqrt{2} < \sqrt{2(\leni+1)^2-2} - \sqrt{2\leni^2-2} < 3/2.
\end{equation}
Since $1 < \sqrt{2}$ and $3/2 < 2$, \eqref{eq: der bounds 2} implies that $\funti(\leni+1)-\funti(\leni)$ can only take the values $1$ or $2$ for $\leni \geq 3$. In addition, $\funti(2)-\funti(1) = \funti(3)-\funti(2) = 2$, and thus the result holds for all $\leni \in \mathbb N$.

Using the first bound in \eqref{eq: der bounds 2} three times,
\begin{equation}
3\sqrt{2} < \sqrt{2(\leni+3)^2-2} - \sqrt{2\leni^2-2}.
\end{equation}
Considering that $4 < 3\sqrt{2}$, this implies that $\funti(\leni+3)-\funti(\leni) \geq 4$ for $\leni \geq 3$. Therefore at least one of the three increments from $\funti(\leni)$ to $\funti(\leni+3)$ is $2$. Since $\funti(2)-\funti(1) = \funti(3)-\funti(2) = 2$, the result holds for all $\leni \in \mathbb N$.

Similarly, using the second bound in \eqref{eq: der bounds 2} twice,
\begin{equation}
\sqrt{2(\leni+2)^2-2} - \sqrt{2\leni^2-2} < 3,
\end{equation}
which implies that $\funti(\leni+2)-\funti(\leni) \leq 3$ for $\leni \geq 3$. Therefore the two increments $\funti(\leni+1)-\funti(\leni)$ and $\funti(\leni+2)-\funti(\leni+1)$ cannot both be $2$ for $\leni \geq 3$.
\end{proof}

The \emph{inverse problem} with integer-length segments can be formulated as follows: given $\tiles \in \mathbb N$, find the \emph{minimum} integer length that allows touching at least $\tiles$ tiles. Observe that in this case, unlike with real-valued lengths, there is indeed a minimum length, because every subset of $\mathbb N$ has a minimum. This can be expressed as a function $\funli: \mathbb N \to \mathbb N$:
\begin{equation}
\label{eq: funli funti}
\funli(\tiles) = \min\{\leni \in \mathbb N \st \funti(\leni) \geq \tiles\},
\end{equation}
which is related to the function $\funl$ corresponding to real-valued lengths by
\begin{equation}
\label{eq: funli funl}
\funli(\tiles) = \lfloor\funl(\tiles)\rfloor + 1.
\end{equation}
The converse to \eqref{eq: funli funti} is (compare to \eqref{eq: funt funl}):
\begin{equation}
\label{eq: funti funli}
\funti(\leni) = \max \{\tiles \in \mathbb N \st \funli(\tiles) \leq \leni\}.
\end{equation}
In view of \eqref{eq: funli funti} and \eqref{eq: funti funli}, $\funti$ and $\funli$ can be considered as ``pseudo-inverse'' sequences of each other.
% ***Is this a good name? Is there an established name for this relationship?

The sequence $\funli(\tiles)$, $\tiles \in \mathbb N$ can be computed using \eqref{eq: theo: funt, sq, form; funl} and \eqref{eq: funli funl}, and has initial values $1, 1, 1, 2, 2, 3, 3, 4, 5 \ldots$, as shown in Figure~\ref{fig: funli}. This is A**** in the On-Line Encyclopedia of Integer Sequences \cite{***}. However, a simpler expression can be obtained from \eqref{eq: funti leni} and \eqref{eq: funli funti}. This is established by the next theorem, which also states some properties of $\funli$, parallel to those of $\funti$.

\begin{theorem}
\label{theo: funli}
For $\tiles \in \mathbb N$,
% $\funli(\tiles)$ can be computed as
\begin{equation}
\label{eq: funli}
\funli(\tiles) = \begin{cases}
\displaystyle
1 & \text{for } \tiles \leq 3 \\[1.3mm]
\displaystyle
\left \lceil \sqrt{\frac{(\tiles-3)^2} 2 + 1} \ \right \rceil & \text{for } \tiles \geq 4.
\end{cases}
\end{equation}
In addition,
\begin{enumerate}
\item
This sequence is non-decreasing. Except for the initial run of $3$ equal values, it is formed by runs of $1$ or $2$ equal values, with an increment of $1$ between consecutive runs.
\item
There can be no more than $3$ different consecutive terms.
\item
A run of $2$ equal values always has $2$ different terms before and $2$ different terms after the run, except for the initial terms $1, 1, 1, 2, 2, 3, 3$.
\end{enumerate}
\end{theorem}

\begin{proof}
Using \eqref{eq: funti leni}, the inequality $\funti(\leni) \geq \tiles$ in \eqref{eq: funli funti} is written as
\begin{equation}
\left\lfloor \sqrt{2\leni^2-2} \right\rfloor + 3 \geq \tiles.
\end{equation}
Since the right-hand side is an integer, this is equivalent to
\begin{equation}
\label{eq: theo funli ineq}
\sqrt{2\leni^2-2} \geq \tiles-3.
\end{equation}
For $\tiles \geq 4$, taking squares and rearranging gives
\begin{equation}
\leni \geq \left \lceil \sqrt{\frac{(\tiles-3)^2} 2 + 1} \ \right \rceil,
\end{equation}
which combined with \eqref{eq: funli funti} yields the second part of \eqref{eq: funli}. The first part results from noting that $\tiles-3 \leq 0$ for $\tiles \leq 3$, and thus $\leni=1$ satisfies \eqref{eq: theo funli ineq}.

The stated properties for $\funli$ follow directly from those of $\funti$ established by Theorem~\ref{theo: funti}.
\end{proof}

\bibliographystyle{plain}
\bibliography{bibliogr/lista_local}

\end{document}

### 2021-07-31 figures/code_average.m

@@ -1,0 +1,119 @@
+%% #1 Average number of touched squares, experimentally
+
+clear, clc
+L = .1:.1:9;
+a = 1.35; b = 1;
+N = 1e5;
+z0 = rand(N, numel(L))*a + 1j*rand(N, numel(L))*b;
+z1 = z0 + L.*exp(1j*2*pi*rand(N, numel(L)));
+ii = ceil(abs(real(z1/a))) + (real(z1)<0); % add 1 for negative
+jj = ceil(abs(imag(z1/b))) + (imag(z1)<0); % add 1 for negative
+result = mean(ii + jj, 1) - 1;
+plot(L, result, '.-'), grid on, axis equal
+
+
+%% #2 Selectable range for theta
+% I see that theta on (0,pi/2) doesn't give the same result. Plottin with something like
+% plot([z0(1:600,60), z1(1:600,60)].', '.-'), samples seem to be missing near the extremes
+
+L = .1:.1:9;
+a = 1.35; b = 1;
+N = 1e5;
+theta_max = 2*pi;   
+unfix = @(x) sign(x).*ceil(abs(x)); % round away from 0; real values only
+z0 = rand(N, numel(L))*a + 1j*rand(N, numel(L))*b;
+z1 = z0 + L.*exp(1j*theta_max*rand(N, numel(L)));
+ii = unfix(real(z1/a)) - (real(z1)<0); % subtract 1 for negative; it's signed
+jj = unfix(imag(z1/b)) - (imag(z1)<0); % subtract 1 for negative; it's signed
+result = mean(abs(ii) + abs(jj), 1) - 1;
+plot(L, result, '.-'), grid on, axis equal
+
+
+%% #3 I define ki, kj signed (page 17, (*)), from which ii, jj are obtained 
+
+L = .1:.1:5;
+a = 1.35; b = 1;
+N = 3e5;%1e5;
+theta_max = 2*pi;
+%z0 = repmat(.42+.73j, N, numel(L));
+z0 = rand(N, numel(L))*a + 1j*rand(N, numel(L))*b;
+z1 = z0 + L.*exp(1j*theta_max*rand(N, numel(L)));
+ki = floor(real(z1/a));
+kj = floor(imag(z1/b));
[…]

### 2021-07-31 figures/figures.m

@@ -1,0 +1,1153 @@
+%% #1. Testing (for the challenge). Squares of side 2, segments of odd length (here I use squares of side 1, segments of length .5, 1.5, ...)
+result_all = [];
+for L = ((1:15)*2-1)/2
+    result = 0;
+    for theta = linspace(0,pi/4,30e3)
+        points = linspace(0,L,3e3*L)*exp(1j*theta);
+        num_squares = numel(unique(floor(points)))+2;
+        result = max(result, num_squares);
+    end
+    result_all(end+1) = result
+end
+
+%3,5,6,7,9,10,12,13,15,16,17,19,20,22,23
+
+
+%% #2. Computing  (for the challenge). Squares of side 2, segments of odd length (here I use squares of side 1, segments of length .5, 1.5, ...)
+clear
+result_all = [];
+for L = ((1:15)*2-1)/2
+    t = (0:L).^2;
+    [ii, jj] = find(t + t.' < L.^2);
+    result_all(end+1) = max(ii+jj)+1; % -2 por 0-based indexing; -1 por los cuadrados hasta (ii,jj);
+                                      % +4 por desplazamiento para coger trocitos de cuadrados en los extremos
+end
+
+
+%% #3. Computing. Squares of size 1, segments of length 1, 2, 3, ...
+clear
+result_all = [];
+for L = 1:50
+    t = (0:L).^2;
+    [ii, jj] = find((t + t.' <= L.^2) & (logical(t)==logical(t.'))); % con igualdad. No contamos las líneas verticales u horizontales, pero sí el origen
+    M = max(ii+jj)-1; % -2 por 0-based indexing; -1 por los cuadrados hasta (ii,jj); +2 por desplazamientos para coger trocitos
+    [ii, jj] = find(t + t.' < L.^2);
[…]

### 2021-07-31 paper.tex
"Make room" in the paper structure for the average-number-of-tiles problem
This involves new section/subsection structure and rewriting the introduction.

I also include some ideas to be developed in the new section for the average number of tiles

@@
 \newcommand{\funli}{\Lambda} % function that gives minimum integer-valued length for a given number of touched tiles
+\newcommand{\funta}{\varphi} % function that gives average number of touched tiles for real-valued length
 \newcommand{\len}{\ell} % length variable, real-valued
@@
 \title{
-On the maximum number of tiles touched\\
-by a line segment on a rectangular grid
+On the number of tiles touched by a\\
+line segment on a rectangular grid
 }
@@
 \author[1]{Luis Mendo}
-\author[2]{**Name**}
+\author[2]{Alex Arkhipov}
 
@@
 The segment is said to \emph{touch} a tile if it intersects its interior.\footnote{
-The definition uses the interior of the tile, excluding the border, to avoid uninteresting results such as a ``zero-length'' segment touching (a vertex of) $4$ tiles.}
-What is the maximum number of tiles that the segment can touch?
+The definition uses the interior of the tile, excluding the border, to avoid uninteresting results such as a ``zero-length'' segment touching (a vertex of) $4$ tiles.} This paper deals with the following problems:
+\begin{itemize}
+\item What is the maximum number of tiles that the segment can touch?
+\item What length should a segment have to touch a given number of tiles?
+\item What is the average number of tiles touched by a uniformly random segment of a given length?
+\end{itemize}
 
-As an example, consider $a=1.35$, $b=1$.
+As an example of the first problem, consider $a=1.35$, $b=1$.
 A segment of unit length can be placed as shown in Figure~\ref{fig: examples} (left) to touch $3$ tiles.
@@
 For real-valued lengths this infimum is not a minimum, because given a segment that touches $\tiles$ tiles, its length can always be reduced by some non-zero amount without changing the number of touched tiles.
-This is a consequence of the interior of the each tile being an open set.
+This is a consequence of the interior of each tile being an open set.
+
+The direct and inverse problems are closely related.
+Namely, if $\len$ is the infimum of all lengths that allow touching at least $\tiles$ tiles (inverse problem), $\tiles$ is the maximum number of tiles that can be touched with lengths slightly greater than $\len$ (direct problem).
+
+The question of determining how many tiles a segment of a given length touches on average will be referred to as the \emph{average-number} problem.
+As will be seen, solving this also answers the inverse question (segment length to touch a given number of tiles on average), which thus need not be considered separately.
+By symmetry, one endpoint of the segment can be assumed to lie in a fixed, reference tile.
+The position of this endpoint will be considered to be \emph{uniformly} distributed on the tile.
+The segment orientation will be have a \emph{uniform} distribution on the set of all possible directions, and is \emph{independent} of the endpoint position.
 
@@
 
-These problems are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem.
+The problems addressed in this paper are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem.
 Some references: ``Buffon's Noodle Problem'', J.F.~Ramaley, 1969.
@@
 
-This paper analyzes both problems for real-valued lengths, and for integer-valued lengths on a unit square grid.
-Section~\ref{part: fund results} presents fundamental results, from which the characterization is obtained in Section~\ref{part: arbitrary grid, real lengths} for an arbitrary grid with real-valued lengths; in Section~\ref{part: square grid, real lengths} for a square grid with real-valued lengths; and in Section~\ref{part: unit square grid, integer lengths} for a unit square grid with integer-valued lengths.
+The rest of the paper is organized as follows.
+\S\ref{part: fund results} presents fundamental results, which form the basis of the subsequent analysis.
+The direct and inverse problems are considered in \S\ref{part: max}, first for real-valued lengths, and then for integer-valued lengths on a unit square grid.
+The average-number problem is analyzed in \S\ref{part: ave}.
 
@@
 In order to derive a general method to build a minimal sufficient set, it is insightful to consider two specific cases first, as will be presented in the following.
-It is also helpful to use the equivalent formulation of the direct problem referred to in Section~\ref{part: intro}, i.e.\ finding the maximum number of tiles touched by segments of length $\len$ \emph{or less}.
+It is also helpful to use the equivalent formulation of the direct problem referred to in \S\ref{part: intro}, i.e.\ finding the maximum number of tiles touched by segments of length $\len$ \emph{or less}.
 
@@
 
-\section{Arbitrary grid with real-valued lengths}
-\label{part: arbitrary grid, real lengths}
+\section{Maximum number of tiles: direct and inverse problems}
+\label{part: max}
+% ***Is it clear enough that "maximum number of tiles" can refer to the inverse problem too? (because the inverse problem is closely related to the direct one, as discussed in the introduction)
+% If not, an alernative would be a more explicit title like: "Maximum number of tiles fo a given length. Conditions on length for a given number of tiles"
 
-For a grid with parameters $a, b \in \mathbb R^+$, the maximum number $\tiles$ of touched tiles for a given real-valued length $\len$ can be represented by a function $\funt: \mathbb R^+ \to \mathbb N$ such that $\tiles = \funt(\len)$.
+The direct and inverse problems for the maximum number of tiles are considered in this section.
+The general case for rectangular grids with real-valued segment lengths is analyzed first, in \S\ref{part: max: arbitrary grid, real lengths}.
+The particular case of square grids is addressed in \S\ref{part: max: square grid, real lengths}, as it allows a specialized formula for the direct problem.
+Lastly, the analysis of a unit square grid with integer-valued lengths is given in \S\ref{part: max: unit square grid, integer lengths}.
+
+\subsection{Arbitrary grid with real-valued lengths}
+\label{part: max: arbitrary grid, real lengths}
+
+For a grid with parameters $a, b \in \mathbb R^+$, the maximum number $\tiles$ of touched tiles with a given real-valued length $\len$ can be represented by a function $\funt: \mathbb R^+ \to \mathbb N$ such that $\tiles = \funt(\len)$.
 Similarly, for the inverse problem a function $\funl: \mathbb N \to \mathbb R^+$ can be defined such that $\funl(\tiles)$ gives the infimum length of all segments that touch at least $\tiles$ tiles.
@@
 
-Theorems~\ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the two problems stated in Section~\ref{part: intro}; they also provide a way to actually position a segment of length slightly greater than $\len$ or $\funl(\tiles)$ so that it touches $\tiles$ or $\funt(\len)$ tiles.
+Theorems~\ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the two problems stated in \S\ref{part: intro}; they also provide a way to actually position a segment of length slightly greater than $\len$ or $\funl(\tiles)$ so that it touches $\tiles$ or $\funt(\len)$ tiles.
 Namely, the segment should have its endpoints respectively in the interior of two tiles separated $\isoli-1$ steps horizontally and $\jsoli-1$ steps vertically, with its exact position and orientation  adjusted to avoid any grid points.
@@
 
-For $a=b$, either from symmetry considerations or particularizing the formulas in this section it stems that the optimal orientation of the segment is close to $45^\circ$.
-This case will be dealt with in Section~\ref{part: square grid, real lengths}, as it lends itself to somewhat simplified formulas.
+For $a=b$, either from symmetry considerations or particularizing the formulas in the previous theorems it stems that the optimal orientation of the segment is close to $45^\circ$.
+This case will be dealt with in \S\ref{part: max: square grid, real lengths}, as it lends itself to somewhat simplified formulas.
 
@@
 
-\section{Square grid with real-valued lengths}
-\label{part: square grid, real lengths}
+\subsection{Square grid with real-valued lengths}
+\label{part: max: square grid, real lengths}
 
-Particularizing the results in Section~\ref{part: arbitrary grid, real lengths} to a square grid, $a=b$, obviously yields simpler formulas.
+Particularizing the results in \S\ref{part: max: arbitrary grid, real lengths} to a square grid, $a=b$, obviously yields simpler formulas.
 
@@
 
-\section{Unit square grid with integer lengths}
-\label{part: unit square grid, integer lengths}
+\subsection{Unit square grid with integer lengths}
+\label{part: max: unit square grid, integer lengths}
 
-A natural variation of the two problems introduced in Section~\ref{part: intro} is to consider $a=b=1$, with the additional restriction that the segment length can only be a positive integer (equivalently, a square grid is considered with step $a$ and the segment length can only be an integer multiple of $a$).
+A natural variation of the two problems introduced in \S\ref{part: intro} is to consider $a=b=1$, with the additional restriction that the segment length can only be a positive integer (equivalently, a square grid is considered with step $a$ and the segment length can only be an integer multiple of $a$).
 
@@
 This is A346232 in the On-Line Encyclopedia of Integer Sequences \cite{OEIS_unitsq_int_dir}.
-For this sequence, a simpler expression can be obtained than those resulting from Section~\ref{part: square grid, real lengths} with $a=1$, and the following properties hold.
+For this sequence, a simpler expression can be obtained than those resulting from \S\ref{part: max: square grid, real lengths} with $a=1$, and the following properties hold.
 
@@
 
+\section{Average number of tiles}
+\label{part: ave}
+
+***A theorem, and maybe one or two propositions. It is interesting to state $\Pr[i \geq i_0]$ (averaged over $x$) in a proposition, as it can be compared with the Buffon experiment.
+
+***Define the average number of tiles touched as a function $\funta: \mathbb R^+ \to \mathbb R^+$.
+
+***The formula for the average number of files is:
+\begin{equation}
+\funta(\len) = \frac{2\len}{\pi}\left(\frac 1 a + \frac 1 b\right) + 1.
+\end{equation}
+
+***Explain that the inverse question is straightforward, because $\funta$ is monotone increasing, thus it has an inverse function (defined on the range of $\funta$).
+In fact, as will be seen, $\funta$ is an affine function, and therefore it is immediate to obtain its inverse.
+
+***Some mention that there is no need to study square grids or the case of integer-valued lengths separately.
+
+***Relation the the problem of the Buffon needle.
+On one hand, the formula for b=infinity gives the average number of lines crossed in the Buffon experiment.
+On the other hand, the intermediate values $\Pr[i \geq i_0]$ (proposition) are a generalization of the probability of crossing at least a line in the Buffon experiment, when the needle endpoint is only allowed to move in a smaller vertical strip.
+
+***Mention particular values: a=b=1 gives...; a=b=4/pi gives len+1. Can we find any ``meaning'' for these particular cases?
+
+***The function is affine, not linear. This is related to the fact that the parts of the segment near the endpoints can touch several tiles
+
+***Maybe a graph comparing maximum and average.
+Or compute/plot their ratio.
+The asymptotic slopes are $\sqrt(1/a^2+1/b^2)$ for $\funt$ and $2/\pi(1/a+1/b)$.
+It may be interesting to consider the ratio of asymptotic slopes.
+Some quick plot of that ratio suggests it is maximum for $a=b$, and equals $sqrt(8)/pi = 0.9003$.
+
+***The distribution would be interesting, not just the average.
+But it is more difficult to obtain, because of the statistical relationship between the random variables i and j.
+This relationship does not affect the \emph{average} of the random variable $i+j-1$ (for a given length), but it does affect the its \emph{distribution}.
+[I think it is interesting to mention this].
+
+\begin{theorem}
+\label{theo: funta, form}
+\end{theorem}
+
 \bibliographystyle{plain}

### 2021-08-01 paper.tex
Quick note for the "to-consider" list
Wording
Make the three problems clearer

@@
 
-The question of determining how many tiles a segment of a given length touches on average will be referred to as the \emph{average-number} problem.
-As will be seen, solving this also answers the inverse question (segment length to touch a given number of tiles on average), which thus need not be considered separately.
+The problem of how many tiles a \emph{random} segment of a given length touches on average is precisely defined as follows.
 By symmetry, one endpoint of the segment can be assumed to lie in a fixed, reference tile.
-The position of this endpoint will be considered to be \emph{uniformly} distributed on the tile.
-The segment orientation will be have a \emph{uniform} distribution on the set of all possible directions, and is \emph{independent} of the endpoint position.
+The position of this endpoint is \emph{uniformly} distributed on the tile.
+The segment orientation has a \emph{uniform} distribution on the set of all possible directions, and is \emph{independent} of the endpoint position.
+As will be seen, solving this problem also answers the inverse question (segment length to touch a given number of tiles on average), which thus need not be considered separately.
 
@@
 
-The problems addressed in this paper are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem.
+The problems studied in this paper are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem.
 Some references: ``Buffon's Noodle Problem'', J.F.~Ramaley, 1969.
@@
 The rest of the paper is organized as follows.
-\S\ref{part: fund results} presents fundamental results, which form the basis of the subsequent analysis.
-The direct and inverse problems are considered in \S\ref{part: max}, first for real-valued lengths, and then for integer-valued lengths on a unit square grid.
-The average-number problem is analyzed in \S\ref{part: ave}.
+Fundamental results are presented in \S\ref{part: fund results}, which form the basis of the subsequent analysis.
+The direct and inverse problems for a deterministic segment are considered in \S\ref{part: max}, first for real-valued lengths, and then for integer-valued lengths on a unit square grid.
+The problem of computing the average number of tiles touched by a random segment is analyzed in \S\ref{part: ave}.
 
@@
 
-\section{Maximum number of tiles: direct and inverse problems}
+\section{Direct and inverse problems for a deterministic segment}
 \label{part: max}
@@
 
-The direct and inverse problems for the maximum number of tiles are considered in this section.
+The direct and inverse problems defined in \S\ref{part: intro}, considering the segment position and orientation as deterministic, are addressed in this section.
 The general case for rectangular grids with real-valued segment lengths is analyzed first, in \S\ref{part: max: arbitrary grid, real lengths}.
@@
 
-\section{Average number of tiles}
+\section{Average number of tiles touched by a random segment}
 \label{part: ave}
@@
 Some quick plot of that ratio suggests it is maximum for $a=b$, and equals $sqrt(8)/pi = 0.9003$.
+The minimum number of tile that can be touched with non-zero probability (corresponding to an n times 1 canonical rectangle if a>=b) can also be plotted.
 

### 2021-08-01 paper.tex
Writing the part about the average number of tiles

@@
 \newcommand{\st}{\mid}%{\,|\,}
+\newcommand{\cond}{\,|\,} % less spacing that \mid
+\newcommand{\diff}{\,\mathrm d}
+\let\Re\relax % remove the command...
+\DeclareMathOperator{\Re}{Re} % ...to define it differently
+\let\Im\relax
+\DeclareMathOperator{\Im}{Im}
+\DeclareMathOperator{\E}{E}
 \newcommand{\funt}{\tau} % function that gives maximum number of touched tiles for real-valued length
@@
 \newcommand{\tiles}{t} % number of tiles (integer-valued)
-\let\Re\relax % remove the command...
-\DeclareMathOperator{\Re}{Re} % ...to define it differently
-\let\Im\relax
-\DeclareMathOperator{\Im}{Im}
 \newcommand{\isolr}{i^+}
@@
 \newcommand{\jsoli}{j^\ast}
+\newcommand{\genvar}{r}
+\newcommand{\orient}{\theta}
 \newcommand{\mss}{M}
@@
 
-In order to prove that $\funti(\leni+1)-\funti(\leni) \in \{1,2\}$, consider the function $f(x) = \sqrt{2x^2-2}$ for $x \in \mathbb R$, $x>1$. Its first derivative is
-\begin{equation}
-f'(x) = \frac {\sqrt{2} \, x} {\sqrt{x^2-1}},
-\end{equation}
-and its second derivative is easily seen to be negative. Therefore $f'(x)$ can be bounded for $x \geq 3$ as
+In order to prove that $\funti(\leni+1)-\funti(\leni) \in \{1,2\}$, consider the function $f(\genvar) = \sqrt{2\genvar^2-2}$ for $\genvar \in \mathbb R$, $\genvar>1$. Its first derivative is
+\begin{equation}
+f'(\genvar) = \frac {\sqrt{2} \, \genvar} {\sqrt{\genvar^2-1}},
+\end{equation}
+and its second derivative is easily seen to be negative. Therefore $f'(\genvar)$ can be bounded for $\genvar \geq 3$ as
 \begin{equation}
 \label{eq: der bounds}
-\lim_{x \rightarrow \infty} f'(x) = \sqrt{2} < f'(x) < f'(3) = 3/2.
-\end{equation}
-For $\leni \geq 2$, by the mean value theorem \cite[section~5.3]{Abbott15}, when $\leni$ is increased to $\leni+1$ the term $\sqrt{2\leni^2-2}$ in \eqref{eq: funti leni} has an increment that equals $f'(x)$ for some $\leni < x < \leni+1$.
+\lim_{\genvar \rightarrow \infty} f'(\genvar) = \sqrt{2} < f'(\genvar) < f'(3) = 3/2.
+\end{equation}
+For $\leni \geq 2$, by the mean value theorem \cite[section~5.3]{Abbott15}, when $\leni$ is increased to $\leni+1$ the term $\sqrt{2\leni^2-2}$ in \eqref{eq: funti leni} has an increment that equals $f'(\genvar)$ for some $\leni < \genvar < \leni+1$.
 Therefore
@@
 
-***A theorem, and maybe one or two propositions. It is interesting to state $\Pr[i \geq i_0]$ (averaged over $x$) in a proposition, as it can be compared with the Buffon experiment.
-
-***Define the average number of tiles touched as a function $\funta: \mathbb R^+ \to \mathbb R^+$.
-
-***The formula for the average number of files is:
-\begin{equation}
+Given $\len \in \mathbb R*+$, consider a segment of length $\len$ with random position and orientation.
+Specifically, the coordinates $x_1$, $y_1$ of the first endpoint are independent random variables uniformly distributed on $(0,a)$ and $(0,b)$ respectively.
+The orientation $\orient$ of the segment, defined as the angle with respect to the horizontal axis of the vector that goes from the first to the second endpoint, is uniformly distributed on $(0,2\pi)$.
+The variables $x_1$, $y_1$ and $\theta$ determine the coordinates $x_2$, $y_2$ of the second endpoint.
+Note that events of probability $0$, such as $x_1=a$ or the segment passing through a grid point, can be disregarded for computing the average number of tiles touched by the segment.
+
+Let the function $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles touched by a random segment of length $\len$, assuming the above distributions.
+
+Each realization of the random segment gives rise to a canonical rectangle.
+The normalized dimensions $i$ and $j$ of the canonical rectangle are thus random variables.
+Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and the segment touches $t = i+j-1$ tiles.
+
+The random variables $i$ and $j$ are not independent.
+As a consequence, while the marginal distributions of $i$ and $j$ are easy to compute, that of $i+j$ is difficult.
+However, the average number of touched tiles can be obtained from the marginal distributions as
+\begin{equation}
+\label{eq: funta E[i] E[j]}
+\funta(\len) = \E[t] = \E[i]+\E[j]+1.
+\end{equation}
+
+For $0 \leq \genvar \leq 1$, let
+\begin{equation}
+\label{eq: g}
+g(\genvar) = \int_0^r \arccos z \diff z = \genvar \arccos \genvar - \sqrt{1-\genvar^2} + 1.
+\end{equation}
+
+\begin{proposition}
+Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above.
+Let the random variables $i$, $j$ represent the normalized dimensions of the canonical rectangle.
+For $n \in \mathbb N$,
+\begin{equation}
+\label{eq: Pr i geq n}
+\Pr[i \geq n] = \begin{cases}
+\displaystyle
+1 &\text{if\ \ } \displaystyle n =1 \\[1 mm]
+\displaystyle
+\frac{2\len}{\pi a} \left(g\left(\frac{a(n-1)}{\len}\right)-g\left(\frac{a(n-2)}{\len} \right)\right) &\text{if\ \ } \displaystyle 2 \leq n < \frac \len a + 1 \\[4 mm]
+\displaystyle
+\frac{2\len}{\pi a}\,\, \left(1 - g\left(\frac{a(n-2)}{\len}\right)\right) &\text{if\ \ } \displaystyle \frac\len a + 1 \leq n < \frac\len a+2 \\[3 mm]
+\displaystyle
+0 &\text{if\ \ } \displaystyle \frac \len a + 2 \leq  n; \\
+\end{cases}
+\end{equation}
+and $\Pr[j \geq n]$ is obtained from \eqref{eq: Pr i geq n} replacing $a$ by $b$.
+\end{proposition}
+
+\begin{proof}
+Clearly, $\Pr[i \geq 1]=1$.
+In the following it will be assumed that $n \geq 2$.
+The basic idea is to compute $\Pr[i \geq n]$ conditioned on $(x_1,y_1)$ (or, as will be seen, only on $y_1$), and then to average over $x_1$ and $y_1$ (actually only over $x_1$).
+
+% **Aq. Prepare and include two subfigures
+Given the coordinates $(x_1, y_1)$ of the first segment endpoint, with $0 \leq x_1 \leq a$, $0 \leq y_1 \leq b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure **ab**.
+The segment orientation is a random angle $\orient$ uniformly distributed on $[0,2\pi)$.
+It is easy to see that $i \geq n$ if and only if either $x_2 \geq a(n-1)$ or $x_2 \leq -a(n-2)$.
+The latter two events are exclusive, and by symmetry have the same probability.
+Therefore
+\begin{equation}
+\label{eq: Pr i = 2 Pr}
+\Pr[i \geq n] = 2 \Pr[x_2 \geq a(n-1)].
+\end{equation}
+
+Consider the event $x_2 \geq a(n-1)$.
+There are three possibilities depending on $x_1$, $n$ and $\len$.
+If $a(n-1) < \len$, the length $\len$ is enough for $x_2$ to exceed $a(n-1)$ for some angles $\orient$, regardless of $x_1$.
+This is depicted in Figure **a**.
+If $\len \leq a(n-1) < \len+a$, the length is enough only if $x_1 \geq a(n-1)-\len$, as seen in Figure **b**.
+Lastly, if $\len+a \leq a(n-1)$ it is impossible that $x_2$ exceeds $a(n-1)$.
+Note that the value $y_1$ is irrelevant for this.
+
+In the first two cases above, the probability that $x_2 \geq a(n-1)$, conditioned on $x_1$, is the length of the arc to the right of the line $x=a(n-1)$ divided by $2\pi\len$; that is,
+\begin{equation}
+\label{eq: Pr x_2 geq a(n-1) cond}
+\Pr[x_2 \geq a(n-1) \cond x_1] = \frac 1 \pi \arccos \frac{a(n-1)-x_1}{\len}.
+\end{equation}
+In the first case $x_1$ has a uniform distribution on $(0,a)$, and $\Pr[x_2 \geq a(n-1)]$ is easily obtained from \eqref{eq: Pr x_2 geq a(n-1) cond} as follows:
+\begin{equation}
+\Pr[x_2 \geq a(n-1)] = \frac 1 {\pi a} \int_0^a\arccos \frac{a(n-1)-x_1}{\len} \diff x_1. 
+\end{equation}
+Integrating by parts, this is seen to be
+\begin{equation}
+\label{eq: Pr x_2 geq a(n-1), first case}
+\Pr[x_2 \geq a(n-1)] = \frac{\len}{\pi a}\,\, \left( g\left( \frac{a(n-1)}{\len} \right) - g\left( \frac{a(n-2)}{\len} \right) \right),
+\end{equation}
+where the function $g$ is defined in \eqref{eq: g}. Substituting into \eqref{eq: Pr i = 2 Pr} yields the result in \eqref{eq: Pr i geq n}, second line.
+
+The second case is similar, but the integration over $x_1$ is from $a(n-1)-\len$ to $a$. Noting that $g(1)=1$, this gives
+\begin{equation}
+\label{eq: Pr x_2 geq a(n-1), second case}
+\Pr[x_2 \geq a(n-1)] = \frac{\len}{\pi a}\,\, \left( 1 - g\left( \frac{a(n-2)}{\len} \right) \right),
+\end{equation}
+which combined with \eqref{eq: Pr i = 2 Pr} yields the expression in \eqref{eq: Pr i geq n}, third line.
+
+The third case obviously gives $\Pr[i \geq n] = 0$, as in \eqref{eq: Pr i geq n}, fourth line.
+
+The analysis above can be analogously applied to $\Pr[j \geq n]$ if the $x$ and $y$ axes are interchanged. Thus the result is the same with $a$ replaced by $b$.
+\end{proof}
+
+\begin{theorem}
+\label{theo: funta, form}
+Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above.
+The average number of tiles touched by the segment, $\funta(\len)$, is
+\begin{equation}
+\label{eq: funta, expression}
 \funta(\len) = \frac{2\len}{\pi}\left(\frac 1 a + \frac 1 b\right) + 1.
 \end{equation}
+\end{theorem}
+
+\begin{proof}
+***
+\end{proof}
 
@@
 
-***Relation the the problem of the Buffon needle.
+***Relation the problem of the Buffon needle.
 On one hand, the formula for b=infinity gives the average number of lines crossed in the Buffon experiment.
@@
 Some quick plot of that ratio suggests it is maximum for $a=b$, and equals $sqrt(8)/pi = 0.9003$.
-The minimum number of tile that can be touched with non-zero probability (corresponding to an n times 1 canonical rectangle if a>=b) can also be plotted.
-
-***The distribution would be interesting, not just the average.
-But it is more difficult to obtain, because of the statistical relationship between the random variables i and j.
-This relationship does not affect the \emph{average} of the random variable $i+j-1$ (for a given length), but it does affect the its \emph{distribution}.
-[I think it is interesting to mention this].
-
-\begin{theorem}
-\label{theo: funta, form}
-\end{theorem}
+The minimum number of tiles that can be touched with non-zero probability (corresponding to an n times 1 canonical rectangle if a>=b) can also be plotted.
 

### 2021-08-02 paper.tex
More writing on the average number

@@
 
-% **Aq. Prepare and include two subfigures
+% *** Prepare and include two subfigures
 Given the coordinates $(x_1, y_1)$ of the first segment endpoint, with $0 \leq x_1 \leq a$, $0 \leq y_1 \leq b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure **ab**.
@@
 \begin{equation}
-\label{eq: funta, expression}
+\label{eq: funta, form}
 \funta(\len) = \frac{2\len}{\pi}\left(\frac 1 a + \frac 1 b\right) + 1.
@@
 \begin{proof}
-***
+The expected value of $i$ can be obtained as
+\begin{equation}
+\label{eq: theo funta form 1}
+\E[i] = \sum_{n=1}^\infty \Pr[i \geq n].
+\end{equation}
+From \eqref{eq: Pr i geq n}, this sum only has a finite number of non-zero terms: $1$ for $n=1$, plus $\lceil \len/a \rceil-1$ terms given by the second line of \eqref{eq: Pr i geq n}, plus one term given by the third line, corresponding to $n = \lceil \len/a \rceil +1 $:
+\begin{equation}
+\label{eq: theo funta form 2}
+\begin{split}
+\E[i] = 1 +
+\sum_{n=2}^{\lceil \len/a \rceil} \frac{2\len}{\pi a} \left(g\left(\frac{a(n-1)}{\len}\right)-g\left(\frac{a(n-2)}{\len} \right)\right) + \\
+\frac{2\len}{\pi a} \left(1-g\left(\frac{a(\lceil \len/a \rceil - 1)}{\len} \right)\right).
+\end{split}
+\end{equation}
+Most terms involving the function $g$ in \eqref{eq: theo funta form 2} cancel, with the result
+\begin{equation}
+\label{eq: theo funta form 3}
+\E[i] = 1+\frac{2\len}{\pi a} (1-g(0))= 1+\frac{2\len}{\pi a}.
+\end{equation}
+Similarly,
+\begin{equation}
+\label{eq: theo funta form 4}
+\E[j] = 1+\frac{2\len}{\pi b}.
+\end{equation}
+Substituting \eqref{eq: theo funta form 3} and \eqref{eq: theo funta form 4} into \eqref{eq: funta E[i] E[j]} gives \eqref{eq: funta, form}.
 \end{proof}

### 2021-08-02 figures/code_average.m
Merged code files

@@ -1,119 +1,0 @@
-%% #1 Average number of touched squares, experimentally
-
-clear, clc
-L = .1:.1:9;
-a = 1.35; b = 1;
-N = 1e5;
-z0 = rand(N, numel(L))*a + 1j*rand(N, numel(L))*b;
-z1 = z0 + L.*exp(1j*2*pi*rand(N, numel(L)));
-ii = ceil(abs(real(z1/a))) + (real(z1)<0); % add 1 for negative
-jj = ceil(abs(imag(z1/b))) + (imag(z1)<0); % add 1 for negative
-result = mean(ii + jj, 1) - 1;
-plot(L, result, '.-'), grid on, axis equal
-
-
-%% #2 Selectable range for theta
-% I see that theta on (0,pi/2) doesn't give the same result. Plottin with something like
-% plot([z0(1:600,60), z1(1:600,60)].', '.-'), samples seem to be missing near the extremes
-
-L = .1:.1:9;
-a = 1.35; b = 1;
-N = 1e5;
-theta_max = 2*pi;   
-unfix = @(x) sign(x).*ceil(abs(x)); % round away from 0; real values only
-z0 = rand(N, numel(L))*a + 1j*rand(N, numel(L))*b;
-z1 = z0 + L.*exp(1j*theta_max*rand(N, numel(L)));
-ii = unfix(real(z1/a)) - (real(z1)<0); % subtract 1 for negative; it's signed
-jj = unfix(imag(z1/b)) - (imag(z1)<0); % subtract 1 for negative; it's signed
-result = mean(abs(ii) + abs(jj), 1) - 1;
-plot(L, result, '.-'), grid on, axis equal
-
-
-%% #3 I define ki, kj signed (page 17, (*)), from which ii, jj are obtained 
-
-L = .1:.1:5;
-a = 1.35; b = 1;
-N = 3e5;%1e5;
-theta_max = 2*pi;
-%z0 = repmat(.42+.73j, N, numel(L));
-z0 = rand(N, numel(L))*a + 1j*rand(N, numel(L))*b;
-z1 = z0 + L.*exp(1j*theta_max*rand(N, numel(L)));
-ki = floor(real(z1/a));
-kj = floor(imag(z1/b));
[…]

### 2021-08-02 figures/figures.m

@@ -1130,7 +1130,7 @@ funl_int(~ind) = ceil(sqrt((G(~ind)-3).^2/2+1));
 isequal(funl_real, funl_int)
 
 
-%%  #47. Figures of the direct-problem and inverse problem sequences (funti and funli)
+%%  #48. Figures of the direct-problem and inverse problem sequences (funti and funli)
 clear
 close all
 figure
@@ -1151,3 +1151,124 @@ stem(G, funli, 'o')
 grid on, % axis equal
 axis([0 25 0 16])
 xlabel t, ylabel([char(hex2dec('039B')) '(t)'])
+
+
+%% #49. Average number of touched squares, experimentally
+
+clear, clc
+L = .1:.1:9;
+a = 1.35; b = 1;
+N = 1e5;
+z0 = rand(N, numel(L))*a + 1j*rand(N, numel(L))*b;
+z1 = z0 + L.*exp(1j*2*pi*rand(N, numel(L)));
+ii = ceil(abs(real(z1/a))) + (real(z1)<0); % add 1 for negative
+jj = ceil(abs(imag(z1/b))) + (imag(z1)<0); % add 1 for negative
+result = mean(ii + jj, 1) - 1;
+plot(L, result, '.-'), grid on, axis equal
+
+
+%% #50. Selectable range for theta
+% I see that theta on (0,pi/2) doesn't give the same result. Plottin with something like
+% plot([z0(1:600,60), z1(1:600,60)].', '.-'), samples seem to be missing near the extremes
+
+L = .1:.1:9;
+a = 1.35; b = 1;
+N = 1e5;
+theta_max = 2*pi;   
+unfix = @(x) sign(x).*ceil(abs(x)); % round away from 0; real values only
+z0 = rand(N, numel(L))*a + 1j*rand(N, numel(L))*b;
+z1 = z0 + L.*exp(1j*theta_max*rand(N, numel(L)));
+ii = unfix(real(z1/a)) - (real(z1)<0); % subtract 1 for negative; it's signed
+jj = unfix(imag(z1/b)) - (imag(z1)<0); % subtract 1 for negative; it's signed
[…]

### 2021-08-02 paper.tex
Added figures for proof of theorem of average number

@@
 
-% *** Prepare and include two subfigures
-Given the coordinates $(x_1, y_1)$ of the first segment endpoint, with $0 \leq x_1 \leq a$, $0 \leq y_1 \leq b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure **ab**.
+Given the coordinates $(x_1, y_1)$ of the first segment endpoint, with $0 \leq x_1 \leq a$, $0 \leq y_1 \leq b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure~\ref{fig: Pr_i}.
 The segment orientation is a random angle $\orient$ uniformly distributed on $[0,2\pi)$.
-It is easy to see that $i \geq n$ if and only if either $x_2 \geq a(n-1)$ or $x_2 \leq -a(n-2)$.
+It is clear from the figure that $i \geq n$ if and only if either $x_2 \geq a(n-1)$ or $x_2 \leq -a(n-2)$.
 The latter two events are exclusive, and by symmetry have the same probability.
@@
 \end{equation}
+
+\begin{figure}
+\centering%
+\subfigure[$i$ can be equal or greater than $n$ for all $x_1$, $0 \leq x_1 \leq a$]{%
+\label{fig: Pr_i_full}%
+\includegraphics[width=.48\textwidth]{Pr_i_full_1p35}%
+}\hfill%
+\subfigure[$i$ can be equal or greater than $n$ only for $a(n-1)-\len \leq x_1 \leq a$]{%
+\label{fig: Pr_i_notfull}%
+\includegraphics[width=.48\textwidth]{Pr_i_notfull_1p35}%
+}%
+\caption{Conditions for the normalized width of the canonical rectangle, $i$, being equal or greater than a given $n$. Example with $a=1.35$, $b=1$
+}%
+\label{fig: Pr_i}
+\end{figure}%
 
@@
 If $a(n-1) < \len$, the length $\len$ is enough for $x_2$ to exceed $a(n-1)$ for some angles $\orient$, regardless of $x_1$.
-This is depicted in Figure **a**.
-If $\len \leq a(n-1) < \len+a$, the length is enough only if $x_1 \geq a(n-1)-\len$, as seen in Figure **b**.
+This is depicted in Figure~\ref{fig: Pr_i_full}, where the thicker line on the arc represents those angles for which $x_2 > a(n-1)$, for a specific $x_1$.
+If $\len \leq a(n-1) < \len+a$, the length is enough only if $x_1 \geq a(n-1)-\len$.
+This can be seen in Figure~\ref{fig: Pr_i_notfull}, where the thicker line again represents the angles for which this happens, for a specific $x_1 \geq a(n-1)-\len$.
 Lastly, if $\len+a \leq a(n-1)$ it is impossible that $x_2$ exceeds $a(n-1)$.

### 2021-08-02 figures/figures.m

@@ -1270,5 +1270,52 @@ hold on
 plot(L, 2*L/pi*(1/a+1/b)+1)
 
 
+%% #53. Figures for proof of Pr[i>=n]
 
-
+clear, clf
+a = 1.35; b = 1;
+x1 = .6*a; y1 = .72*b;
+n = 3;
+min_x = -2; max_x = 3; min_y = -3; max_y = 4;
+flag_full = false;
+if flag_full
+    len = 2.1*a;
+else
+    len = 1.78*a;
+end
+m = max(0, a*(n-1)-len);
+patch([m a a m m], [0 0 b b 0], .88*[1 1 1], 'edgecolor', 'none')
+hold on
+patch([0 a a 0 0], [0 0 b b 0], 1*[1 1 1], 'edgecolor', 'k', 'FaceAlpha', 0)
+plot(x1 + 1j*y1, 'k.', 'markersize', 7)
+grid on, axis equal
+set(gca, 'GridAlpha', .4)
+xticks((min_x:max_x)*a), yticks((min_y:max_y)*b)
+xticklabels([arrayfun(@(n)[num2str(n) char(8201) repmat('a', 1, n~=0)], min_x:max_x, 'UniformOutput', false)]) % thin space
+yticklabels([arrayfun(@(n)[num2str(n) char(8201) repmat('b', 1, n~=0)], min_y:max_y, 'UniformOutput', false)])
+xlabel x, ylabel y
+axis([min_x*a max_x*a min_y*b max_y*b])
+set(gca, 'layer', 'top')
+plot([min_y*b*1j max_y*b*1j], 'k')
+plot(complex([min_x*a max_x*a]), 'k')
+plot(a*(n-1)+[(min_y*b+.7)*1j max_y*b*1j], 'k--', 'linewidth', .75)
+plot(-a*(n-2)+[(min_y*b+.7)*1j max_y*b*1j], 'k--', 'linewidth', .75)
+theta_range = [-63 65]/180*pi;
+theta = linspace(theta_range(1), theta_range(2), 1000);
+z = x1+1j*y1 + len*exp(1j*theta);
+set(gca, 'colororderindex', 1)
+plot(z)
+ind = real(z)>=a*(n-1);
+set(gca, 'colororderindex', 1)
+plot(z(ind), 'linewidth', 1.2)
+ang_quiver = 59.5/180*pi;
[…]

### 2021-08-02 paper.tex
Some discussion about average-number result

@@
 ``On Laplace's Extension of the Buffon Needle Problem'', B.J.~Arnow, 1994.
+\emph{An Introduction to Geometrical Probability}, A.M~Mathai, section 1.1.
 \\
@@
 \begin{proposition}
+\label{eq: prop Pr i}
 Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above.
@@
 If $a(n-1) < \len$, the length $\len$ is enough for $x_2$ to exceed $a(n-1)$ for some angles $\orient$, regardless of $x_1$.
-This is depicted in Figure~\ref{fig: Pr_i_full}, where the thicker line on the arc represents those angles for which $x_2 > a(n-1)$, for a specific $x_1$.
-If $\len \leq a(n-1) < \len+a$, the length is enough only if $x_1 \geq a(n-1)-\len$.
-This can be seen in Figure~\ref{fig: Pr_i_notfull}, where the thicker line again represents the angles for which this happens, for a specific $x_1 \geq a(n-1)-\len$.
+This is depicted in Figure~\ref{fig: Pr_i_full}, where the section of the arc with solid line represents those angles for which $x_2 > a(n-1)$.
+If $\len \leq a(n-1) < \len+a$, the length is enough only if $x_1 \geq a(n-1)-\len$, which corresponds to the shaded region in Figure~\ref{fig: Pr_i_notfull}.
 Lastly, if $\len+a \leq a(n-1)$ it is impossible that $x_2$ exceeds $a(n-1)$.
@@
 
-***Explain that the inverse question is straightforward, because $\funta$ is monotone increasing, thus it has an inverse function (defined on the range of $\funta$).
-In fact, as will be seen, $\funta$ is an affine function, and therefore it is immediate to obtain its inverse.
-
-***Some mention that there is no need to study square grids or the case of integer-valued lengths separately.
-
-***Relation the problem of the Buffon needle.
-On one hand, the formula for b=infinity gives the average number of lines crossed in the Buffon experiment.
-On the other hand, the intermediate values $\Pr[i \geq i_0]$ (proposition) are a generalization of the probability of crossing at least a line in the Buffon experiment, when the needle endpoint is only allowed to move in a smaller vertical strip.
-
-***Mention particular values: a=b=1 gives...; a=b=4/pi gives len+1. Can we find any ``meaning'' for these particular cases?
-
-***The function is affine, not linear. This is related to the fact that the parts of the segment near the endpoints can touch several tiles
+In view of Theorem~\ref{theo: funta, form}, the average number of touches tiles has a remarkably simple dependence on the segment length, namely an affine function.
+Conversely, for any $t>1$ it is immediate to compute the length of a random segment that touches $t$ tiles on average, given as $\funta^{-1}(t)$.
+
+The results in this section make clear the relationship between the problem considered here (average number of squares touched by a random segment of fixes length) and the Buffon's needle problem (probability that a random segment of fixed length crosses a line of a regular structure of parallel strips) \cite[section 1.1]{Mathai99}.
+Firstly, since $\pi$ is involved in \eqref{eq: funta, form}, it is possible to design a simple probabilistic experiment to estimate the value of $\pi$, as in Buffon's original experiment.
+For example, choosing $a=b=\len=1$ gives $\funta(\len) = 1+1/\pi$.
+
+Secondly, a grid with $b \rightarrow \infty$ corresponds to Buffon's arrangement of parallel strips with spacing $a$. Thus
+\begin{equation}
+\label{eq: number of lines crossed, Buffon}
+\lim_{b \rightarrow \infty} (\funta(\len)-1) = \frac{2\len}{pi a})
+\end{equation}
+gives the average number of lines crossed in Buffon's experiment.
+For $\len \leq a$ the segment can only cross one line, and \eqref{eq: number of lines crossed, Buffon} coincides with the probability of crossing \cite[section 1.1]{Mathai99}.
+
+Lastly, $\Pr[i \geq n]$ as computed in Proposition~\ref{eq: prop Pr i} for $n \geq 2$ can be seen as a generalization of the probability of crossing at least one line in Buffon's experiment, when a needle endpoint is only allowed to move in a narrower region of width $n-1$ times smaller than the width of the strips (instead of moving across the full strip as in Buffon's setting).
 

### 2021-08-02 figures/figures.m

@@ -1277,7 +1277,7 @@ a = 1.35; b = 1;
 x1 = .6*a; y1 = .72*b;
 n = 3;
 min_x = -2; max_x = 3; min_y = -3; max_y = 4;
-flag_full = false;
+flag_full = true;
 if flag_full
     len = 2.1*a;
 else
@@ -1304,10 +1304,10 @@ theta_range = [-63 65]/180*pi;
 theta = linspace(theta_range(1), theta_range(2), 1000);
 z = x1+1j*y1 + len*exp(1j*theta);
 set(gca, 'colororderindex', 1)
-plot(z)
+plot(z, '--', 'linewidth', .75)
 ind = real(z)>=a*(n-1);
 set(gca, 'colororderindex', 1)
-plot(z(ind), 'linewidth', 1.2)
+plot(z(ind), '-', 'linewidth', .75)
 ang_quiver = 59.5/180*pi;
 quiver(x1,y1,len*cos(ang_quiver),len*sin(ang_quiver),.99,'k')
 if flag_full

### 2021-08-02 paper.tex
Improved figures and some text

@@
 The segment orientation is a random angle $\orient$ uniformly distributed on $[0,2\pi)$.
-It is clear from the figure that $i \geq n$ if and only if either $x_2 \geq a(n-1)$ or $x_2 \leq -a(n-2)$.
-The latter two events are exclusive, and by symmetry have the same probability.
-Therefore
+It is clear from the figure that $i \geq n$ if and only if $x_2 \geq a(n-1)$ or $x_2 \leq -a(n-2)$; and these events are exclusive for $n \geq 2$.
+Thus
+\begin{equation}
+\label{eq: Pr i cond}
+\Pr[i \geq n \cond x_1] = \Pr[x_2 \geq a(n-1) \cond x_1] + \Pr[x_2 \leq -a(n-2) \cond x_1].
+\end{equation}
+The two conditional probabilities on the right-hand side of \eqref{eq: Pr i cond} are different in general.
+However, averaging over $x_1$ gives, by symmetry, $\Pr[x_2 \geq a(n-1)] = \Pr[x_2 \leq -a(n-2)]$; and consequently, for $n \geq 2$,
 \begin{equation}
@@
 \centering%
-\subfigure[$i$ can be equal or greater than $n$ for all $x_1$, $0 \leq x_1 \leq a$]{%
+\subfigure[$x_2$ can exceed $a(n-1)$ for all $x_1$, $0 \leq x_1 \leq a$]{%
 \label{fig: Pr_i_full}%
-\includegraphics[width=.48\textwidth]{Pr_i_full_1p35}%
-}\hfill%
-\subfigure[$i$ can be equal or greater than $n$ only for $a(n-1)-\len \leq x_1 \leq a$]{%
+\includegraphics[width=.58\textwidth]{Pr_i_full_1p35}%
+}\\%\hfill%
+\subfigure[$x_2$ can exceed $a(n-1)$ only if $a(n-1)-\len \leq x_1 \leq a$]{%
 \label{fig: Pr_i_notfull}%
-\includegraphics[width=.48\textwidth]{Pr_i_notfull_1p35}%
+\includegraphics[width=.58\textwidth]{Pr_i_notfull_1p35}%
 }%
@@
 
-Consider the event $x_2 \geq a(n-1)$.
+Consider the event $x_2 \geq a(n-1)$ conditioned on $x_1$, with $n \geq 2$.
 There are three possibilities depending on $x_1$, $n$ and $\len$.
-If $a(n-1) < \len$, the length $\len$ is enough for $x_2$ to exceed $a(n-1)$ for some angles $\orient$, regardless of $x_1$.
+If $a(n-1) < \len$, regardless of $x_1$ the length $\len$ is enough for $x_2$ to exceed $a(n-1)$ for some angles $\orient$.
 This is depicted in Figure~\ref{fig: Pr_i_full}, where the section of the arc with solid line represents those angles for which $x_2 > a(n-1)$.
-If $\len \leq a(n-1) < \len+a$, the length is enough only if $x_1 \geq a(n-1)-\len$, which corresponds to the shaded region in Figure~\ref{fig: Pr_i_notfull}.
-Lastly, if $\len+a \leq a(n-1)$ it is impossible that $x_2$ exceeds $a(n-1)$.
-Note that the value $y_1$ is irrelevant for this.
-
-In the first two cases above, the probability that $x_2 \geq a(n-1)$, conditioned on $x_1$, is the length of the arc to the right of the line $x=a(n-1)$ divided by $2\pi\len$; that is,
+If $\len \leq a(n-1) < \len+a$, the length can only be enough provided that $x_1 \geq a(n-1)-\len$, which corresponds to the shaded region in Figure~\ref{fig: Pr_i_notfull}.
+Lastly, if $\len+a \leq a(n-1)$ it is not possible for $x_2$ to exceed $a(n-1)$.
+Note that the coordinate $y_1$ is irrelevant to this.
+
+In the first two cases above, the probability that $x_2 \geq a(n-1)$, conditioned on $x_1$, is the length of the arc to the right of the line $x=a(n-1)$ divided by $2\pi\len$, that is,
 \begin{equation}
@@
 \end{equation}
-Integrating by parts, this is seen to be
+Integrating by parts,
 \begin{equation}
@@
 \label{eq: number of lines crossed, Buffon}
-\lim_{b \rightarrow \infty} (\funta(\len)-1) = \frac{2\len}{pi a})
+\lim_{b \rightarrow \infty} \funta(\len)-1 = \frac{2\len}{\pi a}
 \end{equation}
 gives the average number of lines crossed in Buffon's experiment.
-For $\len \leq a$ the segment can only cross one line, and \eqref{eq: number of lines crossed, Buffon} coincides with the probability of crossing \cite[section 1.1]{Mathai99}.
+For $\len \leq a$ the segment can cross at most one line, and \eqref{eq: number of lines crossed, Buffon} coincides with the probability of crossing \cite[section 1.1]{Mathai99}.
 
@@
 The asymptotic slopes are $\sqrt(1/a^2+1/b^2)$ for $\funt$ and $2/\pi(1/a+1/b)$.
-It may be interesting to consider the ratio of asymptotic slopes.
-Some quick plot of that ratio suggests it is maximum for $a=b$, and equals $sqrt(8)/pi = 0.9003$.
+It may be interesting to consider the ratio of asymptotic slopes, as a function of $a/b$.
+It is maximum for $a=b$, and equals $2\sqrt 2/\pi = 0.9003$.
+It tends to $2/\pi$ for $a/b \rightarrow \infty$ and for $a/b \rightarrow 0$.
 The minimum number of tiles that can be touched with non-zero probability (corresponding to an n times 1 canonical rectangle if a>=b) can also be plotted.

### 2021-08-02 figures/figures.m

@@ -1274,10 +1274,11 @@ plot(L, 2*L/pi*(1/a+1/b)+1)
 
 clear, clf
 a = 1.35; b = 1;
-x1 = .6*a; y1 = .72*b;
+x1 = .6*a; y1 = .69*b;
 n = 3;
 min_x = -2; max_x = 3; min_y = -3; max_y = 4;
-flag_full = true;
+flag_full = false;
+flag_segment = true;
 if flag_full
     len = 2.1*a;
 else
@@ -1286,6 +1287,9 @@ end
 m = max(0, a*(n-1)-len);
 patch([m a a m m], [0 0 b b 0], .88*[1 1 1], 'edgecolor', 'none')
 hold on
+if ~flag_full
+    plot([m m+1j*b], 'k:')
+end
 patch([0 a a 0 0], [0 0 b b 0], 1*[1 1 1], 'edgecolor', 'k', 'FaceAlpha', 0)
 plot(x1 + 1j*y1, 'k.', 'markersize', 7)
 grid on, axis equal
@@ -1308,14 +1312,79 @@ plot(z, '--', 'linewidth', .75)
 ind = real(z)>=a*(n-1);
 set(gca, 'colororderindex', 1)
 plot(z(ind), '-', 'linewidth', .75)
-ang_quiver = 59.5/180*pi;
-quiver(x1,y1,len*cos(ang_quiver),len*sin(ang_quiver),.99,'k')
+if flag_segment
+    if flag_full
+        ang_segment = 42.5/180*pi;
+    else
+        ang_segment = 19/180*pi;
+    end
+    plot([x1+1j*y1 x1+1j*y1+len*exp(1j*ang_segment)],'k')
+    plot(x1+1j*y1+len*exp(1j*ang_segment), 'k.', 'markersize', 7)
+else
+    ang_quiver = 59.5/180*pi;
+    quiver(x1,y1,len*cos(ang_quiver),len*sin(ang_quiver),.99,'k')
+end
 if flag_full
     text(0.44, 0.43, 'x_1,y_1')
-    text(1.52, 2.63, char(hex2dec('2113')))
+    if flag_segment
+        text(2.0, 2.28, char(hex2dec('2113')))
+        text(3.13, 2.75, 'x_2,y_2')
+    else
+        text(1.52, 2.63, char(hex2dec('2113')))
+    end
 else
-    text(0.52, 0.43, 'x_1,y_1')
[…]

### 2021-08-03 paper.tex
Minor corrections

@@
 
-Given $\len \in \mathbb R*+$, consider a segment of length $\len$ with random position and orientation.
-Specifically, the coordinates $x_1$, $y_1$ of the first endpoint are independent random variables uniformly distributed on $(0,a)$ and $(0,b)$ respectively.
-The orientation $\orient$ of the segment, defined as the angle with respect to the horizontal axis of the vector that goes from the first to the second endpoint, is uniformly distributed on $(0,2\pi)$.
+Given $\len \in \mathbb R^+$, consider a segment of length $\len$ with random position and orientation.
+Specifically, the coordinates $x_1$, $y_1$ of the first endpoint are independent random variables uniformly distributed on $[0,a)$ and $[0,b)$ respectively.
+The orientation $\orient$ of the segment, defined as the angle with respect to the horizontal axis of the vector that goes from the first to the second endpoint, is uniformly distributed on $[0,2\pi)$.
 The variables $x_1$, $y_1$ and $\theta$ determine the coordinates $x_2$, $y_2$ of the second endpoint.
@@
 
-Let the function $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles touched by a random segment of length $\len$, assuming the above distributions.
+Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles touched by a random segment of length $\len$, with the above distributions.
 
@@
 The random variables $i$ and $j$ are not independent.
-As a consequence, while the marginal distributions of $i$ and $j$ are easy to compute, that of $i+j$ is difficult.
-However, the average number of touched tiles can be obtained from the marginal distributions as
+As a consequence, while the marginal distributions of $i$ and $j$ are easy to compute, the distribution of $i+j$ is difficult.
+However, knowing the marginal distributions is sufficient to compute the average number of touched tiles:
 \begin{equation}
 \label{eq: funta E[i] E[j]}
-\funta(\len) = \E[t] = \E[i]+\E[j]+1.
+\funta(\len) = \E[t] = \E[i]+\E[j]-1.
 \end{equation}
@@
 \begin{proposition}
-\label{eq: prop Pr i}
+\label{prop: Pr i}
 Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above.
@@
 \end{equation}
-and $\Pr[j \geq n]$ is obtained from \eqref{eq: Pr i geq n} replacing $a$ by $b$.
+and $\Pr[j \geq n]$ is given by the same expressions with $a$ replaced by $b$.
 \end{proposition}
@@
 In the following it will be assumed that $n \geq 2$.
-The basic idea is to compute $\Pr[i \geq n]$ conditioned on $(x_1,y_1)$ (or, as will be seen, only on $y_1$), and then to average over $x_1$ and $y_1$ (actually only over $x_1$).
-
-Given the coordinates $(x_1, y_1)$ of the first segment endpoint, with $0 \leq x_1 \leq a$, $0 \leq y_1 \leq b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure~\ref{fig: Pr_i}.
+The basic idea is to compute $\Pr[i \geq n]$ conditioned on $(x_1,y_1)$ (or, as will be seen, only on $x_1$), and then to average over $x_1$ and $y_1$ (actually only over $x_1$).
+
+Given the coordinates $(x_1, y_1)$ of the first endpoint of the segment, with $0 \leq x_1 \leq a$, $0 \leq y_1 \leq b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure~\ref{fig: Pr_i}.
 The segment orientation is a random angle $\orient$ uniformly distributed on $[0,2\pi)$.
@@
 \label{eq: Pr i cond}
-\Pr[i \geq n \cond x_1] = \Pr[x_2 \geq a(n-1) \cond x_1] + \Pr[x_2 \leq -a(n-2) \cond x_1].
+\Pr[i \geq n \cond x_1, y_1] = \Pr[x_2 \geq a(n-1) \cond x_1, y_1] + \Pr[x_2 \leq -a(n-2) \cond x_1, y_1].
 \end{equation}
 The two conditional probabilities on the right-hand side of \eqref{eq: Pr i cond} are different in general.
-However, averaging over $x_1$ gives, by symmetry, $\Pr[x_2 \geq a(n-1)] = \Pr[x_2 \leq -a(n-2)]$; and consequently, for $n \geq 2$,
+However, averaging over $x_1, y_1$ gives, by symmetry, $\Pr[x_2 \geq a(n-1)] = \Pr[x_2 \leq -a(n-2)]$.
+In addition, the coordinate $y_1$ does not have any influence on these events, and therefore conditioning on $x_1, y_1$ is the same as conditioning on $x_1$.
+This implies that, for $n \geq 2$,
 \begin{equation}
 \label{eq: Pr i = 2 Pr}
-\Pr[i \geq n] = 2 \Pr[x_2 \geq a(n-1)].
+\Pr[i \geq n] = 2 \Pr[x_2 \geq a(n-1)] = 2 \E[\Pr[x_2 \geq a(n-1) \cond x_1]].
 \end{equation}
@@
 }%
-\caption{Conditions for the normalized width of the canonical rectangle, $i$, being equal or greater than a given $n$. Example with $a=1.35$, $b=1$
+\caption{Conditions for the normalized width of the canonical rectangle, $i$, to be equal or greater than a given $n$. Example with $a=1.35$, $b=1$, $n=3$
 }%
@@
 This is depicted in Figure~\ref{fig: Pr_i_full}, where the section of the arc with solid line represents those angles for which $x_2 > a(n-1)$.
-If $\len \leq a(n-1) < \len+a$, the length can only be enough provided that $x_1 \geq a(n-1)-\len$, which corresponds to the shaded region in Figure~\ref{fig: Pr_i_notfull}.
-Lastly, if $\len+a \leq a(n-1)$ it is not possible for $x_2$ to exceed $a(n-1)$.
+If $a(n-2) < \len \leq  a(n-1)$, the length will be enough provided that $x_1 \geq a(n-1)-\len$, which corresponds to the shaded region in Figure~\ref{fig: Pr_i_notfull}.
+Lastly, if $\len \leq  a(n-2)$ it is not possible for $x_2$ to exceed $a(n-1)$.
 Note that the coordinate $y_1$ is irrelevant to this.
@@
 \end{equation}
-In the first case $x_1$ has a uniform distribution on $(0,a)$, and $\Pr[x_2 \geq a(n-1)]$ is easily obtained from \eqref{eq: Pr x_2 geq a(n-1) cond} as follows:
-\begin{equation}
-\Pr[x_2 \geq a(n-1)] = \frac 1 {\pi a} \int_0^a\arccos \frac{a(n-1)-x_1}{\len} \diff x_1. 
-\end{equation}
-Integrating by parts,
+In the first case $x_1$ has a uniform distribution on $(0,a)$, and $\Pr[x_2 \geq a(n-1)]$ is easily obtained from \eqref{eq: Pr x_2 geq a(n-1) cond} as
 \begin{equation}
 \label{eq: Pr x_2 geq a(n-1), first case}
-\Pr[x_2 \geq a(n-1)] = \frac{\len}{\pi a}\,\, \left( g\left( \frac{a(n-1)}{\len} \right) - g\left( \frac{a(n-2)}{\len} \right) \right),
+\begin{split}
+\Pr[x_2 \geq a(n-1)] &= \frac 1 {\pi a} \int_0^a\arccos \frac{a(n-1)-x_1}{\len} \diff x_1 \\
+&= \frac{\len}{\pi a}\,\, \left( g\left( \frac{a(n-1)}{\len} \right) - g\left( \frac{a(n-2)}{\len} \right) \right),
+\end{split}
 \end{equation}
@@
 \end{equation}
-From \eqref{eq: Pr i geq n}, this sum only has a finite number of non-zero terms: $1$ for $n=1$, plus $\lceil \len/a \rceil-1$ terms given by the second line of \eqref{eq: Pr i geq n}, plus one term given by the third line, corresponding to $n = \lceil \len/a \rceil +1 $:
+According to \eqref{eq: Pr i geq n}, this sum only has a finite number of non-zero terms: $1$ for $n=1$, plus $\lceil \len/a \rceil-1$ terms given by the second line of \eqref{eq: Pr i geq n}, plus one term given by the third line, corresponding to $n = \lceil \len/a \rceil +1 $:
 \begin{equation}
@@
 
-In view of Theorem~\ref{theo: funta, form}, the average number of touches tiles has a remarkably simple dependence on the segment length, namely an affine function.
+In view of Theorem~\ref{theo: funta, form}, the average number of touched tiles has a remarkably simple dependence on the segment length, namely an affine function.
 Conversely, for any $t>1$ it is immediate to compute the length of a random segment that touches $t$ tiles on average, given as $\funta^{-1}(t)$.
 
-The results in this section make clear the relationship between the problem considered here (average number of squares touched by a random segment of fixes length) and the Buffon's needle problem (probability that a random segment of fixed length crosses a line of a regular structure of parallel strips) \cite[section 1.1]{Mathai99}.
-Firstly, since $\pi$ is involved in \eqref{eq: funta, form}, it is possible to design a simple probabilistic experiment to estimate the value of $\pi$, as in Buffon's original experiment.
+The results in Proposition~\ref{prop: Pr i} and Theorem~\ref{theo: funta, form} make clear the relationship between the problem considered here and Buffon's needle problem, i.e.~the probability that a random segment of fixed length crosses a line in a regular structure of parallel strips \cite[section 1.1]{Mathai99}.
+Firstly, since the number $\pi$ is involved in \eqref{eq: funta, form}, it is possible to design a simple probabilistic experiment to estimate the value of $\pi$, as in Buffon's original experiment.
 For example, choosing $a=b=\len=1$ gives $\funta(\len) = 1+1/\pi$.
@@
 
-Lastly, $\Pr[i \geq n]$ as computed in Proposition~\ref{eq: prop Pr i} for $n \geq 2$ can be seen as a generalization of the probability of crossing at least one line in Buffon's experiment, when a needle endpoint is only allowed to move in a narrower region of width $n-1$ times smaller than the width of the strips (instead of moving across the full strip as in Buffon's setting).
+Lastly, $\Pr[i \geq n]$ as computed in Proposition~\ref{prop: Pr i} for $n \geq 2$ can be seen as a generalization of the probability of crossing at least one line in Buffon's experiment, when a needle endpoint is only allowed to move in a narrower region of width $n-1$ times smaller than the width of the strips (instead of across the full strip as in Buffon's setting).
 

### 2021-08-12 paper.tex
Added comments regarding the word "touch"
"Visit" probably better than "touch"
Replaced "touch" etc by commands
The commands are defined as "visit" etc, but can be easily changed

@@
 \newcommand{\funt}{\tau} % function that gives maximum number of touched tiles for real-valued length
-\newcommand{\funl}{\lambda} % function that gices infimum real-valued length for a given number of touched tiles
+\newcommand{\funl}{\lambda} % function that gives infimum real-valued length for a given number of touched tiles
 \newcommand{\funti}{T} % function that gives maximum number of touched tiles for integer-valued length
@@
 \newcommand{\mss}{M}
+\newcommand{\touch}{visit}
+\newcommand{\touches}{visits}
+\newcommand{\touched}{visited}
+\newcommand{\touching}{visiting}
 
@@
 \title{
-On the number of tiles touched by a\\
+On the number of tiles \touched{} by a\\
 line segment on a rectangular grid
@@
 A line segment of length $\len \in \mathbb R^+$ is located on the plane with arbitrary position and orientation.
-The segment is said to \emph{touch} a tile if it intersects its interior.\footnote{
-The definition uses the interior of the tile, excluding the border, to avoid uninteresting results such as a ``zero-length'' segment touching (a vertex of) $4$ tiles.} This paper deals with the following problems:
+The segment is said to \emph{\touch} a tile if it intersects its interior.\footnote{
+The definition uses the interior of the tile, excluding the border, to avoid uninteresting results such as a ``zero-length'' segment \touching{} (a vertex of) $4$ tiles.} This paper deals with the following problems:
 \begin{itemize}
-\item What is the maximum number of tiles that the segment can touch?
-\item What length should a segment have to touch a given number of tiles?
-\item What is the average number of tiles touched by a uniformly random segment of a given length?
+\item What is the maximum number of tiles that the segment can \touch?
+\item What length should a segment have to \touch{} a given number of tiles?
+\item What is the average number of tiles \touched{} by a uniformly random segment of a given length?
 \end{itemize}
@@
 As an example of the first problem, consider $a=1.35$, $b=1$.
-A segment of unit length can be placed as shown in Figure~\ref{fig: examples} (left) to touch $3$ tiles.
+A segment of unit length can be placed as shown in Figure~\ref{fig: examples} (left) to make it \touch{} $3$ tiles.
 In fact, the solution for $\len=1$ is $3$ tiles.
@@
 An equivalent formulation of the problem is obtained allowing segments of length $\len$ \emph{or smaller}.
-The equivalence is clear from the fact that reducing the length cannot increase the number of touched tiles.
+The equivalence is clear from the fact that reducing the length cannot increase the number of \touched{} tiles.
 Either of these formulations will be referred to as the \emph{direct} problem.
 
-The \emph{inverse} problem is, given $\tiles \in \mathbb N$, to determine the infimum length of all segments that touch at least $\tiles$ tiles.
-For real-valued lengths this infimum is not a minimum, because given a segment that touches $\tiles$ tiles, its length can always be reduced by some non-zero amount without changing the number of touched tiles.
+The \emph{inverse} problem is, given $\tiles \in \mathbb N$, to determine the infimum length of all segments that \touch{} at least $\tiles$ tiles.
+For real-valued lengths this infimum is not a minimum, because given a segment that \touches{} $\tiles$ tiles, its length can always be reduced by some non-zero amount without changing the number of \touched{} tiles.
 This is a consequence of the interior of each tile being an open set.
@@
 The direct and inverse problems are closely related.
-Namely, if $\len$ is the infimum of all lengths that allow touching at least $\tiles$ tiles (inverse problem), $\tiles$ is the maximum number of tiles that can be touched with lengths slightly greater than $\len$ (direct problem).
-
-The problem of how many tiles a \emph{random} segment of a given length touches on average is precisely defined as follows.
+Namely, if $\len$ is the infimum of all lengths that allow \touching{} at least $\tiles$ tiles (inverse problem), $\tiles$ is the maximum number of tiles that can be \touched{} with lengths slightly greater than $\len$ (direct problem).
+
+The problem of how many tiles a \emph{random} segment of a given length \touches{} on average is precisely defined as follows.
 By symmetry, one endpoint of the segment can be assumed to lie in a fixed, reference tile.
@@
 The segment orientation has a \emph{uniform} distribution on the set of all possible directions, and is \emph{independent} of the endpoint position.
-As will be seen, solving this problem also answers the inverse question (segment length to touch a given number of tiles on average), which thus need not be considered separately.
+As will be seen, solving this problem also answers the inverse question (segment length to \touch{} a given number of tiles on average), which thus need not be considered separately.
 
@@
 The direct and inverse problems for a deterministic segment are considered in \S\ref{part: max}, first for real-valued lengths, and then for integer-valued lengths on a unit square grid.
-The problem of computing the average number of tiles touched by a random segment is analyzed in \S\ref{part: ave}.
+The problem of computing the average number of tiles \touched{} by a random segment is analyzed in \S\ref{part: ave}.
 
@@
 Two examples are illustrated in Figure~\ref{fig: canonical rectangle and touched tiles}, both with $i=5$, $j=4$.
-Clearly, all tiles touched by the segment are contained in the canonical rectangle.
+Clearly, all tiles \touched{} by the segment are contained in the canonical rectangle.
 Note also that the canonical rectangle can have $i=0$ or $j=0$ if the segment coincides with part of a grid line.
@@
 }%
-\caption{Canonical rectangle and touched tiles
+\caption{Canonical rectangle and \touched{} tiles
 }%
@@
 Consider an arbitrary segment, and let $i, j$ respectively denote the normalized width and height of its canonical rectangle.
-If $i, j \geq 1$, the number of tiles touched by the segment is at most $i+j-1$.
+If $i, j \geq 1$, the number of tiles \touched{} by the segment is at most $i+j-1$.
 This bound is attained if and only if the segment does not pass through any grid point in the interior of the rectangle.
@@
 \begin{proof}
-The segment touches, by definition, two tiles in opposite corners of the canonical rectangle.
+The segment \touches{}, by definition, two tiles in opposite corners of the canonical rectangle.
 It can be assumed, without loss of generality, that those tiles are in the lower-left and upper-right corners of the rectangle, as in Figure~\ref{fig: canonical rectangle and touched tiles}.
-The touched tiles can be thought of as following a path within the canonical rectangle.
+The \touched{} tiles can be thought of as following a path within the canonical rectangle.
 Starting at the lower-left tile, the next tile can be the one to the left, the one above, or the one above and to the left.
@@
 
-Since the segment follows a straight line, once it ``leaves'' a row of tiles in its path from the lower-left to the upper-right corner, it can never touch again a tile from that row.
+Since the segment follows a straight line, once it ``leaves'' a row of tiles in its path from the lower-left to the upper-right corner, it can never \touch{} any more tiles from that row.
 The same observation applies for the columns.
 
-This implies that the maximum number of touched tiles is $i+j-1$, which is attained if and only if the segment avoids all grid points in the interior of the canonical rectangle, as in Figure~\ref{fig: S_nogridpoints}.
+This implies that the maximum number of \touched{} tiles is $i+j-1$, which is attained if and only if the segment avoids all grid points in the interior of the canonical rectangle, as in Figure~\ref{fig: S_nogridpoints}.
 Note that grid points at the corners of the rectangle do not count for this; and that the segment cannot pass through any other grid points on the rectangle border, because that would imply $i=0$ or $j=0$.
@@
 
-According to Proposition~\ref{prop: i+j-1}, in order to maximize the number of touched tiles for a given length, the position and orientation of the segment should be chosen to obtain $i+j-1$ as large as possible, where $i$ and $j$ are the normalized dimensions of its canonical rectangle.
+According to Proposition~\ref{prop: i+j-1}, in order to maximize the number of \touched{} tiles for a given length, the position and orientation of the segment should be chosen to obtain $i+j-1$ as large as possible, where $i$ and $j$ are the normalized dimensions of its canonical rectangle.
 On the other hand, Proposition~\ref{prop: len ineq i j} restricts the $i, j$ values that can be achieved with a given length.
@@
 
-For instance, it is intuitively clear from Figure~\ref{fig: examples} that segment orientations near the vertical or horizontal directions (resulting in $i=1$ with large $j$, or $j=1$ with large $i$) will not maximize the number of touched tiles.
+For instance, it is intuitively clear from Figure~\ref{fig: examples} that segment orientations near the vertical or horizontal directions (resulting in $i=1$ with large $j$, or $j=1$ with large $i$) will not maximize the number of \touched{} tiles.
 % This is true for any a, b: 2 can achieved instead of 1 without compromising the other dimension.
@@
 In order to derive a general method to build a minimal sufficient set, it is insightful to consider two specific cases first, as will be presented in the following.
-It is also helpful to use the equivalent formulation of the direct problem referred to in \S\ref{part: intro}, i.e.\ finding the maximum number of tiles touched by segments of length $\len$ \emph{or less}.
-
-Thus each choice of $(i, j)$ implies a maximum of $i+j-1$ touched tiles on one hand; and sets a lower bound on the required length $\len$ on the other hand.
+It is also helpful to use the equivalent formulation of the direct problem referred to in \S\ref{part: intro}, i.e.\ finding the maximum number of tiles \touched{} by segments of length $\len$ \emph{or less}.
+
+Thus each choice of $(i, j)$ implies a maximum of $i+j-1$ \touched{} tiles on one hand; and sets a lower bound on the required length $\len$ on the other hand.
 
@@
 \includegraphics[width=.7\textwidth]{ijLS_1}%
-\caption{Relationship between segment length and number of touched tiles with the width and height of the canonical rectangle, for $a=b$
+\caption{Relationship between segment length and number of \touched{} tiles with the width and height of the canonical rectangle, for $a=b$
 }%
@@
 For a given $\len \in \mathbb R^+$, the $(i,j)$ pairs that can be achieved with segments of length not exceeding $\len$ are, by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, those whose distance from $(2a,2b)$ is less than $\len$, i.e.~those delimited by the corresponding arc.
-It is clear from the figure that for $i+j-1$ odd, if some point $(i_1,j_1)$ with $|i_1-j_1|>1$ is achievable with the length bound $\len$, then necessarily there is a point $(i_0,i_0)$ with the same number of touched tiles $i+j-1$ (same diagonal line) that is also achievable with the same length restriction (because it has smaller distance to $(2a,2b)$).
+It is clear from the figure that for $i+j-1$ odd, if some point $(i_1,j_1)$ with $|i_1-j_1|>1$ is achievable with the length bound $\len$, then necessarily there is a point $(i_0,i_0)$ with the same number of \touched{} tiles $i+j-1$ (same diagonal line) that is also achievable with the same length restriction (because it has smaller distance to $(2a,2b)$).
 Similarly, for $i+j-1$ even, if a point $(i_1,j_1)$ with $|i_1-j_1|>1$ is achievable, then there is a point $(i_0,i_0+1)$ in the same diagonal that is also achievable with the same length restriction.
@@
 The pairs in the minimal sufficient set are marked with filled circles in Figure~\ref{fig: ijLS_1}.
-Given a length, the maximum number of touched tiles will be achieved with one of those pairs, namely that in the diagonal line furthest from $(2a,2b)$ that is within the circle with radius $\len$ centered at that point.
+Given a length, the maximum number of \touched{} tiles will be achieved with one of those pairs, namely that in the diagonal line furthest from $(2a,2b)$ that is within the circle with radius $\len$ centered at that point.
 
@@
 \includegraphics[width=.7\textwidth]{ijLS_1p35}%
-\caption{Relationship between segment length and number of touched tiles with the width and height of the canonical rectangle, for $a=1.35b$
+\caption{Relationship between segment length and number of \touched{} tiles with the width and height of the canonical rectangle, for $a=1.35b$
 }%
@@
 For each $\tiles$, the pair $(i_\tiles,j_\tiles)$ of the minimal sufficient set should be chosen as that on the line $i+j-1=\tiles$ which minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$.
-This allows maximizing the sum $i+j-1$, and thus the number of touched tiles, for a given length restriction (direct problem); or minimizing the required lengths for a specified number of touched files (inverse problem).
+This allows maximizing the sum $i+j-1$, and thus the number of \touched{} tiles, for a given length restriction (direct problem); or minimizing the required lengths for a specified number of \touched{} files (inverse problem).
 
@@
 Therefore the optimal slope for long segments is approximately $a/b$.
-This formalizes the intuition that to maximize the number of touched tiles, the segment should follow a direction along which the perceived ``length'' of the tile is smaller.
+This formalizes the intuition that to maximize the number of \touched{} tiles, the segment should follow a direction along which the perceived ``length'' of the tile is smaller.
 
@@
 \label{part: max}
-% ***Is it clear enough that "maximum number of tiles" can refer to the inverse problem too? (because the inverse problem is closely related to the direct one, as discussed in the introduction)
-% If not, an alernative would be a more explicit title like: "Maximum number of tiles fo a given length. Conditions on length for a given number of tiles"
 
@@
 
-For a grid with parameters $a, b \in \mathbb R^+$, the maximum number $\tiles$ of touched tiles with a given real-valued length $\len$ can be represented by a function $\funt: \mathbb R^+ \to \mathbb N$ such that $\tiles = \funt(\len)$.
-Similarly, for the inverse problem a function $\funl: \mathbb N \to \mathbb R^+$ can be defined such that $\funl(\tiles)$ gives the infimum length of all segments that touch at least $\tiles$ tiles.
-
-The function $\funt$ can be obtained from $\funl$ by noting that the maximum number of tiles that can be touched by a segment of length $\len$ is the index of the largest term of the sequence $\funl(\tiles)$ that is less than $\len$:
+For a grid with parameters $a, b \in \mathbb R^+$, the maximum number $\tiles$ of \touched{} tiles with a given real-valued length $\len$ can be represented by a function $\funt: \mathbb R^+ \to \mathbb N$ such that $\tiles = \funt(\len)$.
+Similarly, for the inverse problem a function $\funl: \mathbb N \to \mathbb R^+$ can be defined such that $\funl(\tiles)$ gives the infimum length of all segments that \touch{} at least $\tiles$ tiles.
+
+The function $\funt$ can be obtained from $\funl$ by noting that the maximum number of tiles that can be \touched{} by a segment of length $\len$ is the index of the largest term of the sequence $\funl(\tiles)$ that is less than $\len$:
 \begin{equation}
@@
 Then there is $\epsilon > 0$ for which $\funt(\len+\epsilon) = \tiles = \funt(\len-\epsilon)$.
-This means that there exists a segment with length $\len-\epsilon$ that touches $\tiles$ tiles, and thus $\funl(\tiles) \leq \len-\epsilon < \len$, in contradiction with \eqref{eq: theo funt, form: prop 1}.
+This means that there exists a segment with length $\len-\epsilon$ that \touches{} $\tiles$ tiles, and thus $\funl(\tiles) \leq \len-\epsilon < \len$, in contradiction with \eqref{eq: theo funt, form: prop 1}.
 Therefore $\funt$ is discontinuous, from the right, at $\len$.
@@
 \begin{proof}
-The $(i_\tiles,j_\tiles)$ pair in the minimal sufficient set $\mss$ corresponds to a maximum of $\tiles$ touched tiles.
-By construction of this set, any segment that touches $\tiles$ tiles must have length greater than $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$.
+The $(i_\tiles,j_\tiles)$ pair in the minimal sufficient set $\mss$ corresponds to a maximum of $\tiles$ \touched{} tiles.
+By construction of this set, any segment that \touches{} $\tiles$ tiles must have length greater than $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$.
 
@@
 
-Theorems~\ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the two problems stated in \S\ref{part: intro}; they also provide a way to actually position a segment of length slightly greater than $\len$ or $\funl(\tiles)$ so that it touches $\tiles$ or $\funt(\len)$ tiles.
+Theorems~\ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the two problems stated in \S\ref{part: intro}; they also provide a way to actually position a segment of length slightly greater than $\len$ or $\funl(\tiles)$ so that it \touches{} $\tiles$ or $\funt(\len)$ tiles.
 Namely, the segment should have its endpoints respectively in the interior of two tiles separated $\isoli-1$ steps horizontally and $\jsoli-1$ steps vertically, with its exact position and orientation  adjusted to avoid any grid points.
@@
 % *!*
-it is seen that for long segments the number of touched tiles and the segment length are approximately proportional, with
+it is seen that for long segments the number of \touched{} tiles and the segment length are approximately proportional, with
 \begin{equation}
@@
 
-The \emph{inverse problem} with integer-length segments can be formulated as follows: given $\tiles \in \mathbb N$, find the \emph{minimum} integer length that allows touching at least $\tiles$ tiles.
+The \emph{inverse problem} with integer-length segments can be formulated as follows: given $\tiles \in \mathbb N$, find the \emph{minimum} integer length that allows \touching{} at least $\tiles$ tiles.
 Observe that in this case, unlike with real-valued lengths, there is indeed a minimum length, because every subset of $\mathbb N$ has a minimum.
@@
 
-\section{Average number of tiles touched by a random segment}
+\section{Average number of tiles \touched{} by a random segment}
 \label{part: ave}
@@
 The variables $x_1$, $y_1$ and $\theta$ determine the coordinates $x_2$, $y_2$ of the second endpoint.
-Note that events of probability $0$, such as $x_1=a$ or the segment passing through a grid point, can be disregarded for computing the average number of tiles touched by the segment.
-
-Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles touched by a random segment of length $\len$, with the above distributions.
+Note that events of probability $0$, such as $x_1=a$ or the segment passing through a grid point, can be disregarded for computing the average number of tiles \touched{} by the segment.
+
+Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles \touched{} by a random segment of length $\len$, with the above distributions.
 
@@
 The normalized dimensions $i$ and $j$ of the canonical rectangle are thus random variables.
-Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and the segment touches $t = i+j-1$ tiles.
+Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and the segment \touches{} $t = i+j-1$ tiles.
 
@@
 As a consequence, while the marginal distributions of $i$ and $j$ are easy to compute, the distribution of $i+j$ is difficult.
-However, knowing the marginal distributions is sufficient to compute the average number of touched tiles:
+However, knowing the marginal distributions is sufficient to compute the average number of \touched{} tiles:
 \begin{equation}
@@
 Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above.
-The average number of tiles touched by the segment, $\funta(\len)$, is
+The average number of tiles \touched{} by the segment, $\funta(\len)$, is
 \begin{equation}
@@
 
-In view of Theorem~\ref{theo: funta, form}, the average number of touched tiles has a remarkably simple dependence on the segment length, namely an affine function.
-Conversely, for any $t>1$ it is immediate to compute the length of a random segment that touches $t$ tiles on average, given as $\funta^{-1}(t)$.
+In view of Theorem~\ref{theo: funta, form}, the average number of \touched{} tiles has a remarkably simple dependence on the segment length, namely an affine function.
+Conversely, for any $t>1$ it is immediate to compute the length of a random segment that \touches{} $t$ tiles on average, given as $\funta^{-1}(t)$.
 
@@
 It tends to $2/\pi$ for $a/b \rightarrow \infty$ and for $a/b \rightarrow 0$.
-The minimum number of tiles that can be touched with non-zero probability (corresponding to an n times 1 canonical rectangle if a>=b) can also be plotted.
+The minimum number of tiles that can be \touched{} with non-zero probability (corresponding to an n times 1 canonical rectangle if a>=b) can also be plotted.
 

### 2021-08-15 figures/figures.m
Average number of visited cells, experimentally, for arbitrary number of dimensions

@@ -1388,3 +1388,25 @@ for k = 1:numel(a_all)
 end
 xlabel(char(hex2dec('2113'))), ylabel([char(hex2dec('03C4')) '(' char(hex2dec('2113')) ')']), grid on
 
+
+%% #55. Like #49 but for arbitrary number of dimensions: average number of visited cells, experimentally
+
+clear, clc
+L = .1:.1:9;
+a = [1.35 1 .8];
+N = 1e5;
+ndim = numel(a);
+result = NaN(size(L));
+for iter_L = 1:numel(L)
+    dir = 2*rand(N, ndim)-1;
+    ind = sum(dir.^2, 2) <= 1;
+    dir = dir(ind,:);
+    dir = dir./sqrt(sum(dir.^2, 2));
+    %plot3(dir(:,1), dir(:,2), dir(:,3), '.', 'markersize', 2), axis equal
+    z0 = rand(size(dir)).*a(:).';
+    z1 = z0 + L(iter_L)*dir;
+    ii = ceil(abs(z1./a)) + (z1<0); % add 1 for negative
+    result(iter_L) = mean(sum(ii, 2)) - ndim + 1;
+end
+plot(L, result, '.-'), grid on, axis equal
+

### 2021-08-15 figures/figures.m
Counting average number of visited cells in d dimensions; another method

@@ -1394,7 +1394,7 @@ xlabel(char(hex2dec('2113'))), ylabel([char(hex2dec('03C4')) '(' char(hex2dec('2
 clear, clc
 L = .1:.1:9;
 a = [1.35 1 .8];
-N = 1e5;
+N = 1e4;
 ndim = numel(a);
 result = NaN(size(L));
 for iter_L = 1:numel(L)
@@ -1407,6 +1407,35 @@ for iter_L = 1:numel(L)
     z1 = z0 + L(iter_L)*dir;
     ii = ceil(abs(z1./a)) + (z1<0); % add 1 for negative
     result(iter_L) = mean(sum(ii, 2)) - ndim + 1;
+end
+plot(L, result, '.-'), grid on, axis equal
+
+
+%% #56. Like #55 but cells are counted using deterministic uniform sampling along the segment
+% Checked: gives simular results as #55
+
+clear, clc
+L = .1:.1:5;
+a = [1.35 1 .8];
+N = 1e4;
+n_points_along = 1000;
+ndim = numel(a);
+t = linspace(0, 1, n_points_along);
+result = NaN(size(L));
+for iter_L = 1:numel(L)
+    dir = 2*rand(N, ndim)-1;
+    ind = sum(dir.^2, 2) <= 1;
+    dir = dir(ind,:);
+    dir = dir./sqrt(sum(dir.^2, 2));
+    %plot3(dir(:,1), dir(:,2), dir(:,3), '.', 'markersize', 2), axis equal
+    z0 = rand(size(dir)).*a(:).';
+    z1 = z0 + L(iter_L)*dir;
+    size_unique  = 0;
+    for n = 1:size(z0,1)
+        size_unique = size_unique + ...
+            size(unique(floor(z0(n,:).*(1-t(:)) + z1(n,:).*t(:)), 'rows'), 1);
+    end
+    result(iter_L) = size_unique/size(z0,1);
 end
 plot(L, result, '.-'), grid on, axis equal
 

### 2021-08-19 figures/figures.m
Probability of maximum number of files: code

@@ -1444,6 +1444,157 @@ case 2
 case 3
     plot(L, 1+L/2*sum(1./a), '-')
 end
+% There's a general formula for the coefficient (Jacuwes, 18; found by Alex)
 
+
+%% #57. Probability of visiting the maximum number of tiles, experimentally. Based on #47 and #49
+
+clear, clc
+L = 3.8:.1:4.2; %3.8:.1:4.2;%3:.1:3.5;
+N = 1e7;
+ii = ceil(L/sqrt(2)) + 1;
+jj = ceil(sqrt(L.^2 - (ii-2).^2 )) + 1;
+funt_real = ii+jj-1 % formula valid for real-valued lengths
+z0 = rand(N, numel(L)) + 1j*rand(N, numel(L));
+z1 = z0 + L.*exp(1j*2*pi*rand(N, numel(L)));
+ii = ceil(abs(real(z1))) + (real(z1)<0); % add 1 for negative
+jj = ceil(abs(imag(z1))) + (imag(z1)<0); % add 1 for negative
+result = mean(ii+jj-1 == funt_real , 1)
+
+
+%% #58. Computation for #57, for odd values of the maximum number of tiles. Page 21. 
+% This should be run after #57. funt_real must contain odd values
+% Checked with #57
+
+clc
+result_computed = NaN(size(result));
+for n = 1:numel(L)
+    k = funt_real(n)-1;
+    assert(mod(k,2)==0)
+    %alpha = acos((k/2-1)/L(n))-pi/4;
+    %result_computed(n) = (alpha*(k/2-1)^2 - L(n)^2/4*(cos(2*(pi/4+alpha))-cos(2*pi/4)) - L(n)*(k/2-1)*(-cos(pi/4+alpha)+sin(pi/4+alpha)+cos(pi/4)-sin(pi/4))) * 4/pi;
+    %result_computed(n) = integral(@(theta) (1-k/2+L(n)*sin(theta)).*(1-k/2+L(n)*cos(theta)), pi/4, pi/4+alpha) * 4/pi;
+    alphap = acos((k/2-1)/L(n));
+    %result_computed(n) = ((alphap-pi/4)*(k/2-1)^2 - L(n)^2/4*cos(2*alphap) + L(n)*(k/2-1)*(cos(alphap)-sin(alphap))) * 4/pi;
[…]

### 2021-08-19 paper.tex
Ranges of length in the square case
This will be used for the probability of maximum

@@
 
+From Theorem~\ref{theo: funt, sq, form} it is easily seen that odd values of $\funt(\len)$ correspond to $\isoli = \jsoli$, whereas even values are achieved with $\jsoli = \isoli+1$.
+Given $\tiles \geq 3$ with $t$ odd, $\funt(\len) = \tiles$ if and only if $\len \in ((\tiles-3)/\sqrt{2}, \sqrt{(\tiles^2-4\tiles+5)/2}]$.
+For $t$ even, $\funt(\len) = \tiles$ if and only if $\len \in (\sqrt{(\tiles^2-6\tiles+10)/2}, (\tiles-2)/\sqrt{2}]$.
+
 \subsection{Unit square grid with integer lengths}

### 2021-08-21 paper.tex
Get rid of "\orient" command

--- 2021-08-12 paper.tex
@@
 \newcommand{\genvar}{r}
-\newcommand{\orient}{\theta}
 \newcommand{\mss}{M}
@@
 
+From Theorem~\ref{theo: funt, sq, form} it is easily seen that odd values of $\funt(\len)$ correspond to $\isoli = \jsoli$, whereas even values are achieved with $\jsoli = \isoli+1$.
+Given $\tiles \geq 3$ with $t$ odd, $\funt(\len) = \tiles$ if and only
+\begin{equation}
+\label{eq: square, odd max tiles, lengths}
+\frac \len a \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{ \sqrt{(\tiles^2-2)^2+1}}{\sqrt{2}} \right].
+\end{equation}
+For $\tiles \geq 3$, $t$ even, $\funt(\len) = \tiles$ if and only if
+\begin{equation}
+\label{eq: square, even max tiles, lengths}
+\frac \len a \in \left(\frac{ \sqrt{(\tiles^2-3)^2+1}}{\sqrt{2}}, \frac{\tiles-2}{\sqrt{2}} \right].
+\end{equation}
+
 \subsection{Unit square grid with integer lengths}
@@
 Specifically, the coordinates $x_1$, $y_1$ of the first endpoint are independent random variables uniformly distributed on $[0,a)$ and $[0,b)$ respectively.
-The orientation $\orient$ of the segment, defined as the angle with respect to the horizontal axis of the vector that goes from the first to the second endpoint, is uniformly distributed on $[0,2\pi)$.
+The orientation $\theta$ of the segment, defined as the angle with respect to the horizontal axis of the vector that goes from the first to the second endpoint, is uniformly distributed on $[0,2\pi)$.
 The variables $x_1$, $y_1$ and $\theta$ determine the coordinates $x_2$, $y_2$ of the second endpoint.
@@
 Given the coordinates $(x_1, y_1)$ of the first endpoint of the segment, with $0 \leq x_1 \leq a$, $0 \leq y_1 \leq b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure~\ref{fig: Pr_i}.
-The segment orientation is a random angle $\orient$ uniformly distributed on $[0,2\pi)$.
+The segment orientation is a random angle $\theta$ uniformly distributed on $[0,2\pi)$.
 It is clear from the figure that $i \geq n$ if and only if $x_2 \geq a(n-1)$ or $x_2 \leq -a(n-2)$; and these events are exclusive for $n \geq 2$.
@@
 There are three possibilities depending on $x_1$, $n$ and $\len$.
-If $a(n-1) < \len$, regardless of $x_1$ the length $\len$ is enough for $x_2$ to exceed $a(n-1)$ for some angles $\orient$.
+If $a(n-1) < \len$, regardless of $x_1$ the length $\len$ is enough for $x_2$ to exceed $a(n-1)$ for some angles $\theta$.
 This is depicted in Figure~\ref{fig: Pr_i_full}, where the section of the arc with solid line represents those angles for which $x_2 > a(n-1)$.

### 2021-08-21 paper.tex
Removed the exact definition of the segment orientation
because a different one will be used in the proof of the "probability of maximum" result, and the exact definition is not important here
Changed function names. There will be a "g" in the "probability of maximum" result

@@
 \newcommand{\jsoli}{j^\ast}
+\newcommand{\genfun}{u}
 \newcommand{\genvar}{r}
@@
 
-In order to prove that $\funti(\leni+1)-\funti(\leni) \in \{1,2\}$, consider the function $f(\genvar) = \sqrt{2\genvar^2-2}$ for $\genvar \in \mathbb R$, $\genvar>1$. Its first derivative is
-\begin{equation}
-f'(\genvar) = \frac {\sqrt{2} \, \genvar} {\sqrt{\genvar^2-1}},
-\end{equation}
-and its second derivative is easily seen to be negative. Therefore $f'(\genvar)$ can be bounded for $\genvar \geq 3$ as
+In order to prove that $\funti(\leni+1)-\funti(\leni) \in \{1,2\}$, consider the function $\genfun(\genvar) = \sqrt{2\genvar^2-2}$ for $\genvar \in \mathbb R$, $\genvar>1$. Its first derivative is
+\begin{equation}
+\genfun'(\genvar) = \frac {\sqrt{2} \, \genvar} {\sqrt{\genvar^2-1}},
+\end{equation}
+and its second derivative is easily seen to be negative. Therefore $\genfun'(\genvar)$ can be bounded for $\genvar \geq 3$ as
 \begin{equation}
 \label{eq: der bounds}
-\lim_{\genvar \rightarrow \infty} f'(\genvar) = \sqrt{2} < f'(\genvar) < f'(3) = 3/2.
-\end{equation}
-For $\leni \geq 2$, by the mean value theorem \cite[section~5.3]{Abbott15}, when $\leni$ is increased to $\leni+1$ the term $\sqrt{2\leni^2-2}$ in \eqref{eq: funti leni} has an increment that equals $f'(\genvar)$ for some $\leni < \genvar < \leni+1$.
+\lim_{\genvar \rightarrow \infty} \genfun'(\genvar) = \sqrt{2} < \genfun'(\genvar) < \genfun'(3) = 3/2.
+\end{equation}
+For $\leni \geq 2$, by the mean value theorem \cite[section~5.3]{Abbott15}, when $\leni$ is increased to $\leni+1$ the term $\sqrt{2\leni^2-2}$ in \eqref{eq: funti leni} has an increment that equals $\genfun'(\genvar)$ for some $\leni < \genvar < \leni+1$.
 Therefore
@@
 Specifically, the coordinates $x_1$, $y_1$ of the first endpoint are independent random variables uniformly distributed on $[0,a)$ and $[0,b)$ respectively.
-The orientation $\theta$ of the segment, defined as the angle with respect to the horizontal axis of the vector that goes from the first to the second endpoint, is uniformly distributed on $[0,2\pi)$.
+The orientation $\theta$ of the segment is uniformly distributed on $[0,2\pi)$.
 The variables $x_1$, $y_1$ and $\theta$ determine the coordinates $x_2$, $y_2$ of the second endpoint.
@@
 \begin{equation}
-\label{eq: g}
-g(\genvar) = \int_0^r \arccos z \diff z = \genvar \arccos \genvar - \sqrt{1-\genvar^2} + 1.
+\label{eq: f}
+f(\genvar) = \int_0^r \arccos z \diff z = \genvar \arccos \genvar - \sqrt{1-\genvar^2} + 1.
 \end{equation}
@@
 \displaystyle
-\frac{2\len}{\pi a} \left(g\left(\frac{a(n-1)}{\len}\right)-g\left(\frac{a(n-2)}{\len} \right)\right) &\text{if\ \ } \displaystyle 2 \leq n < \frac \len a + 1 \\[4 mm]
+\frac{2\len}{\pi a} \left(f\left(\frac{a(n-1)}{\len}\right)-f\left(\frac{a(n-2)}{\len} \right)\right) &\text{if\ \ } \displaystyle 2 \leq n < \frac \len a + 1 \\[4 mm]
 \displaystyle
-\frac{2\len}{\pi a}\,\, \left(1 - g\left(\frac{a(n-2)}{\len}\right)\right) &\text{if\ \ } \displaystyle \frac\len a + 1 \leq n < \frac\len a+2 \\[3 mm]
+\frac{2\len}{\pi a}\,\, \left(1 - f\left(\frac{a(n-2)}{\len}\right)\right) &\text{if\ \ } \displaystyle \frac\len a + 1 \leq n < \frac\len a+2 \\[3 mm]
 \displaystyle
@@
 \Pr[x_2 \geq a(n-1)] &= \frac 1 {\pi a} \int_0^a\arccos \frac{a(n-1)-x_1}{\len} \diff x_1 \\
-&= \frac{\len}{\pi a}\,\, \left( g\left( \frac{a(n-1)}{\len} \right) - g\left( \frac{a(n-2)}{\len} \right) \right),
+&= \frac{\len}{\pi a}\,\, \left( f\left( \frac{a(n-1)}{\len} \right) - f\left( \frac{a(n-2)}{\len} \right) \right),
 \end{split}
 \end{equation}
-where the function $g$ is defined in \eqref{eq: g}. Substituting into \eqref{eq: Pr i = 2 Pr} yields the result in \eqref{eq: Pr i geq n}, second line.
-
-The second case is similar, but the integration over $x_1$ is from $a(n-1)-\len$ to $a$. Noting that $g(1)=1$, this gives
+where the function $f$ is defined in \eqref{eq: f}. Substituting into \eqref{eq: Pr i = 2 Pr} yields the result in \eqref{eq: Pr i geq n}, second line.
+
+The second case is similar, but the integration over $x_1$ is from $a(n-1)-\len$ to $a$. Noting that $f(1)=1$, this gives
 \begin{equation}
 \label{eq: Pr x_2 geq a(n-1), second case}
-\Pr[x_2 \geq a(n-1)] = \frac{\len}{\pi a}\,\, \left( 1 - g\left( \frac{a(n-2)}{\len} \right) \right),
+\Pr[x_2 \geq a(n-1)] = \frac{\len}{\pi a}\,\, \left( 1 - f\left( \frac{a(n-2)}{\len} \right) \right),
 \end{equation}
@@
 \E[i] = 1 +
-\sum_{n=2}^{\lceil \len/a \rceil} \frac{2\len}{\pi a} \left(g\left(\frac{a(n-1)}{\len}\right)-g\left(\frac{a(n-2)}{\len} \right)\right) + \\
-\frac{2\len}{\pi a} \left(1-g\left(\frac{a(\lceil \len/a \rceil - 1)}{\len} \right)\right).
+\sum_{n=2}^{\lceil \len/a \rceil} \frac{2\len}{\pi a} \left(f\left(\frac{a(n-1)}{\len}\right)-f\left(\frac{a(n-2)}{\len} \right)\right) + \\
+\frac{2\len}{\pi a} \left(1-f\left(\frac{a(\lceil \len/a \rceil - 1)}{\len} \right)\right).
 \end{split}
 \end{equation}
-Most terms involving the function $g$ in \eqref{eq: theo funta form 2} cancel, with the result
+Most terms involving the function $f$ in \eqref{eq: theo funta form 2} cancel, with the result
 \begin{equation}
 \label{eq: theo funta form 3}
-\E[i] = 1+\frac{2\len}{\pi a} (1-g(0))= 1+\frac{2\len}{\pi a}.
+\E[i] = 1+\frac{2\len}{\pi a} (1-f(0))= 1+\frac{2\len}{\pi a}.
 \end{equation}

### 2021-08-21 figures/figures.m
Updated figures for the proof of "probability of maximum" result

@@ -1622,21 +1622,21 @@ for k = [0 1+3j 2+2j 3+1j 1-3j 2-2j 3-1j -3+1j -2+2j -1+3j -3-1j -2-2j -1-3j]
 end
 
 
-%% #62. Figure for proof of probmax: funt odd
+%% #63. Figure for proof of probmax: funt odd, diagonal
 clf
 hold on
 set(gca, 'layer', 'top')
 axis equal
 grid on
-plot([-.4 4.6], [0 0], 'k')
-plot([-.4j 4.6j], 'k')
+plot([-.4 5.55], [0 0], 'k')
+plot([-.4j 5.55j], 'k')
 set(gca, 'GridAlpha', .4)
-xticks(-1:4), yticks(-1:4)
+xticks(-1:5), yticks(-1:5)
 xticklabels({'' '0' '1' '' ['(t' char(hex2dec('2212')) '1)/2'] '' ''})
 yticklabels({'' '0' '1' '' ['(t' char(hex2dec('2212')) '1)/2'] '' ''})
 xlabel x 
 ylabel y
-axis([-1 5 -1 5])
+axis([-1 6 -1 6])
 L = 3.37;
 theta_0 = asin(2/L);
 theta = 3*pi/2 - linspace(30*pi/180,theta_0,50);
@@ -1649,8 +1649,8 @@ theta = 3*pi/2 - linspace(theta_0,theta_1,50);
 set(gca, 'colororderindex', 1)
 plot(3+3j + L*exp(1j*theta), '-', 'linewidth', .75)
 plot([3+3j], 'k.', 'markersize', 7)
-theta = theta_0 + (theta_1-theta_0)*.67;
-endpoint = 3+3j + L*exp(1j*(3*pi/2 - theta));
+theta_example = theta_0 + (theta_1-theta_0)*.67;
+endpoint = 3+3j + L*exp(1j*(3*pi/2 - theta_example));
 plot(endpoint, 'k.', 'markersize', 7)
 plot([endpoint real(endpoint)+1j], 'k:')
 plot([endpoint 1+1j*imag(endpoint)], 'k:')
@@ -1663,7 +1663,56 @@ for k = [0 3+3j]
         plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k--', 'linewidth', .75)
     end
 end
-text(1.69, 2.17, char(hex2dec('2113')))
+text(1.65, 2.27, char(hex2dec('2113')))
[…]

### 2021-08-22 figures/figures.m
Checked formula for probmax with g function

@@ -1716,3 +1716,41 @@ plot([2+4j 2+3.45j], 'k:')
 theta = 3*pi/2 - linspace(0,theta_example);
 plot(2+4j + .35*exp(1j*theta), 'k-')
 text(1.81, 3.45, char(hex2dec('03B8')))
+
+
+%% #65. Like #60 but with the g function
+% Checked
+
+clear, clc
+a = 1;
+L = .1:.1:10; %3.8:.1:4.2;%3:.1:3.5;
+N = 3e5;
+ii = ceil(L/sqrt(2)) + 1;
+jj = ceil(sqrt(L.^2 - (ii-2).^2 )) + 1;
+funt = ii+jj-1 % formula valid for real-valued lengths
+z0 = rand(N, numel(L)) + 1j*rand(N, numel(L));
+z1 = z0 + L.*exp(1j*2*pi*rand(N, numel(L)));
+ii = ceil(abs(real(z1))) + (real(z1)<0); % add 1 for negative
+jj = ceil(abs(imag(z1))) + (imag(z1)<0); % add 1 for negative
+result = mean(ii+jj-1 == funt , 1);
+plot(L, result, 'o', 'markersize', 4), hold on, grid on
+
+g = @(r,u,v) ((acos(u/2./r)-asin(v/2./r)).*u.*v + 2*r.^2 + u.^2/2 + v.^2/2 - u.*sqrt(4*r.^2-v.^2) - v.*sqrt(4*r.^2-u.^2))/2/pi;
+L_fine = linspace(.01,10,1e4);
+ii = ceil(L_fine/sqrt(2)) + 1;
+jj = ceil(sqrt(L_fine.^2 - (ii-2).^2 )) + 1;
+funt_fine = ii+jj-1; % formula valid for real-valued lengths
+result_computed = NaN(size(funt_fine));
+ind_odd = (mod(funt_fine,2)==1) & (L_fine > (funt_fine-3)/sqrt(2)) & (L_fine <= sqrt((funt_fine-5).^2+(funt_fine-1).^2)/2);
+result_computed(ind_odd) = g(L_fine(ind_odd)/a, funt_fine(ind_odd)-3, funt_fine(ind_odd)-3);
+ind_odd2 = (mod(funt_fine,2)==1) & (L_fine > sqrt((funt_fine-5).^2+(funt_fine-1).^2)/2) & (L_fine <= sqrt((funt_fine-3).^2+(funt_fine-1).^2)/2);
[…]

### 2021-08-22 figures/figures.m
Limits of probmax: Maple

@@ -1574,7 +1574,7 @@ result_computed(ind_odd) = ( (alpha_odd-pi/4).*(funt_fine(ind_odd)-3).^2 + (funt
 result_computed(ind_odd2) = result_computed(ind_odd2) + ...
     ( (alphap_odd2-betap_odd2).*(funt_fine(ind_odd2)-1).*(funt_fine(ind_odd2)-5) + (funt_fine(ind_odd2)-1).^2/2 + (funt_fine(ind_odd2)-5).^2/2 + ...
     2*L_fine(ind_odd2).^2 - (funt_fine(ind_odd2)-1).*sqrt(4*L_fine(ind_odd2).^2 - (funt_fine(ind_odd2)-5).^2) - (funt_fine(ind_odd2)-5).*sqrt(4*L_fine(ind_odd2).^2 - (funt_fine(ind_odd2)-1).^2) ) / pi;
-plot(L_fine, result_computed.*L_fine), grid on
+plot(L_fine, result_computed.*funt_fine), grid on
 
 
 %% #61. Probabilities at the end of the intervals with even maximum number of segments
@@ -1628,15 +1628,15 @@ hold on
 set(gca, 'layer', 'top')
 axis equal
 grid on
-plot([-.4 5.55], [0 0], 'k')
-plot([-.4j 5.55j], 'k')
+plot([-.4 5], [0 0], 'k')
+plot([-.4j 5j], 'k')
 set(gca, 'GridAlpha', .4)
 xticks(-1:5), yticks(-1:5)
 xticklabels({'' '0' '1' '' ['(t' char(hex2dec('2212')) '1)/2'] '' ''})
 yticklabels({'' '0' '1' '' ['(t' char(hex2dec('2212')) '1)/2'] '' ''})
 xlabel x 
 ylabel y
-axis([-1 6 -1 6])
+axis([-1 5 -1 5])
 L = 3.37;
 theta_0 = asin(2/L);
 theta = 3*pi/2 - linspace(30*pi/180,theta_0,50);
@@ -1676,15 +1676,15 @@ hold on
 set(gca, 'layer', 'top')
 axis equal
 grid on
-plot([-.4 5.55], [0 0], 'k')
-plot([-.4j 5.55j], 'k')
+plot([-.4 5], [0 0], 'k')
+plot([-.4j 5j], 'k')
[…]

### 2021-08-23 paper.tex
Errata
"Reword" expressions
Checked formula for probmax with g function
Got rid of commands "\touch" etc

@@
 \newcommand{\funta}{\varphi} % function that gives average number of touched tiles for real-valued length
+\newcommand{\probmax}{\zeta} % function that gives the probability of achieving the maximum number of touched files in the square case for real-valued length
 \newcommand{\len}{\ell} % length variable, real-valued
@@
 \newcommand{\jsoli}{j^\ast}
-\newcommand{\genfun}{u}
+\newcommand{\genfun}{q}
 \newcommand{\genvar}{r}
 \newcommand{\mss}{M}
-\newcommand{\touch}{visit}
-\newcommand{\touches}{visits}
-\newcommand{\touched}{visited}
-\newcommand{\touching}{visiting}
 
@@
 \title{
-On the number of tiles \touched{} by a\\
+On the number of tiles visited by a\\
 line segment on a rectangular grid
@@
 A line segment of length $\len \in \mathbb R^+$ is located on the plane with arbitrary position and orientation.
-The segment is said to \emph{\touch} a tile if it intersects its interior.\footnote{
-The definition uses the interior of the tile, excluding the border, to avoid uninteresting results such as a ``zero-length'' segment \touching{} (a vertex of) $4$ tiles.} This paper deals with the following problems:
+The segment is said to \emph{visit} a tile if it intersects its interior.\footnote{
+The definition uses the interior of the tile, excluding the border, to avoid uninteresting results such as a ``zero-length'' segment visiting (a vertex of) $4$ tiles.} This paper deals with the following problems:
 \begin{itemize}
-\item What is the maximum number of tiles that the segment can \touch?
-\item What length should a segment have to \touch{} a given number of tiles?
-\item What is the average number of tiles \touched{} by a uniformly random segment of a given length?
+\item What is the maximum number of tiles that the segment can visit?
+\item What length should a segment have to visit a given number of tiles?
+\item What is the average number of tiles visited by a uniformly random segment of a given length?
 \end{itemize}
@@
 As an example of the first problem, consider $a=1.35$, $b=1$.
-A segment of unit length can be placed as shown in Figure~\ref{fig: examples} (left) to make it \touch{} $3$ tiles.
+A segment of unit length can be placed as shown in Figure~\ref{fig: examples} (left) to make it visit $3$ tiles.
 In fact, the solution for $\len=1$ is $3$ tiles.
@@
 An equivalent formulation of the problem is obtained allowing segments of length $\len$ \emph{or smaller}.
-The equivalence is clear from the fact that reducing the length cannot increase the number of \touched{} tiles.
+The equivalence is clear from the fact that reducing the length cannot increase the number of visited tiles.
 Either of these formulations will be referred to as the \emph{direct} problem.
 
-The \emph{inverse} problem is, given $\tiles \in \mathbb N$, to determine the infimum length of all segments that \touch{} at least $\tiles$ tiles.
-For real-valued lengths this infimum is not a minimum, because given a segment that \touches{} $\tiles$ tiles, its length can always be reduced by some non-zero amount without changing the number of \touched{} tiles.
+The \emph{inverse} problem is, given $\tiles \in \mathbb N$, to determine the infimum length of all segments that visit at least $\tiles$ tiles.
+For real-valued lengths this infimum is not a minimum, because given a segment that visits $\tiles$ tiles, its length can always be reduced by some non-zero amount without changing the number of visited tiles.
 This is a consequence of the interior of each tile being an open set.
@@
 The direct and inverse problems are closely related.
-Namely, if $\len$ is the infimum of all lengths that allow \touching{} at least $\tiles$ tiles (inverse problem), $\tiles$ is the maximum number of tiles that can be \touched{} with lengths slightly greater than $\len$ (direct problem).
-
-The problem of how many tiles a \emph{random} segment of a given length \touches{} on average is precisely defined as follows.
+Namely, if $\len$ is the infimum of all lengths that allow visiting at least $\tiles$ tiles (inverse problem), $\tiles$ is the maximum number of tiles that can be visited with lengths slightly greater than $\len$ (direct problem).
+
+The problem of how many tiles a \emph{random} segment of a given length visits on average is precisely defined as follows.
 By symmetry, one endpoint of the segment can be assumed to lie in a fixed, reference tile.
@@
 The segment orientation has a \emph{uniform} distribution on the set of all possible directions, and is \emph{independent} of the endpoint position.
-As will be seen, solving this problem also answers the inverse question (segment length to \touch{} a given number of tiles on average), which thus need not be considered separately.
+As will be seen, solving this problem also answers the inverse question (segment length to visit a given number of tiles on average), which thus need not be considered separately.
 
@@
 The direct and inverse problems for a deterministic segment are considered in \S\ref{part: max}, first for real-valued lengths, and then for integer-valued lengths on a unit square grid.
-The problem of computing the average number of tiles \touched{} by a random segment is analyzed in \S\ref{part: ave}.
+The problem of computing the average number of tiles visited by a random segment is analyzed in \S\ref{part: ave}.
 
@@
 Two examples are illustrated in Figure~\ref{fig: canonical rectangle and touched tiles}, both with $i=5$, $j=4$.
-Clearly, all tiles \touched{} by the segment are contained in the canonical rectangle.
+Clearly, all tiles visited by the segment are contained in the canonical rectangle.
 Note also that the canonical rectangle can have $i=0$ or $j=0$ if the segment coincides with part of a grid line.
@@
 }%
-\caption{Canonical rectangle and \touched{} tiles
+\caption{Canonical rectangle and visited tiles
 }%
@@
 Consider an arbitrary segment, and let $i, j$ respectively denote the normalized width and height of its canonical rectangle.
-If $i, j \geq 1$, the number of tiles \touched{} by the segment is at most $i+j-1$.
+If $i, j \geq 1$, the number of tiles visited by the segment is at most $i+j-1$.
 This bound is attained if and only if the segment does not pass through any grid point in the interior of the rectangle.
@@
 \begin{proof}
-The segment \touches{}, by definition, two tiles in opposite corners of the canonical rectangle.
+The segment visits, by definition, two tiles in opposite corners of the canonical rectangle.
 It can be assumed, without loss of generality, that those tiles are in the lower-left and upper-right corners of the rectangle, as in Figure~\ref{fig: canonical rectangle and touched tiles}.
-The \touched{} tiles can be thought of as following a path within the canonical rectangle.
+The visited tiles can be thought of as following a path within the canonical rectangle.
 Starting at the lower-left tile, the next tile can be the one to the left, the one above, or the one above and to the left.
@@
 
-Since the segment follows a straight line, once it ``leaves'' a row of tiles in its path from the lower-left to the upper-right corner, it can never \touch{} any more tiles from that row.
+Since the segment follows a straight line, once it ``leaves'' a row of tiles in its path from the lower-left to the upper-right corner, it can never visit any more tiles from that row.
 The same observation applies for the columns.
 
-This implies that the maximum number of \touched{} tiles is $i+j-1$, which is attained if and only if the segment avoids all grid points in the interior of the canonical rectangle, as in Figure~\ref{fig: S_nogridpoints}.
+This implies that the maximum number of visited tiles is $i+j-1$, which is attained if and only if the segment avoids all grid points in the interior of the canonical rectangle, as in Figure~\ref{fig: S_nogridpoints}.
 Note that grid points at the corners of the rectangle do not count for this; and that the segment cannot pass through any other grid points on the rectangle border, because that would imply $i=0$ or $j=0$.
@@
 
-According to Proposition~\ref{prop: i+j-1}, in order to maximize the number of \touched{} tiles for a given length, the position and orientation of the segment should be chosen to obtain $i+j-1$ as large as possible, where $i$ and $j$ are the normalized dimensions of its canonical rectangle.
+According to Proposition~\ref{prop: i+j-1}, in order to maximize the number of visited tiles for a given length, the position and orientation of the segment should be chosen to obtain $i+j-1$ as large as possible, where $i$ and $j$ are the normalized dimensions of its canonical rectangle.
 On the other hand, Proposition~\ref{prop: len ineq i j} restricts the $i, j$ values that can be achieved with a given length.
@@
 
-For instance, it is intuitively clear from Figure~\ref{fig: examples} that segment orientations near the vertical or horizontal directions (resulting in $i=1$ with large $j$, or $j=1$ with large $i$) will not maximize the number of \touched{} tiles.
+For instance, it is intuitively clear from Figure~\ref{fig: examples} that segment orientations near the vertical or horizontal directions (resulting in $i=1$ with large $j$, or $j=1$ with large $i$) will not maximize the number of visited tiles.
 % This is true for any a, b: 2 can achieved instead of 1 without compromising the other dimension.
@@
 In order to derive a general method to build a minimal sufficient set, it is insightful to consider two specific cases first, as will be presented in the following.
-It is also helpful to use the equivalent formulation of the direct problem referred to in \S\ref{part: intro}, i.e.\ finding the maximum number of tiles \touched{} by segments of length $\len$ \emph{or less}.
-
-Thus each choice of $(i, j)$ implies a maximum of $i+j-1$ \touched{} tiles on one hand; and sets a lower bound on the required length $\len$ on the other hand.
+It is also helpful to use the equivalent formulation of the direct problem referred to in \S\ref{part: intro}, i.e.\ finding the maximum number of tiles visited by segments of length $\len$ \emph{or less}.
+
+Thus each choice of $(i, j)$ implies a maximum of $i+j-1$ visited tiles on one hand; and sets a lower bound on the required length $\len$ on the other hand.
 
@@
 \includegraphics[width=.7\textwidth]{ijLS_1}%
-\caption{Relationship between segment length and number of \touched{} tiles with the width and height of the canonical rectangle, for $a=b$
+\caption{Relationship between segment length and number of visited tiles with the width and height of the canonical rectangle, for $a=b$
 }%
@@
 For a given $\len \in \mathbb R^+$, the $(i,j)$ pairs that can be achieved with segments of length not exceeding $\len$ are, by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, those whose distance from $(2a,2b)$ is less than $\len$, i.e.~those delimited by the corresponding arc.
-It is clear from the figure that for $i+j-1$ odd, if some point $(i_1,j_1)$ with $|i_1-j_1|>1$ is achievable with the length bound $\len$, then necessarily there is a point $(i_0,i_0)$ with the same number of \touched{} tiles $i+j-1$ (same diagonal line) that is also achievable with the same length restriction (because it has smaller distance to $(2a,2b)$).
+It is clear from the figure that for $i+j-1$ odd, if some point $(i_1,j_1)$ with $|i_1-j_1|>1$ is achievable with the length bound $\len$, then necessarily there is a point $(i_0,i_0)$ with the same number of visited tiles $i+j-1$ (same diagonal line) that is also achievable with the same length restriction (because it has smaller distance to $(2a,2b)$).
 Similarly, for $i+j-1$ even, if a point $(i_1,j_1)$ with $|i_1-j_1|>1$ is achievable, then there is a point $(i_0,i_0+1)$ in the same diagonal that is also achievable with the same length restriction.
@@
 The pairs in the minimal sufficient set are marked with filled circles in Figure~\ref{fig: ijLS_1}.
-Given a length, the maximum number of \touched{} tiles will be achieved with one of those pairs, namely that in the diagonal line furthest from $(2a,2b)$ that is within the circle with radius $\len$ centered at that point.
+Given a length, the maximum number of visited tiles will be achieved with one of those pairs, namely that in the diagonal line furthest from $(2a,2b)$ that is within the circle with radius $\len$ centered at that point.
 
@@
 \includegraphics[width=.7\textwidth]{ijLS_1p35}%
-\caption{Relationship between segment length and number of \touched{} tiles with the width and height of the canonical rectangle, for $a=1.35b$
+\caption{Relationship between segment length and number of visited tiles with the width and height of the canonical rectangle, for $a=1.35b$
 }%
@@
 For each $\tiles$, the pair $(i_\tiles,j_\tiles)$ of the minimal sufficient set should be chosen as that on the line $i+j-1=\tiles$ which minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$.
-This allows maximizing the sum $i+j-1$, and thus the number of \touched{} tiles, for a given length restriction (direct problem); or minimizing the required lengths for a specified number of \touched{} files (inverse problem).
+This allows maximizing the sum $i+j-1$, and thus the number of visited tiles, for a given length restriction (direct problem); or minimizing the required lengths for a specified number of visited files (inverse problem).
 
@@
 Therefore the optimal slope for long segments is approximately $a/b$.
-This formalizes the intuition that to maximize the number of \touched{} tiles, the segment should follow a direction along which the perceived ``length'' of the tile is smaller.
+This formalizes the intuition that to maximize the number of visited tiles, the segment should follow a direction along which the perceived ``length'' of the tile is smaller.
 
@@
 
-For a grid with parameters $a, b \in \mathbb R^+$, the maximum number $\tiles$ of \touched{} tiles with a given real-valued length $\len$ can be represented by a function $\funt: \mathbb R^+ \to \mathbb N$ such that $\tiles = \funt(\len)$.
-Similarly, for the inverse problem a function $\funl: \mathbb N \to \mathbb R^+$ can be defined such that $\funl(\tiles)$ gives the infimum length of all segments that \touch{} at least $\tiles$ tiles.
-
-The function $\funt$ can be obtained from $\funl$ by noting that the maximum number of tiles that can be \touched{} by a segment of length $\len$ is the index of the largest term of the sequence $\funl(\tiles)$ that is less than $\len$:
+For a grid with parameters $a, b \in \mathbb R^+$, the maximum number $\tiles$ of visited tiles with a given real-valued length $\len$ can be represented by a function $\funt: \mathbb R^+ \to \mathbb N$ such that $\tiles = \funt(\len)$.
+Similarly, for the inverse problem a function $\funl: \mathbb N \to \mathbb R^+$ can be defined such that $\funl(\tiles)$ gives the infimum length of all segments that visit at least $\tiles$ tiles.
+
+The function $\funt$ can be obtained from $\funl$ by noting that the maximum number of tiles that can be visited by a segment of length $\len$ is the index of the largest term of the sequence $\funl(\tiles)$ that is less than $\len$:
 \begin{equation}
@@
 Then there is $\epsilon > 0$ for which $\funt(\len+\epsilon) = \tiles = \funt(\len-\epsilon)$.
-This means that there exists a segment with length $\len-\epsilon$ that \touches{} $\tiles$ tiles, and thus $\funl(\tiles) \leq \len-\epsilon < \len$, in contradiction with \eqref{eq: theo funt, form: prop 1}.
+This means that there exists a segment with length $\len-\epsilon$ that visits $\tiles$ tiles, and thus $\funl(\tiles) \leq \len-\epsilon < \len$, in contradiction with \eqref{eq: theo funt, form: prop 1}.
 Therefore $\funt$ is discontinuous, from the right, at $\len$.
@@
 \begin{proof}
-The $(i_\tiles,j_\tiles)$ pair in the minimal sufficient set $\mss$ corresponds to a maximum of $\tiles$ \touched{} tiles.
-By construction of this set, any segment that \touches{} $\tiles$ tiles must have length greater than $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$.
+The $(i_\tiles,j_\tiles)$ pair in the minimal sufficient set $\mss$ corresponds to a maximum of $\tiles$ visited tiles.
+By construction of this set, any segment that visits $\tiles$ tiles must have length greater than $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$.
 
@@
 
-Theorems~\ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the two problems stated in \S\ref{part: intro}; they also provide a way to actually position a segment of length slightly greater than $\len$ or $\funl(\tiles)$ so that it \touches{} $\tiles$ or $\funt(\len)$ tiles.
+Theorems~\ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the two problems stated in \S\ref{part: intro}; they also provide a way to actually position a segment of length slightly greater than $\len$ or $\funl(\tiles)$ so that it visits $\tiles$ or $\funt(\len)$ tiles.
 Namely, the segment should have its endpoints respectively in the interior of two tiles separated $\isoli-1$ steps horizontally and $\jsoli-1$ steps vertically, with its exact position and orientation  adjusted to avoid any grid points.
@@
 % *!*
-it is seen that for long segments the number of \touched{} tiles and the segment length are approximately proportional, with
+it is seen that for long segments the number of visited tiles and the segment length are approximately proportional, with
 \begin{equation}
@@
 \label{eq: square, odd max tiles, lengths}
-\frac \len a \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{ \sqrt{(\tiles^2-2)^2+1}}{\sqrt{2}} \right].
+\frac \len a \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{ \sqrt{(\tiles-1)^2+(\tiles-3)^2}}{2} \right].
 \end{equation}
@@
 \label{eq: square, even max tiles, lengths}
-\frac \len a \in \left(\frac{ \sqrt{(\tiles^2-3)^2+1}}{\sqrt{2}}, \frac{\tiles-2}{\sqrt{2}} \right].
+\frac \len a \in \left(\frac{ \sqrt{(\tiles-2)^2+(\tiles-4)^2}}{2}, \frac{\tiles-2}{\sqrt{2}} \right].
 \end{equation}
@@
 
-The \emph{inverse problem} with integer-length segments can be formulated as follows: given $\tiles \in \mathbb N$, find the \emph{minimum} integer length that allows \touching{} at least $\tiles$ tiles.
+The \emph{inverse problem} with integer-length segments can be formulated as follows: given $\tiles \in \mathbb N$, find the \emph{minimum} integer length that allows visiting at least $\tiles$ tiles.
 Observe that in this case, unlike with real-valued lengths, there is indeed a minimum length, because every subset of $\mathbb N$ has a minimum.
@@
 In view of \eqref{eq: funli funti} and \eqref{eq: funti funli}, $\funti$ and $\funli$ can be considered as ``pseudo-inverse'' sequences of each other.
-% ***Is this a good name? Is there an established name for this relationship?
 
@@
 
-\section{Average number of tiles \touched{} by a random segment}
+\section{Average number of tiles visited by a random segment}
 \label{part: ave}
@@
 The variables $x_1$, $y_1$ and $\theta$ determine the coordinates $x_2$, $y_2$ of the second endpoint.
-Note that events of probability $0$, such as $x_1=a$ or the segment passing through a grid point, can be disregarded for computing the average number of tiles \touched{} by the segment.
-
-Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles \touched{} by a random segment of length $\len$, with the above distributions.
+Note that events of probability $0$, such as $x_1=a$ or the segment passing through a grid point, can be disregarded for computing the average number of tiles visited by the segment.
+
+Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles visited by a random segment of length $\len$, with the above distributions.
 
@@
 The normalized dimensions $i$ and $j$ of the canonical rectangle are thus random variables.
-Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and the segment \touches{} $t = i+j-1$ tiles.
+Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and the segment visits $t = i+j-1$ tiles.
 
@@
 As a consequence, while the marginal distributions of $i$ and $j$ are easy to compute, the distribution of $i+j$ is difficult.
-However, knowing the marginal distributions is sufficient to compute the average number of \touched{} tiles:
+However, knowing the marginal distributions is sufficient to compute the average number of visited tiles:
 \begin{equation}
@@
 Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above.
-The average number of tiles \touched{} by the segment, $\funta(\len)$, is
+The average number of tiles visited by the segment, $\funta(\len)$, is
 \begin{equation}
@@
 
-In view of Theorem~\ref{theo: funta, form}, the average number of \touched{} tiles has a remarkably simple dependence on the segment length, namely an affine function.
-Conversely, for any $t>1$ it is immediate to compute the length of a random segment that \touches{} $t$ tiles on average, given as $\funta^{-1}(t)$.
+In view of Theorem~\ref{theo: funta, form}, the average number of visited tiles has a remarkably simple dependence on the segment length, namely an affine function.
+Conversely, for any $t>1$ it is immediate to compute the length of a random segment that visits $t$ tiles on average, given as $\funta^{-1}(t)$.
 
@@
 It tends to $2/\pi$ for $a/b \rightarrow \infty$ and for $a/b \rightarrow 0$.
-The minimum number of tiles that can be \touched{} with non-zero probability (corresponding to an n times 1 canonical rectangle if a>=b) can also be plotted.
+The minimum number of tiles that can be visited with non-zero probability (corresponding to an n times 1 canonical rectangle if a>=b) can also be plotted.
 

### 2021-08-23 paper.tex
Minor changes in tex file

@@
 \newcommand{\funta}{\varphi} % function that gives average number of touched tiles for real-valued length
-\newcommand{\probmax}{\zeta} % function that gives the probability of achieving the maximum number of touched files in the square case for real-valued length
+\newcommand{\probmax}{\rho} % function that gives the probability of achieving the maximum number of touched files in the square case for real-valued length
 \newcommand{\len}{\ell} % length variable, real-valued
@@
 \includegraphics[width=.7\textwidth]{ijLS_1p35_detail_inverse}%
-\caption{Obtaining $(\isoli,\jsoli)$ in Proposition~\ref{prop: min suff set, form}. Example with $a=1.35, b=1$.%
+\caption{Obtaining $(\isoli,\jsoli)$ in Proposition~\ref{prop: min suff set, form}. Example with $a=1.35, b=1$%
 }%
@@
 }%
-\caption{Obtaining $(\isoli,\jsoli)$ in Theorem~\ref{theo: funt, form}. Examples with $a=1.35, b=1$.
+\caption{Obtaining $(\isoli,\jsoli)$ in Theorem~\ref{theo: funt, form}. Examples with $a=1.35, b=1$%
 }%
@@
 \end{equation}
-For $\tiles \geq 3$, $t$ even, $\funt(\len) = \tiles$ if and only if
+For $\tiles \geq 4$, $t$ even, $\funt(\len) = \tiles$ if and only if
 \begin{equation}

### 2021-08-24 paper.tex
Change introduction and section structure to make room for the new results on the probability of visiting the maximum number of tiles

@@
 \item What length should a segment have to visit a given number of tiles?
-\item What is the average number of tiles visited by a uniformly random segment of a given length?
+\item What is the average number of tiles visited by a uniformly random segment of a given length? How often does the random segment visit the maximum number of tiles?
 \end{itemize}
@@
 
-The problem of how many tiles a \emph{random} segment of a given length visits on average is precisely defined as follows.
+The problem of how many tiles a \emph{random} segment of a given length visits is precisely defined as follows.
 By symmetry, one endpoint of the segment can be assumed to lie in a fixed, reference tile.
@@
 The segment orientation has a \emph{uniform} distribution on the set of all possible directions, and is \emph{independent} of the endpoint position.
-As will be seen, solving this problem also answers the inverse question (segment length to visit a given number of tiles on average), which thus need not be considered separately.
+As will be seen, solving the problem of how many tiles the segment visits on average also answers the inverse question (segment length to visit a given number of tiles on average).
+In addition, it is interesting to know the probability that the segment visits the maximum number of tiles.
 
@@
 The direct and inverse problems for a deterministic segment are considered in \S\ref{part: max}, first for real-valued lengths, and then for integer-valued lengths on a unit square grid.
-The problem of computing the average number of tiles visited by a random segment is analyzed in \S\ref{part: ave}.
+The analysis for the random segment is carried out in \S\ref{part: rand}.
+The average number of tiles is computed in general, and the probability that the segment visits the maximum number of tiles is obtained for a square grid.
 
@@
 
-\section{Average number of tiles visited by a random segment}
-\label{part: ave}
+\section{Probabilistic characterization for a random segment}
+\label{part: rand}
 
@@
 
-Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles visited by a random segment of length $\len$, with the above distributions.
+\section{Arbitrary grid with real-valued lengths}
+
+Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles visited by a random segment of length $\len$, with the distributions specified in \S\ref{part: rand}.
 
@@
 
+% ***This will go after Theorem~\ref{theo: funta, form}
 \begin{proposition}
@@
 
+***Clarify here that computing the probability of visiting the maximum number of tiles in the general case is a difficult problem, and it will be dealt with in the square-grid case only
+
+\section{Square grid with real-valued lengths}
+
+***
+
 \bibliographystyle{plain}

### 2021-08-24 paper.tex
Results on probability of maximum

@@
 
-\section{Square grid with real-valued lengths}
-
-***
+\subsection{Square grid with real-valued lengths}
+
+Consider a square grid with spacing $a \in \mathbb R^+$.
+For a random segment with length $\len \in \mathbb R^+$, let the function $\probmax: \mathbb R^+ \to [0,1]$ be defined such that $\probmax(\len)$ gives the probability that the segment visits the maximum possible number of tiles, $\funt(\len)$.
+This function is characterized by the next theorem.
+
+For $u, v \in \mathbb N$, $\genvar \in \mathbb R^+$, let
+\begin{equation}
+\label{eq: g def}
+g(\genvar, u, v) = \int_{\arcsin \frac u {2r}}^{\arccos \frac u {2r}} \left( r \sin \theta - \frac u 2 \right) \left( r \cos \theta - \frac v 2 \right) \diff \theta.
+\end{equation}
+Using the identifies $\cos^2 \alpha = 1-\sin^2 \alpha$ and $\cos(2\alpha) = 2\cos^2\alpha-1 = 1 - 2\sin^2\alpha$, the integral in \eqref{eq: g def} is computed as
+\begin{equation}
+\label{eq: g fin}
+\begin{split}
+g(\genvar, u, v) &= \frac 1 {2\pi} \biggl(
+\left(\arccos\frac{u}{2\genvar}-\arcsin\frac{v}{2\genvar}\right) u v + 2\genvar^2 + \frac{u^2} 2 + \frac{v^2} 2 \\
+& \quad  - u \sqrt{4\genvar^2-v^2} - v \sqrt{4\genvar^2-u^2} \biggr).
+\end{split}
+\end{equation}
+
+\begin{theorem}
+\label{theo: probmax, sq}
+Given $a, \len \in \mathbb R^+$, consider a square grid with spacing $a$ and a uniformly random segment of length $\len$, as previously defined.
+The probability $\probmax(\len)$ that the segment visits the maximum number of tiles $\funt(\len)$ is
+\begin{equation}
+\label{eq: probmax}
+\begin{split}
+\probmax(\len) & = g\left(\frac{\len} a, 
+\tiles-3, \tiles-3 \right) \quad\ \text{for } t \text{ odd, } \textstyle \frac \len a \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2 \right]; \\
+\probmax(\len) & = g\left(\frac{\len} a, \tiles-3, \tiles-3 \right) + 2 g\left(\frac{\len} a, \tiles-5, \tiles-1 \right) \\
+& \text{for } t \text{ odd, } \textstyle \frac \len a \in \left (\frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2,  \frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2 \right]; \\
+\probmax(\len) & = 2 g\left(\frac{\len} a, \tiles-4, \tiles-2 \right) \quad \text{for } t \text{ even, } \textstyle \frac \len a \in \left( \frac{\sqrt{(\tiles-4)^2+(\tiles-2)^2}} 2, \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2 \right]; \\
+\probmax(\len) & = 2 g\left(\frac{\len} a, \tiles-4, \tiles-2 \right) + 2 g\left(\frac{\len} a, \tiles-6, \tiles \right) \\
+& \text{for } t \text{ even, } \textstyle \frac \len a \in \left( \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2, \frac{\tiles-2}{\sqrt{2}} \right].
+\end{split}
+\end{equation}
+\end{theorem}
+
+\begin{proof}
+The approach is similar to that used in the proof of Proposition~\ref{prop: Pr i}, but conditioning on the segment orientation $\theta$ instead of on the location of its first endpoint.
+
+Consider a length $\len$ with corresponding maximum number of tiles $\tiles = \funt(\len)$ (given by Corollary~\ref{cor: funt, sq, form}).
+The first endpoint of the segment is assumed to be in the reference tile with lower-left corner $(0,0)$.
+The segment visits $\tiles$ tiles almost surely if and only if its second endpoint is in a tile with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, $i+j-1=\tiles$, or in a tile symmetrical to this in the other three quadrants of the plane.
+Figure~\ref{fig: probmax_possible_tiles} shows the reference tile, with solid line, and some of the square tiles that fulfil this condition, with dashed line, for $\tiles=5$.
+The desired probability can be computed by considering only the tiles in the first quadrant and multiplying by $4$.
+
+% ***Does this figure add something, or is it better to remove it?
+\begin{figure}
+\centering%
+\includegraphics[width=.52\textwidth]{probmax_possible_tiles}%
+\caption{Tiles in which the second endpoint can be for a given number of visited tiles $\tiles$ in a square grid (example for $\tiles=5$)
+}%
+\label{fig: probmax_possible_tiles}%
+\end{figure}%
+
+For any $i,j$ with $i+j-1=\tiles$ and for a given orientation $\theta$, the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ if and only if shifting the segment so that its second endpoint has coordinates $(i-1,j-1)$ results in the first endpoint being in the reference tile.
+It cannot be the case that the segment ``overshoots'' past this tile, because that would result in a number of visited tiles exceeding the assumed maximum $\tiles$.
+
+\begin{figure}
+\centering%
+\subfigure[The second endpoint is in the tile with lower-left corner $((t-1)/2,(t-1)/2)$]{%
+\label{fig: probmax_funt_odd_diag}%
+\includegraphics[width=.56\textwidth]{probmax_funt_odd_diag}%
+}\\%\hfill%
+\subfigure[The second endpoint is in the tile with lower-left corner $((t-3)/2,(t+1)/2)$]{%
+\label{fig: probmax_funt_odd_offdiag}%
+\includegraphics[width=.56\textwidth]{probmax_funt_odd_offdiag}%
+}%
+\caption{Segment orientations and positions of the first endpoint within the reference tile for a given number $\tiles$ of visited tiles. Example for odd $\tiles$
+}%
+\label{fig: probmax_funt_odd}%
+\end{figure}%
+
+There is a range of possible values of $\theta$ for which the above condition is satisfied.
+This range corresponds to the part of the arc with solid line in Figure~\ref{fig: probmax_funt_odd}.
+For convenience, the angle $\theta$ is defined clockwise from the downward direction, as seen in the figure.
+For each $\theta$ in the allowed range, the valid positions for the first endpoint of the (shifted) segment are in the shaded rectangle, whose dimensions depend on $\theta$.
+The fraction of the reference tile occupied by this rectangle is the probability $\probmax_{i-1,j-1}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$.
+The desired probability, $\probmax(\len)$, is obtained by averaging $\probmax_{i-1,j-1}(\len, \theta)$ over $\theta$ and summing over all $i-1,j-1$ that can be reached from the reference tile.
+
+The situation is different depending on whether $\tiles$ is odd or even.
+Consider $\tiles$ odd first.
+Under this assumption, the range of lengths corresponding to $\tiles$ is given by \eqref{eq: square, odd max tiles, lengths}.
+For any length in this interval, there are some positions within the reference file from which the tile with lower-left coordinate $((t-1)/2, (t-1)/2)$ can be reached.
+Equivalently, the shifted segment with one endpoint at $((t-1)/2, (t-1)/2)$ can reach the reference tile.
+This is illustrated in Figure~\ref{fig: probmax_funt_odd_diag}.
+In addition, if
+\begin{equation}
+\label{eq: probmax len odd2}
+\len > \frac{\sqrt{(\tiles-5)^2 + (\tiles-1)^2}}{2}
+\end{equation}
+the two tiles with lower-left coordinates $((t-3)/2, (t+1)/2)$ and $((t+1)/2, (t-3)/2)$ are also reachable from the reference tile; or equivalently the shifted segment can reach the reference tile, as shown in Figure~\ref{fig: probmax_funt_odd_offdiag}.
+This can only happen for $\tiles \geq 5$ (for $\tiles=3$ the inequality \eqref{eq: probmax len odd2} cannot be satisfied for lengths in the range given by \eqref{eq: square, odd max tiles, lengths}).
+In this case, by symmetry, each of those two tiles contributes the same probability.
+Other tiles corresponding to the same $\tiles$ need not be considered, because they cannot be reached with lengths satisfying \eqref{eq: square, odd max tiles, lengths}.
+
+For $\tiles$ even, which corresponds to lengths in the interval \eqref{eq: square, even max tiles, lengths}, the two tiles with lower-left coordinates $(t/2-1, t/2)$ and $(t/2, t/2-1)$ can always be reached from some positions within the reference tile.
+In addition, if
+\begin{equation}
+\label{eq: probmax len even2}
+\len > \frac{\sqrt{((\tiles-6)^2 + \tiles^2}}{2}
+\end{equation}
+the two tiles with lower-left coordinates $(\tiles/2-2, \tiles/2+1)$ and $(\tiles/2+1, \tiles/2-2)$ have to be considered too.
+This can only happen for $\tiles \geq 8$.
+No other tiles can be reached from the reference tile.
+
+Therefore, two cases need to be distinguished for $\tiles$ odd, depending on whether \eqref{eq: probmax len odd2} is satisfied or not; and similarly two cases exist for $\tiles$ even, depending on whether \eqref{eq: probmax len even2} holds or not.
+These are, respectively, the four cases in \eqref{eq: probmax}.
+
+In the first case, as has been argued, only the tile with lower-left coordinates $((t-1)/2, (t-1)/2)$ is reachable.
+For a given $\theta$, the conditional probability $\probmax_{(t-1)/2, (t-1)/2}(\len, \theta)$ is the area of the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_diag} divided by $a^2$.
+The horizontal and vertical sides of this rectangle are $\len\sin\theta - a(\tiles-3)/2$ and $\len\cos\theta - a(\tiles-3)/2$ respectively, and thus
+\begin{equation}
+\label{eq: probmax: probmax theta odd}
+\probmax_{(t-1)/2, (t-1)/2}(\len, \theta) = \frac{(\len\sin\theta - a(\tiles-3)/2) (\len\cos\theta - a(\tiles-3)/2)} {a^2}
+\end{equation}
+The range of allowed values for $\theta$, as can be seen from Figure~\ref{fig: probmax_funt_odd_diag}, is $(\theta_0, \theta_1)$ with
+\begin{align}
+\label{eq: odd theta 0}
+\theta_0 & = \arcsin\frac{a(\tiles-3)}{2\len}, \\
+\label{eq: odd theta 1}
+\theta_1 & = \arccos\frac{a(\tiles-3)}{2\len}.
+\end{align}
+Computing the average of \eqref{eq: probmax: probmax theta odd} over $\theta$, multiplying by $4$ to include all quadrants and comparing with \eqref{eq: g def} gives $\probmax(\len)$ as
+\begin{equation}
+\label{eq: probmax: probmax odd}
+\begin{split}
+\probmax(\len) & = 4 \int_{\theta_0}^{\theta_1} \frac{\probmax_{(t-1)/2, (t-1)/2}(\len, \theta)}{2\pi} \diff \theta \\
+& = \frac 2 \pi \int_{\arcsin\frac{a(\tiles-3)}{2\len}}^{\arccos\frac{a(\tiles-3)}{2\len}} \left( \frac{\len\sin\theta}a - \frac{\tiles-3} 2 \right)  \left( \frac{\len\cos\theta}a - \frac{\tiles-3} 2 \right) \diff \theta \\
+& = g\left( \frac \len a, t-3, t-3 \right ),
+\end{split}
+\end{equation}
+in accordance with \eqref{eq: probmax}.
+
+For the second case in \eqref{eq: probmax}, $\probmax(\len)$ is obtained as in the preceding paragraph plus an additional term corresponding to the contribution of the tile with lower-left corner $((t-3)/2, (t+1)/2)$, multiplied by $2$ to account for its symmetrical tile $((t+1)/2, (t-3)/2)$ and by $4$ to include all quadrants.
+The sides of the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_offdiag} are $\len\sin\theta - a(\tiles-5)/2$ and $\len\cos\theta - a(\tiles-1)/2$, and therefore
+\begin{equation}
+\label{eq: probmax: probmax theta odd2}
+\probmax_{(t-3)/2, (t+1)/2}(\len, \theta) = \frac{(\len\sin\theta - a(\tiles-5)/2) (\len\cos\theta - a(\tiles-1)/2)} {a^2}.
+\end{equation}
+The integration interval $(\theta_0, \theta_1)$ is in this case
+\begin{align}
+\theta_0 & = \arcsin\frac{a(\tiles-5)}{2\len}, \\
+\theta_1 & = \arccos\frac{a(\tiles-1)}{2\len}.
+\end{align}
+The additional contribution to $\probmax(\len)$, given by
+\begin{multline}
+8 \int_{\theta_0}^{\theta_1} \frac{\probmax_{(t-3)/2, (t+1)/2}(\len, \theta)}{2\pi} \diff \theta \\
+= \frac 4 \pi \int_{\arcsin\frac{a(\tiles-5)}{2\len}}^{\arccos\frac{a(\tiles-1)}{2\len}} \left( \frac{\len\sin\theta}a - \frac{\tiles-5} 2 \right)  \left( \frac{\len\cos\theta}a - \frac{\tiles-1} 2 \right) \diff \theta,
+\end{multline}
+is recognized to be $2 g(\len/a, \tiles-5, \tiles-1)$. This proves the second equality in \eqref{eq: probmax}.
+
+The expressions for the third and fourth cases in \eqref{eq: probmax}, corresponding to $\tiles$ even, are established analogously.
+\end{proof}
+
+Figure~\ref{fig: probmax_len} shows the probability $\probmax(\len)$ of visiting the maximum number of tiles on a square grid with spacing $a$, as given by \eqref{eq: probmax}.
+As $\len$ grows, $\probmax(\len)$ has a jump when the maximum number of tiles that can be visited increases by $1$.
+This happens when $\len$ equals the right endpoint of the interval \eqref{eq: square, odd max tiles, lengths} for $\tiles$ odd or of \eqref{eq: square, even max tiles, lengths} for $\tiles$ even.
+Let these length values be denoted as $\len_\tiles$:
+\begin{equation}
+\len_\tiles = \begin{cases}
+\displaystyle
+\frac{a \sqrt{(\tiles-3)^2 + (\tiles-1)^2}} {2} & \text{ for $\tiles$ odd, $\tiles \geq 3$} \\
+\displaystyle
+\frac{a (\tiles-2)}{\sqrt{2}} & \text{ for $\tiles$ odd, $\tiles \geq 4$}.
+\end{cases}
+\end{equation}
+For example, the first continuous section seen in the figure corresponds to a maximum number of tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, with $\len_3= a$.
+The second corresponds to $\tiles=4$, for lengths in $(\len_3, \len_4]$, with $\len_4= a\sqrt{2}$.
+Within each continuous section the probability monotonically increases from $0$ to a maximum value.
+The heights of the maxima follow a different trend*** for odd and even $\tiles$, as seen in the figure; but in each case they are asymptotically proportional to $1/\len_\tiles$, as the next result establishes.
+
+\begin{figure}%
+\centering%
+\includegraphics[width=.75\textwidth]{probmax_len}%
+\caption{Probability $\probmax(\len)$ that a random segment of length $\len$ visits the maximum number of tiles%
+}%
+\label{fig: probmax_len}%
+\end{figure}%
+
+\begin{proposition}
+The function $\probmax$ has the following asymptotic behaviour:
+\begin{align}
+\label{eq: probmax liminf}
+\liminf_{\len \rightarrow \infty}\, \probmax(\len) & = 0 \\ 
+\label{eq: probmax lim odd}
+\lim_{\substack{\tiles \rightarrow \infty \\
+\tiles \ \mathrm{odd}}}\, \tiles \probmax(\len_\tiles) & = \frac 2 \pi \\
+\label{eq: probmax lim even}
+\lim_{\substack{\tiles \rightarrow \infty \\ \tiles \ \mathrm{even}}}\, \tiles \probmax(\len_\tiles) & = \frac 8 {3\pi} \\
+\label{eq: probmax limsup}
+\limsup_{\len \rightarrow \infty}\, \len \probmax(\len) & = \frac {4\sqrt{2} a}{3\pi}.
+\end{align}
+\end{proposition}
+
+\begin{proof}
+The probability $\probmax(\len)$ becomes $0$ whenever the length causes the maximum number of visited tiles to increase by $1$. This establishes \eqref{eq: probmax liminf}.
+
+For \eqref{eq: probmax lim odd} it suffices to note that
+\begin{align}
+\label{eq: tedious limit, odd 1}
+\lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2, \tiles-3, \tiles-3 \right) &= \frac 2 {3\pi} \\
+\label{eq: tedious limit, odd 2}
+\lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2, \tiles-5, \tiles-1 \right)
+& = \frac 2 {3\pi},
+\end{align}
+as can be checked using L'H\^opital's rule \cite[section~5.3]{Abbott15}.
+%***I think.
+It's very tedious by hand.
+I used Maple to compute the limit, and then checked it numerically.
+Maybe one or two intermediate steps from applying L'H\^opital's rule should be included.
+Perhaps those can be obtained with Maple too, and this part can be rewritten.
+Combining these expressions with the second equality in \eqref{eq: probmax} yields \eqref{eq: probmax lim odd}.
+
+Similarly,
+\begin{align}
+\label{eq: tedious limit, even 1}
+\lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\tiles-2}{\sqrt 2}, \tiles-4, \tiles-2 \right) &= \frac 2 {3\pi} \\
+\label{eq: tedious limit, even 2}
+\lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\tiles-2}{\sqrt 2}, \tiles-6, \tiles \right)
+& = \frac 2 {3\pi},
+\end{align}
+and combining with the fourth equality in \eqref{eq: probmax} gives \eqref{eq: probmax lim even}.
+
+For large $\tiles$, $\len_\tiles$ asymptotically approaches $a\tiles/\sqrt{2}$.
+Using this into \eqref{eq: probmax lim even}, which is greater than \eqref{eq: probmax lim odd}, establishes \eqref{eq: probmax limsup}.
+\end{proof}
 

### 2021-08-24 paper.tex
Wording in introduction

@@
 \begin{itemize}
-\item What is the maximum number of tiles that the segment can visit?
-\item What length should a segment have to visit a given number of tiles?
+\item What is the maximum number of tiles that the segment can visit? Conversely, what length should a segment have to visit a given number of tiles?
 \item What is the average number of tiles visited by a uniformly random segment of a given length? How often does the random segment visit the maximum number of tiles?
@@
 
-The problem of how many tiles a \emph{random} segment of a given length visits is precisely defined as follows.
+The notion of a \emph{random} segment of a given length is precisely defined as follows.
 By symmetry, one endpoint of the segment can be assumed to lie in a fixed, reference tile.
@@
 The segment orientation has a \emph{uniform} distribution on the set of all possible directions, and is \emph{independent} of the endpoint position.
-As will be seen, solving the problem of how many tiles the segment visits on average also answers the inverse question (segment length to visit a given number of tiles on average).
-In addition, it is interesting to know the probability that the segment visits the maximum number of tiles.
+Solving the problem of how many tiles the segment visits on average, as will be seen, also answers the inverse question (segment length to visit a given number of tiles on average).
+In addition, a natural question is with that probability the segment visits the maximum number of tiles.
 

### 2021-08-24 paper.tex
Inverse sequence in OEIS
Minor corrections ("t" replaced by "\tiles"); and clarifications

@@
 From Theorem~\ref{theo: funt, sq, form} it is easily seen that odd values of $\funt(\len)$ correspond to $\isoli = \jsoli$, whereas even values are achieved with $\jsoli = \isoli+1$.
-Given $\tiles \geq 3$ with $t$ odd, $\funt(\len) = \tiles$ if and only
+Given $\tiles \geq 3$ with $\tiles$ odd, $\funt(\len) = \tiles$ if and only
 \begin{equation}
@@
 \end{equation}
-For $\tiles \geq 4$, $t$ even, $\funt(\len) = \tiles$ if and only if
+For $\tiles \geq 4$, $\tiles$ even, $\funt(\len) = \tiles$ if and only if
 \begin{equation}
@@
 The sequence $\funli(\tiles)$, $\tiles \in \mathbb N$ can be computed using \eqref{eq: theo: funt, sq, form; funl} and \eqref{eq: funli funl}, and has initial values $1, 1, 1, 2, 2, 3, 3, 4, 5 \ldots$, as shown in Figure~\ref{fig: funli}.
-This is A**** in the On-Line Encyclopedia of Integer Sequences \cite{***}.
+This is A346693 in the On-Line Encyclopedia of Integer Sequences \cite{OEIS_unitsq_int_inv}.
 However, a simpler expression can be obtained from \eqref{eq: funti leni} and \eqref{eq: funli funti}.
@@
 
-\section{Arbitrary grid with real-valued lengths}
+Each realization of the random segment gives rise to a canonical rectangle, whose normalized dimensions $i$ and $j$ are thus random variables.
+Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and the segment visits $i+j-1$ tiles.
+Note that $i$ and $j$ are not statistically independent.
+
+\subsection{Arbitrary grid with real-valued lengths}
 
@@
 
-Each realization of the random segment gives rise to a canonical rectangle.
-The normalized dimensions $i$ and $j$ of the canonical rectangle are thus random variables.
-Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and the segment visits $t = i+j-1$ tiles.
-
-The random variables $i$ and $j$ are not independent.
-As a consequence, while the marginal distributions of $i$ and $j$ are easy to compute, the distribution of $i+j$ is difficult.
-However, knowing the marginal distributions is sufficient to compute the average number of visited tiles:
+Although the random variables $i$ and $j$ are dependent, their marginal distributions can be obtained easily, and are sufficient to compute the average number of visited tiles:
 \begin{equation}
 \label{eq: funta E[i] E[j]}
-\funta(\len) = \E[t] = \E[i]+\E[j]-1.
+\funta(\len) = \E[\tiles] = \E[i]+\E[j]-1.
 \end{equation}
@@
 In view of Theorem~\ref{theo: funta, form}, the average number of visited tiles has a remarkably simple dependence on the segment length, namely an affine function.
-Conversely, for any $t>1$ it is immediate to compute the length of a random segment that visits $t$ tiles on average, given as $\funta^{-1}(t)$.
+Conversely, for any $\tiles>1$ it is immediate to compute the length of a random segment that visits $\tiles$ tiles on average, given as $\funta^{-1}(\tiles)$.
 
@@
 
-***Clarify here that computing the probability of visiting the maximum number of tiles in the general case is a difficult problem, and it will be dealt with in the square-grid case only
+The probability of visiting the maximum number of tiles for a rectangular grid is difficult to compute, due to the irregularity of the minimal sufficient set (discussed in \S\ref{part: fund results}).
+In the square case, however, the simplicity of this set makes the problem tractable.
 
@@
 \probmax(\len) & = g\left(\frac{\len} a, 
-\tiles-3, \tiles-3 \right) \quad\ \text{for } t \text{ odd, } \textstyle \frac \len a \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2 \right]; \\
+\tiles-3, \tiles-3 \right) \quad\ \text{for } \tiles \text{ odd, } \textstyle \frac \len a \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2 \right]; \\
 \probmax(\len) & = g\left(\frac{\len} a, \tiles-3, \tiles-3 \right) + 2 g\left(\frac{\len} a, \tiles-5, \tiles-1 \right) \\
-& \text{for } t \text{ odd, } \textstyle \frac \len a \in \left (\frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2,  \frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2 \right]; \\
-\probmax(\len) & = 2 g\left(\frac{\len} a, \tiles-4, \tiles-2 \right) \quad \text{for } t \text{ even, } \textstyle \frac \len a \in \left( \frac{\sqrt{(\tiles-4)^2+(\tiles-2)^2}} 2, \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2 \right]; \\
+& \text{for } \tiles \text{ odd, } \textstyle \frac \len a \in \left (\frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2,  \frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2 \right]; \\
+\probmax(\len) & = 2 g\left(\frac{\len} a, \tiles-4, \tiles-2 \right) \quad \text{for } \tiles \text{ even, } \textstyle \frac \len a \in \left( \frac{\sqrt{(\tiles-4)^2+(\tiles-2)^2}} 2, \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2 \right]; \\
 \probmax(\len) & = 2 g\left(\frac{\len} a, \tiles-4, \tiles-2 \right) + 2 g\left(\frac{\len} a, \tiles-6, \tiles \right) \\
-& \text{for } t \text{ even, } \textstyle \frac \len a \in \left( \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2, \frac{\tiles-2}{\sqrt{2}} \right].
+& \text{for } \tiles \text{ even, } \textstyle \frac \len a \in \left( \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2, \frac{\tiles-2}{\sqrt{2}} \right].
 \end{split}
@@
 \centering%
-\subfigure[The second endpoint is in the tile with lower-left corner $((t-1)/2,(t-1)/2)$]{%
+\subfigure[The second endpoint is in the tile with lower-left corner $((\tiles-1)/2,(\tiles-1)/2)$]{%
 \label{fig: probmax_funt_odd_diag}%
@@
 }\\%\hfill%
-\subfigure[The second endpoint is in the tile with lower-left corner $((t-3)/2,(t+1)/2)$]{%
+\subfigure[The second endpoint is in the tile with lower-left corner $((\tiles-3)/2,(\tiles+1)/2)$]{%
 \label{fig: probmax_funt_odd_offdiag}%
@@
 Under this assumption, the range of lengths corresponding to $\tiles$ is given by \eqref{eq: square, odd max tiles, lengths}.
-For any length in this interval, there are some positions within the reference file from which the tile with lower-left coordinate $((t-1)/2, (t-1)/2)$ can be reached.
-Equivalently, the shifted segment with one endpoint at $((t-1)/2, (t-1)/2)$ can reach the reference tile.
+For any length in this interval, there are some positions within the reference file from which the tile with lower-left coordinate $((\tiles-1)/2, (\tiles-1)/2)$ can be reached.
+Equivalently, the shifted segment with one endpoint at $((\tiles-1)/2, (\tiles-1)/2)$ can reach the reference tile.
 This is illustrated in Figure~\ref{fig: probmax_funt_odd_diag}.
@@
 \end{equation}
-the two tiles with lower-left coordinates $((t-3)/2, (t+1)/2)$ and $((t+1)/2, (t-3)/2)$ are also reachable from the reference tile; or equivalently the shifted segment can reach the reference tile, as shown in Figure~\ref{fig: probmax_funt_odd_offdiag}.
+the two tiles with lower-left coordinates $((\tiles-3)/2, (\tiles+1)/2)$ and $((\tiles+1)/2, (\tiles-3)/2)$ are also reachable from the reference tile; or equivalently the shifted segment can reach the reference tile, as shown in Figure~\ref{fig: probmax_funt_odd_offdiag}.
 This can only happen for $\tiles \geq 5$ (for $\tiles=3$ the inequality \eqref{eq: probmax len odd2} cannot be satisfied for lengths in the range given by \eqref{eq: square, odd max tiles, lengths}).
@@
 
-For $\tiles$ even, which corresponds to lengths in the interval \eqref{eq: square, even max tiles, lengths}, the two tiles with lower-left coordinates $(t/2-1, t/2)$ and $(t/2, t/2-1)$ can always be reached from some positions within the reference tile.
+For $\tiles$ even, which corresponds to lengths in the interval \eqref{eq: square, even max tiles, lengths}, the two tiles with lower-left coordinates $(\tiles/2-1, \tiles/2)$ and $(\tiles/2, \tiles/2-1)$ can always be reached from some positions within the reference tile.
 In addition, if
@@
 
-In the first case, as has been argued, only the tile with lower-left coordinates $((t-1)/2, (t-1)/2)$ is reachable.
-For a given $\theta$, the conditional probability $\probmax_{(t-1)/2, (t-1)/2}(\len, \theta)$ is the area of the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_diag} divided by $a^2$.
+In the first case, as has been argued, only the tile with lower-left coordinates $((\tiles-1)/2, (\tiles-1)/2)$ is reachable.
+For a given $\theta$, the conditional probability $\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)$ is the area of the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_diag} divided by $a^2$.
 The horizontal and vertical sides of this rectangle are $\len\sin\theta - a(\tiles-3)/2$ and $\len\cos\theta - a(\tiles-3)/2$ respectively, and thus
@@
 \label{eq: probmax: probmax theta odd}
-\probmax_{(t-1)/2, (t-1)/2}(\len, \theta) = \frac{(\len\sin\theta - a(\tiles-3)/2) (\len\cos\theta - a(\tiles-3)/2)} {a^2}
+\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta) = \frac{(\len\sin\theta - a(\tiles-3)/2) (\len\cos\theta - a(\tiles-3)/2)} {a^2}
 \end{equation}
@@
 \begin{split}
-\probmax(\len) & = 4 \int_{\theta_0}^{\theta_1} \frac{\probmax_{(t-1)/2, (t-1)/2}(\len, \theta)}{2\pi} \diff \theta \\
+\probmax(\len) & = 4 \int_{\theta_0}^{\theta_1} \frac{\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)}{2\pi} \diff \theta \\
 & = \frac 2 \pi \int_{\arcsin\frac{a(\tiles-3)}{2\len}}^{\arccos\frac{a(\tiles-3)}{2\len}} \left( \frac{\len\sin\theta}a - \frac{\tiles-3} 2 \right)  \left( \frac{\len\cos\theta}a - \frac{\tiles-3} 2 \right) \diff \theta \\
-& = g\left( \frac \len a, t-3, t-3 \right ),
+& = g\left( \frac \len a, \tiles-3, \tiles-3 \right ),
 \end{split}
@@
 
-For the second case in \eqref{eq: probmax}, $\probmax(\len)$ is obtained as in the preceding paragraph plus an additional term corresponding to the contribution of the tile with lower-left corner $((t-3)/2, (t+1)/2)$, multiplied by $2$ to account for its symmetrical tile $((t+1)/2, (t-3)/2)$ and by $4$ to include all quadrants.
+For the second case in \eqref{eq: probmax}, $\probmax(\len)$ is obtained as in the preceding paragraph plus an additional term corresponding to the contribution of the tile with lower-left corner $((\tiles-3)/2, (\tiles+1)/2)$, multiplied by $2$ to account for its symmetrical tile $((\tiles+1)/2, (\tiles-3)/2)$ and by $4$ to include all quadrants.
 The sides of the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_offdiag} are $\len\sin\theta - a(\tiles-5)/2$ and $\len\cos\theta - a(\tiles-1)/2$, and therefore
@@
 \label{eq: probmax: probmax theta odd2}
-\probmax_{(t-3)/2, (t+1)/2}(\len, \theta) = \frac{(\len\sin\theta - a(\tiles-5)/2) (\len\cos\theta - a(\tiles-1)/2)} {a^2}.
+\probmax_{(\tiles-3)/2, (\tiles+1)/2}(\len, \theta) = \frac{(\len\sin\theta - a(\tiles-5)/2) (\len\cos\theta - a(\tiles-1)/2)} {a^2}.
 \end{equation}
@@
 \begin{multline}
-8 \int_{\theta_0}^{\theta_1} \frac{\probmax_{(t-3)/2, (t+1)/2}(\len, \theta)}{2\pi} \diff \theta \\
+8 \int_{\theta_0}^{\theta_1} \frac{\probmax_{(\tiles-3)/2, (\tiles+1)/2}(\len, \theta)}{2\pi} \diff \theta \\
 = \frac 4 \pi \int_{\arcsin\frac{a(\tiles-5)}{2\len}}^{\arccos\frac{a(\tiles-1)}{2\len}} \left( \frac{\len\sin\theta}a - \frac{\tiles-5} 2 \right)  \left( \frac{\len\cos\theta}a - \frac{\tiles-1} 2 \right) \diff \theta,
@@
 \end{proof}
+
+It should be noted that the technique used in the proof of Theorem~\ref{theo: probmax, sq} could in principle be used for obtaining the probability that the number of tiles visited by the segment equals or exceeds any given value $t \funt(\len)$.
+The resulting expressions, however, are cumbersome, because more cases (more subintervals for the length) need to be distinguished depending on how many tiles with $i+j-1=t$ can be reached from the reference tile.
 

### 2021-08-26 figures/figures.m
Check for new formula found by Alex (square grid, real-valued length)

@@ -1812,3 +1812,28 @@ result_computed(ind_even2) = result_computed(ind_even2) + ...
     2*L_fine(ind_even2).^2/a^2 - funt_fine(ind_even2).*sqrt(4*L_fine(ind_even2).^2/a^2 - (funt_fine(ind_even2)-6).^2) - (funt_fine(ind_even2)-6).*sqrt(4*L_fine(ind_even2).^2/a^2 - funt_fine(ind_even2).^2) ) / pi;
 plot(L_fine, result_computed.*L_fine), grid on
 
+
+%% #69. Like the first part of #46 (that is, funt) but including formula found by Alex for funt in the square case with real-valued lengths (e-mail 26viii21)
+
+clear
+clc
+L = .001:.001:300;
+G = 1:35;
+a = 1.5754;
+b = a;
+
+ii = ceil((a + b*real(sqrt(4*L.^2/(a^2+b^2)-1))) / 2/a) + 1;
+jj = ceil(sqrt( L.^2 - (ii-2).^2*a^2 )/b) + 1;
+funt_rect = ii+jj-1;
+
+ii = ceil((1 + real(sqrt(2*L.^2/a^2-1)))/2) + 1;
+jj = ceil(sqrt(L.^2/a^2 - (ii-2).^2 )) + 1;
+funt_sq = ii+jj-1;
+
+ii = ceil(L/a/sqrt(2)) + 1;
+jj = ceil(sqrt(L.^2/a^2 - (ii-2).^2 )) + 1;
+funt_sq_2 = ii+jj-1;
+
+funt_sq_3 = floor(sqrt(2*ceil(L.^2/a^2) - 2)) + 3; % formula found by Alex
+isequal(funt_rect, funt_sq, funt_sq_2, funt_sq_3)
+

### 2021-08-26 paper.tex
Got rid of "a" in the square grid with real-valued lengths

@@
 In order to derive a general method to build a minimal sufficient set, it is insightful to consider two specific cases first, as will be presented in the following.
-It is also helpful to use the equivalent formulation of the direct problem referred to in \S\ref{part: intro}, i.e.\ finding the maximum number of tiles visited by segments of length $\len$ \emph{or less}.
+It is also helpful to use the equivalent formulation of the direct problem referred to in \S\ref{part: intro}, i.e.~finding the maximum number of tiles visited by segments of length $\len$ \emph{or less}.
 
@@
 
-\subsection{Square grid with real-valued lengths}
+\subsection{Unit square grid with real-valued lengths}
 \label{part: max: square grid, real lengths}
 
-Particularizing the results in \S\ref{part: max: arbitrary grid, real lengths} to a square grid, $a=b$, obviously yields simpler formulas.
+A square grid has $a=b$. If the segment lengths are real-valued it can be further assumed that $a=1$. For $a \neq 1$ the results to be obtained apply with $\len$ replaced by $\len/a$.
+
+Particularizing the results in \S\ref{part: max: arbitrary grid, real lengths} to $a=b=1$ obviously yields simpler formulas.
 
@@
 \label{cor: funt, sq, form}
-For a square grid with $a \in \mathbb R^+$, and for $\len \in \mathbb R^+$,
+For a unit square grid with $\len \in \mathbb R^+$,
 \begin{equation}
@@
 \label{eq: cor: funt, sq, form; i}
-\isoli &= \left\lceil \frac{1 + \Re \sqrt{2\len^2/a^2-1}}{2} \right\rceil + 1, \\
+\isoli &= \left\lceil \frac{1 + \Re \sqrt{2\len^2-1}}{2} \right\rceil + 1, \\
 \label{eq: cor: funt, sq, form; j}
-\jsoli &= \left\lceil \sqrt{\len^2/a^2-(\isoli-2)^2} \right\rceil + 1,
+\jsoli &= \left\lceil \sqrt{\len^2-(\isoli-2)^2} \right\rceil + 1,
 \end{align}
@@
 \label{theo: funl, sq, form}
-For a square grid with $a \in \mathbb R^+$, and for $\tiles \in \mathbb N$,
+For a unit square grid, and for $\tiles \in \mathbb N$,
 \begin{equation}
 \label{eq: theo: funt, sq, form; funl}\\
-\frac{\funl(\tiles)}{a} = \begin{cases}
+\funl(\tiles) = \begin{cases}
 \displaystyle
@@
 \displaystyle
-\sqrt{\frac{(\tiles-3)^2+1} {2}} & \text{for } \tiles \text{ even, } \tiles \geq 3.
+\sqrt{\frac{(\tiles-3)^2+1} {2}} & \text{for } \tiles \text{ even, } \tiles \geq 4.
 \end{cases}
@@
 \label{theo: funt, sq, form}
-For a square grid with $a \in \mathbb R^+$, and for $\len \in \mathbb R^+$,
+For a unit square grid with $\len \in \mathbb R^+$,
 \begin{equation}
@@
 \label{eq: theo: funt, sq, form; i}
-\isoli &= \left\lceil \frac{\len}{a \sqrt{2}} \right\rceil + 1, \\
+\isoli &= \left\lceil \frac{\len}{\sqrt{2}} \right\rceil + 1, \\
 \label{eq: theo: funt, sq, form; j}
-\jsoli &= \left\lceil \sqrt{\len^2/a^2-(\isoli-2)^2} \right\rceil + 1.
+\jsoli &= \left\lceil \sqrt{\len^2-(\isoli-2)^2} \right\rceil + 1.
 \end{align}
@@
 \begin{proof}
-For $a=b$ the set $\mss$ from Proposition~\ref{prop: min suff set} consists of points of the form $(i,i)$ and $(i+1,i)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}.
+For $a=b=1$ the set $\mss$ from Proposition~\ref{prop: min suff set} consists of points of the form $(i,i)$ and $(i+1,i)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}.
 By symmetry, replacing each point $(i+1,i)$ by $(i,i+1)$ gives a set $\mss'$ which is also minimal sufficient.
 For this new set, the lower bounding line \eqref{eq: lower bound, line} can be replaced by the simpler $j=i$.
-The same approach used in the proof of Theorem~\ref{theo: funt, form} can be used, but using this line.
+The same approach used in the proof of Theorem~\ref{theo: funt, form} can be followed, but using this line.
 Thus $(\isolr, \jsolr)$ is obtained from
 \begin{align}
-(\isolr-2)^2 a^2 + (\jsolr-2)^2 a^2 &= \len^2, \\
+(\isolr-2)^2 + (\jsolr-2)^2  &= \len^2, \\
 \jsolr &= \isoli,
@@
 \label{eq: theo: funt, sq, form; proof 1}
-\isolr = \frac{\len}{a \sqrt{2}} + 2.
+\isolr = \len/\sqrt{2} + 2.
 \end{equation}
@@
 \label{eq: square, odd max tiles, lengths}
-\frac \len a \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{ \sqrt{(\tiles-1)^2+(\tiles-3)^2}}{2} \right].
+\len \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{ \sqrt{(\tiles-1)^2+(\tiles-3)^2}}{2} \right].
 \end{equation}
@@
 \label{eq: square, even max tiles, lengths}
-\frac \len a \in \left(\frac{ \sqrt{(\tiles-2)^2+(\tiles-4)^2}}{2}, \frac{\tiles-2}{\sqrt{2}} \right].
+\len \in \left(\frac{ \sqrt{(\tiles-2)^2+(\tiles-4)^2}}{2}, \frac{\tiles-2}{\sqrt{2}} \right].
 \end{equation}
@@
 
-A natural variation of the two problems introduced in \S\ref{part: intro} is to consider $a=b=1$, with the additional restriction that the segment length can only be a positive integer (equivalently, a square grid is considered with step $a$ and the segment length can only be an integer multiple of $a$).
+A natural variation of the first two problems introduced in \S\ref{part: intro} is to consider $a=b=1$, with the additional restriction that the segment length can only be a positive integer (equivalently, a square grid is considered with step $a$ and the segment length can only be an integer multiple of $a$).
 

### 2021-08-26 paper.tex
Got rid of "a" in for probmax too

@@
 
-\subsection{Square grid with real-valued lengths}
-
-Consider a square grid with spacing $a \in \mathbb R^+$.
-For a random segment with length $\len \in \mathbb R^+$, let the function $\probmax: \mathbb R^+ \to [0,1]$ be defined such that $\probmax(\len)$ gives the probability that the segment visits the maximum possible number of tiles, $\funt(\len)$.
+\subsection{Unit square grid with real-valued lengths}
+
+Consider a square grid with unit spacing, $a=b=1$. Again, the results to follow can be applied to a square grid with spacing $a \neq 1$ if $\len$ is replaced by $\len/a$.
+
+For a random segment with length $\len \in \mathbb R^+$ on a unit square grid, let the function $\probmax: \mathbb R^+ \to [0,1]$ be defined such that $\probmax(\len)$ gives the probability that the segment visits the maximum possible number of tiles, $\funt(\len)$.
 This function is characterized by the next theorem.
@@
 \label{theo: probmax, sq}
-Given $a, \len \in \mathbb R^+$, consider a square grid with spacing $a$ and a uniformly random segment of length $\len$, as previously defined.
+Given $\len \in \mathbb R^+$, consider a unit square grid and a uniformly random segment of length $\len$, as previously defined.
 The probability $\probmax(\len)$ that the segment visits the maximum number of tiles $\funt(\len)$ is
@@
 \begin{split}
-\probmax(\len) & = g\left(\frac{\len} a, 
-\tiles-3, \tiles-3 \right) \quad\ \text{for } \tiles \text{ odd, } \textstyle \frac \len a \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2 \right]; \\
-\probmax(\len) & = g\left(\frac{\len} a, \tiles-3, \tiles-3 \right) + 2 g\left(\frac{\len} a, \tiles-5, \tiles-1 \right) \\
-& \text{for } \tiles \text{ odd, } \textstyle \frac \len a \in \left (\frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2,  \frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2 \right]; \\
-\probmax(\len) & = 2 g\left(\frac{\len} a, \tiles-4, \tiles-2 \right) \quad \text{for } \tiles \text{ even, } \textstyle \frac \len a \in \left( \frac{\sqrt{(\tiles-4)^2+(\tiles-2)^2}} 2, \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2 \right]; \\
-\probmax(\len) & = 2 g\left(\frac{\len} a, \tiles-4, \tiles-2 \right) + 2 g\left(\frac{\len} a, \tiles-6, \tiles \right) \\
-& \text{for } \tiles \text{ even, } \textstyle \frac \len a \in \left( \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2, \frac{\tiles-2}{\sqrt{2}} \right].
+\probmax(\len) & = g\left(\len, 
+\tiles-3, \tiles-3 \right) \quad\ \text{for } \tiles \text{ odd, } \textstyle \len \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2 \right]; \\
+\probmax(\len) & = g\left(\len, \tiles-3, \tiles-3 \right) + 2 g\left(\len, \tiles-5, \tiles-1 \right) \\
+& \text{for } \tiles \text{ odd, } \textstyle \len \in \left (\frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2,  \frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2 \right]; \\
+\probmax(\len) & = 2 g\left(\len, \tiles-4, \tiles-2 \right) \quad \text{for } \tiles \text{ even, } \textstyle \len \in \left( \frac{\sqrt{(\tiles-4)^2+(\tiles-2)^2}} 2, \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2 \right]; \\
+\probmax(\len) & = 2 g\left(\len, \tiles-4, \tiles-2 \right) + 2 g\left(\len, \tiles-6, \tiles \right) \\
+& \text{for } \tiles \text{ even, } \textstyle \len \in \left( \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2, \frac{\tiles-2}{\sqrt{2}} \right].
 \end{split}
@@
 For each $\theta$ in the allowed range, the valid positions for the first endpoint of the (shifted) segment are in the shaded rectangle, whose dimensions depend on $\theta$.
-The fraction of the reference tile occupied by this rectangle is the probability $\probmax_{i-1,j-1}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$.
+The area occupied by this rectangle is the probability $\probmax_{i-1,j-1}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$.
 The desired probability, $\probmax(\len)$, is obtained by averaging $\probmax_{i-1,j-1}(\len, \theta)$ over $\theta$ and summing over all $i-1,j-1$ that can be reached from the reference tile.
@@
 In the first case, as has been argued, only the tile with lower-left coordinates $((\tiles-1)/2, (\tiles-1)/2)$ is reachable.
-For a given $\theta$, the conditional probability $\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)$ is the area of the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_diag} divided by $a^2$.
-The horizontal and vertical sides of this rectangle are $\len\sin\theta - a(\tiles-3)/2$ and $\len\cos\theta - a(\tiles-3)/2$ respectively, and thus
+For a given $\theta$, the conditional probability $\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)$ is the area of the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_diag}:
 \begin{equation}
 \label{eq: probmax: probmax theta odd}
-\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta) = \frac{(\len\sin\theta - a(\tiles-3)/2) (\len\cos\theta - a(\tiles-3)/2)} {a^2}
+\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta) = \left(\len\sin\theta - \frac{\tiles-3}2\right) \left(\len\cos\theta - \frac{\tiles-3}2\right).
 \end{equation}
@@
 \label{eq: odd theta 0}
-\theta_0 & = \arcsin\frac{a(\tiles-3)}{2\len}, \\
+\theta_0 & = \arcsin\frac{\tiles-3}{2\len}, \\
 \label{eq: odd theta 1}
-\theta_1 & = \arccos\frac{a(\tiles-3)}{2\len}.
+\theta_1 & = \arccos\frac{\tiles-3}{2\len}.
 \end{align}
@@
 \probmax(\len) & = 4 \int_{\theta_0}^{\theta_1} \frac{\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)}{2\pi} \diff \theta \\
-& = \frac 2 \pi \int_{\arcsin\frac{a(\tiles-3)}{2\len}}^{\arccos\frac{a(\tiles-3)}{2\len}} \left( \frac{\len\sin\theta}a - \frac{\tiles-3} 2 \right)  \left( \frac{\len\cos\theta}a - \frac{\tiles-3} 2 \right) \diff \theta \\
-& = g\left( \frac \len a, \tiles-3, \tiles-3 \right ),
+& = \frac 2 \pi \int_{\arcsin\frac{\tiles-3}{2\len}}^{\arccos\frac{\tiles-3}{2\len}} \left( \len\sin\theta - \frac{\tiles-3} 2 \right)  \left( \len\cos\theta - \frac{\tiles-3} 2 \right) \diff \theta \\
+& = g\left( \len, \tiles-3, \tiles-3 \right ),
 \end{split}
@@
 For the second case in \eqref{eq: probmax}, $\probmax(\len)$ is obtained as in the preceding paragraph plus an additional term corresponding to the contribution of the tile with lower-left corner $((\tiles-3)/2, (\tiles+1)/2)$, multiplied by $2$ to account for its symmetrical tile $((\tiles+1)/2, (\tiles-3)/2)$ and by $4$ to include all quadrants.
-The sides of the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_offdiag} are $\len\sin\theta - a(\tiles-5)/2$ and $\len\cos\theta - a(\tiles-1)/2$, and therefore
+Observing the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_offdiag},
 \begin{equation}
 \label{eq: probmax: probmax theta odd2}
-\probmax_{(\tiles-3)/2, (\tiles+1)/2}(\len, \theta) = \frac{(\len\sin\theta - a(\tiles-5)/2) (\len\cos\theta - a(\tiles-1)/2)} {a^2}.
+\probmax_{(\tiles-3)/2, (\tiles+1)/2}(\len, \theta) = \left(\len\sin\theta - \frac{\tiles-5}2\right) \left(\len\cos\theta - \frac{\tiles-1}2\right).
 \end{equation}
@@
 \begin{align}
-\theta_0 & = \arcsin\frac{a(\tiles-5)}{2\len}, \\
-\theta_1 & = \arccos\frac{a(\tiles-1)}{2\len}.
+\theta_0 & = \arcsin\frac{\tiles-5}{2\len}, \\
+\theta_1 & = \arccos\frac{\tiles-1}{2\len}.
 \end{align}
@@
 8 \int_{\theta_0}^{\theta_1} \frac{\probmax_{(\tiles-3)/2, (\tiles+1)/2}(\len, \theta)}{2\pi} \diff \theta \\
-= \frac 4 \pi \int_{\arcsin\frac{a(\tiles-5)}{2\len}}^{\arccos\frac{a(\tiles-1)}{2\len}} \left( \frac{\len\sin\theta}a - \frac{\tiles-5} 2 \right)  \left( \frac{\len\cos\theta}a - \frac{\tiles-1} 2 \right) \diff \theta,
+= \frac 4 \pi \int_{\arcsin\frac{\tiles-5}{2\len}}^{\arccos\frac{\tiles-1}{2\len}} \left( \len\sin\theta - \frac{\tiles-5} 2 \right)  \left( \len\cos\theta - \frac{\tiles-1} 2 \right) \diff \theta,
 \end{multline}
-is recognized to be $2 g(\len/a, \tiles-5, \tiles-1)$. This proves the second equality in \eqref{eq: probmax}.
+is recognized to be $2 g(\len, \tiles-5, \tiles-1)$. This proves the second equality in \eqref{eq: probmax}.
 
@@
 
-It should be noted that the technique used in the proof of Theorem~\ref{theo: probmax, sq} could in principle be used for obtaining the probability that the number of tiles visited by the segment equals or exceeds any given value $t \funt(\len)$.
+It should be noted that the technique used in the proof of Theorem~\ref{theo: probmax, sq} could in principle be used for obtaining the probability that the number of tiles visited by the segment equals or exceeds any given value $t < \funt(\len)$.
 The resulting expressions, however, are cumbersome, because more cases (more subintervals for the length) need to be distinguished depending on how many tiles with $i+j-1=t$ can be reached from the reference tile.
 
-Figure~\ref{fig: probmax_len} shows the probability $\probmax(\len)$ of visiting the maximum number of tiles on a square grid with spacing $a$, as given by \eqref{eq: probmax}.
+Figure~\ref{fig: probmax_len} shows the probability $\probmax(\len)$ of visiting the maximum number of tiles on a unit square grid, as given by \eqref{eq: probmax}.
 As $\len$ grows, $\probmax(\len)$ has a jump when the maximum number of tiles that can be visited increases by $1$.
@@
 \displaystyle
-\frac{a \sqrt{(\tiles-3)^2 + (\tiles-1)^2}} {2} & \text{ for $\tiles$ odd, $\tiles \geq 3$} \\
+\frac{\sqrt{(\tiles-3)^2 + (\tiles-1)^2}} {2} & \text{ for $\tiles$ odd, $\tiles \geq 3$} \\
 \displaystyle
-\frac{a (\tiles-2)}{\sqrt{2}} & \text{ for $\tiles$ odd, $\tiles \geq 4$}.
+\frac{\tiles-2}{\sqrt{2}} & \text{ for $\tiles$ odd, $\tiles \geq 4$}.
 \end{cases}
 \end{equation}
-For example, the first continuous section seen in the figure corresponds to a maximum number of tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, with $\len_3= a$.
-The second corresponds to $\tiles=4$, for lengths in $(\len_3, \len_4]$, with $\len_4= a\sqrt{2}$.
+For example, the first continuous section seen in the figure corresponds to a maximum number of tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, with $\len_3= 1$.
+The second corresponds to $\tiles=4$, for lengths in $(\len_3, \len_4]$, with $\len_4= \sqrt{2}$.
 Within each continuous section the probability monotonically increases from $0$ to a maximum value.
-The heights of the maxima follow a different trend*** for odd and even $\tiles$, as seen in the figure; but in each case they are asymptotically proportional to $1/\len_\tiles$, as the next result establishes.
+The heights of the maxima follow a different trend**better word?** for odd and even $\tiles$, as seen in the figure; but in each case they are asymptotically proportional to $1/\len_\tiles$, as the next result establishes.
 
@@
 \label{eq: probmax limsup}
-\limsup_{\len \rightarrow \infty}\, \len \probmax(\len) & = \frac {4\sqrt{2} a}{3\pi}.
+\limsup_{\len \rightarrow \infty}\, \len \probmax(\len) & = \frac {4\sqrt{2}}{3\pi}.
 \end{align}
@@
 
-For large $\tiles$, $\len_\tiles$ asymptotically approaches $a\tiles/\sqrt{2}$.
+For large $\tiles$, $\len_\tiles$ asymptotically approaches $\tiles/\sqrt{2}$.
 Using this into \eqref{eq: probmax lim even}, which is greater than \eqref{eq: probmax lim odd}, establishes \eqref{eq: probmax limsup}.

### 2021-08-26 figures/figures.m

@@ -1788,7 +1788,7 @@ set(gca, 'colororderindex', 1)
 plot(L_fine.*(ind_even|ind_even2)./(ind_even|ind_even2), result_computed.*(ind_even|ind_even2)./(ind_even|ind_even2), '-')
 %plot(L_fine, result_computed, '-')
 legend({'Odd maximum number of squares' 'Even maximum number of squares'})
-xlabel([char(hex2dec('2113')) '/a']), ylabel([char(hex2dec('03C1')) '(' char(hex2dec('2113')) ')'])
+xlabel(char(hex2dec('2113'))), ylabel([char(hex2dec('03C1')) '(' char(hex2dec('2113')) ')'])
 
 
 %% #68. Like #61 but including "a" (grid spacing)

### 2021-08-27 paper.tex
Changes in structure

@@
 \includegraphics[width=.7\textwidth]{ijLS_1}%
-\caption{Relationship between segment length and number of visited tiles with the width and height of the canonical rectangle, for $a=b$
+\caption{Relationship of segment length and number of visited tiles with the width and height of the canonical rectangle, for $a=b$
 }%
@@
 \end{figure}%
+% ***Combine this figure and the next into one figure with two subfigures (stacked vertically, size similar to now)? This has been done in others
+% ***Are these figures necessary?
+They were created with the aim of helping understand the first of the two propositions on $\mss$.
+Now that proposition has been removed.
+They are probably still useful, but maybe the text about these figures needs to be rewritten.
 
@@
 \includegraphics[width=.7\textwidth]{ijLS_1p35}%
-\caption{Relationship between segment length and number of visited tiles with the width and height of the canonical rectangle, for $a=1.35b$
+\caption{Relationship of segment length and number of visited tiles with the width and height of the canonical rectangle, for $a=1.35b$
 }%
@@
 
-In general, a systematic way to obtain a minimal sufficient set is to gradually increase $\len$ starting from $0$, which gives an arc of increasing radius in the figure; and to take note of the first $(i,j)$ pair that is reached for each line of constant $i+j-1$.
-If two such pairs are reached at the same time, any of them can be used.
-In Figure~\ref{fig: ijLS_1p35} this rule produces the following pairs, in order: $(2,2)$, $(2,3)$, $(3,3)$, $(3,4)$, $(3,5)$, $(4,5)$, \ldots, as indicated by the filled circles.
-
-Based on the above, an explicit procedure that produces a minimal sufficient set can be given. This set will be denoted as $\mss = \{(i_3,j_3), (i_4.j_4), \ldots\}$.
+The following proposition gives an explicit method to obtain a minimal sufficient set, $\mss = \{(i_3,j_3), (i_4.j_4), \ldots\}$, where the pair $(i_\tiles,j_\tiles)$ corresponds to $i+j-1 = \tiles$.
 
 \begin{proposition}
-\label{prop: min suff set}
-Given grid parameters $a, b \in \mathbb R^+$, a minimal sufficient set $\mss$ can be obtained as follows.
-The chosen pair for $i+j-1=3$ is $(2,2)$.
-For each $\tiles \in \mathbb N, \tiles \geq 3$, let $(i_\tiles, j_\tiles)$ denote the pair chosen for $i+j-1=\tiles$.
-Then $(i_{\tiles+1}, j_{\tiles+1})$ is either $(i_\tiles+1, j_\tiles)$ or $(i_\tiles, j_\tiles+1)$, whichever minimizes $(i_{\tiles+1}-2)^2 a^2 + (j_{\tiles+1}-2)^2 b^2$; if both are minimizers $(i_\tiles+1, j_\tiles)$ is taken.
-Equivalently, $(i_{\tiles+1}, j_{\tiles+1})$ is $(i_\tiles, j_\tiles+1)$ if
-\begin{equation}
-\label{eq: cond for incrementing j}
-j_\tiles < \frac{i_\tiles a^2}{b^2} - \frac{3a^2}{2b^2} + \frac 3 2,
-\end{equation}
-or $(i_\tiles+1, j_\tiles)$ otherwise.
+\label{prop: min suff set, form}
+Given $a, b \in \mathbb R^+$ and $\tiles \in \mathbb N$, $\tiles \geq 3$, a minimal sufficient set $\mss = \{(i_3,j_3), (i_4.j_4), \ldots\}$ can be obtained as
+\begin{align}
+\label{eq: min suff set, form, i}
+i_\tiles &= \left\lfloor \frac {(2\tiles-1)b^2+5a^2}{2(a^2+b^2)} \right\rfloor, \\
+\label{eq: min suff set, form, j}
+j_\tiles &= \left\lceil \frac {2\tiles a^2 + 3(b^2-a^2)}{2(a^2+b^2)} \right\rceil,
+\end{align}
+where $i_\tiles + j_\tiles-1 = \tiles$. All pairs $(i_\tiles,j_\tiles)$ are strictly below the line
+\begin{equation}
+\label{eq: upper bound, line}
+j = \frac{i a^2}{b^2} - \frac{3a^2}{2b^2} + \frac 5 2,
+\end{equation}
+and above or on the line
+\begin{equation}
+\label{eq: lower bound, line}
+j = \frac{i a^2}{b^2} - \frac{5a^2}{2b^2} + \frac 3 2.
+\end{equation}
 \end{proposition}
@@
 \begin{proof}
-For each $\tiles$, the pair $(i_\tiles,j_\tiles)$ of the minimal sufficient set should be chosen as that on the line $i+j-1=\tiles$ which minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$.
-This allows maximizing the sum $i+j-1$, and thus the number of visited tiles, for a given length restriction (direct problem); or minimizing the required lengths for a specified number of visited files (inverse problem).
-
-The fact that $(i_t,j_t)$ minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$ among all pairs with $i+j-1=\tiles$ can be expressed as follows: for any $k \in \mathbb N$,
-\begin{align}
-\label{eq: chosen contiguous ineq g, 1}
-(i_\tiles-2)^2 a^2 + (j_\tiles-2)^2 b^2 &\leq (i_\tiles-2-k)^2 a^2 + (j_\tiles-2+k)^2 b^2, \\
-\label{eq: chosen contiguous ineq g, 2}
-(i_\tiles-2)^2 a^2 + (j_\tiles-2)^2 b^2 &\leq (i_\tiles-2+k)^2 a^2 + (j_\tiles-2-k)^2 b^2.
-\end{align}
-To see that $(i_{\tiles+1},j_{\tiles+1})$ is either $(i_\tiles,j_\tiles+1)$ or $(i_\tiles+1,j_\tiles)$ it suffices to prove that, for any $k \in \mathbb N$,
-\begin{align}
-\label{eq: chosen contiguous ineq g+1, 1}
-(i_{\tiles}-2)^2 a^2 + (j_{\tiles}-1)^2 b^2 &\leq (i_{\tiles}-2-k)^2 a^2 + (j_{\tiles}-1+k)^2 b^2, \\
-\label{eq: chosen contiguous ineq g+1, 2}
-(i_{\tiles}-1)^2 a^2 + (j_{\tiles}-2)^2 b^2 &\leq (i_{\tiles}-1+k)^2 a^2 + (j_{\tiles}-2-k)^2 b^2.
-\end{align}
-With reference to Figure~\ref{fig: ijLS_1p35}, consider for instance $(i_7,j_7) = (3,5)$.
-Graphically it can be seen that $(i_8,j_8)$ could only be $(3,6)$ or $(4,5)$, because for example $(2a,7b)$ is further from $(2a,2b)$ than $(3a,6b)$ is (this corresponds to \eqref{eq: chosen contiguous ineq g+1, 1} with $k=1$), and $(5a,4b)$ is also further from $(2a,2b)$ than $(4a,5b)$ is (this corresponds to \eqref{eq: chosen contiguous ineq g+1, 2} with $k=1$).
-
-The inequality \eqref{eq: chosen contiguous ineq g+1, 1} can be rewritten as
-\begin{equation}
-\label{eq: chosen contiguous ineq g+1, 1b}
-\begin{split}
-& (i_{\tiles}-2)^2 a^2 + (j_{\tiles}-2)^2 b^2 + 2(j_{\tiles}-2)b^2 + b^2 \\
-& \quad \leq (i_{\tiles}-2-k)^2 a^2 + (j_{\tiles}-2+k)^2 b^2 + 2(j_{\tiles}-2+k) b^2 + b^2.
-\end{split}
-\end{equation}
-Clearly,
-\begin{equation}
-\label{eq: chosen contiguous ineq g+1, step}
-2(j_{\tiles+1}-2)b^2 + b^2 < 2(j_{\tiles+1}-2+k) b^2 + b^2.
-\end{equation}
-Adding \eqref{eq: chosen contiguous ineq g, 1} and \eqref{eq: chosen contiguous ineq g+1, step} yields \eqref{eq: chosen contiguous ineq g+1, 1b}.
-Therefore \eqref{eq: chosen contiguous ineq g+1, 1} holds.
-A similar procedure can be used to establish \eqref{eq: chosen contiguous ineq g+1, 2}.
-
-The condition for choosing $(i_t,j_t+1)$ over $(i_t+1,j_t)$ is that
-\begin{equation}
-\label{eq: chosen contiguous criterion 1}
-(i_\tiles-2)^2 a^2 + (j_\tiles-1)^2 b^2 < (i_\tiles-1)^2 a^2 + (j_\tiles-2)^2 b^2.
-\end{equation}
-This can be expressed as
-\begin{equation}
-\label{eq: chosen contiguous criterion 2}
-2(j_\tiles-2)b^2 + b^2 < 2(i_\tiles-2)a^2 + a^2.
-\end{equation}
-Rearranging terms, \eqref{eq: chosen contiguous criterion 2} is seen to be the same as \eqref{eq: cond for incrementing j}.
+***New proof based on treating i,j as continuous along the line i+j=constant, then quantizing to the nearest allowed point along that line.
+I think we should the notation $\isolr$, $\jsolr$ for the continuous coordinates, similarly to other proofs that use continuous coordinates.
+% ***The continuous argument gives something like $i-2 = b^2/(a^2+b^2)*(t-3)$, $j = a^2/(a^2+b^2)*(t-3)$.
+We then compute how much the distance from $(i,j)$ to $(2,2)$ is increased by quantizing along the i+j line, to argue that the nearest quantized point should be chosen.
+To ensue that in case of a tie we choose the point with larger $i$, we apply floor(...+1/2) to i and ceiling(...-1/2) to j.
+
+***The bounding lines are a direct consequence of how $i_\tiles$ and $j_\tiles$ are obtained.
+
+%\begin{figure}%
+%\centering%
+%\includegraphics[width=.7\textwidth]{ijLS_1p35_detail_inverse}%
+%\caption{***Figure illustrating the proof of Proposition~\ref{prop: min suff set, form}. Example with $a=1.35, b=1$%
+%}%
+%\label{fig: ***}%
+%\end{figure}%
 \end{proof}
 
-The condition \eqref{eq: cond for incrementing j} for incrementing $j_\tiles$ in Proposition~\ref{prop: min suff set} can be expressed in terms of two bounding lines,
-\begin{align}
-\label{eq: upper bound, line}
-j &= \frac{i a^2}{b^2} - \frac{3a^2}{2b^2} + \frac 5 2, \\
-\label{eq: lower bound, line}
-j &= \frac{i a^2}{b^2} - \frac{5a^2}{2b^2} + \frac 3 2.
-\end{align}
-All pairs $(i,j)$ in $\mss$ are strictly below the upper line \eqref{eq: upper bound, line} and above or on the lower line \eqref{eq: lower bound, line}.
-This is shown in Figure~\ref{fig: pairs_bounds}, using three different pairs of grid parameters $a$, $b$ as examples.
+The bounding lines in Proposition~\ref{prop: min suff set, form} are shown in Figure~\ref{fig: pairs_bounds}, using three different pairs of grid parameters $a$, $b$ as examples.
 Given $(i_\tiles,j_\tiles) \in \mss$, the next pair $(i_{\tiles+1},j_{\tiles+1})$ is obtained by incrementing $j$ if that results in a point below \eqref{eq: upper bound, line}.
@@
 
-The construction of $\mss$ in Proposition~\ref{prop: min suff set} computes the pairs $(i_\tiles,i_\tiles)$ \emph{sequentially}: $(i_3,i_3), \ldots, (i_{\tiles-1},i_{\tiles-1})$ need to be previously computed in order to obtain $(i_\tiles,i_\tiles)$.
-Proposition~\ref{prop: min suff set, form} avoids this by providing a direct formula.
-Its derivation is based on finding the intersection point of \eqref{eq: lower bound, line} with  $i+j-1=\tiles$ and appropriately rounding to integer coordinates in order to satisfy \eqref{eq: upper bound, line}.
-% The proof method is briefly described here because in a later Theorem it is said (in the text, not in the proof) that the method is similar to this
-
-\begin{proposition}
-\label{prop: min suff set, form}
-Given $a, b \in \mathbb R^+$ and $\tiles \in \mathbb N, \tiles \geq 3$, the pair $(i_\tiles,j_\tiles)$ from the set $\mss$ defined in Proposition~\ref{prop: min suff set} can be computed as
-\begin{align}
-\label{eq: min suff set, form, i}
-i_\tiles &= \left\lfloor \frac {(2\tiles-1)b^2+5a^2}{2(a^2+b^2)} \right\rfloor, \\
-\label{eq: min suff set, form, j}
-j_\tiles &= \left\lceil \frac {2\tiles a^2 + 3(b^2-a^2)}{2(a^2+b^2)} \right\rceil. 
-\end{align}
-\end{proposition}
-
-\begin{proof}
-Let $(\isolr, \jsolr)$ denote the intersection of \eqref{eq: lower bound, line} and the line $i+j-1=\tiles$. This is easily seen to be
-\begin{align}
-\label{eq: min suff set, form, isolr}
-\isolr &= \frac {(2\tiles-1)b^2+5a^2}{2(a^2+b^2)}, \\
-\label{eq: min suff set, form, jsolr}
-\jsolr &= \frac {2\tiles a^2 + 3(b^2-a^2)}{2(a^2+b^2)}. 
-\end{align}
-This point is shown with a square marker in Figure \ref{fig: ijLS_1p35_detail_inverse}, which also shows the elements of $\mss$ as filled circles, as well as the two bounding lines.
-The line \eqref{eq: upper bound, line} is \eqref{eq: lower bound, line} shifted one step left in $i$ and one step up in $j$, i.e.~it is shifted by a diagonal step.
-Since all points of $\mss$ are between these two lines or on the lower line, the only possibility for $(i_\tiles,j_\tiles)$, as illustrated in the figure, is to be the point with integer coordinates closest to $(\isolr,\jsolr)$ along the line $i+j-1=\tiles$ in the left, upwards direction.
-This gives \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}.
-
-\begin{figure}%
-\centering%
-\includegraphics[width=.7\textwidth]{ijLS_1p35_detail_inverse}%
-\caption{Obtaining $(\isoli,\jsoli)$ in Proposition~\ref{prop: min suff set, form}. Example with $a=1.35, b=1$%
-}%
-\label{fig: ijLS_1p35_detail_inverse}%
-\end{figure}%
-\end{proof}
-
-For $a^2/b^2$ arbitrary, the number of pairs with the same $i$ coordinate in the set $\mss$ is in general irregular, because the line \eqref{eq: lower bound, line} does not follow a ``natural'' direction of the grid.
+For $a^2/b^2$ arbitrary, the number of pairs with the same $i$ coordinate in the set $\mss$ is in general irregular, because the line \eqref{eq: lower bound, line} does not follow a ``natural''
+% ***it may be better to say ``simple'' instead of ``natural'', specially if we rename ``canonical`'' to ``natural'' or ``natural bounding''
+direction of the grid.
 This happens for instance in Figure~\ref{fig: pairs_bounds_1p35}, where the number of pairs for the first $i$ values equals either $2$ or $3$ without a clear pattern.\footnote{Strictly, there is a periodic pattern whenever $a^2/b^2$ is rational, which is the case in Figure~\ref{fig: pairs_bounds_1p35}.
@@
 
-A segment whose canonical rectangle has normalized width $i$ and height $j$ is oriented with approximate slope $jb/(ia)$ with respect to the $x$ axis (See Figure~\ref{fig: len ineq i j}); and the approximation becomes better for greater segment lengths.
+A segment whose canonical rectangle has normalized width $i$ and height $j$ is oriented with approximate slope $jb/(ia)$ with respect to the $x$ axis (see Figure~\ref{fig: len ineq i j}); and the approximation becomes better for greater segment lengths.
 From \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} it can be seen that the positions of the pairs $(i,j) \in \mss$ have $j/i \approx a^2/b^2$ for large $i, j$.
@@
 The general case for rectangular grids with real-valued segment lengths is analyzed first, in \S\ref{part: max: arbitrary grid, real lengths}.
-The particular case of square grids is addressed in \S\ref{part: max: square grid, real lengths}, as it allows a specialized formula for the direct problem.
+The particular case of square grids is addressed in \S\ref{part: max: unit square grid, real lengths}, as it allows a specialized formula for the direct problem.
 Lastly, the analysis of a unit square grid with integer-valued lengths is given in \S\ref{part: max: unit square grid, integer lengths}.
@@
 In addition, it is possible to obtain a direct formula also for $\funt$ using an approach similar to that used in Proposition~\ref{prop: min suff set, form}.
+% **this approach will change. Remove the above comment about the approach begin similar, and briefly describe the approach at the beginning of the proof here. Well, that's more or less done already**
 These formulas are given in Theorems \ref{theo: funt, form} and \ref{theo: funl, form}.
 
-% ***Should i-1, j-2 be used instead of i, j? I don't think so. The formulas would be simplified a little, but their interpretation would be more difficult
 \begin{theorem}
@@
 \begin{proof}
-The general idea is similar to that used in Proposition~\ref{prop: min suff set, form}: find the intersection $(\isolr, \jsolr)$ of the line \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ for $i, j \geq 2$, and from that obtain the actual pair $(\isoli, \jsoli)$ of integer numbers which yields the solution.
+The general idea is similar to that used in Proposition~\ref{prop: min suff set, form}**this may have changed**: find the intersection $(\isolr, \jsolr)$ of the line \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ for $i, j \geq 2$, and from that obtain the actual pair $(\isoli, \jsoli)$ of integer numbers which yields the solution.
 
@@
 \end{align}
+Equivalently,
+\begin{equation}
+\text{***new formula by Alex; unless we can unify with the above}
+\end{equation}
 This function is monotone increasing for $\tiles \geq 3$.
@@
 For $\tiles \in \{1, 2\}$ both \eqref{eq: theo: funl, form; i} and \eqref{eq: theo: funl, form; j} equal $2$, and \eqref{eq: theo: funl, form; funl} gives $0$, which is the correct result.
+
+***Proof of new formula
 
@@
 For $a=b$, either from symmetry considerations or particularizing the formulas in the previous theorems it stems that the optimal orientation of the segment is close to $45^\circ$.
-This case will be dealt with in \S\ref{part: max: square grid, real lengths}, as it lends itself to somewhat simplified formulas.
+This case will be dealt with in \S\ref{part: max: unit square grid, real lengths}, as it lends itself to simplified formulas.
 
@@
 \subsection{Unit square grid with real-valued lengths}
-\label{part: max: square grid, real lengths}
+\label{part: max: unit square grid, real lengths}
 
@@
 \begin{equation}
-\label{eq: theo: funt, sq, form; funl}\\
 \funl(\tiles) = \begin{cases}
@@
 \displaystyle
-\sqrt{\frac{(\tiles-3)^2+1} {2}} & \text{for } \tiles \text{ even, } \tiles \geq 4.
+\sqrt{\frac{(\tiles-3)^2+1} {2}} & \text{for } \tiles \text{ even, } \tiles \geq 4,
 \end{cases}
 \end{equation}
+or equivalently
+\begin{equation}
+\label{eq: theo: funt, sq, form; funl}
+\funl(\tiles) = \begin{cases}
+\displaystyle
+0 & \text{for } \tiles =1, 2 \\[1.4mm]
+\displaystyle
+\sqrt{\left\lceil \frac{(\tiles-3)^2} {2} \right\rceil} & \text{for } \tiles \geq 3.
+\end{cases}
+\end{equation}
+
 \end{corollary}
 
-Furthermore, an even simpler formula can be obtained for $\funt$. Its derivation uses a variation of the minimal sufficient set $\mss$ which is  more suited to this case.
+Furthermore, an even simpler formula can be obtained for $\funt$, as the next theorem shows.
 
@@
 \begin{equation}
-\label{eq: theo: funt, sq, form; funt}
+\label{eq: theo: funt, sq, form; funt, with i, j}
 \funt(\len) = \isoli+\jsoli-1
@@
 \end{align}
+Equivalently,
+\begin{equation}
+\label{eq: theo: funt, sq, form; funt, simpler}
+\funt(\len) = \left\lfloor \sqrt{2 \left\lceil \len^2 \right\rceil - 2} \right\rfloor + 3.
+\end{equation}
 \end{theorem}
@@
 \begin{proof}
-For $a=b=1$ the set $\mss$ from Proposition~\ref{prop: min suff set} consists of points of the form $(i,i)$ and $(i+1,i)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}.
+The proof of \eqref{eq: theo: funt, sq, form; funt, with i, j} uses a variation of the minimal sufficient set $\mss$ which is more suited to this situation.
+
+For $a=b=1$ the set $\mss$ consists of points of the form $(i,i)$ and $(i+1,i)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}.
 By symmetry, replacing each point $(i+1,i)$ by $(i,i+1)$ gives a set $\mss'$ which is also minimal sufficient.
@@
 The pair $(\isoli,\jsoli)$ resulting from \eqref{eq: theo: funt, sq, form; proof 1} is in $\mss'$, and is given in \eqref{eq: theo: funt, sq, form; i} and \eqref{eq: theo: funt, sq, form; j}.
+This establishes \eqref{eq: theo: funt, sq, form; funt, with i, j}.
+
+***Alex
+***Derivation of \eqref{eq: theo: funt, sq, form; funt, simpler} \ldots
+***It looks like we may have to use different techniques for deriving \eqref{eq: theo: funt, sq, form; funt, with i, j} and \eqref{eq: theo: funt, sq, form; funt, simpler}.
+I don't see a way to use a common proof method.
+And \eqref{eq: theo: funt, sq, form; funt, with i, j} is needed because of $\isoli$, $\jsoli$, which are used in (the text below in preparation for) \S\ref{part: probmax: unit square grid, real lengths}.
 \end{proof}
@@
 
-A natural variation of the first two problems introduced in \S\ref{part: intro} is to consider $a=b=1$, with the additional restriction that the segment length can only be a positive integer (equivalently, a square grid is considered with step $a$ and the segment length can only be an integer multiple of $a$).
+A natural variation of the first two problems introduced in \S\ref{part: intro} is to consider $a=b=1$ with the additional restriction that the segment length can only be a positive integer (equivalently, a square grid is considered with step $a$ and the segment length can only be an integer multiple of $a$).
 
@@
 This is A346232 in the On-Line Encyclopedia of Integer Sequences \cite{OEIS_unitsq_int_dir}.
-For this sequence, a simpler expression can be obtained than those resulting from \S\ref{part: max: square grid, real lengths} with $a=1$, and the following properties hold.
+For this sequence, the expression in Theorem~\ref{theo: funt, sq, form} simplifies in the obvious way, and the following properties hold.
 
@@
 \begin{proof}
-***The derivation of \eqref{eq: funti leni} in CGCC can probably be adapted.
-$i$, $j$ there are $i-2$, $j-2$ here.
-(It cannot be expressed with $\lceil \cdot \rceil$ instead of $\lfloor \cdot \rfloor$, which would be more ``aesthetically aligned'' with previous expressions, because for example $\len=3$ makes $2\len^2-2$ a square).
-
-%***Otherwise, perhaps we can prove the formula for the inverse problem first (starting from the one for real lenghts), and derive this from that
+The equality \eqref{eq: funti leni} stems from \eqref{eq: theo: funt, sq, form; funt} noting that $\len$ is an integer.
 
@@
 
-Each realization of the random segment gives rise to a canonical rectangle, whose normalized dimensions $i$ and $j$ are thus random variables.
-Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and the segment visits $i+j-1$ tiles.
+Each realization of the random segment gives rise to a canonical rectangle, whose normalized dimensions $i$ and $j$ are thus random variables, as is the number $\tiles$ of visited tiles.
+Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and $\tiles = i+j-1$ tiles.
 Note that $i$ and $j$ are not statistically independent.
@@
 \subsection{Arbitrary grid with real-valued lengths}
+\label{part: probmax: arbitrary grid, real lengths}
 
 Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles visited by a random segment of length $\len$, with the distributions specified in \S\ref{part: rand}.
-
-Although the random variables $i$ and $j$ are dependent, their marginal distributions can be obtained easily, and are sufficient to compute the average number of visited tiles:
+Using the fact that
 \begin{equation}
 \label{eq: funta E[i] E[j]}
-\funta(\len) = \E[\tiles] = \E[i]+\E[j]-1.
-\end{equation}
-
-For $0 \leq \genvar \leq 1$, let
+\funta(\len) = \E[\tiles] = \E[i]+\E[j]-1,
+\end{equation}
+the average number of tiles can be easily obtained.
+
+\begin{theorem}
+\label{theo: funta, form}
+Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above.
+The average number of tiles visited by the segment, $\funta(\len)$, is
+\begin{equation}
+\label{eq: funta, form}
+\funta(\len) = \frac{2\len}{\pi}\left(\frac 1 a + \frac 1 b\right) + 1.
+\end{equation}
+\end{theorem}
+
+\begin{proof}
+***New proof without using the proposition, which has been moved after the theorem
+\end{proof}
+
+In view of Theorem~\ref{theo: funta, form}, the average number of visited tiles has a remarkably simple dependence on the segment length, namely an affine function.
+Conversely, for any $\tiles>1$ it is immediate to compute the length of a random segment that visits $\tiles$ tiles on average, given as $\funta^{-1}(\tiles)$.
+
+Although the random variables $i$ and $j$ are dependent, their marginal distributions can be obtained easily, as established by the next proposition. For $0 \leq \genvar \leq 1$, let
 \begin{equation}
@@
 
-% ***This will go after Theorem~\ref{theo: funta, form}
 \begin{proposition}
@@
 
-\begin{theorem}
-\label{theo: funta, form}
-Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above.
-The average number of tiles visited by the segment, $\funta(\len)$, is
-\begin{equation}
-\label{eq: funta, form}
-\funta(\len) = \frac{2\len}{\pi}\left(\frac 1 a + \frac 1 b\right) + 1.
-\end{equation}
-\end{theorem}
-
-\begin{proof}
-The expected value of $i$ can be obtained as
-\begin{equation}
-\label{eq: theo funta form 1}
-\E[i] = \sum_{n=1}^\infty \Pr[i \geq n].
-\end{equation}
-According to \eqref{eq: Pr i geq n}, this sum only has a finite number of non-zero terms: $1$ for $n=1$, plus $\lceil \len/a \rceil-1$ terms given by the second line of \eqref{eq: Pr i geq n}, plus one term given by the third line, corresponding to $n = \lceil \len/a \rceil +1 $:
-\begin{equation}
-\label{eq: theo funta form 2}
-\begin{split}
-\E[i] = 1 +
-\sum_{n=2}^{\lceil \len/a \rceil} \frac{2\len}{\pi a} \left(f\left(\frac{a(n-1)}{\len}\right)-f\left(\frac{a(n-2)}{\len} \right)\right) + \\
-\frac{2\len}{\pi a} \left(1-f\left(\frac{a(\lceil \len/a \rceil - 1)}{\len} \right)\right).
-\end{split}
-\end{equation}
-Most terms involving the function $f$ in \eqref{eq: theo funta form 2} cancel, with the result
-\begin{equation}
-\label{eq: theo funta form 3}
-\E[i] = 1+\frac{2\len}{\pi a} (1-f(0))= 1+\frac{2\len}{\pi a}.
-\end{equation}
-Similarly,
-\begin{equation}
-\label{eq: theo funta form 4}
-\E[j] = 1+\frac{2\len}{\pi b}.
-\end{equation}
-Substituting \eqref{eq: theo funta form 3} and \eqref{eq: theo funta form 4} into \eqref{eq: funta E[i] E[j]} gives \eqref{eq: funta, form}.
-\end{proof}
-
-In view of Theorem~\ref{theo: funta, form}, the average number of visited tiles has a remarkably simple dependence on the segment length, namely an affine function.
-Conversely, for any $\tiles>1$ it is immediate to compute the length of a random segment that visits $\tiles$ tiles on average, given as $\funta^{-1}(\tiles)$.
-
-The results in Proposition~\ref{prop: Pr i} and Theorem~\ref{theo: funta, form} make clear the relationship between the problem considered here and Buffon's needle problem, i.e.~the probability that a random segment of fixed length crosses a line in a regular structure of parallel strips \cite[section 1.1]{Mathai99}.
+The results in Theorem~\ref{theo: funta, form} and Proposition~\ref{prop: Pr i} and make clear the relationship between the problem considered here and Buffon's needle problem, i.e.~the probability that a random segment of fixed length crosses a line in a regular structure of parallel strips \cite[section 1.1]{Mathai99}.
 Firstly, since the number $\pi$ is involved in \eqref{eq: funta, form}, it is possible to design a simple probabilistic experiment to estimate the value of $\pi$, as in Buffon's original experiment.
@@
 \subsection{Unit square grid with real-valued lengths}
+\label{part: probmax: unit square grid, real lengths}
 

### 2021-08-28 paper.tex
Updated "to do" list

@@
 Fundamental results are presented in \S\ref{part: fund results}, which form the basis of the subsequent analysis.
-The direct and inverse problems for a deterministic segment are considered in \S\ref{part: max}, first for real-valued lengths, and then for integer-valued lengths on a unit square grid.
+The direct and inverse problems for a deterministic segment are considered in \S\ref{part: max}, first for arbitrary grids and then for a square grid.
 The analysis for the random segment is carried out in \S\ref{part: rand}.
-The average number of tiles is computed in general, and the probability that the segment visits the maximum number of tiles is obtained for a square grid.
+The average number of tiles is computed for arbitrary grids, and the probability that the segment visits the maximum number of tiles is obtained for a square grid.
 
@@
 \end{figure}%
-% ***Combine this figure and the next into one figure with two subfigures (stacked vertically, size similar to now)? This has been done in others
+% ***Maybe combine this figure and the next into one figure with two subfigures (stacked vertically, size similar to now). Not sure. It has been done in others
 % ***Are these figures necessary?
@@
 \end{proof}
+
+***Comment about how the new formula has two terms with definite meanings: The first term is the distance-squared to the closest point on the diagonal, The second term is additional distance-squared incurred from rounding (i,j) to integers
 

### 2021-08-28 paper.tex
Partial reading and rewriting/fixing

@@
 
-As an example of the first problem, consider $a=1.35$, $b=1$.
+As an example of the first question, consider $a=1.35$, $b=1$.
 A segment of unit length can be placed as shown in Figure~\ref{fig: examples} (left) to make it visit $3$ tiles.
-In fact, the solution for $\len=1$ is $3$ tiles.
+In fact, this is the maximum number for $\len=1$.
 The figure also illustrates that the solution for length $2.4$ is $5$ (center), and for $4.7$ it is $8$ (right).
@@
 Solving the problem of how many tiles the segment visits on average, as will be seen, also answers the inverse question (segment length to visit a given number of tiles on average).
-In addition, a natural question is with that probability the segment visits the maximum number of tiles.
+A natural, related question is with that probability the segment visits the maximum number of tiles.
 
@@
 For a grid with horizontal spacing $a$ and vertical spacing $b$, lines of the form $x = ka$ or $y = kb$ with $k \in \mathbb Z$ will be called \emph{grid lines}.
-A \emph{tile} is delimited by two pairs of consecutive horizontal an vertical grid lines.
+A \emph{tile} is delimited by two pairs of consecutive horizontal and vertical grid lines.
 The intersection points of horizontal and vertical grid lines will be called \emph{grid points}.
@@
 Since the segment follows a straight line, once it ``leaves'' a row of tiles in its path from the lower-left to the upper-right corner, it can never visit any more tiles from that row.
-The same observation applies for the columns.
+The same observation applies to the columns.
 
@@
 
-\ref*{prop: len ineq i j: ineq, exist} ``\eqref{eq: len > sqrt} $\Rightarrow$ there is a segment\ldots''.
-Assume that \eqref{eq: len > sqrt} holds.
+\ref*{prop: len ineq i j: ineq, exist} ``\eqref{eq: len > sqrt} $\Rightarrow$ there is a segment\ldots'': Assume that \eqref{eq: len > sqrt} holds.
 It is always possible to choose a length equal to or smaller than $\len$ such that both \eqref{eq: len > sqrt} and \eqref{eq: len leq sqrt} hold.
@@
 
-``There is a segment\ldots $\Rightarrow$ \eqref{eq: len > sqrt} ''.
-Assume that a segment exists with length not exceeding $\len$ and with a canonical rectangle with normalized dimensions $i$, $j$.
-From part~\ref{prop: len ineq i j: ineqs}, \eqref{eq: len > sqrt} holds for that segment length, and thus for $\len$.
+``There is a segment\ldots $\Rightarrow$ \eqref{eq: len > sqrt} '': Assume that a segment exists with length not exceeding $\len$ and with a canonical rectangle with normalized dimensions $i$, $j$.
+From part~\ref{prop: len ineq i j: ineqs}, inequality \eqref{eq: len > sqrt} holds for that segment length, and thus for $\len$.
 \end{proof}
@@
 A relevant question is: are there any $(i,j)$ pairs that can be disregarded irrespective of the length $\len$?
-In other words, what is the ``smallest'' subset of $\mathbb N^2$ such that the $(i,j)$ pair that maximizes the number of tiles for any length can always be found within that subset?
+In other words, what is the ``smallest'' subset of $\mathbb N^2$ such that the $(i,j)$ pair that maximizes the number of tiles for any given length can always be found within that subset?
 A subset that contains a maximizing $(i,j)$ for any length will be called a \emph{sufficient} set.
-Clearly, this set must contain at least one such pair for each possible value of $i+j-1$, so that the set can produce that value as solution (for the corresponding lengths).
-A sufficient set that contains exactly one pair $(i,j)$ for each value of $i+j-1$ will be termed a \emph{minimal sufficient} set.
+Clearly, this set must contain at least one such pair for each possible value of $i+j-1$, so that the set can produce that value as solution (for certain lengths).
+A sufficient set that contains exactly one pair $(i,j)$ for each value of $i+j-1$ will be called a \emph{minimal sufficient} set.
 
@@
 Knowing a minimal sufficient set of $(i,j)$ pairs facilitates the solution of both the direct an inverse problems, because it reduces the number of pairs that need to be tried.
-In order to derive a general method to build a minimal sufficient set, it is insightful to consider two specific cases first, as will be presented in the following.
-It is also helpful to use the equivalent formulation of the direct problem referred to in \S\ref{part: intro}, i.e.~finding the maximum number of tiles visited by segments of length $\len$ \emph{or less}.
-
-Thus each choice of $(i, j)$ implies a maximum of $i+j-1$ visited tiles on one hand; and sets a lower bound on the required length $\len$ on the other hand.
-
-Consider first the case $a=b=1$.
+In order to derive a general method to build a minimal sufficient set, it is insightful to consider two specific cases first.
+It is also convenient to use the equivalent formulation of the direct problem referred to in \S\ref{part: intro}, i.e.~finding the maximum number of tiles visited by segments of length $\len$ \emph{or less}.
+
+% Thus each choice of $(i, j)$ implies a maximum of $i+j-1$ visited tiles on one hand; and sets a lower bound on the required length $\len$ on the other hand.
+
+Consider first $a=b=1$.
 This is illustrated in Figure~\ref{fig: ijLS_1}.
@@
 For a given $\len \in \mathbb R^+$, the $(i,j)$ pairs that can be achieved with segments of length not exceeding $\len$ are, by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, those whose distance from $(2a,2b)$ is less than $\len$, i.e.~those delimited by the corresponding arc.
-It is clear from the figure that for $i+j-1$ odd, if some point $(i_1,j_1)$ with $|i_1-j_1|>1$ is achievable with the length bound $\len$, then necessarily there is a point $(i_0,i_0)$ with the same number of visited tiles $i+j-1$ (same diagonal line) that is also achievable with the same length restriction (because it has smaller distance to $(2a,2b)$).
-Similarly, for $i+j-1$ even, if a point $(i_1,j_1)$ with $|i_1-j_1|>1$ is achievable, then there is a point $(i_0,i_0+1)$ in the same diagonal that is also achievable with the same length restriction.
+It is clear from the figure that for $i+j-1$ odd, if some point $(i',j')$ with $|i'-j'|>1$ is achievable with the length bound $\len$, then necessarily there is a point $(i'',i'')$ with the same number of visited tiles $i+j-1$ (same diagonal line) that is also achievable with the same length restriction (because it has smaller distance to $(2a,2b)$).
+Similarly, for $i+j-1$ even, if a point $(i',j')$ with $|i'-j'|>1$ is achievable, then there is a point $(i'',i''-1)$ in the same diagonal that is also achievable with the same length restriction.
 
 Thus for $a=b=1$, a minimal sufficient set is formed by the pairs $(i,j)$ with $j=i$ or $j=i+1$.
-Observe that due to symmetry, any point $(i_0,i_0+1)$ could be replaced by $(i_0+1,i_0)$.
+Observe that due to symmetry, any point $(i'',i''-1)$ could be replaced by $(i''-1,i'')$.
 This illustrates that the minimal sufficient set is not unique in general.
@@
 The pairs in the minimal sufficient set are marked with filled circles in Figure~\ref{fig: ijLS_1}.
-Given a length, the maximum number of visited tiles will be achieved with one of those pairs, namely that in the diagonal line furthest from $(2a,2b)$ that is within the circle with radius $\len$ centered at that point.
+Given a length $\len$, the maximum number of visited tiles will be achieved with one of those pairs, namely that in the diagonal line furthest from $(2a,2b)$ that is within the circle with radius $\len$ centered at that point.
 
@@
 ***New proof based on treating i,j as continuous along the line i+j=constant, then quantizing to the nearest allowed point along that line.
-I think we should the notation $\isolr$, $\jsolr$ for the continuous coordinates, similarly to other proofs that use continuous coordinates.
+We should the notation $\isolr$, $\jsolr$ for the continuous coordinates, in conformity to other proofs that use continuous coordinates.
 % ***The continuous argument gives something like $i-2 = b^2/(a^2+b^2)*(t-3)$, $j = a^2/(a^2+b^2)*(t-3)$.
@@
 Given $(i_\tiles,j_\tiles) \in \mss$, the next pair $(i_{\tiles+1},j_{\tiles+1})$ is obtained by incrementing $j$ if that results in a point below \eqref{eq: upper bound, line}.
-Else $j$ cannot be increased and $i$ is incremented instead, and the new pair is guaranteed to be above or on \eqref{eq: lower bound, line}.
+Else $i$ is incremented instead, and the new pair is guaranteed to be above or on \eqref{eq: lower bound, line}.
 
@@
 
-For $a^2/b^2$ arbitrary, the number of pairs with the same $i$ coordinate in the set $\mss$ is in general irregular, because the line \eqref{eq: lower bound, line} does not follow a ``natural''
+For $a^2/b^2$ arbitrary, the number of pairs with the same $i$ coordinate, or with the same $j$, in the set $\mss$ is in general irregular, because the lines \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} do not follow a ``natural''
 % ***it may be better to say ``simple'' instead of ``natural'', specially if we rename ``canonical`'' to ``natural'' or ``natural bounding''
@@
 However, this is not easily discernible unless $a^2/b^2$ is a ratio of small numbers.}
-On the other hand, a simple pattern arises for $a^2/b^2 \in \mathbb N$, as seen in Figures~\ref{fig: pairs_bounds_1} and \ref{fig: pairs_bounds_sqrt2}.
+On the other hand, a simple pattern arises when $a^2/b^2$ or $b^2/a^2$ is a natural number, as seen in Figures~\ref{fig: pairs_bounds_1} and \ref{fig: pairs_bounds_sqrt2}.
 
@@
 This formalizes the intuition that to maximize the number of visited tiles, the segment should follow a direction along which the perceived ``length'' of the tile is smaller.
+% but not ``smallest'', if that were the case it would always be either vertical or horizontal
 
@@
 The particular case of square grids is addressed in \S\ref{part: max: unit square grid, real lengths}, as it allows a specialized formula for the direct problem.
-Lastly, the analysis of a unit square grid with integer-valued lengths is given in \S\ref{part: max: unit square grid, integer lengths}.
+Lastly, the analysis of a unit square grid with integer-valued lengths is presented in \S\ref{part: max: unit square grid, integer lengths}.
 
@@
 For $\funl$ the analogous method gives a direct formula.
-In addition, it is possible to obtain a direct formula also for $\funt$ using an approach similar to that used in Proposition~\ref{prop: min suff set, form}.
-% **this approach will change. Remove the above comment about the approach begin similar, and briefly describe the approach at the beginning of the proof here. Well, that's more or less done already**
+In addition, it is possible to obtain a direct formula also for $\funt$ using a different approach.
 These formulas are given in Theorems \ref{theo: funt, form} and \ref{theo: funl, form}.
@@
 \begin{proof}
-The general idea is similar to that used in Proposition~\ref{prop: min suff set, form}**this may have changed**: find the intersection $(\isolr, \jsolr)$ of the line \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ for $i, j \geq 2$, and from that obtain the actual pair $(\isoli, \jsoli)$ of integer numbers which yields the solution.
+The idea of the proof is loosely similar to that used in Proposition~\ref{prop: min suff set, form}:
+% ***is it appropriate to say that it is loosely similar? Both are based on a continuous approximation which is then quantized. Does ``loosely'' have a negative connotation?
+find the intersection $(\isolr, \jsolr)$ of the line \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ for $i, j \geq 2$, and from that obtain the actual pair $(\isoli, \jsoli)$ of integer numbers that yields the solution.
 
@@
 
-Although Theorem~\ref{theo: funt, form} is valid for $a \geq b$, the result could also be applied for $a < b$ by simply swapping the values of $a$ and $b$.
+Although Theorem~\ref{theo: funt, form} is valid for $a \geq b$, the result could be applied for $a < b$ by simply swapping the values of $a$ and $b$.
 In other words, \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j}
@@
 \begin{equation}
-\text{***new formula by Alex; unless we can unify with the above}
+\text{***new formula found by Alex; unless we can unify with the above}
 \end{equation}
@@
 
-***Comment about how the new formula has two terms with definite meanings: The first term is the distance-squared to the closest point on the diagonal, The second term is additional distance-squared incurred from rounding (i,j) to integers
+***Comment on how the new formula has two terms with definite meanings: The first term is the distance-squared to the closest point on the diagonal, The second term is additional distance-squared incurred from rounding (i,j) to integers
 
@@
 The graphs illustrate some of the observations of the previous paragraphs.
-Indeed, the asymptotic slope in Figure~\ref{fig: funt examples}, or the inverse of the asymptotic slope in Figure~\ref{fig: funl examples}, is approximately $\sqrt{2}$ for $a, b=1$, and $1/b$ for the case $a=5,b=1$, and even for $a=5,b=1.5$ and $a=10,b=3$.
-Comparing the latter two cases it is also seen that scaling $a$, $b$ and $\len$ by the same factor does not alter $\funt(\len)$; or scaling $a$, $b$ results in $\funl(\tiles)$ being scaled by the same factor.
+Indeed, the asymptotic slope in Figure~\ref{fig: funt examples}, or the inverse of the asymptotic slope in Figure~\ref{fig: funl examples}, is approximately $\sqrt{2}$ for $a, b=1$; and it is $1/b$ for the case $a=5,b=1$, or even for $a=5,b=1.5$ or $a=10,b=3$.
+Comparing the latter two cases it is also seen that scaling $a$, $b$ and $\len$ by the same factor does not alter $\funt(\len)$, and results in $\funl(\tiles)$ being scaled by that factor.
 
@@
 
-A square grid has $a=b$. If the segment lengths are real-valued it can be further assumed that $a=1$. For $a \neq 1$ the results to be obtained apply with $\len$ replaced by $\len/a$.
+A square grid has $a=b$. For real-values segment lengths it can be further assumed that $a=1$. For $a \neq 1$ the results to be obtained apply with $\len$ replaced by $\len/a$.
 
@@
 \end{equation}
-
 \end{corollary}
@@
 \begin{proof}
-The proof of \eqref{eq: theo: funt, sq, form; funt, with i, j} uses a variation of the minimal sufficient set $\mss$ which is more suited to this situation.
-
-For $a=b=1$ the set $\mss$ consists of points of the form $(i,i)$ and $(i+1,i)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}.
-By symmetry, replacing each point $(i+1,i)$ by $(i,i+1)$ gives a set $\mss'$ which is also minimal sufficient.
+The proof of \eqref{eq: theo: funt, sq, form; funt, with i, j} uses a variation of the minimal sufficient set $\mss$ defined in Proposition~\ref{prop: min suff set, form} that is more suited to this situation.
+% ***``situation'' or ``case''?
+
+For $a=b=1$ the set $\mss$ consists of points of the form $(i,i)$ and $(i,i-1)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}.
+By symmetry, replacing each point $(i,i-1)$ by $(i-1,i)$ gives a set $\mss'$ which is also minimal sufficient.
 For this new set, the lower bounding line \eqref{eq: lower bound, line} can be replaced by the simpler $j=i$.
@@
 
+% ***Aq
 \subsection{Unit square grid with integer lengths}

### 2021-08-28 paper.tex
Done with reading and rewriting/fixing

@@
 
-% ***Aq
 \subsection{Unit square grid with integer lengths}
@@
 
-A natural variation of the first two problems introduced in \S\ref{part: intro} is to consider $a=b=1$ with the additional restriction that the segment length can only be a positive integer (equivalently, a square grid is considered with step $a$ and the segment length can only be an integer multiple of $a$).
+A natural variation of the first two problems introduced in \S\ref{part: intro} is to consider $a=b=1$ with the additional restriction that the segment length can only be a positive integer (equivalently, the square grid has spacing $a$ and the segment length can only be an integer multiple of $a$).
 
@@
 This will be denoted as a function $\funti: \mathbb N \to \mathbb N$ for greater clarity, although obviously $\funti(\leni) = \funt(\leni)$ for all $\leni \in \mathbb N$.
-The sequence $\funti(\leni)$, $\leni \in \mathbb N$ takes values $3, 5, 7, 8, 9, \ldots$, and is depicted in Figure~\ref{fig: funti}.
+The sequence $\funti(\leni)$, $\leni \in \mathbb N$, which takes values $3, 5, 7, 8, 9, \ldots$, is depicted in Figure~\ref{fig: funti}.
 This is A346232 in the On-Line Encyclopedia of Integer Sequences \cite{OEIS_unitsq_int_dir}.
@@
 \begin{proof}
-The equality \eqref{eq: funti leni} stems from \eqref{eq: theo: funt, sq, form; funt} noting that $\len$ is an integer.
+The equality \eqref{eq: funti leni} stems from \eqref{eq: theo: funt, sq, form; funt, simpler} noting that $\len$ is an integer.
 
@@
 This is A346693 in the On-Line Encyclopedia of Integer Sequences \cite{OEIS_unitsq_int_inv}.
-However, a simpler expression can be obtained from \eqref{eq: funti leni} and \eqref{eq: funli funti}.
+However, a slightly simpler expression can be obtained from \eqref{eq: funti leni} and \eqref{eq: funli funti}.
 This is established by the next theorem, which also states some properties of $\funli$, parallel to those of $\funti$.
@@
 The variables $x_1$, $y_1$ and $\theta$ determine the coordinates $x_2$, $y_2$ of the second endpoint.
-Note that events of probability $0$, such as $x_1=a$ or the segment passing through a grid point, can be disregarded for computing the average number of tiles visited by the segment.
 
@@
 Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles visited by a random segment of length $\len$, with the distributions specified in \S\ref{part: rand}.
-Using the fact that
+This function can be easily computed using the fact that
 \begin{equation}
@@
 \end{equation}
-the average number of tiles can be easily obtained.
+and noting that events of probability $0$, such as $x_1=a$ or the segment passing through a grid point, can be disregarded.
 
@@
 \begin{proof}
-***New proof without using the proposition, which has been moved after the theorem
+***New proof without using the proposition (which has been moved after the theorem)
 \end{proof}
@@
 
-Although the random variables $i$ and $j$ are dependent, their marginal distributions can be obtained easily, as established by the next proposition. For $0 \leq \genvar \leq 1$, let
+In spite of the dependence between the random variables $i$ and $j$, their marginal distributions have relatively simple analytic expressions, as established by the next proposition.
+
+For $0 \leq \genvar \leq 1$, let
 \begin{equation}
@@
 
-The analysis above can be analogously applied to $\Pr[j \geq n]$ if the $x$ and $y$ axes are interchanged. Thus the result is the same with $a$ replaced by $b$.
+The above arguments can be applied to $\Pr[j \geq n]$ if the $x$ and $y$ axes are interchanged. Thus the result is the same with $a$ replaced by $b$.
 \end{proof}
 
-The results in Theorem~\ref{theo: funta, form} and Proposition~\ref{prop: Pr i} and make clear the relationship between the problem considered here and Buffon's needle problem, i.e.~the probability that a random segment of fixed length crosses a line in a regular structure of parallel strips \cite[section 1.1]{Mathai99}.
+The results in Theorem~\ref{theo: funta, form} and Proposition~\ref{prop: Pr i} make clear the relationship between the problem considered in this section here and Buffon's needle problem, i.e.~the probability that a random segment of fixed length crosses a line in a regular structure of parallel strips \cite[section 1.1]{Mathai99}.
 Firstly, since the number $\pi$ is involved in \eqref{eq: funta, form}, it is possible to design a simple probabilistic experiment to estimate the value of $\pi$, as in Buffon's original experiment.
@@
 
-The probability of visiting the maximum number of tiles for a rectangular grid is difficult to compute, due to the irregularity of the minimal sufficient set (discussed in \S\ref{part: fund results}).
-In the square case, however, the simplicity of this set makes the problem tractable.
+The probability of visiting the maximum number of tiles for a rectangular grid is difficult to compute, due to the irregularity of the minimal sufficient set (see \S\ref{part: fund results}).
+In the square case, however, the simplicity of the set makes the problem more tractable.
 
@@
 \end{equation}
-Using the identifies $\cos^2 \alpha = 1-\sin^2 \alpha$ and $\cos(2\alpha) = 2\cos^2\alpha-1 = 1 - 2\sin^2\alpha$, the integral in \eqref{eq: g def} is computed as
+Using the identities $\cos^2 \alpha = 1-\sin^2 \alpha$ and $\cos(2\alpha) = 2\cos^2\alpha-1 = 1 - 2\sin^2\alpha$, the integral in \eqref{eq: g def} is computed as
 \begin{equation}
@@
 
-Consider a length $\len$ with corresponding maximum number of tiles $\tiles = \funt(\len)$ (given by Corollary~\ref{cor: funt, sq, form}).
+Consider a length $\len$ with corresponding maximum number of tiles $\tiles = \funt(\len)$.
 The first endpoint of the segment is assumed to be in the reference tile with lower-left corner $(0,0)$.
 The segment visits $\tiles$ tiles almost surely if and only if its second endpoint is in a tile with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, $i+j-1=\tiles$, or in a tile symmetrical to this in the other three quadrants of the plane.
-Figure~\ref{fig: probmax_possible_tiles} shows the reference tile, with solid line, and some of the square tiles that fulfil this condition, with dashed line, for $\tiles=5$.
+Figure~\ref{fig: probmax_possible_tiles} shows the reference tile, with solid line, and some of the tiles that fulfil this condition, with dashed line, for $\tiles=5$.
 The desired probability can be computed by considering only the tiles in the first quadrant and multiplying by $4$.
@@
 This range corresponds to the part of the arc with solid line in Figure~\ref{fig: probmax_funt_odd}.
-For convenience, the angle $\theta$ is defined clockwise from the downward direction, as seen in the figure.
+For convenience, the angle $\theta$ is defined clockwise from the downward direction.
 For each $\theta$ in the allowed range, the valid positions for the first endpoint of the (shifted) segment are in the shaded rectangle, whose dimensions depend on $\theta$.
@@
 
-The situation is different depending on whether $\tiles$ is odd or even.
+There are two situations depending on whether $\tiles$ is odd or even.
 Consider $\tiles$ odd first.
@@
 
-Therefore, two cases need to be distinguished for $\tiles$ odd, depending on whether \eqref{eq: probmax len odd2} is satisfied or not; and similarly two cases exist for $\tiles$ even, depending on whether \eqref{eq: probmax len even2} holds or not.
+Thus, two cases need to be distinguished for $\tiles$ odd, depending on whether \eqref{eq: probmax len odd2} is satisfied or not; and similarly two cases exist for $\tiles$ even, depending on whether \eqref{eq: probmax len even2} holds or not.
 These are, respectively, the four cases in \eqref{eq: probmax}.
@@
 \end{equation}
-The range of allowed values for $\theta$, as can be seen from Figure~\ref{fig: probmax_funt_odd_diag}, is $(\theta_0, \theta_1)$ with
+The range of allowed values for $\theta$, as can be seen from the figure, is $(\theta_0, \theta_1)$ with
 \begin{align}
@@
 
-It should be noted that the technique used in the proof of Theorem~\ref{theo: probmax, sq} could in principle be used for obtaining the probability that the number of tiles visited by the segment equals or exceeds any given value $t < \funt(\len)$.
-The resulting expressions, however, are cumbersome, because more cases (more subintervals for the length) need to be distinguished depending on how many tiles with $i+j-1=t$ can be reached from the reference tile.
+The technique used in the proof of Theorem~\ref{theo: probmax, sq} could in principle be used for obtaining the probability that the number of tiles visited by the segment equals or exceeds any given value $t < \funt(\len)$.
+The resulting expressions, however, are cumbersome, because more cases (more subintervals for the length) need to be distinguished depending on how many tiles with lower-left corner $(i-1,j-1)$ such that $i+j-1=t$ can be reached from the reference tile.
 
@@
 \end{equation}
-For example, the first continuous section seen in the figure corresponds to a maximum number of tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, with $\len_3= 1$.
+For example, the first continuous section seen in Figure~\ref{fig: probmax_len} corresponds to a maximum number of tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, with $\len_3= 1$.
 The second corresponds to $\tiles=4$, for lengths in $(\len_3, \len_4]$, with $\len_4= \sqrt{2}$.
 Within each continuous section the probability monotonically increases from $0$ to a maximum value.
-The heights of the maxima follow a different trend**better word?** for odd and even $\tiles$, as seen in the figure; but in each case they are asymptotically proportional to $1/\len_\tiles$, as the next result establishes.
+The heights of the maxima follow a different trend**better word?** for odd and even $\tiles$, as can be seen in the figure; but in each case they are asymptotically proportional to $1/\len_\tiles$.
+This is established by the next result.
 
@@
 For large $\tiles$, $\len_\tiles$ asymptotically approaches $\tiles/\sqrt{2}$.
-Using this into \eqref{eq: probmax lim even}, which is greater than \eqref{eq: probmax lim odd}, establishes \eqref{eq: probmax limsup}.
+Using this into \eqref{eq: probmax lim even}, which is greater than \eqref{eq: probmax lim odd}, yields \eqref{eq: probmax limsup}.
 \end{proof}

### 2021-08-28 figures/figures.m

@@ -1615,9 +1615,9 @@ xlabel y
 axis([-4 5 -4 5])
 for k = [0 1+3j 2+2j 3+1j 1-3j 2-2j 3-1j -3+1j -2+2j -1+3j -3-1j -2-2j -1-3j]
     if k==0
-        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k-', 'linewidth', .75)
+        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k-', 'linewidth', 1)
     else
-        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k--', 'linewidth', .75)
+        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k--', 'linewidth', 1)
     end
 end
 
@@ -1658,9 +1658,9 @@ patch([real(endpoint) 1 1 real(endpoint) real(endpoint)], [imag(endpoint) imag(e
 plot([3+3j endpoint], 'k')
 for k = [0 3+3j]
     if k==0
-        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k-', 'linewidth', .75)
+        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k-', 'linewidth', 1)
     else
-        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k--', 'linewidth', .75)
+        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k--', 'linewidth', 1)
     end
 end
 text(1.65, 2.27, char(hex2dec('2113')))
@@ -1706,9 +1706,9 @@ patch([real(endpoint) 1 1 real(endpoint) real(endpoint)], [imag(endpoint) imag(e
 plot([2+4j endpoint], 'k')
 for k = [0 2+4j]
     if k==0
-        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k-', 'linewidth', .75)
+        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k-', 'linewidth', 1)
     else
-        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k--', 'linewidth', .75)
[…]

### 2021-09-02 paper.tex
Added note
Alex proofs of maximum and expected tiles touched
Added proofs the tiles-touched formula (Theorem 3) and expected tiles visited via Buffon's Needle argument (Theorem 6).

@@
 
+% *** AA: Maybe add a transition that now we're switching to looking at the average number of tiles visited by a random segment.
 The notion of a \emph{random} segment of a given length is precisely defined as follows.
@@
 
+% ***Is the text from here to the proposition too long? Shorten it?
+
+% ***AA: Do we actually need the concept of a minimal sufficient set, given that there's a single value at each diagonal (tiebreaking a certain way)? Perhaps rename to suggest it's one value?
+
 Knowing a minimal sufficient set of $(i,j)$ pairs facilitates the solution of both the direct an inverse problems, because it reduces the number of pairs that need to be tried.
@@
 \text{***new formula found by Alex; unless we can unify with the above}
+% *** AA: Wrote a shot at how a unification might look in e-mail.
 \end{equation}
@@
 
-For $a=b=1$ the set $\mss$ consists of points of the form $(i,i)$ and $(i,i-1)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}.
+For $a=b=1$, the set $\mss$ consists of points of the form $(i,i)$ and $(i,i-1)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}.
 By symmetry, replacing each point $(i,i-1)$ by $(i-1,i)$ gives a set $\mss'$ which is also minimal sufficient.
@@
 And \eqref{eq: theo: funt, sq, form; funt, with i, j} is needed because of $\isoli$, $\jsoli$, which are used in (the text below in preparation for) \S\ref{part: probmax: unit square grid, real lengths}.
+
+To show \eqref{eq: theo: funt, sq, form; funt, simpler}, we refer back to \eqref{eq: theo: funt, sq, form; funl} in Corollary~\ref{theo: funl, sq, form} for $\tiles \geq 3$:
+\begin{equation}
+\funl(\tiles) = \sqrt{\left\lceil \frac{(\tiles-3)^2} {2} \right\rceil}
+\end{equation}
+Then, the pseudo-inverse function $\funt(\len)$ is given by the largest positive integer $t$ such that 
+\begin{equation}
+\left\lceil \frac{(\tiles-3)^2} {2} \right\rceil < \len ^ 2
+\end{equation}
+Since the left-hand side is an integer, the condition of being strictly less than $\len^2$ is equivalent to: 
+\begin{equation}
+\left\lceil \frac{(\tiles-3)^2} {2} \right\rceil  \leq \left\lceil \len^2 \right\rceil  - 1 
+\end{equation}
+and removing the now-superfluous ceiling on the left side gives:
+\begin{equation}
+\frac{(\tiles-3)^2} {2}\leq  \left\lceil \len^2 \right\rceil  - 1 
+\end{equation}
+Finally, we solve for $\tiles$:
+\begin{equation}
+\tiles \leq \sqrt{2 \left\lceil \len^2 \right\rceil -2 } + 3
+\end{equation}
+The desired quantity $\funt(\len)$, that is the largest positive integer $\tiles$ satisfying the above, is thus the floor of the right-hand side as given by \eqref{eq: theo: funt, sq, form; funt, simpler}.
 \end{proof}
@@
 
-Given $\len \in \mathbb R^+$, consider a segment of length $\len$ with random position and orientation.
+Given $\len \in \mathbb R^+$, consider a segment of length $\len$ with uniformly random position and orientation.
 Specifically, the coordinates $x_1$, $y_1$ of the first endpoint are independent random variables uniformly distributed on $[0,a)$ and $[0,b)$ respectively.
@@
 \begin{proof}
-***New proof without using the proposition (which has been moved after the theorem)
+% *** Is this writeup too informal/non-rigorous?
+The classic Buffon's Needle result (*** cite Buffon's 1777 paper?) says that if we take infinite ruled paper with parallel lines a distance of $1$ apart, and we toss a needle of length $1$ with uniformly random position and orientation, the probability that it crosses any line is $2/\pi$.
+
+This is also the expected (that is, average) number of crossings, since this needle cannot reach more than one line\footnote{We may ignore the probability-zero event of it lying exactly along a line.}.
+For a needle of length $2$, the expected number of crossings is double the amount of $2/\pi$, since expectation is additive and so unchanged if we instead break this needle into two length-$1$ halves
+that we toss separately and count their total crossings. Conversely, for a needle of length $1/2$, it would be half the original expectation.
+
+A similar argument extends this proportionality to all rational lengths, and by monotonicity all real lengths: for a needle of length $l$, the expected number of crossings is
+$\frac{2\len}{\pi}$.
+If the lines are a distance $a$ apart, this is equivalent to shrinking the needle by a factor of $a$, giving expected crossings
+$\frac{2\len}{\pi a}$.
+
+Now, consider a rectangular grid paper with parameters of width $a$ and height $b$.
+Equivalently, this is paper ruled with horizontal lines $b$ apart and vertical lines $a$ apart.
+The distribution of the number of crossings of a needle tossed onto this grid is known as the Buffon-Laplace Needle Problem (*** cite).
+This number decomposes into the counts of crossings of horizontal and vertical grid lines, so the expected number of crossings is the sum of those for lined paper of width $a$ and $b$ respectively:
+\begin{equation}
+\frac{2\len}{\pi}\left(\frac \len a + \frac \len b\right)
+\end{equation}
+As noted in Proposition \ref{prop: i+j-1}, the number of grid cells visited by a segment is the count of its grid line crossings plus one -- unless, that is, it exactly crosses any grid points, but that occurs with probability zero and so doesn't affect the expectation.
+So, $\funta(\len)$ is the above quantity plus one.
+
 \end{proof}

### 2021-09-02 paper.tex
Minor wording edits

--- 2021-08-28 paper.tex
@@
 
-The notion of a \emph{random} segment of a given length is precisely defined as follows.
+To address the remaining two questions, the notion of a \emph{random} segment of a given length needs to be precisely defined.
+This is done as follows.
 By symmetry, one endpoint of the segment can be assumed to lie in a fixed, reference tile.
@@
 \label{eq: min suff set, form, j}
-j_\tiles &= \left\lceil \frac {2\tiles a^2 + 3(b^2-a^2)}{2(a^2+b^2)} \right\rceil,
+j_\tiles &= \left\lceil \frac {(2\tiles-3) a^2 + 3b^2}{2(a^2+b^2)} \right\rceil,
+% ***change expressions to the following to match the statement of Theorem~\ref{theo: funl, form}:
+%i_\tiles &= \left\lfloor \frac{b^2 (\tiles-3)}{a^2+b^2} + \frac 5 2 \right\rfloor \\
+%j_\tiles &= \left\lceil  \frac{a^2 (\tiles-3)}{a^2+b^2} + \frac 3 2 \right\rceil
 \end{align}
@@
 
-***Alex
-***Derivation of \eqref{eq: theo: funt, sq, form; funt, simpler} \ldots
-***It looks like we may have to use different techniques for deriving \eqref{eq: theo: funt, sq, form; funt, with i, j} and \eqref{eq: theo: funt, sq, form; funt, simpler}.
-I don't see a way to use a common proof method.
-And \eqref{eq: theo: funt, sq, form; funt, with i, j} is needed because of $\isoli$, $\jsoli$, which are used in (the text below in preparation for) \S\ref{part: probmax: unit square grid, real lengths}.
+To show \eqref{eq: theo: funt, sq, form; funt, simpler}, it is first noted that for $\tiles \geq 3$ Corollary~\ref{theo: funl, sq, form} gives
+\begin{equation}
+\funl(\tiles) = \sqrt{\left\lceil \frac{(\tiles-3)^2} {2} \right\rceil}.
+\end{equation}
+According to \eqref{eq: funt funl}, the function $\funt(\len)$ is given by the largest positive integer $t$ such that 
+\begin{equation}
+\label{eq: theo: funt, sq, form; proof 2}
+\left\lceil \frac{(\tiles-3)^2} {2} \right\rceil < \len ^ 2.
+\end{equation}
+Since the left-hand side of \eqref{eq: theo: funt, sq, form; proof 2} is an integer, the condition of being strictly less than $\len^2$ is equivalent to
+\begin{equation}
+\left\lceil \frac{(\tiles-3)^2} {2} \right\rceil  \leq \left\lceil \len^2 \right\rceil  - 1,
+\end{equation}
+which in turn is the same as
+\begin{equation}
+\frac{(\tiles-3)^2} {2}\leq  \left\lceil \len^2 \right\rceil  - 1. 
+\end{equation}
+Finally, solving for $\tiles$ gives
+\begin{equation}
+\label{eq: theo: funt, sq, form; proof 3}
+\tiles \leq \sqrt{2 \left\lceil \len^2 \right\rceil -2 } + 3.
+\end{equation}
+The desired quantity $\funt(\len)$, that is the largest positive integer $\tiles$ satisfying \eqref{eq: theo: funt, sq, form; proof 3}, is thus the right-hand side rounded down, as given by \eqref{eq: theo: funt, sq, form; funt, simpler}.
 \end{proof}
@@
 \begin{proof}
-***New proof without using the proposition (which has been moved after the theorem)
+% *** Is this writeup too informal/non-rigorous?
+% ***LM: I would make the rational-length extension more explicit, and invoke a suitable theorem (with the corresponding book reference) for the real-length case.
+Or better yet: the first reference linked as a comment in the introduction (https://www.jstor.org/stable/2317945?seq=1#metadata_info_tab_contents) already gives the $2\len/(\pi a)$ formula.
+We can directly start from that formula, referencing that paper.
+The classic Buffon's Needle result (**AA cite Buffon's 1777 paper?** **LM I think maybe something more modern such as the references that are linked in the Introduction section**) states that if a needle of length $1$ is tossed with uniformly random position and orientation on
+infinite ruled paper with parallel lines a distance of $1$ apart the probability that it crosses any line is $2/\pi$.
+
+This is also the expected (that is, average) number of crossings, since the needle cannot reach more than one line, ignoring the event of it lying exactly along a line, which has probability $0$.
+For a needle of length $2$, the expected number of crossings is double the amount of $2/\pi$, since expectation is additive and thus unchanged if the needle is broken into two length-$1$ halves that are tossed separately, and their total crossings are counted.
+Conversely, for a needle of length $1/2$ the expectation would be half the original one.
+
+A similar argument extends this proportionality to all rational lengths, and by monotonicity all real lengths.
+Thus for a needle of length $\len$, the expected number of crossings is $2\len/\pi$.
+If the lines are a distance $a$ apart, this is equivalent to shrinking the needle by a factor of $a$, giving $2\len/(\pi a)$ as expected number of crossings.
+
+Consider a rectangular grid with horizontal spacing $a$ and vertical spacing $b$.
+Equivalently, this is paper ruled with horizontal lines $b$ apart and vertical lines $a$ apart.
+The distribution of the number of crossings of a needle tossed onto this grid is known as the Buffon-Laplace Needle Problem (*** cite).
+This number decomposes into the counts of crossings of horizontal and vertical grid lines, and thus the expected number of crossings is the sum of those for lined paper of width $a$ and $b$ respectively:
+\[
+\frac{2\len}{\pi}\left(\frac 1 a + \frac 1 b\right)
+\]
+As noted in Proposition \ref{prop: i+j-1}, the number of tiles visited by a segment is the count of its grid line crossings plus $1$, unless it exactly crosses any grid points, but again that occurs with probability $0$ and therefore it does not affect the expectation.
+Thus, $\funta(\len)$ is the above quantity plus $1$.
 \end{proof}
 
-In view of Theorem~\ref{theo: funta, form}, the average number of visited tiles has a remarkably simple dependence on the segment length, namely an affine function.
+In view of Theorem~\ref{theo: funta, form}, the average number of visited tiles as a function of the segment length has a very simple form, namely an affine function.
 Conversely, for any $\tiles>1$ it is immediate to compute the length of a random segment that visits $\tiles$ tiles on average, given as $\funta^{-1}(\tiles)$.

### 2021-09-02 paper.tex
Typos
Typo
New figure for proof of proposition on the minimal sufficient set

@@
 \includegraphics[width=.85\textwidth]{examples_1p35}%
-\caption{Examples for $a=1.35, b=1$; $\len=1$, $\len=2.4$ and $\len=4.7$
+\caption{Examples for $a=1.35$, $b=1$; $\len=1$, $\len=2.4$ and $\len=4.7$
 }%
@@
 ***New proof based on treating i,j as continuous along the line i+j=constant, then quantizing to the nearest allowed point along that line.
-We should the notation $\isolr$, $\jsolr$ for the continuous coordinates, in conformity to other proofs that use continuous coordinates.
-% ***The continuous argument gives something like $i-2 = b^2/(a^2+b^2)*(t-3)$, $j = a^2/(a^2+b^2)*(t-3)$.
+We should use the notation $\isolr$, $\jsolr$ for the continuous coordinates, in conformity to other proofs that use continuous coordinates.
+% ***The continuous argument gives something like $i-2 = b^2/(a^2+b^2)*(t-3)$, $j-2 = a^2/(a^2+b^2)*(t-3)$.
 We then compute how much the distance from $(i,j)$ to $(2,2)$ is increased by quantizing along the i+j line, to argue that the nearest quantized point should be chosen.
-To ensue that in case of a tie we choose the point with larger $i$, we apply floor(...+1/2) to i and ceiling(...-1/2) to j.
+To ensure that in case of a tie we choose the point with larger $i$, we apply floor(...+1/2) to i and ceiling(...-1/2) to j.
+% ***Reference the figure in the text.
+The figure is an example with $a=1.35$, $b=1$, $\tiles=7$.
+It contains the line j-2 = a^2/b^2*(i-2), whose intersection with i+j-1=t (which is perpendicular to it) gives the continuous solution $(\isolr, \jsolr)$ (marked with a square).
 
@@
 
-%\begin{figure}%
-%\centering%
-%\includegraphics[width=.7\textwidth]{ijLS_1p35_detail_inverse}%
-%\caption{***Figure illustrating the proof of Proposition~\ref{prop: min suff set, form}. Example with $a=1.35, b=1$%
-%}%
-%\label{fig: ***}%
-%\end{figure}%
+\begin{figure}%
+\centering%
+\includegraphics[width=.7\textwidth]{mss}%
+\caption{Obtaining $(\isolr, \jsolr)$ and $(i_\tiles, j_\tiles)$ in  Proposition~\ref{prop: min suff set, form}. Example with $a=1.35$, $b=1$, $\tiles=7$%
+}%
+\label{fig: min suff set, form}%
+\end{figure}%
 \end{proof}
@@
 
-The two possibilities are illustrated in Figure~\ref{fig: ijLS_1p35_detail} for $a=1.35, b=1$.
+The two possibilities are illustrated in Figure~\ref{fig: ijLS_1p35_detail} for $a=1.35$, $b=1$.
 In each case, the arc displayed in the graph is centered at $(2a,2b)$ and has radius $\len$.
@@
 }%
-\caption{Obtaining $(\isoli,\jsoli)$ in Theorem~\ref{theo: funt, form}. Examples with $a=1.35, b=1$%
+\caption{Obtaining $(\isoli,\jsoli)$ in Theorem~\ref{theo: funt, form}. Examples with $a=1.35$, $b=1$%
 }%

### 2021-09-02 figures/figures.m

@@ -1837,3 +1837,47 @@ funt_sq_2 = ii+jj-1;
 funt_sq_3 = floor(sqrt(2*ceil(L.^2/a^2) - 2)) + 3; % formula found by Alex
 isequal(funt_rect, funt_sq, funt_sq_2, funt_sq_3)
 
+
+%% #70. Figure for proof of {the proposition that gives the formula of the minimum sufficient set} based on a continuous solution that is then rounded (Alex)
+% Based on #43 (with flag_hexagram = false; flag_fun = false; flag_pairs = true)
+
+clear, clf, clc
+flag_pairs = true;
+a = 1.35; b = 1;
+hold on
+set(gca, 'layer', 'top')
+grid on
+plot([2*a+2j*b 2*a+5.7j*b], 'k') % "secondary" axes
+plot([2*a+2j*b 5.7*a+2j*b], 'k')
+chosen = [2*a+2j*b 2*a+3j*b 3*a+3j*b 3*a+4j*b 3*a+5j*b 4*a+5j*b];
+chosen_2 = 4*a+4j*b;
+ms = 5;
+plot(chosen, 'ko', 'markerfacecolor', 'k', 'markersize', ms) % chosen (i,j) pairs
+plot(chosen_2, 'ko', 'markerfacecolor', 'w', 'markersize', ms) % chosen 2
+plot(setdiff((2:5)*a + 1j*(2:5).'*b, [chosen chosen_2]), 'ko', 'markerfacecolor', 'none', 'markersize', ms) % non-chosen (i,j) pairs
+axis equal %pbaspect([1 1 1])
+axis([1*a 6*a 1*b 6*b])
+plot([1.8 2.4]*a, [2.2 1.6]*b, 'k--')
+plot([1.8 3.4]*a, [3.2 1.6]*b, 'k--')
+plot([1.8 4.4]*a, [4.2 1.6]*b, 'k--')
+plot([1.8 5.4]*a, [5.2 1.6]*b, 'k--')
+plot([2.5 5.4]*a, [5.5 2.6]*b, 'k--') % limit to secondary axes
+plot([3.5 5.4]*a, [5.5 3.6]*b, 'k--')
+plot([4.5 5.4]*a, [5.5 4.6]*b, 'k--')
+xl = xlim; yl = ylim;
+set(gca, 'GridAlpha', .4, 'XTick', xl(1):xl(2), 'YTick', yl(1):yl(2))
[…]

### 2021-09-07 paper.tex
Wording

@@
 The function $\funt$ is piecewise constant and left-continuous, with unit-height jumps.
-A jump occurs at $\len$ if and only if $\len = \funl(\tiles)$ for some $\tiles \in \mathbb N, \tiles \geq 4$; and then $\funt(\len) = \tiles-1$, $\lim_{h \rightarrow 0+} \funt(\len+h) = \tiles$.
+A jump occurs at $\len$ if and only if $\len = \funl(\tiles)$ for some $\tiles \in \mathbb N$, $\tiles \geq 4$; and then $\funt(\len) = \tiles-1$, $\lim_{h \rightarrow 0+} \funt(\len+h) = \tiles$.
 \end{theorem}
@@
 \begin{proof}
-The idea of the proof is loosely similar to that used in Proposition~\ref{prop: min suff set, form}:
-% ***is it appropriate to say that it is loosely similar? Both are based on a continuous approximation which is then quantized. Does ``loosely'' have a negative connotation?
-find the intersection $(\isolr, \jsolr)$ of the line \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ for $i, j \geq 2$, and from that obtain the actual pair $(\isoli, \jsoli)$ of integer numbers that yields the solution.
-
-Specifically, assuming first that \eqref{eq: lower bound, line} holds, the largest $i$ that can be achieved with lengths not exceeding $\len$ is obtained, as given by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}.
+The approach is similar to that used in Proposition~\ref{prop: min suff set, form}.
+Namely, for $i, j \geq 2$ the intersection $(\isolr, \jsolr)$ of the line \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ is found, and from that the actual integer solution $(\isoli, \jsoli)$ is computed.
+
+More specifically, assuming that \eqref{eq: lower bound, line} holds, the largest $i$ that can be achieved with lengths not exceeding $\len$ is obtained, as given by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}.
 This value, $\isoli$, is computed by rounding down $\isolr$.
@@
 
-As for the properties of $\funt$, it follows from \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j} that it is piecewise constant and left-continuous.
-From the procedure described in the previous paragraphs for obtaining $(\isoli, \jsoli)$ it is clear that $\isoli+\jsoli-1$ increases in steps of $1$ when $\len$ is increased continuously; that is, $\funt$ has jumps of unit value.
-
-Consider an arbitrary $\len$ such that for some $\tiles \in \mathbb N, \tiles \geq 4$
+As for the properties of $\funt$, it follows from \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j} that this function is piecewise constant and left-continuous.
+From the procedure described in the previous paragraphs for obtaining $(\isoli, \jsoli)$ it is clear that $\isoli+\jsoli-1$ increases in steps of $1$ when $\len$ is increased continuously; that is, $\funt$ has jumps of unit height.
+
+Consider an arbitrary $\len$ such that for some $\tiles \in \mathbb N$, $\tiles \geq 4$
 \begin{equation}
@@
 
-By Theorem~\ref{theo: funt, form}, $\funt$ has unit-height jumps at the values $\funl(\tiles)$, $\tiles \in \mathbb N, \tiles \geq 4$.
+By Theorem~\ref{theo: funt, form}, $\funt$ has unit-height jumps at the values $\funl(\tiles)$, $\tiles \in \mathbb N$, $\tiles \geq 4$.
 This implies that $\funl$ is monotone increasing for $\tiles \geq 3$.

### 2021-09-18 paper.tex
Added references on Buffon's needle problem
Typos

@@
 \newcommand{\funta}{\varphi} % function that gives average number of touched tiles for real-valued length
-\newcommand{\probmax}{\rho} % function that gives the probability of achieving the maximum number of touched files in the square case for real-valued length
+\newcommand{\probmax}{\rho} % function that gives the probability of achieving the maximum number of touched tiles in the square case for real-valued length
 \newcommand{\len}{\ell} % length variable, real-valued
@@
 The problems studied in this paper are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem.
-Some references: ``Buffon's Noodle Problem'', J.F.~Ramaley, 1969.
-``On Laplace's Extension of the Buffon Needle Problem'', B.J.~Arnow, 1994.
-\emph{An Introduction to Geometrical Probability}, A.M~Mathai, section 1.1.
-\\
-\url{https://www.jstor.org/stable/2317945} \\
-\url{https://www.jstor.org/stable/2687085}
+Some references: \cite{Ramaley69}, \cite{Arnow94}, \cite{Mathai99}.
+%\url{https://www.jstor.org/stable/2317945}
+%\url{https://www.jstor.org/stable/2687085}
 
@@
 \begin{proof}
-% *** Is this writeup too informal/non-rigorous?
-% ***LM: I would make the rational-length extension more explicit, and invoke a suitable theorem (with the corresponding book reference) for the real-length case.
-Or better yet: the first reference linked as a comment in the introduction (https://www.jstor.org/stable/2317945?seq=1#metadata_info_tab_contents) already gives the $2\len/(\pi a)$ formula.
-We can directly start from that formula, referencing that paper.
+% ***AA: Is this writeup too informal/non-rigorous?
+% ***LM: \cite{Ramaley69} already gives the $2\len/(\pi a)$ formula. We can directly start from that formula, referencing that paper, which simplies the proof.
 The classic Buffon's Needle result (**AA cite Buffon's 1777 paper?** **LM I think maybe something more modern such as the references that are linked in the Introduction section**) states that if a needle of length $1$ is tossed with uniformly random position and orientation on
@@
 Under this assumption, the range of lengths corresponding to $\tiles$ is given by \eqref{eq: square, odd max tiles, lengths}.
-For any length in this interval, there are some positions within the reference file from which the tile with lower-left coordinate $((\tiles-1)/2, (\tiles-1)/2)$ can be reached.
+For any length in this interval, there are some positions within the reference tile from which the tile with lower-left coordinate $((\tiles-1)/2, (\tiles-1)/2)$ can be reached.
 Equivalently, the shifted segment with one endpoint at $((\tiles-1)/2, (\tiles-1)/2)$ can reach the reference tile.

### 2021-09-21 paper.tex
Incorporated Buffon citations
Simplified proof using Buffon's results to use citations, and added introductory paragraph on the paper relating to past work.

@@
 
-The problems studied in this paper are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem.
-Some references: \cite{Ramaley69}, \cite{Arnow94}, \cite{Mathai99}.
+The problems studied in this paper are related to Buffon's needle problem and Laplace's extension of it.
+Buffon's original 1733 problem considers tossing a needle of length $\len$ onto ruled paper with parallel lines a height $b$ apart, and asks what is the probability of the needle crossing a line if it lands in a uniformly random position and orientation.
+For a needle of length $\len = b$, this is $2/\pi$, and hence repeated trials of this experiment can be used to estimate $\pi$.
+This is generalized in \cite{Ramaley69} to $\frac{2\len}{\pi b}$ for the expected number of crossings of a needle which might multiple lines.
+
+The Buffon-Laplace needle problem considers a needle tossed onto a grid of rectangular tiles of width $a$ and height $b$.
+The number of visited tiles equals one plus the number of crossings with probability 1, as shown in Proposition \ref{prop:
+i+j-1}.
+The probability of the needle staying within a single rectangle, that is visiting only one tile, is computed in \cite{Arnow94} for the case where $\len < \min(a,b)$.
+This work considers the complementary question of the chance that the needle visits the maximum number of tiles possible for its length, described by a formula in Theorem~\ref{theo: probmax, sq}.
+
+%  \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem. Some references: \cite{Ramaley69}, \cite{Arnow94}, \cite{Mathai99}. 
+% Alex: Mentioned Ramaley69 and Arnow94, not sure where Mathai99 fits in
 %\url{https://www.jstor.org/stable/2317945}
@@
 \begin{proof}
-% ***AA: Is this writeup too informal/non-rigorous?
-% ***LM: \cite{Ramaley69} already gives the $2\len/(\pi a)$ formula. We can directly start from that formula, referencing that paper, which simplies the proof.
-The classic Buffon's Needle result (**AA cite Buffon's 1777 paper?** **LM I think maybe something more modern such as the references that are linked in the Introduction section**) states that if a needle of length $1$ is tossed with uniformly random position and orientation on
-infinite ruled paper with parallel lines a distance of $1$ apart the probability that it crosses any line is $2/\pi$.
-
-This is also the expected (that is, average) number of crossings, since the needle cannot reach more than one line, ignoring the event of it lying exactly along a line, which has probability $0$.
-For a needle of length $2$, the expected number of crossings is double the amount of $2/\pi$, since expectation is additive and thus unchanged if the needle is broken into two length-$1$ halves that are tossed separately, and their total crossings are counted.
-Conversely, for a needle of length $1/2$ the expectation would be half the original one.
-
-A similar argument extends this proportionality to all rational lengths, and by monotonicity all real lengths.
-Thus for a needle of length $\len$, the expected number of crossings is $2\len/\pi$.
-If the lines are a distance $a$ apart, this is equivalent to shrinking the needle by a factor of $a$, giving $2\len/(\pi a)$ as expected number of crossings.
-
-Consider a rectangular grid with horizontal spacing $a$ and vertical spacing $b$.
-Equivalently, this is paper ruled with horizontal lines $b$ apart and vertical lines $a$ apart.
-The distribution of the number of crossings of a needle tossed onto this grid is known as the Buffon-Laplace Needle Problem (*** cite).
-This number decomposes into the counts of crossings of horizontal and vertical grid lines, and thus the expected number of crossings is the sum of those for lined paper of width $a$ and $b$ respectively:
+First, suppose the grid is simplified to just horizontal lines a height of $b$ apart.
+This matches the setup of the 1733 Buffon's needle problem, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines.
+As shown in \cite{Ramaley69}, the expected (i.e., average) number of lines crossed equals
+$$ \frac{2\len}{\pi b}.$$
+Where $b<\len$, this expectation matches the probability of the needle crossing a grid line, since it cannot cross more than one.
+
+Now consider again a grid of width $a$ and height $b$, as studied by Laplace as an extension to Buffon's original needle problem \cite{Arnow94}.
+The grid crossings decompose into crossings of horizontal and vertical grid lines, respectively.
+By linearity of expectation, the expected number of crossings is the sum of the respective expectations for parallel lines of width $a$ and $b$ respectively, giving:
 \[

### 2021-09-24 figures/figures.m
Thicker lines in figure

@@ -1617,7 +1617,7 @@ for k = [0 1+3j 2+2j 3+1j 1-3j 2-2j 3-1j -3+1j -2+2j -1+3j -3-1j -2-2j -1-3j]
     if k==0
         plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k-', 'linewidth', 1)
     else
-        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k--', 'linewidth', 1)
+        plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k--', 'linewidth', 1.4)
     end
 end
 

### 2021-09-24 figures/figures.m
Corrected figure: "a" no longer needed in tick labels

@@ -1608,8 +1608,8 @@ plot([-3.7 4.7], [0 0], 'k')
 plot([-3.7j 4.7j], 'k')
 set(gca, 'GridAlpha', .4)
 xticks(-3:4), yticks(-3:4)
-xticklabels([arrayfun(@(n)[num2str(n) char(8201) 'a'], -3:-1, 'UniformOutput', false) {'0'}  arrayfun(@(n)[num2str(n) char(8201) 'a'], 1:4, 'UniformOutput', false)]) % thin space
-yticklabels([arrayfun(@(n)[num2str(n) char(8201) 'a'], -3:-1, 'UniformOutput', false) {'0'}  arrayfun(@(n)[num2str(n) char(8201) 'a'], 1:4, 'UniformOutput', false)]) % thin space
+%xticklabels([arrayfun(@(n)[num2str(n) char(8201) 'a'], -3:-1, 'UniformOutput', false) {'0'}  arrayfun(@(n)[num2str(n) char(8201) 'a'], 1:4, 'UniformOutput', false)]) % thin space
+%yticklabels([arrayfun(@(n)[num2str(n) char(8201) 'a'], -3:-1, 'UniformOutput', false) {'0'}  arrayfun(@(n)[num2str(n) char(8201) 'a'], 1:4, 'UniformOutput', false)]) % thin space
 xlabel x 
 xlabel y
 axis([-4 5 -4 5])

### 2021-09-24 paper.tex
Rewritten proof of \ref{theo: probmax, sq} to clarify symmetry of considered tiles

@@
 
-A square grid has $a=b$. For real-values segment lengths it can be further assumed that $a=1$. For $a \neq 1$ the results to be obtained apply with $\len$ replaced by $\len/a$.
+A square grid has $a=b$. For real-valued segment lengths it can be further assumed that $a=1$. For $a \neq 1$ the results to be obtained apply with $\len$ replaced by $\len/a$.
 
@@
 
-Consider a length $\len$ with corresponding maximum number of tiles $\tiles = \funt(\len)$.
 The first endpoint of the segment is assumed to be in the reference tile with lower-left corner $(0,0)$.
-The segment visits $\tiles$ tiles almost surely if and only if its second endpoint is in a tile with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, $i+j-1=\tiles$, or in a tile symmetrical to this in the other three quadrants of the plane.
-Figure~\ref{fig: probmax_possible_tiles} shows the reference tile, with solid line, and some of the tiles that fulfil this condition, with dashed line, for $\tiles=5$.
-The desired probability can be computed by considering only the tiles in the first quadrant and multiplying by $4$.
-
-% ***Does this figure add something, or is it better to remove it?
-\begin{figure}
-\centering%
-\includegraphics[width=.52\textwidth]{probmax_possible_tiles}%
-\caption{Tiles in which the second endpoint can be for a given number of visited tiles $\tiles$ in a square grid (example for $\tiles=5$)
-}%
-\label{fig: probmax_possible_tiles}%
-\end{figure}%
-
-For any $i,j$ with $i+j-1=\tiles$ and for a given orientation $\theta$, the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ if and only if shifting the segment so that its second endpoint has coordinates $(i-1,j-1)$ results in the first endpoint being in the reference tile.
-It cannot be the case that the segment ``overshoots'' past this tile, because that would result in a number of visited tiles exceeding the assumed maximum $\tiles$.
+For an arbitrary value $\tiles$, the segment visits $\tiles$ tiles almost surely if and only if its second endpoint is in a tile with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, $i+j-1=\tiles$, or in a tile symmetrical to this with respect to the line $x=1/2$ (with $i \leq 0$), or $y=1/2$ (with $j \leq 0$), or both.
+
+Consider $i,j\geq 1$ for the moment, and let $\tiles = \funt(\len)$.
+For a given orientation $\theta$, the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ if and only if shifting the segment so that its second endpoint has coordinates $(i-1,j-1)$ results in the first endpoint being in the reference tile.
+It cannot be the case that the segment ``overshoots'' past this tile, because that would result in a number of visited tiles exceeding the assumed maximum, $\tiles$.
 
@@
 The area occupied by this rectangle is the probability $\probmax_{i-1,j-1}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$.
-The desired probability, $\probmax(\len)$, is obtained by averaging $\probmax_{i-1,j-1}(\len, \theta)$ over $\theta$ and summing over all $i-1,j-1$ that can be reached from the reference tile.
-
-There are two situations depending on whether $\tiles$ is odd or even.
+The desired probability, $\probmax(\len)$, will be obtained by averaging $\probmax_{i-1,j-1}(\len, \theta)$ over $\theta$, summing over all $i, j \geq 1$ such that the tile with lower-left corner $(i-1, j-1)$ can be reached from the reference tile, and then using symmetry to take into account the tiles with $i \leq 0$ or $j \leq 0$.
+
+The situation is different depending on whether $\tiles$ is odd or even.
 Consider $\tiles$ odd first.
-Under this assumption, the range of lengths corresponding to $\tiles$ is given by \eqref{eq: square, odd max tiles, lengths}.
+Under this assumption, the range of lengths corresponding to the condition $\tiles = \funt(\len)$ is given by \eqref{eq: square, odd max tiles, lengths}.
 For any length in this interval, there are some positions within the reference tile from which the tile with lower-left coordinate $((\tiles-1)/2, (\tiles-1)/2)$ can be reached.
@@
 In this case, by symmetry, each of those two tiles contributes the same probability.
-Other tiles corresponding to the same $\tiles$ need not be considered, because they cannot be reached with lengths satisfying \eqref{eq: square, odd max tiles, lengths}.
+Other tiles with $i,j\geq 1$ corresponding to the same $\tiles$ do not need to be considered, because they cannot be reached with lengths satisfying \eqref{eq: square, odd max tiles, lengths}.
 
@@
 This can only happen for $\tiles \geq 8$.
-No other tiles can be reached from the reference tile.
-
-Thus, two cases need to be distinguished for $\tiles$ odd, depending on whether \eqref{eq: probmax len odd2} is satisfied or not; and similarly two cases exist for $\tiles$ even, depending on whether \eqref{eq: probmax len even2} holds or not.
+No other tiles with $i,j\geq 1$ can be reached from the reference tile.
+
+The above two paragraphs show that, among all possible tiles with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, those whose centers lie on the lines $x=1/2$ or $y=1/2$ (i.e.~those with $i=1$ of $j=1$) are never reachable from the reference tile, in the sense that a segment that touches the maximum number of tiles cannot have its second endpoint in those tiles.
+This implies that $\probmax(\len)$ can be computed by considering only the tiles with $i,j\geq 1$ (actually $i,j \geq 2$), and then multiplying by $4$ to include the symmetrical tiles with respect to $x=1/2$ and $y=1/2$.
+As an example, Figure~\ref{fig: probmax_possible_tiles} shows the reference tile (solid line) and the tiles that can contain the second endpoint (dotted lines) for $\funt(\len) = 5$.
+
+\begin{figure}
+\centering%
+\includegraphics[width=.52\textwidth]{probmax_possible_tiles}%
+\caption{Tiles that can contain the second endpoint of a segment that visits the maximum number of tiles $\funt(\len)$ in a square grid (example for $\funt(\len)=5$)%
+}%
+\label{fig: probmax_possible_tiles}%
+\end{figure}%
+
+According to the preceding analysis, two cases need to be distinguished for $\tiles$ odd, depending on whether \eqref{eq: probmax len odd2} is satisfied or not; and similarly two cases exist for $\tiles$ even, depending on whether \eqref{eq: probmax len even2} holds or not.
 These are, respectively, the four cases in \eqref{eq: probmax}.
 
-In the first case, as has been argued, only the tile with lower-left coordinates $((\tiles-1)/2, (\tiles-1)/2)$ is reachable.
+In the first case indicated in \eqref{eq: probmax}, as argued, $\tiles$ is odd and only the tile with lower-left coordinates $((\tiles-1)/2, (\tiles-1)/2)$ is reachable.
 For a given $\theta$, the conditional probability $\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)$ is the area of the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_diag}:
@@
 
-For the second case in \eqref{eq: probmax}, $\probmax(\len)$ is obtained as in the preceding paragraph plus an additional term corresponding to the contribution of the tile with lower-left corner $((\tiles-3)/2, (\tiles+1)/2)$, multiplied by $2$ to account for its symmetrical tile $((\tiles+1)/2, (\tiles-3)/2)$ and by $4$ to include all quadrants.
+For the second case in \eqref{eq: probmax}, $\probmax(\len)$ is obtained as above plus an additional term corresponding to the tile with lower-left corner $((\tiles-3)/2, (\tiles+1)/2)$, multiplied by $2$ to account for its symmetrical tile about the line $y=x$, namely that with lower-left corner $((\tiles+1)/2, (\tiles-3)/2)$, and by $4$ to include all quadrants.
 Observing the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_offdiag},

### 2021-09-25 paper.tex
New proof of Proposition 3 (\ref{prop: min suff set, form})

@@
 \newcommand{\genfun}{q}
-\newcommand{\genvar}{r}
+\newcommand{\genvar}{s}
+\newcommand{\roundterm}{r}
 \newcommand{\mss}{M}
@@
 \label{eq: min suff set, form, i}
-i_\tiles &= \left\lfloor \frac {(2\tiles-1)b^2+5a^2}{2(a^2+b^2)} \right\rfloor, \\
+i_\tiles &= \left\lfloor \frac{(\tiles-3) b^2}{a^2+b^2} + \frac 5 2 \right\rfloor, \\
 \label{eq: min suff set, form, j}
-j_\tiles &= \left\lceil \frac {(2\tiles-3) a^2 + 3b^2}{2(a^2+b^2)} \right\rceil,
-% ***change expressions to the following to match the statement of Theorem~\ref{theo: funl, form}:
-%i_\tiles &= \left\lfloor \frac{b^2 (\tiles-3)}{a^2+b^2} + \frac 5 2 \right\rfloor \\
-%j_\tiles &= \left\lceil  \frac{a^2 (\tiles-3)}{a^2+b^2} + \frac 3 2 \right\rceil
+j_\tiles &= \left\lceil  \frac{(\tiles-3) a^2}{a^2+b^2} + \frac 3 2 \right\rceil,
 \end{align}
@@
 \begin{proof}
-***New proof based on treating i,j as continuous along the line i+j=constant, then quantizing to the nearest allowed point along that line.
-We should use the notation $\isolr$, $\jsolr$ for the continuous coordinates, in conformity to other proofs that use continuous coordinates.
-% ***The continuous argument gives something like $i-2 = b^2/(a^2+b^2)*(t-3)$, $j-2 = a^2/(a^2+b^2)*(t-3)$.
-We then compute how much the distance from $(i,j)$ to $(2,2)$ is increased by quantizing along the i+j line, to argue that the nearest quantized point should be chosen.
-To ensure that in case of a tie we choose the point with larger $i$, we apply floor(...+1/2) to i and ceiling(...-1/2) to j.
-% ***Reference the figure in the text.
-The figure is an example with $a=1.35$, $b=1$, $\tiles=7$.
-It contains the line j-2 = a^2/b^2*(i-2), whose intersection with i+j-1=t (which is perpendicular to it) gives the continuous solution $(\isolr, \jsolr)$ (marked with a square).
-
-***The bounding lines are a direct consequence of how $i_\tiles$ and $j_\tiles$ are obtained.
+For each $\tiles$, the pair $(i_\tiles,j_\tiles)$ of the minimal sufficient set should be chosen as that on the line $i+j-1=\tiles$ which minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$.
+This allows maximizing the sum $i+j-1$, and thus the number of visited tiles, for a given length restriction (direct problem); or minimizing the required lengths for a specified number of visited files (inverse problem).
+
+Consider, for the moment, $i$, $j$ as if they were real-valued variables, and denote $x=ia$, $y=jb$. The line $i+j-1 = \tiles$ then becomes
+\begin{equation}
+\label{eq: line 1}
+\frac x a + \frac y b = \tiles + 1,
+\end{equation}
+and $(i-2)^2 a^2 + (j-2)^2 b^2$ is expressed as $(x-2a)^2+(y-2b)^2$.
+The point minimizing this quadratic function along \eqref{eq: line 1} is the intersection of this line with the perpendicular line passing through $(2a,2b)$, that is,
+\begin{equation}
+\label{eq: line 2}
+y = \frac {a (x-2a)} b + 2b.
+\end{equation}
+This is depicted for $a=1.35$, $b=1$, $\tiles=7$ in Figure~\ref{fig: min suff set, form}, where \eqref{eq: line 2} is shown with solid line.
+Solving the system of equations \eqref{eq: line 1} and \eqref{eq: line 2} gives
+\begin{align}
+\frac x a - 2 &= \frac{(t-3)b^2}{a^2+b^2}, \\
+\frac y b - 2 &= \frac{(t-3)a^2}{a^2+b^2}.
+\end{align}
+In terms of the real-valued variables $i$, $j$, the solution $(\isolr, \jsolr)$ is
+\begin{align}
+\label{eq: prop: min suff set, form: isolr}
+\isolr &= \frac{(t-3)b^2}{a^2+b^2} + 2, \\
+\label{eq: prop: min suff set, form: jsolr}
+\jsolr &= \frac{(t-3)a^2}{a^2+b^2} + 2.
+\end{align}
+This point is shown in Figure~\ref{fig: min suff set, form} with a square marker.
 
@@
 \end{figure}%
+
+The variables $i$, $j$ are actually limited to integer values.
+The point $(i_\tiles, j_\tiles)$ that minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$ along the line $i+j-1=\tiles$ restricted to integer values is either $(\lceil \isolr \rceil, \lfloor \jsolr \rfloor)$ or $(\lfloor \isolr \rfloor, \lceil \jsolr \rceil)$, whichever is closest to $(\isolr, \jsolr)$, as can be seen in Figure~\ref{fig: min suff set, form}.
+This can be expressed as
+\begin{align}
+\label{eq: prop: min suff set, form: isoli}
+i_\tiles &= \left\lfloor \isolr + \frac 1 2 \right\rfloor, \\
+\label{eq: prop: min suff set, form: jsoli}
+j_\tiles &= \left\lceil \jsolr - \frac 1 2 \right\rceil,
+\end{align}
+which in case of tie takes the first of the two points indicated above. Combining \eqref{eq: prop: min suff set, form: isolr}--\eqref{eq: prop: min suff set, form: jsoli}
+% *!*
+yields \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}.
+
+From \eqref{eq: prop: min suff set, form: isolr} and \eqref{eq: prop: min suff set, form: isoli},
+\begin{equation}
+\label{eq: prop: min suff set, form: isoli ineq}
+\frac{(t-3)b^2}{a^2+b^2} + \frac 3 2 < i_\tiles \leq \frac{(t-3)b^2}{a^2+b^2} + \frac 5 2,
+\end{equation}
+and similarly, from \eqref{eq: prop: min suff set, form: jsolr} and \eqref{eq: prop: min suff set, form: jsoli},
+\begin{equation}
+\label{eq: prop: min suff set, form: jsoli ineq}
+\frac{(t-3)a^2}{a^2+b^2} + \frac 3 2 \leq j_\tiles < \frac{(t-3)a^2}{a^2+b^2} + \frac 5 2.
+\end{equation}
+Considering the first inequality in \eqref{eq: prop: min suff set, form: isoli ineq} and the second in \eqref{eq: prop: min suff set, form: jsoli ineq} as equalities and eliminating $\tiles$ gives \eqref{eq: upper bound, line}.
+The point $(i_\tiles, j_\tiles)$ is strictly below that line because the considered inequalities are strict.
+Similarly, the bounding line \eqref{eq: lower bound, line} results from the second inequality in \eqref{eq: prop: min suff set, form: isoli ineq} and the first in \eqref{eq: prop: min suff set, form: jsoli ineq}, and the fact that both inequalities are not strict implies that the bound \eqref{eq: lower bound, line} can actually be attained.
 \end{proof}
@@
 \end{equation}
+with **define i+, such that i* = round-half-up(i+), and similarly for j+**
+\begin{align}
+\label{eq: theo: funl, form; isolr}
+\isolr &= \max\left\{ \frac{(\tiles-3) b^2}{a^2+b^2}, 0 \right\} + 2, \\
+\label{eq: theo: funl, form; jsolr}
+\jsolr &= \max\left\{ \frac{(\tiles-3) a^2}{a^2+b^2}, 0 \right\} + 2, \\
+\label{eq: theo: funl, form; isoli}
+\isoli &= \left\lfloor \isolr + \frac 1 2 \right\rfloor, \\
+\label{eq: theo: funl, form; jsoli}
+\jsoli &= \left\lceil \jsolr - \frac 1 2 \right\rceil.
+\end{align}
+Equivalently,
+\begin{equation}
+\label{eq: theo: funl, form; funl 2}
+\funl(\tiles) = \sqrt{\frac{(t-3) a^2 b^2}{a^2+b^2} + \roundterm^2(a^2+b^2)}
+\end{equation}
 with
-\begin{align}
-\label{eq: theo: funl, form; i}
-\isoli &= \max\left\{ \left\lfloor \frac {(2\tiles-1)b^2+5a^2}{2(a^2+b^2)} \right\rfloor, 2 \right\}, \\
-\label{eq: theo: funl, form; j}
-\jsoli &= \max\left\{ \left\lceil \frac {2\tiles a^2 + 3(b^2-a^2)}{2(a^2+b^2)} \right\rceil, 2 \right\}.
-\end{align}
-Equivalently,
-\begin{equation}
-\text{***new formula found by Alex; unless we can unify with the above}
-% *** AA: Wrote a shot at how a unification might look in e-mail.
+\begin{equation}
+\label{eq: roundterm}
+\roundterm = |\isoli - \isolr| = |\jsoli - \jsolr|.
 \end{equation}
@@
 By construction of this set, any segment that visits $\tiles$ tiles must have length greater than $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$.
-
-For $\tiles \geq 3$ the expressions \eqref{eq: theo: funl, form; i} and \eqref{eq: theo: funl, form; j} coincide with \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}.
+For $\tiles \geq 3$ the expressions \eqref{eq: theo: funl, form; isoli} and \eqref{eq: theo: funl, form; jsoli} coincide with \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}.
 Therefore \eqref{eq: theo: funl, form; funl} gives the desired result.
-
-For $\tiles \in \{1, 2\}$ both \eqref{eq: theo: funl, form; i} and \eqref{eq: theo: funl, form; j} equal $2$, and \eqref{eq: theo: funl, form; funl} gives $0$, which is the correct result.
-
-***Proof of new formula
+For $\tiles \in \{1, 2\}$ both \eqref{eq: theo: funl, form; isoli} and \eqref{eq: theo: funl, form; jsoli} equal $2$, and \eqref{eq: theo: funl, form; funl} gives $0$, which is the correct result.
+
+***Proof of new formula. I think it's better to do it once the proof of \ref{prop: min suff set, form} is done.
 
@@
 % *!*
-and from \eqref{eq: theo: funl, form; funl}--\eqref{eq: theo: funl, form; j}
+and from \eqref{eq: theo: funl, form; funl}--\eqref{eq: theo: funl, form; jsoli}
 % *!*
@@
 \label{eq: f}
-f(\genvar) = \int_0^r \arccos z \diff z = \genvar \arccos \genvar - \sqrt{1-\genvar^2} + 1.
+f(\genvar) = \int_0^\genvar \arccos z \diff z = \genvar \arccos \genvar - \sqrt{1-\genvar^2} + 1.
 \end{equation}
@@
 \label{eq: g def}
-g(\genvar, u, v) = \int_{\arcsin \frac u {2r}}^{\arccos \frac u {2r}} \left( r \sin \theta - \frac u 2 \right) \left( r \cos \theta - \frac v 2 \right) \diff \theta.
+g(\genvar, u, v) = \int_{\arcsin \frac u {2r}}^{\arccos \frac u {2r}} \left( \genvar \sin \theta - \frac u 2 \right) \left( \genvar \cos \theta - \frac v 2 \right) \diff \theta.
 \end{equation}

### 2021-09-25 paper.tex
Some advances towards proof of Theorem 2, second formula

@@
 \end{equation}
-with **define i+, such that i* = round-half-up(i+), and similarly for j+**
+with
 \begin{align}
@@
 \end{align}
-Equivalently,
+Equivalently, **I think this is valid for t geq 3 only, and needs a modification**
 \begin{equation}
 \label{eq: theo: funl, form; funl 2}
-\funl(\tiles) = \sqrt{\frac{(t-3) a^2 b^2}{a^2+b^2} + \roundterm^2(a^2+b^2)}
+\funl(\tiles) = \sqrt{\frac{(t-3)^2 a^2 b^2}{a^2+b^2} + \roundterm^2(a^2+b^2)}
 \end{equation}
@@
 
-***Proof of new formula. I think it's better to do it once the proof of \ref{prop: min suff set, form} is done.
+***Proof of new formula. Perhaps it will reuse Figure~\ref{fig: min suff set, form}.
 
@@
 ***Comment on how the new formula has two terms with definite meanings: The first term is the distance-squared to the closest point on the diagonal, The second term is additional distance-squared incurred from rounding (i,j) to integers
+***Figure for that? \ref{fig: min suff set, form} would do, but it's within a proof and uses different notation for the final point
 

### 2021-09-25 figures/figures.m

@@ -1881,3 +1881,15 @@ plot(xl*a, yl*b, 's', 'markersize', ms)
 text(4.78, 4.69, '(i^+,j^+)')
 text(3.52, 4.82, '(i_t,j_t)')
 
+
+%% #71. Check for new formula for funt:
+
+t = 12;
+a = 1.56;
+b = 1.15;
+iplus = max((t-3)*b^2/(a^2+b^2), 0)+2;
+jplus = max((t-3)*a^2/(a^2+b^2), 0)+2;
+istar = floor(iplus+1/2);
+jstar = ceil(jplus-1/2);
+lambda = sqrt((istar-2)^2*a^2 + (jstar-2)^2*b^2) % original formula
+lambda_2 = sqrt( (t-3)^2*a^2*b^2/(a^2+b^2) + (istar-iplus)^2*a^2 + (jstar-jplus)^2*b^2 ) % new formula, corrected

### 2021-09-25 figures/figures.m
Confirmed new formula in Theorem 2 (\ref{eq: theo: funl, form; jsoli})

@@ -1884,7 +1884,8 @@ text(3.52, 4.82, '(i_t,j_t)')
 
 %% #71. Check for new formula for funt:
 
-t = 12;
+clear, clc
+t = 18;
 a = 1.56;
 b = 1.15;
 iplus = max((t-3)*b^2/(a^2+b^2), 0)+2;
@@ -1892,4 +1893,5 @@ jplus = max((t-3)*a^2/(a^2+b^2), 0)+2;
 istar = floor(iplus+1/2);
 jstar = ceil(jplus-1/2);
 lambda = sqrt((istar-2)^2*a^2 + (jstar-2)^2*b^2) % original formula
-lambda_2 = sqrt( (t-3)^2*a^2*b^2/(a^2+b^2) + (istar-iplus)^2*a^2 + (jstar-jplus)^2*b^2 ) % new formula, corrected
+%lambda_2 = sqrt( (t-3)^2*a^2*b^2/(a^2+b^2) + (istar-iplus)^2*a^2 + (jstar-jplus)^2*b^2 ) % new formula, corrected
+lambda_2 = sqrt( (t-3)^2*a^2*b^2/(a^2+b^2) + (istar-iplus)^2*(a^2+b^2) ) % new formula, corrected

### 2021-09-25 paper.tex
Confirmed new formula in Theorem 2 (\ref{eq: theo: funl, form; jsoli})
Proof of new formula in Theorem 2 (\ref{theo: funl, form})

@@
 
+% ***This figure and Figure~\ref{fig: ijLS_1p35_detail} label the points as (i,j) instead of (ia,jb). Clarify in the text, or change the figures
 \begin{figure}%
@@
 \end{align}
-Equivalently, **I think this is valid for t geq 3 only, and needs a modification**
+Equivalently, for $\tiles \geq 3$,
 \begin{equation}
@@
 By construction of this set, any segment that visits $\tiles$ tiles must have length greater than $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$.
-For $\tiles \geq 3$ the expressions \eqref{eq: theo: funl, form; isoli} and \eqref{eq: theo: funl, form; jsoli} coincide with \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}.
-Therefore \eqref{eq: theo: funl, form; funl} gives the desired result.
-For $\tiles \in \{1, 2\}$ both \eqref{eq: theo: funl, form; isoli} and \eqref{eq: theo: funl, form; jsoli} equal $2$, and \eqref{eq: theo: funl, form; funl} gives $0$, which is the correct result.
-
-***Proof of new formula. Perhaps it will reuse Figure~\ref{fig: min suff set, form}.
+For $\tiles \geq 3$ the variables $\isoli$, $\jsoli$ in \eqref{eq: theo: funl, form; isoli} and \eqref{eq: theo: funl, form; jsoli} respectively coincide with $i_\tiles$, $j_\tiles$ as given by \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}.
+Therefore \eqref{eq: theo: funl, form; funl} gives the correct result.
+For $\tiles \in \{1, 2\}$ both \eqref{eq: theo: funl, form; isoli} and \eqref{eq: theo: funl, form; jsoli} equal $2$, and \eqref{eq: theo: funl, form; funl} gives $0$, which is again the correct result.
+
+For $\tiles \geq 3$, the term $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$ can be interpreted geometrically as the distance between $(i_\tiles, j_\tiles)$ and $(2a, 2b)$.
+As can be seen with the help of Figure~\ref{fig: min suff set, form}, this distance is the hypotenuse of a right triangle whose other two sides extend from $(2a, 2b)$ to $(\isolr a, \jsolr b)$ and from $(\isolr a, \jsolr b)$ to $(i_\tiles, j_\tiles)$ respectively.
+Therefore,
+\begin{equation}
+\label{eq: distance as hypotenuse}
+\begin{split}
+(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2 &= (\isolr-2)^2a^2 + (\jsolr-2)^2b^2 \\
+&\quad + (i_\tiles-\isolr)^2a^2 + (j_\tiles-\jsolr)^2b^2.
+\end{split}
+\end{equation}
+For $\tiles \geq 3$ it stems from \eqref{eq: theo: funl, form; isolr}
+and \eqref{eq: theo: funl, form; jsolr} that
+\begin{equation}
+\label{eq: distance as hypotenuse, first leg}
+(\isolr-2)^2a^2 + (\jsolr-2)^2b^2 =
+\frac{(\tiles-3)^2 (b^4a^2+a^2b^4)}{(a^2+b^2)^2} =
+\frac{(\tiles-3)^2 a^2b^2}{a^2+b^2}.
+\end{equation}
+The fact that both $(\isolr, \jsolr)$ and $(i_\tiles, j_\tiles)$ are on the line $i+j-1=\tiles$ implies that $\isolr+\jsolr = i_\tiles+ j_\tiles$, and thus $|i_\tiles-\isolr| = |j_\tiles-\jsolr| = \roundterm$.
+Consequently,
+\begin{equation}
+\label{eq: distance as hypotenuse, second leg}
+(i_\tiles-\isolr)^2a^2 + (j_\tiles-\jsolr)^2b^2 = \roundterm^2 (a^2+b^2).
+\end{equation}
+Substituting \eqref{eq: distance as hypotenuse, first leg} and \eqref{eq: distance as hypotenuse, second leg} into \eqref{eq: distance as hypotenuse} yields \eqref{eq: theo: funl, form; funl 2}.
 
@@
 ***Comment on how the new formula has two terms with definite meanings: The first term is the distance-squared to the closest point on the diagonal, The second term is additional distance-squared incurred from rounding (i,j) to integers
-***Figure for that? \ref{fig: min suff set, form} would do, but it's within a proof and uses different notation for the final point
+***I think we can do without a figure (Figure~\ref{fig: min suff set, form} is inside a proof and uses different notation for the final point).
 

### 2021-10-06 paper.tex
Add Mathai citations
Add Mathai citations

@@
 For a needle of length $\len = b$, this is $2/\pi$, and hence repeated trials of this experiment can be used to estimate $\pi$.
-This is generalized in \cite{Ramaley69} to $\frac{2\len}{\pi b}$ for the expected number of crossings of a needle which might multiple lines.
-
-The Buffon-Laplace needle problem considers a needle tossed onto a grid of rectangular tiles of width $a$ and height $b$.
+This is generalized in \cite{Ramaley69} to $\frac{2\len}{\pi b}$ for the expected number of crossings of a needle which might intersect multiple lines.
+
+The Buffon-Laplace needle problem, described in \cite{Mathai99}, considers a needle tossed onto a grid of rectangular tiles of width $a$ and height $b$.
 The number of visited tiles equals one plus the number of crossings with probability 1, as shown in Proposition \ref{prop:
@@
 First, suppose the grid is simplified to just horizontal lines a height of $b$ apart.
-This matches the setup of the 1733 Buffon's needle problem, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines.
+This matches the setup of the 1733 Buffon's needle problem described in \cite{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines.
 As shown in \cite{Ramaley69}, the expected (i.e., average) number of lines crossed equals

### 2021-10-22 paper.tex
Renamed "canonical rectangle" to "discrete bounding rectangle"

@@
 
-Every segment has an associated \emph{canonical rectangle}, which is the minimum-size rectangle that is formed by grid lines and contains the segment.
-More specifically, if the segment has endpoints $(x_1,y_1)$, $(x_2,y_2)$, where $x_1, x_2, y_1, y_2 \in \mathbb R$, its canonical rectangle has lower-left corner and upper-right corner given as
+Every segment has an associated \emph{discrete bounding rectangle}, which is the minimum-size rectangle that is formed by grid lines and contains the segment.
+More specifically, if the segment has endpoints $(x_1,y_1)$, $(x_2,y_2)$, where $x_1, x_2, y_1, y_2 \in \mathbb R$, its discrete bounding rectangle has lower-left corner and upper-right corner given as
 \begin{align*}
@@
 \end{align*}
-The dimensions of the canonical rectangle, normalized to the tile width and height respectively, are two integer numbers $i$, $j$.
-Two examples are illustrated in Figure~\ref{fig: canonical rectangle and touched tiles}, both with $i=5$, $j=4$.
-Clearly, all tiles visited by the segment are contained in the canonical rectangle.
-Note also that the canonical rectangle can have $i=0$ or $j=0$ if the segment coincides with part of a grid line.
+The dimensions of the discrete bounding rectangle, normalized to the tile width and height respectively, are two integer numbers $i$, $j$.
+Two examples are illustrated in Figure~\ref{fig: discrete bounding rectangle and touched tiles}, both with $i=5$, $j=4$.
+Clearly, all tiles visited by the segment are contained in the discrete bounding rectangle.
+Note also that the rectangle can have $i=0$ or $j=0$ if the segment coincides with part of a grid line.
 
@@
 }%
-\caption{Canonical rectangle and visited tiles
-}%
-\label{fig: canonical rectangle and touched tiles}%
+\caption{Discrete bounding rectangle and visited tiles
+}%
+\label{fig: discrete bounding rectangle and touched tiles}%
 \end{figure}%
@@
 \label{prop: i+j-1}
-Consider an arbitrary segment, and let $i, j$ respectively denote the normalized width and height of its canonical rectangle.
+Consider an arbitrary segment, and let $i, j$ respectively denote the normalized width and height of its discrete bounding rectangle.
 If $i, j \geq 1$, the number of tiles visited by the segment is at most $i+j-1$.
@@
 \begin{proof}
-The segment visits, by definition, two tiles in opposite corners of the canonical rectangle.
-It can be assumed, without loss of generality, that those tiles are in the lower-left and upper-right corners of the rectangle, as in Figure~\ref{fig: canonical rectangle and touched tiles}.
-The visited tiles can be thought of as following a path within the canonical rectangle.
+The segment visits, by definition, two tiles in opposite corners of the discrete bounding rectangle.
+It can be assumed, without loss of generality, that those tiles are in the lower-left and upper-right corners of the rectangle, as in Figure~\ref{fig: discrete bounding rectangle and touched tiles}.
+The visited tiles can be thought of as following a path within the discrete bounding rectangle.
 Starting at the lower-left tile, the next tile can be the one to the left, the one above, or the one above and to the left.
@@
 
-This implies that the maximum number of visited tiles is $i+j-1$, which is attained if and only if the segment avoids all grid points in the interior of the canonical rectangle, as in Figure~\ref{fig: S_nogridpoints}.
+This implies that the maximum number of visited tiles is $i+j-1$, which is attained if and only if the segment avoids all grid points in the interior of the discrete bounding rectangle, as in Figure~\ref{fig: S_nogridpoints}.
 Note that grid points at the corners of the rectangle do not count for this; and that the segment cannot pass through any other grid points on the rectangle border, because that would imply $i=0$ or $j=0$.
@@
 \label{prop: len ineq i j: ineqs}
-The following inequalities hold for any segment with length $\len$ whose canonical rectangle has normalized dimensions $i, j$:
+The following inequalities hold for any segment with length $\len$ whose discrete bounding rectangle has normalized dimensions $i, j$:
 \begin{align}
@@
 \label{prop: len ineq i j: exist}
-Conversely, if $\len$, $i$, $j$ satisfy \eqref{eq: len > sqrt} and \eqref{eq: len leq sqrt} there exists a segment of length $\len$ whose canonical rectangle has normalized dimensions $i$ and $j$.
+Conversely, if $\len$, $i$, $j$ satisfy \eqref{eq: len > sqrt} and \eqref{eq: len leq sqrt} there exists a segment of length $\len$ whose discrete bounding rectangle has normalized dimensions $i$ and $j$.
 \item
 \label{prop: len ineq i j: ineq, exist}
-There is a segment of length not exceeding $\len$ that has a canonical rectangle with normalized dimensions $i, j$ if and only if \eqref{eq: len > sqrt} holds.
+There is a segment of length not exceeding $\len$ that has a discrete bounding rectangle with normalized dimensions $i, j$ if and only if \eqref{eq: len > sqrt} holds.
 \end{enumerate}
@@
 \begin{proof}
-\ref*{prop: len ineq i j: ineqs} The inequalities follow from the fact that the segment endpoints lie in the interiors or on the outer edges of two tiles in opposite corners of the canonical rectangle.
+\ref*{prop: len ineq i j: ineqs} The inequalities follow from the fact that the segment endpoints lie in the interiors or on the outer edges of two tiles in opposite corners of the discrete bounding rectangle.
 This is illustrated in Figures~\ref{fig: len ineq 4 3} and \ref{fig: len ineq 4 2} for two specific $(i,j)$ pairs respectively.
@@
 
-\ref*{prop: len ineq i j: exist} For $a$, $b$, $\len$, $i$, $j$ satisfying the two inequalities, a segment of length $\len$ can be found that has its endpoints in the interiors or on the outer edges of the two shaded tiles of a rectangle of normalized dimensions $i$ and $j$, as in Figure~\ref{fig: len ineq i j}.
-Therefore this segment has the given rectangle as canonical.
+\ref*{prop: len ineq i j: exist} For $a$, $b$, $\len$, $i$, $j$ satisfying the two inequalities, a segment of length $\len$ can be found that has its endpoints in the interiors or on the outer edges of the two shaded tiles of a rectangle of normalized dimensions $i$ and $j$, as in Figure~\ref{fig: len ineq i j}, which is thus the discrete bounding rectangle of that segment.
 \begin{figure}
@@
 }%
-\caption{Relationship between segment length $\len$ and dimensions $i$, $j$ of the canonical rectangle
+\caption{Relationship between segment length $\len$ and dimensions $i$, $j$ of the discrete bounding rectangle
 }%
@@
 
-``There is a segment\ldots $\Rightarrow$ \eqref{eq: len > sqrt} '': Assume that a segment exists with length not exceeding $\len$ and with a canonical rectangle with normalized dimensions $i$, $j$.
+``There is a segment\ldots $\Rightarrow$ \eqref{eq: len > sqrt} '': Assume that a segment exists with length not exceeding $\len$ and with a discrete bounding rectangle with normalized dimensions $i$, $j$.
 From part~\ref{prop: len ineq i j: ineqs}, inequality \eqref{eq: len > sqrt} holds for that segment length, and thus for $\len$.
@@
 
-According to Proposition~\ref{prop: i+j-1}, in order to maximize the number of visited tiles for a given length, the position and orientation of the segment should be chosen to obtain $i+j-1$ as large as possible, where $i$ and $j$ are the normalized dimensions of its canonical rectangle.
+According to Proposition~\ref{prop: i+j-1}, in order to maximize the number of visited tiles for a given length, the position and orientation of the segment should be chosen to obtain $i+j-1$ as large as possible, where $i$ and $j$ are the normalized dimensions of its discrete bounding rectangle.
 On the other hand, Proposition~\ref{prop: len ineq i j} restricts the $i, j$ values that can be achieved with a given length.
@@
 \includegraphics[width=.7\textwidth]{ijLS_1}%
-\caption{Relationship of segment length and number of visited tiles with the width and height of the canonical rectangle, for $a=b$
+\caption{Relationship of segment length and number of visited tiles with the width and height of the discrete bounding rectangle, for $a=b$
 }%
@@
 \includegraphics[width=.7\textwidth]{ijLS_1p35}%
-\caption{Relationship of segment length and number of visited tiles with the width and height of the canonical rectangle, for $a=1.35b$
+\caption{Relationship of segment length and number of visited tiles with the width and height of the discrete bounding rectangle, for $a=1.35b$
 }%
@@
 
-For $a^2/b^2$ arbitrary, the number of pairs with the same $i$ coordinate, or with the same $j$, in the set $\mss$ is in general irregular, because the lines \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} do not follow a ``natural''
-% ***it may be better to say ``simple'' instead of ``natural'', specially if we rename ``canonical`'' to ``natural'' or ``natural bounding''
-direction of the grid.
+For $a^2/b^2$ arbitrary, the number of pairs with the same $i$ coordinate, or with the same $j$, in the set $\mss$ is in general irregular, because the lines \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} do not follow a ``natural'' direction of the grid.
 This happens for instance in Figure~\ref{fig: pairs_bounds_1p35}, where the number of pairs for the first $i$ values equals either $2$ or $3$ without a clear pattern.\footnote{Strictly, there is a periodic pattern whenever $a^2/b^2$ is rational, which is the case in Figure~\ref{fig: pairs_bounds_1p35}.
@@
 
-A segment whose canonical rectangle has normalized width $i$ and height $j$ is oriented with approximate slope $jb/(ia)$ with respect to the $x$ axis (see Figure~\ref{fig: len ineq i j}); and the approximation becomes better for greater segment lengths.
+A segment whose discrete bounding rectangle has normalized width $i$ and height $j$ is oriented with approximate slope $jb/(ia)$ with respect to the $x$ axis (see Figure~\ref{fig: len ineq i j}); and the approximation becomes better for greater segment lengths.
 From \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} it can be seen that the positions of the pairs $(i,j) \in \mss$ have $j/i \approx a^2/b^2$ for large $i, j$.
@@
 
-As for $a \gg b$, in this case the optimal canonical rectangle has $\isoli = 2$, and $\jsoli$ as large as allowed by $\len$ (direct problem) or as required by $\tiles$ (inverse problem), corresponding to an almost vertical segment.
+As for $a \gg b$, in this case the optimal discrete bounding rectangle has $\isoli = 2$, and $\jsoli$ as large as allowed by $\len$ (direct problem) or as required by $\tiles$ (inverse problem), corresponding to an almost vertical segment.
 Obviously, for $a \gg b$ the length of the segment is best invested in increasing the number of tiles traversed vertically (but the segment should be slightly tilted to cross a vertical edge), and the asymptotic value of \eqref{eq: asympt} is approximately $1/b$.
@@
 
-Each realization of the random segment gives rise to a canonical rectangle, whose normalized dimensions $i$ and $j$ are thus random variables, as is the number $\tiles$ of visited tiles.
+Each realization of the random segment gives rise to a discrete bounding rectangle, whose normalized dimensions $i$ and $j$ are thus random variables, as is the number $\tiles$ of visited tiles.
 Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and $\tiles = i+j-1$ tiles.
@@
 Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above.
-Let the random variables $i$, $j$ represent the normalized dimensions of the canonical rectangle.
+Let the random variables $i$, $j$ represent the normalized dimensions of the discrete bounding rectangle.
 For $n \in \mathbb N$,
@@
 }%
-\caption{Conditions for the normalized width of the canonical rectangle, $i$, to be equal or greater than a given $n$. Example with $a=1.35$, $b=1$, $n=3$
+\caption{Conditions for the normalized width of the discrete bounding rectangle, $i$, to be equal or greater than a given $n$. Example with $a=1.35$, $b=1$, $n=3$
 }%
@@
 It tends to $2/\pi$ for $a/b \rightarrow \infty$ and for $a/b \rightarrow 0$.
-The minimum number of tiles that can be visited with non-zero probability (corresponding to an n times 1 canonical rectangle if a>=b) can also be plotted.
+The minimum number of tiles that can be visited with non-zero probability (corresponding to an n times 1 discrete bounding rectangle if a>=b) can also be plotted.
 

### 2021-10-22 paper.tex
Update about to-do graph

@@
 
-***Maybe a graph comparing maximum and average.
-Or compute/plot their ratio.
-The asymptotic slopes are $\sqrt(1/a^2+1/b^2)$ for $\funt$ and $2/\pi(1/a+1/b)$.
+***Graph of asymptotic slopes: $\sqrt(1/a^2+1/b^2)$ for $\funt$ and $2/\pi(1/a+1/b)$.
 It may be interesting to consider the ratio of asymptotic slopes, as a function of $a/b$.
@@
 It tends to $2/\pi$ for $a/b \rightarrow \infty$ and for $a/b \rightarrow 0$.
-The minimum number of tiles that can be visited with non-zero probability (corresponding to an n times 1 discrete bounding rectangle if a>=b) can also be plotted.
 

### 2021-10-22 paper.tex
Rephrasing and rewriting to remove the name "minimal sufficient set"

@@
 In other words, what is the ``smallest'' subset of $\mathbb N^2$ such that the $(i,j)$ pair that maximizes the number of tiles for any given length can always be found within that subset?
-A subset that contains a maximizing $(i,j)$ for any length will be called a \emph{sufficient} set.
-Clearly, this set must contain at least one such pair for each possible value of $i+j-1$, so that the set can produce that value as solution (for certain lengths).
-A sufficient set that contains exactly one pair $(i,j)$ for each value of $i+j-1$ will be called a \emph{minimal sufficient} set.
-
-For instance, it is intuitively clear from Figure~\ref{fig: examples} that segment orientations near the vertical or horizontal directions (resulting in $i=1$ with large $j$, or $j=1$ with large $i$) will not maximize the number of visited tiles.
-% This is true for any a, b: 2 can achieved instead of 1 without compromising the other dimension.
-The corresponding $(i,j)$ pairs can be left out of the sufficient set.
-In general, the pairs included in a minimal sufficient set will be different depending on $a$ and $b$.
-Two specific examples of minimal sufficient sets will be presented later in this section.
-
-% ***Is the text from here to the proposition too long? Shorten it?
-
-% ***AA: Do we actually need the concept of a minimal sufficient set, given that there's a single value at each diagonal (tiebreaking a certain way)? Perhaps rename to suggest it's one value?
-
-Knowing a minimal sufficient set of $(i,j)$ pairs facilitates the solution of both the direct an inverse problems, because it reduces the number of pairs that need to be tried.
-In order to derive a general method to build a minimal sufficient set, it is insightful to consider two specific cases first.
-It is also convenient to use the equivalent formulation of the direct problem referred to in \S\ref{part: intro}, i.e.~finding the maximum number of tiles visited by segments of length $\len$ \emph{or less}.
-
-% Thus each choice of $(i, j)$ implies a maximum of $i+j-1$ visited tiles on one hand; and sets a lower bound on the required length $\len$ on the other hand.
-
-Consider first $a=b=1$.
+
+For instance, it is clear from Figure~\ref{fig: examples} that segment orientations near the vertical or horizontal directions (resulting in $i=1$ with large $j$, or $j=1$ with large $i$) will not maximize the number of visited tiles, and thus the corresponding $(i,j)$ pairs can be discarded.
+On the other hand, the set of optimal $(i,j)$ pairs must contain one such pair for each possible value of $i+j-1$, so that the set can produce that value as the solution (maximum number of visited tiles) for certain lengths.
+It is insightful to examine two specific examples before giving an explicit formula for the coordinates of the optimal pairs.
+
+Consider $a=b=1$ first.
 This is illustrated in Figure~\ref{fig: ijLS_1}.
 Note that in this and in the next figures the axes represent $ia$ and $jb$ (not $i$ and $j$).
-Each dashed diagonal line joins $(i, j)$ pairs with the same $i+j-1$.
-The radius of each arc represents the lower bound on $\len$ given by \eqref{eq: len > sqrt}, for certain $(i,j)$ pairs, which are marked with filled circles (these pairs form a minimal sufficient set, as will be seen).
+It is convenient, however, to refer to the point $(ia,jb)$ by its normalized coordinates $(i,j)$.
+In this graph, each dashed diagonal line contains pairs $(i,j)$ with the same $i+j-1$; and the lower bound \eqref{eq: len > sqrt} corresponds to an arc centered at normalized coordinates $(2,2)$.
 
@@
 % ***Maybe combine this figure and the next into one figure with two subfigures (stacked vertically, size similar to now). Not sure. It has been done in others
-% ***Are these figures necessary?
-They were created with the aim of helping understand the first of the two propositions on $\mss$.
-Now that proposition has been removed.
-They are probably still useful, but maybe the text about these figures needs to be rewritten.
-
-For a given $\len \in \mathbb R^+$, the $(i,j)$ pairs that can be achieved with segments of length not exceeding $\len$ are, by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, those whose distance from $(2a,2b)$ is less than $\len$, i.e.~those delimited by the corresponding arc.
-It is clear from the figure that for $i+j-1$ odd, if some point $(i',j')$ with $|i'-j'|>1$ is achievable with the length bound $\len$, then necessarily there is a point $(i'',i'')$ with the same number of visited tiles $i+j-1$ (same diagonal line) that is also achievable with the same length restriction (because it has smaller distance to $(2a,2b)$).
-Similarly, for $i+j-1$ even, if a point $(i',j')$ with $|i'-j'|>1$ is achievable, then there is a point $(i'',i''-1)$ in the same diagonal that is also achievable with the same length restriction.
-
-Thus for $a=b=1$, a minimal sufficient set is formed by the pairs $(i,j)$ with $j=i$ or $j=i+1$.
-Observe that due to symmetry, any point $(i'',i''-1)$ could be replaced by $(i''-1,i'')$.
-This illustrates that the minimal sufficient set is not unique in general.
-
-The pairs in the minimal sufficient set are marked with filled circles in Figure~\ref{fig: ijLS_1}.
-Given a length $\len$, the maximum number of visited tiles will be achieved with one of those pairs, namely that in the diagonal line furthest from $(2a,2b)$ that is within the circle with radius $\len$ centered at that point.
-
-As a second example, consider $a=1.35$, $b=1$. This is depicted in Figure~\ref{fig: ijLS_1p35}. In this case the minimal sufficient set does not follow a rule as simple as in the previous example.
+
+For a given $\len \in \mathbb R^+$, the $(i,j)$ pairs that can be achieved with segments of length not exceeding $\len$ are, by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, those that satisfy \eqref{eq: len > sqrt}.
+Thus for a given value of $i+j-1$ the optimal $(i,j)$ pair (the one that can be achieved with the smallest possible length) is that with minimum distance to the point with normalized coordinates $(2,2)$.
+Such pairs are depicted as filled circles in the figure.
+The displayed arcs represent the inequality \eqref{eq: len > sqrt} for each of the optimal pairs.
+Conversely, given a length $\len$, the maximum number of visited tiles will be achieved with one of these pairs, namely that situated on the highest (or rightmost) diagonal line that still lies within the circle with radius $\len$ centered at normalized coordinates $(2,2)$.
+
+The optimal pairs in this specific case have $j=i$ or $j=i-1$, as seen in the figure.
+Due to symmetry, any pair $(i,i-1)$ could be replaced by $(i-1,i)$.
+This illustrates that the set of optimal pairs is not unique in general.
+
+As a second example, consider $a=1.35$, $b=1$.
+This is depicted in Figure~\ref{fig: ijLS_1p35}.
+Again, the optimal pair $(i,j)$ for each diagonal is that closest to $(2,2)$; but in this case the $i,j$ coordinates of these pairs do not follow a rule as simple as in the previous example.
 
@@
 
-The following proposition gives an explicit method to obtain a minimal sufficient set, $\mss = \{(i_3,j_3), (i_4.j_4), \ldots\}$, where the pair $(i_\tiles,j_\tiles)$ corresponds to $i+j-1 = \tiles$.
+The following proposition gives an explicit method to obtain a set of optimal $(i,j)$ pairs.
+This set will be denoted as $\mss = \{(i_3,j_3), (i_4.j_4), \ldots\}$, where the pair $(i_\tiles,j_\tiles)$ corresponds to $i+j-1 = \tiles$.
 
@@
 \label{prop: min suff set, form}
-Given $a, b \in \mathbb R^+$ and $\tiles \in \mathbb N$, $\tiles \geq 3$, a minimal sufficient set $\mss = \{(i_3,j_3), (i_4.j_4), \ldots\}$ can be obtained as
+Given $a, b \in \mathbb R^+$ and $\tiles \in \mathbb N$, $\tiles \geq 3$, a set of optimal pairs $\mss = \{(i_3,j_3), (i_4.j_4), \ldots\}$ can be obtained as
 \begin{align}
@@
 \begin{proof}
-For each $\tiles$, the pair $(i_\tiles,j_\tiles)$ of the minimal sufficient set should be chosen as that on the line $i+j-1=\tiles$ which minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$.
+For each $\tiles$, the pair $(i_\tiles,j_\tiles)$ should be chosen as that on the line $i+j-1=\tiles$ which minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$.
 This allows maximizing the sum $i+j-1$, and thus the number of visited tiles, for a given length restriction (direct problem); or minimizing the required lengths for a specified number of visited files (inverse problem).
@@
 
-% ***This figure and Figure~\ref{fig: ijLS_1p35_detail} label the points as (i,j) instead of (ia,jb). Clarify in the text, or change the figures
+% ***This figure and Figure~\ref{fig: ijLS_1p35_detail} label the points as (i,j) instead of (ia,jb). Clarify in the text, or change the figures ***Maybe use the concept of normalized coordinates
 \begin{figure}%
@@
 }%
-\caption{Minimal sufficient set $\mss$ and bounding lines
+\caption{Set of optimal pairs $\mss$, and bounding lines
 }%
@@
 \begin{proof}
-The $(i_\tiles,j_\tiles)$ pair in the minimal sufficient set $\mss$ corresponds to a maximum of $\tiles$ visited tiles.
+The $(i_\tiles,j_\tiles)$ pair in $\mss$ corresponds to a maximum of $\tiles$ visited tiles.
 By construction of this set, any segment that visits $\tiles$ tiles must have length greater than $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$.
@@
 \begin{proof}
-The proof of \eqref{eq: theo: funt, sq, form; funt, with i, j} uses a variation of the minimal sufficient set $\mss$ defined in Proposition~\ref{prop: min suff set, form} that is more suited to this situation.
-% ***``situation'' or ``case''?
+The proof of \eqref{eq: theo: funt, sq, form; funt, with i, j} uses a variation of the set $\mss$ defined in Proposition~\ref{prop: min suff set, form} that is more suited to this situation.
 
 For $a=b=1$, the set $\mss$ consists of points of the form $(i,i)$ and $(i,i-1)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}.
-By symmetry, replacing each point $(i,i-1)$ by $(i-1,i)$ gives a set $\mss'$ which is also minimal sufficient.
+By symmetry, replacing each point $(i,i-1)$ by $(i-1,i)$ gives a set $\mss'$ which is also optimal.
 For this new set, the lower bounding line \eqref{eq: lower bound, line} can be replaced by the simpler $j=i$.
@@
 
-The probability of visiting the maximum number of tiles for a rectangular grid is difficult to compute, due to the irregularity of the minimal sufficient set (see \S\ref{part: fund results}).
+The probability of visiting the maximum number of tiles for a rectangular grid is difficult to compute, due to the irregularity of the optimal $(i,j)$ pairs (see \S\ref{part: fund results}).
 In the square case, however, the simplicity of the set makes the problem more tractable.

### 2021-10-22 paper.tex
Rewording

@@
 This is illustrated in Figure~\ref{fig: ijLS_1}.
-Note that in this and in the next figures the axes represent $ia$ and $jb$ (not $i$ and $j$).
-It is convenient, however, to refer to the point $(ia,jb)$ by its normalized coordinates $(i,j)$.
+Note that in this and in the next figures the axes represent $ia$ and $jb$, rather than normalized coordinates $i$ and $j$.
 In this graph, each dashed diagonal line contains pairs $(i,j)$ with the same $i+j-1$; and the lower bound \eqref{eq: len > sqrt} corresponds to an arc centered at normalized coordinates $(2,2)$.

### 2021-10-29 paper.tex
Cleaning
of repeated or obsolete things
Rewriting to avoid inconsistency of coordinates
In some parts of the text and figures the "actual" coordinates (ia, jb) were used, whereas in others the normalized coordinates (i,j) were used. They were even mixed in the same sentence of figure. This has been unified in the text for consistency. The actual coordinates are now used. The choice had to be this, otherwise the intuition of Euclidean distance was lost. Pending: correct labels in Figures 6 and 8 according to this

@@
 
-%  \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem. Some references: \cite{Ramaley69}, \cite{Arnow94}, \cite{Mathai99}. 
-% Alex: Mentioned Ramaley69 and Arnow94, not sure where Mathai99 fits in
-%\url{https://www.jstor.org/stable/2317945}
-%\url{https://www.jstor.org/stable/2687085}
-
 The rest of the paper is organized as follows.
@@
 
-Consider $a=b=1$ first.
-This is illustrated in Figure~\ref{fig: ijLS_1}.
-Note that in this and in the next figures the axes represent $ia$ and $jb$, rather than normalized coordinates $i$ and $j$.
-In this graph, each dashed diagonal line contains pairs $(i,j)$ with the same $i+j-1$; and the lower bound \eqref{eq: len > sqrt} corresponds to an arc centered at normalized coordinates $(2,2)$.
+Consider $a=b=1$ first. This is illustrated in Figure~\ref{fig: ijLS_1}. Note that in this and in the next figures the axes represent $ia$ and $jb$, rather than
+% normalized coordinates
+$i$ and $j$. In this graph, each dashed diagonal line contains pairs $(ia,jb)$ with the same $i+j-1$; and the lower bound \eqref{eq: len > sqrt} corresponds to an arc centered at $(2a,2b)$.
 
@@
 For a given $\len \in \mathbb R^+$, the $(i,j)$ pairs that can be achieved with segments of length not exceeding $\len$ are, by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, those that satisfy \eqref{eq: len > sqrt}.
-Thus for a given value of $i+j-1$ the optimal $(i,j)$ pair (the one that can be achieved with the smallest possible length) is that with minimum distance to the point with normalized coordinates $(2,2)$.
+Thus for a given value of $i+j-1$ the optimal $(i,j)$ pair (the one that can be achieved with the smallest possible length) is determined by the condition that the point $(ia,jb)$ minimizes the distance to $(2a,2b)$.
 Such pairs are depicted as filled circles in the figure.
 The displayed arcs represent the inequality \eqref{eq: len > sqrt} for each of the optimal pairs.
-Conversely, given a length $\len$, the maximum number of visited tiles will be achieved with one of these pairs, namely that situated on the highest (or rightmost) diagonal line that still lies within the circle with radius $\len$ centered at normalized coordinates $(2,2)$.
+Conversely, given a length $\len$, the maximum number of visited tiles will be achieved with one of these pairs, namely the pair $(i,j)$ such that $(ia,jb)$ is on the highest (or rightmost) diagonal line while still being contained in the circle with radius $\len$ centered at $(2a,2b)$.
+% ***rephrase this better?
 
@@
 This is depicted in Figure~\ref{fig: ijLS_1p35}.
-Again, the optimal pair $(i,j)$ for each diagonal is that closest to $(2,2)$; but in this case the $i,j$ coordinates of these pairs do not follow a rule as simple as in the previous example.
+Again, the optimal pair $(i,j)$ for each diagonal is that for which the point $(ia,jb)$ is closest to $(2a,2b)$; but in this case the $i$, $j$ coordinates of these pairs do not follow a rule as simple as in the previous example.
 
@@
 and $(i-2)^2 a^2 + (j-2)^2 b^2$ is expressed as $(x-2a)^2+(y-2b)^2$.
-The point minimizing this quadratic function along \eqref{eq: line 1} is the intersection of this line with the perpendicular line passing through $(2a,2b)$, that is,
+The point minimizing this quadratic function along \eqref{eq: line 1} is the intersection of that line with the perpendicular line passing through $(2a,2b)$, that is,
 \begin{equation}
@@
 \end{align}
-This point is shown in Figure~\ref{fig: min suff set, form} with a square marker.
-
-% ***This figure and Figure~\ref{fig: ijLS_1p35_detail} label the points as (i,j) instead of (ia,jb). Clarify in the text, or change the figures ***Maybe use the concept of normalized coordinates
+The corresponding point $(\isolr a, \jsolr b)$ is shown in Figure~\ref{fig: min suff set, form} with a square marker.
+
+% ***Correct the two labels in the figure
 \begin{figure}%
@@
 The variables $i$, $j$ are actually limited to integer values.
-The point $(i_\tiles, j_\tiles)$ that minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$ along the line $i+j-1=\tiles$ restricted to integer values is either $(\lceil \isolr \rceil, \lfloor \jsolr \rfloor)$ or $(\lfloor \isolr \rfloor, \lceil \jsolr \rceil)$, whichever is closest to $(\isolr, \jsolr)$, as can be seen in Figure~\ref{fig: min suff set, form}.
+The pair $(i_\tiles, j_\tiles)$ that minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$ along the line $i+j-1=\tiles$ restricted to integer values is either $(\lceil \isolr \rceil, \lfloor \jsolr \rfloor)$ or $(\lfloor \isolr \rfloor, \lceil \jsolr \rceil)$, whichever gives $(i_\tiles a, j_\tiles b)$ closest to $(\isolr a, \jsolr b)$, as illustrated in Figure~\ref{fig: min suff set, form}.
 This can be expressed as
@@
 \end{align}
-which in case of tie takes the first of the two points indicated above. Combining \eqref{eq: prop: min suff set, form: isolr}--\eqref{eq: prop: min suff set, form: jsoli}
+which in case of tie takes the first of the two pairs indicated above. Combining \eqref{eq: prop: min suff set, form: isolr}--\eqref{eq: prop: min suff set, form: jsoli}
 % *!*
@@
 Considering the first inequality in \eqref{eq: prop: min suff set, form: isoli ineq} and the second in \eqref{eq: prop: min suff set, form: jsoli ineq} as equalities and eliminating $\tiles$ gives \eqref{eq: upper bound, line}.
-The point $(i_\tiles, j_\tiles)$ is strictly below that line because the considered inequalities are strict.
+The pair $(i_\tiles, j_\tiles)$ is strictly below the line \eqref{eq: upper bound, line} because the used inequalities are strict.
 Similarly, the bounding line \eqref{eq: lower bound, line} results from the second inequality in \eqref{eq: prop: min suff set, form: isoli ineq} and the first in \eqref{eq: prop: min suff set, form: jsoli ineq}, and the fact that both inequalities are not strict implies that the bound \eqref{eq: lower bound, line} can actually be attained.
@@
 
-For $a^2/b^2$ arbitrary, the number of pairs with the same $i$ coordinate, or with the same $j$, in the set $\mss$ is in general irregular, because the lines \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} do not follow a ``natural'' direction of the grid.
-This happens for instance in Figure~\ref{fig: pairs_bounds_1p35}, where the number of pairs for the first $i$ values equals either $2$ or $3$ without a clear pattern.\footnote{Strictly, there is a periodic pattern whenever $a^2/b^2$ is rational, which is the case in Figure~\ref{fig: pairs_bounds_1p35}.
-However, this is not easily discernible unless $a^2/b^2$ is a ratio of small numbers.}
+For $a^2/b^2$ arbitrary, the number of pairs in $\mss$ with the same $i$ coordinate, or with the same $j$, is in general irregular, because the lines \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} do not follow a ``natural'' direction of the grid.
+This happens for instance in Figure~\ref{fig: pairs_bounds_1p35}, where the number of pairs for each $i$ equals either $2$ or $3$ without a clear pattern.\footnote{Strictly, there is a periodic pattern whenever $a^2/b^2$ is rational, which is the case in Figure~\ref{fig: pairs_bounds_1p35}.
+However, it is not easily discernible unless $a^2/b^2$ is a ratio of small numbers.}
 On the other hand, a simple pattern arises when $a^2/b^2$ or $b^2/a^2$ is a natural number, as seen in Figures~\ref{fig: pairs_bounds_1} and \ref{fig: pairs_bounds_sqrt2}.
 
-A segment whose discrete bounding rectangle has normalized width $i$ and height $j$ is oriented with approximate slope $jb/(ia)$ with respect to the $x$ axis (see Figure~\ref{fig: len ineq i j}); and the approximation becomes better for greater segment lengths.
+A segment whose discrete bounding rectangle has normalized width $i$ and height $j$ is oriented with approximate slope $jb/(ia)$ with respect to the $x$ axis (see Figure~\ref{fig: len ineq i j}); and this approximation becomes better for greater segment lengths.
 From \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} it can be seen that the positions of the pairs $(i,j) \in \mss$ have $j/i \approx a^2/b^2$ for large $i, j$.
@@
 The approach is similar to that used in Proposition~\ref{prop: min suff set, form}.
-Namely, for $i, j \geq 2$ the intersection $(\isolr, \jsolr)$ of the line \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ is found, and from that the actual integer solution $(\isoli, \jsoli)$ is computed.
+Namely, for $i, j \geq 2$ the intersection $(\isolr a, \jsolr b)$ of the line defined by \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ is found, and from that the actual integer solution $(\isoli, \jsoli)$ is computed.
 
@@
 Then, assuming that $i = \isoli$, the largest $j$ allowed by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, $\jsoli$, is obtained.
-As will be seen, in some cases the resulting $(\isoli,\jsoli)$ is in $\mss$, and in other cases it is not.
+As will be seen, in some cases the resulting $(\isoli,\jsoli)$ is in $\mss$ as defined by Proposition~\ref{prop: min suff set, form}, and in other cases it is not.
 However, in either case $(\isoli,\jsoli)$ has the largest $i+j-1$ sum that can be achieved with segments of length up to $\len$.
@@
 In each case, the arc displayed in the graph is centered at $(2a,2b)$ and has radius $\len$.
-The inner region defined by the arc contains all $(i,j)$ pairs that are achievable according to Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}.
-As in previous figures, filled circles represent $(i,j)$ pairs that are in $\mss$, and empty circles are those that do not belong to $\mss$.
-The solid line is \eqref{eq: lower bound, line}.
-The intersection point $(\isolr, \jsolr)$ is displayed with a square marker.
-
+The inner region defined by the arc contains all $(ia,jb)$ points such that the pair $(i,j)$ is achievable according to Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}.
+As in previous figures, filled circles represent points $(ia,jb)$ such that $(i,j) \in \mss$.
+The solid line is the bound \eqref{eq: lower bound, line}.
+The intersection point $(\isolr a, \jsolr b)$ is displayed with a square marker.
+
+% ***Correct two labels in each of the two graphs
 \begin{figure}
@@
 
-The point $(\isolr, \jsolr)$ can be obtained as a solution of the equation system
+The pair $(\isolr, \jsolr)$ can be obtained as a solution of the equation system
 \begin{align}
@@
 \end{equation}
-which can have zero, one or two real-valued solutions (this respectively means that the line is exterior, tangent or secant to the circle \eqref{eq: icont jcont 1}).
+which can have zero, one or two real-valued solutions (this respectively means that the line \eqref{eq: icont jcont 2} is exterior, tangent or secant to the circle \eqref{eq: icont jcont 1}).
 Only the solution with $\isolr, \jsolr \geq 2$, if any, is of interest.
@@
 This gives \eqref{eq: theo: funt, form; i}.
-Once $\isoli$ is known, $\jsoli$ is computed, according to \eqref{eq: theo: funt, form; j}, as the greatest integer such that $(\isoli,\jsoli)$ is within the arc determined by $\len$.
+Once $\isoli$ is known, $\jsoli$ is computed, according to \eqref{eq: theo: funt, form; j}, as the greatest integer such that $(\isoli a,\jsoli b)$ is within the circle determined by $\len$.
 This ensures that $(\isoli,\jsoli)$ is achievable with lengths less than $\len$.
 
-There are two possibilities for the point $(\isoli,\jsoli)$, as stated at the outset.
+There are two possibilities for the pair $(\isoli,\jsoli)$, as stated at the outset.
 These are illustrated in Figures~\ref{fig: ijLS_1p35_detail_in} and \ref{fig: ijLS_1p35_detail_notin} respectively.
@@
 The second possibility is that $(\isoli,\jsoli) \notin \mss$ (Figure~\ref{fig: ijLS_1p35_detail_notin}).
-This happens when the point from $\mss$ that has $i = \isoli$ ($(4,5)$ in the figure) is outside the arc, i.e.~it would require a length greater than $\len$.
-The selected point  $(\isoli,\jsoli)$ ($(4,4)$ in the figure), however, has the same $i+j-1$ sum as the point from $\mss$ that ``should'' be used, which is $(\isoli-1,\jsoli+1)$ ($(3,5)$ in the figure); and therefore gives the same result.
+This happens when the pair from $\mss$ that has $i = \isoli$ ($(4,5)$ in the figure) is outside the circle, i.e.~it would require a length greater than $\len$.
+The selected $(\isoli,\jsoli)$ ($(4,4)$ in the figure), however, has the same $i+j-1$ sum as the pair from $\mss$ that ``should'' be used, which is $(\isoli-1,\jsoli+1)$ ($(3,5)$ in the figure); and therefore gives the same result.
 This is always the case, because  $(\isoli,\jsoli+1) \in \mss$ (it is above or on the bounding line) and $(\isoli,\jsoli) \notin \mss$ (it is below the line), and due to how $\mss$ has been constructed, this implies that $(\isoli-1,\jsoli+1) \in \mss$ and $(\isoli-1,\jsoli+k) \notin \mss$ for $k =2, 3, \ldots$.
@@
 
-Therefore, whether $(\isoli,\jsoli) \in \mss$ or $(\isoli,\jsoli) \notin \mss$, equations \eqref{eq: theo: funt, form; i} and \eqref{eq: theo: funt, form; j} give $\isoli+\jsoli-1$ equal to $\funt(\len)$, as claimed.
+Therefore, regardless of whether $(\isoli,\jsoli)$ is in $\mss$ or not, \eqref{eq: theo: funt, form; i} and \eqref{eq: theo: funt, form; j} give $\isoli+\jsoli-1$ equal to $\funt(\len)$, as claimed.
 

### 2021-11-04 paper.tex
Interpretation of floor and ceiling as rounding with tiebreaking in opposite directions; and interpretation of the two terms in the new formula for funl

@@
 
-% ***Correct the two labels in the figure
 \begin{figure}%
@@
 The pair $(i_\tiles, j_\tiles)$ that minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$ along the line $i+j-1=\tiles$ restricted to integer values is either $(\lceil \isolr \rceil, \lfloor \jsolr \rfloor)$ or $(\lfloor \isolr \rfloor, \lceil \jsolr \rceil)$, whichever gives $(i_\tiles a, j_\tiles b)$ closest to $(\isolr a, \jsolr b)$, as illustrated in Figure~\ref{fig: min suff set, form}.
+In case of a tie the first of the two pairs is chosen.
 This can be expressed as
@@
 \end{align}
-which in case of tie takes the first of the two pairs indicated above. Combining \eqref{eq: prop: min suff set, form: isolr}--\eqref{eq: prop: min suff set, form: jsoli}
+which corresponds to rounding $\isolr$ and $\jsolr$ to the closest integers, with ties resolved in opposite directions.
+Combining \eqref{eq: prop: min suff set, form: isolr}--\eqref{eq: prop: min suff set, form: jsoli}
 % *!*
@@
 
-***Comment on how the new formula has two terms with definite meanings: The first term is the distance-squared to the closest point on the diagonal, The second term is additional distance-squared incurred from rounding (i,j) to integers
-***I think we can do without a figure (Figure~\ref{fig: min suff set, form} is inside a proof and uses different notation for the final point).
+The expression \eqref{eq: theo: funl, form; funl 2} allows a neat interpretation of $\funl(\tiles)$ (as stems from the arguments used in the proof).
+Namely, $\funl^2(\tiles)$ is the sum of the two terms within the square root in that expression.
+The first term is the squared distance from $(2a, 2b)$ to the diagonal line defined by $i+j-1=t$; and the second term is additional squared distance incurred from rounding $i$, $j$ to integer values.
 

### 2021-11-04 paper.tex
Style uniformity

@@
 \newcommand{\funta}{\varphi} % function that gives average number of touched tiles for real-valued length
+\newcommand{\funca}{\psi} % function that gives average number of grid line crossings for real-valued length
 \newcommand{\probmax}{\rho} % function that gives the probability of achieving the maximum number of touched tiles in the square case for real-valued length
@@
 Solving the problem of how many tiles the segment visits on average, as will be seen, also answers the inverse question (segment length to visit a given number of tiles on average).
-A natural, related question is with that probability the segment visits the maximum number of tiles.
+A natural, related question is with what probability the segment visits the maximum number of tiles.
 
@@
 The problems studied in this paper are related to Buffon's needle problem and Laplace's extension of it.
-Buffon's original 1733 problem considers tossing a needle of length $\len$ onto ruled paper with parallel lines a height $b$ apart, and asks what is the probability of the needle crossing a line if it lands in a uniformly random position and orientation.
+Buffon's original problem considers tossing a needle of length $\len$ onto ruled paper with parallel lines a distance $b$ apart, and asks what is the probability of the needle crossing a line if it lands in a uniformly random position and orientation.
 For a needle of length $\len = b$, this is $2/\pi$, and hence repeated trials of this experiment can be used to estimate $\pi$.
-This is generalized in \cite{Ramaley69} to $\frac{2\len}{\pi b}$ for the expected number of crossings of a needle which might intersect multiple lines.
-
-The Buffon-Laplace needle problem, described in \cite{Mathai99}, considers a needle tossed onto a grid of rectangular tiles of width $a$ and height $b$.
-The number of visited tiles equals one plus the number of crossings with probability 1, as shown in Proposition \ref{prop:
-i+j-1}.
+This is generalized in \cite{Ramaley69} to $2\len /(\pi b)$ for the expected number of crossings of a needle with arbitrary length.
+
+The Buffon-Laplace needle problem \cite{Mathai99} considers a needle tossed onto a grid of rectangular tiles of width $a$ and height $b$.
+The number of visited tiles equals one plus the number of crossings almost surely.
 The probability of the needle staying within a single rectangle, that is visiting only one tile, is computed in \cite{Arnow94} for the case where $\len < \min(a,b)$.
-This work considers the complementary question of the chance that the needle visits the maximum number of tiles possible for its length, described by a formula in Theorem~\ref{theo: probmax, sq}.
+One of the results in this paper, as has been mentioned, is the complementary question of the probability that the needle visits the maximum number of tiles possible for its length.
 
@@
 \begin{proof}
-First, suppose the grid is simplified to just horizontal lines a height of $b$ apart.
-This matches the setup of the 1733 Buffon's needle problem described in \cite{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines.
-As shown in \cite{Ramaley69}, the expected (i.e., average) number of lines crossed equals
-$$ \frac{2\len}{\pi b}.$$
-Where $b<\len$, this expectation matches the probability of the needle crossing a grid line, since it cannot cross more than one.
-
-Now consider again a grid of width $a$ and height $b$, as studied by Laplace as an extension to Buffon's original needle problem \cite{Arnow94}.
-The grid crossings decompose into crossings of horizontal and vertical grid lines, respectively.
-By linearity of expectation, the expected number of crossings is the sum of the respective expectations for parallel lines of width $a$ and $b$ respectively, giving:
-\[
-\frac{2\len}{\pi}\left(\frac 1 a + \frac 1 b\right)
-\]
-As noted in Proposition \ref{prop: i+j-1}, the number of tiles visited by a segment is the count of its grid line crossings plus $1$, unless it exactly crosses any grid points, but again that occurs with probability $0$ and therefore it does not affect the expectation.
-Thus, $\funta(\len)$ is the above quantity plus $1$.
+Suppose first that the grid is simplified to only horizontal lines with spacing $b$.
+This matches the setup of Buffon's original needle problem \cite{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines.
+As shown in \cite{Ramaley69}, the expected number of lines crossed equals $2\len/(\pi b)$.
+% When $\len<b$ this expectation matches the probability of the needle crossing a grid line, since it cannot cross more than one.
+
+Consider again a grid with horizontal spacing $a$ and vertical spacing $b$.
+The grid crossings decompose into crossings of horizontal and vertical grid lines.
+By linearity of expectation, the expected number of crossings $\funca(\len)$ is the sum of the expectations for parallel lines with spacing $a$ and $b$ respectively, which gives
+\begin{equation}
+\label{eq: funca}
+\funca(\len) = \frac{2\len}{\pi}\left(\frac 1 a + \frac 1 b\right).
+\end{equation}
+As noted in Proposition \ref{prop: i+j-1}, the number of tiles visited by a segment is the count of its grid line crossings plus $1$, unless it exactly passes through any grid points, but that occurs with probability $0$ and therefore does not affect the expectation.
+Thus, $\funta(\len)$ is given by the right-hand side of \eqref{eq: funca} plus $1$.
 \end{proof}

### 2021-11-07 paper.tex
Some more items from the "to do" list

@@
 
+\begin{figure}%
+\centering%
+\includegraphics[width=.75\textwidth]{probmax_len}%
+\caption{Probability $\probmax(\len)$ that a random segment of length $\len$ visits the maximum number of tiles%
+}%
+\label{fig: probmax_len}%
+\end{figure}%
+
 Figure~\ref{fig: probmax_len} shows the probability $\probmax(\len)$ of visiting the maximum number of tiles on a unit square grid, as given by \eqref{eq: probmax}.
-As $\len$ grows, $\probmax(\len)$ has a jump when the maximum number of tiles that can be visited increases by $1$.
+As $\len$ grows, $\probmax(\len)$ has a jump when the maximum number of tiles that can be visited increases by $1$ (and there the function takes the value approached from the left).
 This happens when $\len$ equals the right endpoint of the interval \eqref{eq: square, odd max tiles, lengths} for $\tiles$ odd or of \eqref{eq: square, even max tiles, lengths} for $\tiles$ even.
@@
 \end{equation}
-For example, the first continuous section seen in Figure~\ref{fig: probmax_len} corresponds to a maximum number of tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, with $\len_3= 1$.
-The second corresponds to $\tiles=4$, for lengths in $(\len_3, \len_4]$, with $\len_4= \sqrt{2}$.
+For example, the first continuous section seen in Figure~\ref{fig: probmax_len} corresponds to a maximum number of tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, where $\len_3= 1$.
+The second corresponds to $\tiles=4$, for lengths in $(\len_3, \len_4]$, where $\len_4= \sqrt{2}$.
 Within each continuous section the probability monotonically increases from $0$ to a maximum value.
-The heights of the maxima follow a different trend**better word?** for odd and even $\tiles$, as can be seen in the figure; but in each case they are asymptotically proportional to $1/\len_\tiles$.
+The heights of the maxima follow different trends for odd and even $\tiles$ respectively, as can be seen in the figure; but in each case they are asymptotically proportional to $1/\len_\tiles$.
 This is established by the next result.
-
-\begin{figure}%
-\centering%
-\includegraphics[width=.75\textwidth]{probmax_len}%
-\caption{Probability $\probmax(\len)$ that a random segment of length $\len$ visits the maximum number of tiles%
-}%
-\label{fig: probmax_len}%
-\end{figure}%
 
@@
 \end{align}
-as can be checked using L'H\^opital's rule \cite[section~5.3]{Abbott15}.
+as can be (somewhat tediously) checked using L'H\^opital's rule \cite[section~5.3]{Abbott15}.
 %***I think.

### 2021-11-12 paper.tex
Graph and discussion of ratio of asymptotic slopes

@@
 \newcommand{\probmax}{\rho} % function that gives the probability of achieving the maximum number of touched tiles in the square case for real-valued length
+\newcommand{\ras}{\sigma}
 \newcommand{\len}{\ell} % length variable, real-valued
@@
 The displayed arcs represent the inequality \eqref{eq: len > sqrt} for each of the optimal pairs.
-Conversely, given a length $\len$, the maximum number of visited tiles will be achieved with one of these pairs, namely the pair $(i,j)$ such that $(ia,jb)$ is on the highest (or rightmost) diagonal line while still being contained in the circle with radius $\len$ centered at $(2a,2b)$.
-% ***rephrase this better?
+Conversely, given a length $\len$, the maximum number of visited tiles will be achieved with one of these pairs, namely the pair $(i,j)$ such that $(ia,jb)$ is on the highest (or rightmost) diagonal line while still being contained in the circle of radius $\len$ centered at $(2a,2b)$.
 
@@
 
-% ***Correct two labels in each of the two graphs
 \begin{figure}
@@
 \begin{equation}
-\label{eq: asympt}
+\label{eq: asympt slope}
 \lim_{\len \rightarrow \infty} \frac{\funt(\len)}{\len}
@@
 As for $a \gg b$, in this case the optimal discrete bounding rectangle has $\isoli = 2$, and $\jsoli$ as large as allowed by $\len$ (direct problem) or as required by $\tiles$ (inverse problem), corresponding to an almost vertical segment.
-Obviously, for $a \gg b$ the length of the segment is best invested in increasing the number of tiles traversed vertically (but the segment should be slightly tilted to cross a vertical edge), and the asymptotic value of \eqref{eq: asympt} is approximately $1/b$.
+Obviously, for $a \gg b$ the length of the segment is best invested in increasing the number of tiles traversed vertically (but the segment should be slightly tilted to cross a vertical edge), and the asymptotic value of \eqref{eq: asympt slope} is approximately $1/b$.
 
@@
 
-***Graph of asymptotic slopes: $\sqrt(1/a^2+1/b^2)$ for $\funt$ and $2/\pi(1/a+1/b)$.
-It may be interesting to consider the ratio of asymptotic slopes, as a function of $a/b$.
-It is maximum for $a=b$, and equals $2\sqrt 2/\pi = 0.9003$.
-It tends to $2/\pi$ for $a/b \rightarrow \infty$ and for $a/b \rightarrow 0$.
+For $\len \rightarrow \infty$, the average number of visited tiles $\funta$ has the following asymptotic slope, as stems from \eqref{eq: funta, form}:
+\begin{equation}
+\label{eq: asympt slope ave}
+\lim_{\len \rightarrow \infty} \frac{\funta(\len)}{\len} = \frac 2 \pi \left(\frac 1 a + \frac 1 b \right).
+\end{equation}
+On the other hand, the maximum number of visited tiles $\funt$ has an asymptotic slope given by \eqref{eq: asympt slope}.
+It is interesting to consider the ratio of these values, which also gives the asymptotic ratio of average and maximum numbers of visited tiles.
+This depends only on $a/b$:
+\begin{equation}
+\label{eq: ras}
+\ras(a/b) = \lim_{\len \rightarrow \infty}
+\frac { \funta(\len) } { \funt(\len) }
+= \frac
+{ \lim_{\len \rightarrow \infty} \funta(\len)/\len }
+{ \lim_{\len \rightarrow \infty} \funt(\len)/\len }
+= \frac{2\left(1 + a/b\right)}{\pi\sqrt{1 + \left(a/b\right)^2}}.
+\end{equation}
+The function $\ras$ satisfies $\ras(x) = \ras(1/x)$, and is therefore symmetric in logarithmic scale, as can be seen in Figure~\ref{fig: ratio asympt slopes}.
+It is unimodal with maximum value $2\sqrt 2/\pi \approx 0.9003$ for $a/b=1$, and its limit when $a/b$ tends to $0$ or $\infty$ is $2/\pi \approx 0.6366$.
+This implies that for $\len$ large the average number of tiles $\funta(\len)$ cannot very small compared with the maximum number of tiles $\funt(\len)$.
+For example, $\funta(\len)$ asymptotically exceeds $0.8\funt(\len)$ for $1/3 \leq a/b \leq 3$.
+
+\begin{figure}
+\centering%
+\includegraphics[width=.7\textwidth]{ratio_asympt_slopes}%
+\caption{Asymptotic ratio of average and maximum numbers of visited tiles
+}%
+\label{fig: ratio asympt slopes}%
+\end{figure}%
 

### 2021-11-12 figures/figures.m

@@ -1903,3 +1903,29 @@ jstar = ceil(jplus-1/2);
 lambda = sqrt((istar-2)^2*a^2 + (jstar-2)^2*b^2) % original formula
 %lambda_2 = sqrt( (t-3)^2*a^2*b^2/(a^2+b^2) + (istar-iplus)^2*a^2 + (jstar-jplus)^2*b^2 ) % new formula, corrected
 lambda_2 = sqrt( (t-3)^2*a^2*b^2/(a^2+b^2) + (istar-iplus)^2*(a^2+b^2) ) % new formula, corrected
+
+
+%% #72. Ratio of asymptotic slopes (with respect to length) for maximum and average numbers of visited tiles, as a function of a/b
+
+clear, clc, close all
+ratio_ab = logspace(-2,2,300);
+as_ratio_slopes = 2/pi*(1+ratio_ab)./hypot(1, ratio_ab);
+plot(ratio_ab, as_ratio_slopes)
+set(gca, 'xscale', 'log'), grid on, xticklabels(xticks), xlabel('a/b'), ylabel([char(hex2dec('03C3')) '(a/b)']), set(gcf, 'Position', [360 300 560 310])
+
+
+%% #73. Maximum and average numbers of visited tiles as a function of length, with (a,b) as a parameter.
+% Part taken from #42
+
+clear, clc, close all
+L = .01:.01:200;
+a = 2.35; b = 1;
+ii = ceil((a + b*real(sqrt(4*L.^2/(a^2+b^2)-1))) / 2/a) + 1;
+jj = ceil(sqrt( L.^2 - (ii-2).^2*a^2 )/b) + 1;
+funt = ii+jj-1;
+funta = 2*L/pi*(1/a+1/b)+1;
+as_ratio_slopes = 2/pi*(1+a/b)./hypot(1, a/b)
+%plot(L, funt), hold on, plot(L, funta);
+plot(funta./funt), hold on, plot(xlim, repmat(as_ratio_slopes, 1, 2))
+
+

### 2021-11-12 paper.tex
Joined two figures as subfigures of a single figure

@@
 On the other hand, the set of optimal $(i,j)$ pairs must contain one such pair for each possible value of $i+j-1$, so that the set can produce that value as the solution (maximum number of visited tiles) for certain lengths.
-It is insightful to examine two specific examples before giving an explicit formula for the coordinates of the optimal pairs.
+It is insightful to examine two specific examples (Figure~\ref{fig: ijLS}) before giving an explicit formula for the coordinates of the optimal pairs.
 
@@
 \centering%
+\subfigure[$a=b$]{%
+\label{fig: ijLS_1}
 \includegraphics[width=.7\textwidth]{ijLS_1}%
-\caption{Relationship of segment length and number of visited tiles with the width and height of the discrete bounding rectangle, for $a=b$
-}%
-\label{fig: ijLS_1}%
+}\\%
+\centering%
+\subfigure[$a=1.35b$]{%
+\label{fig: ijLS_1p35}
+\includegraphics[width=.7\textwidth]{ijLS_1p35}%
+}%
+\caption{Relationship of segment length and number of visited tiles with the width and height of the discrete bounding rectangle
+}%
+\label{fig: ijLS}%
 \end{figure}%
-% ***Maybe combine this figure and the next into one figure with two subfigures (stacked vertically, size similar to now). Not sure. It has been done in others
 
@@
 Again, the optimal pair $(i,j)$ for each diagonal is that for which the point $(ia,jb)$ is closest to $(2a,2b)$; but in this case the $i$, $j$ coordinates of these pairs do not follow a rule as simple as in the previous example.
-
-\begin{figure}
-\centering%
-\includegraphics[width=.7\textwidth]{ijLS_1p35}%
-\caption{Relationship of segment length and number of visited tiles with the width and height of the discrete bounding rectangle, for $a=1.35b$
-}%
-\label{fig: ijLS_1p35}%
-\end{figure}%
 
@@
 as can be (somewhat tediously) checked using L'H\^opital's rule \cite[section~5.3]{Abbott15}.
-%***I think.
-It's very tedious by hand.
+% It's very tedious by hand.
 I used Maple to compute the limit, and then checked it numerically.

### 2021-11-17 paper.tex
Revising of introduction section

@@
 The segment is said to \emph{visit} a tile if it intersects its interior.\footnote{
-The definition uses the interior of the tile, excluding the border, to avoid uninteresting results such as a ``zero-length'' segment visiting (a vertex of) $4$ tiles.} This paper deals with the following problems:
-\begin{itemize}
-\item What is the maximum number of tiles that the segment can visit? Conversely, what length should a segment have to visit a given number of tiles?
-\item What is the average number of tiles visited by a uniformly random segment of a given length? How often does the random segment visit the maximum number of tiles?
-\end{itemize}
-
-As an example of the first question, consider $a=1.35$, $b=1$.
+The definition uses the interior of the tile, excluding the border, to avoid uninteresting results such as a ``zero-length'' segment visiting (a vertex of) $4$ tiles.}
+
+This paper deals with the following problems, which relate the length of the segment with the number of visited tiles.
+Two settings are considered, which correspond respectively to the segment position and orientation being deterministic or random.
+In the \emph{deterministic} case, the relevant questions are:
+\begin{enumerate}
+\renewcommand{\labelenumi}{(1\alph{enumi})}
+\renewcommand{\theenumi}{(1\alph{enumi})}
+\item
+\label{prob: determ dir}
+What is the maximum number of tiles that the segment can visit given its length?
+\item
+\label{prob: determ inv}
+Conversely, what length should a segment have to visit a given number of tiles?
+\end{enumerate}
+
+In the \emph{random} setting, if the segment position and orientation are uniformly distributed (this will be precisely defined later),
+\begin{enumerate}
+\renewcommand{\labelenumi}{(2\alph{enumi})}
+\renewcommand{\theenumi}{(2\alph{enumi})}
+\item
+\label{prob: rand ave}
+What is the average number of tiles visited by a segment of a given length?
+\item
+\label{prob: rand prob}
+How often does the random segment visit the maximum number of tiles?
+\end{enumerate}
+
+As an example of the first question \ref{prob: determ dir}, consider $a=1.35$, $b=1$.
 A segment of unit length can be placed as shown in Figure~\ref{fig: examples} (left) to make it visit $3$ tiles.
@@
 
-The \emph{inverse} problem is, given $\tiles \in \mathbb N$, to determine the infimum length of all segments that visit at least $\tiles$ tiles.
-For real-valued lengths this infimum is not a minimum, because given a segment that visits $\tiles$ tiles, its length can always be reduced by some non-zero amount without changing the number of visited tiles.
+The \emph{inverse} problem \ref{prob: determ inv} is, given $\tiles \in \mathbb N$, to determine the infimum length of all segments that visit at least $\tiles$ tiles.
+If the length can take any real value the infimum is not a minimum, because given any segment can be shortened by some small amount without changing the number of visited tiles.
 This is a consequence of the interior of each tile being an open set.
@@
 The position of this endpoint is \emph{uniformly} distributed on the tile.
-The segment orientation has a \emph{uniform} distribution on the set of all possible directions, and is \emph{independent} of the endpoint position.
-Solving the problem of how many tiles the segment visits on average, as will be seen, also answers the inverse question (segment length to visit a given number of tiles on average).
-A natural, related question is with what probability the segment visits the maximum number of tiles.
+The segment orientation has a \emph{uniform} distribution on $[0,2\pi)$, and is \emph{independent} of the endpoint position.
+Solving the problem of how many tiles the segment visits on average \ref{prob: rand ave} also answers, as will be seen, the inverse question (segment length to visit a given number of tiles on average).
+A natural, related question \ref{prob: rand prob} is with what probability the segment visits the maximum number of tiles.
 
@@
 
-The problems studied in this paper are related to Buffon's needle problem and Laplace's extension of it.
+The questions studied in this paper are related to Buffon's needle problem and Laplace's extension of it.
 Buffon's original problem considers tossing a needle of length $\len$ onto ruled paper with parallel lines a distance $b$ apart, and asks what is the probability of the needle crossing a line if it lands in a uniformly random position and orientation.
-For a needle of length $\len = b$, this is $2/\pi$, and hence repeated trials of this experiment can be used to estimate $\pi$.
+For a needle of length $\len = b$, this is $2/\pi$ (and hence repeated trials of this experiment can be used to estimate $\pi$).
 This is generalized in \cite{Ramaley69} to $2\len /(\pi b)$ for the expected number of crossings of a needle with arbitrary length.
@@
 The number of visited tiles equals one plus the number of crossings almost surely.
-The probability of the needle staying within a single rectangle, that is visiting only one tile, is computed in \cite{Arnow94} for the case where $\len < \min(a,b)$.
-One of the results in this paper, as has been mentioned, is the complementary question of the probability that the needle visits the maximum number of tiles possible for its length.
+The probability of the needle staying within a single tile is computed in \cite{Arnow94} for the case where $\len < \min(a,b)$.
+One of the problems considered in this paper, as mentioned above, is the complementary question \ref{prob: rand prob} of the probability that the needle visits the maximum number of tiles possible for its length.
 
@@
 
-\section{Direct and inverse problems for a deterministic segment}
+\section{Deterministic segment: direct and inverse problems}
 \label{part: max}
@@
 
-\section{Probabilistic characterization for a random segment}
+\section{Random segment: probabilistic characterization}
 \label{part: rand}

### 2021-11-25 paper.tex
Rewriting

@@
 %Avenida Complutense, 30. 28040 Madrid, Spain.\\
-\affil[2]{\small **Affiliation, e-mail**}
+\affil[2]{\small \texttt{arkhipov@alum.mit.edu}}
 
@@
 
-For a grid with horizontal spacing $a$ and vertical spacing $b$, lines of the form $x = ka$ or $y = kb$ with $k \in \mathbb Z$ will be called \emph{grid lines}.
+For a grid with horizontal spacing $a$ and vertical spacing $b$, lines $x = ka$ or $y = kb$ with $k \in \mathbb Z$ will be called \emph{grid lines}.
 A \emph{tile} is delimited by two pairs of consecutive horizontal and vertical grid lines.
@@
 Every segment has an associated \emph{discrete bounding rectangle}, which is the minimum-size rectangle that is formed by grid lines and contains the segment.
-More specifically, if the segment has endpoints $(x_1,y_1)$, $(x_2,y_2)$, where $x_1, x_2, y_1, y_2 \in \mathbb R$, its discrete bounding rectangle has lower-left corner and upper-right corner given as
+More specifically, if the segment has endpoints $(x_1,y_1)$, $(x_2,y_2) \in \mathbb R^2$, its discrete bounding rectangle has lower-left and upper-right corners respectively given as
 \begin{align*}
@@
 Two examples are illustrated in Figure~\ref{fig: discrete bounding rectangle and touched tiles}, both with $i=5$, $j=4$.
-Clearly, all tiles visited by the segment are contained in the discrete bounding rectangle.
+All tiles visited by the segment are contained in the discrete bounding rectangle.
 Note also that the rectangle can have $i=0$ or $j=0$ if the segment coincides with part of a grid line.
@@
 For each $(i,j)$, segments are shown with lengths close to either of the two bounds.
-Note that \eqref{eq: len > sqrt} is valid even for $i=2$, $j=2$, in which case it reduces to $\len>0$.
-
-\ref*{prop: len ineq i j: exist} For $a$, $b$, $\len$, $i$, $j$ satisfying the two inequalities, a segment of length $\len$ can be found that has its endpoints in the interiors or on the outer edges of the two shaded tiles of a rectangle of normalized dimensions $i$ and $j$, as in Figure~\ref{fig: len ineq i j}, which is thus the discrete bounding rectangle of that segment.
+Note that inequality \eqref{eq: len > sqrt} is valid even for $i=2$, $j=2$, in which case it reduces to $\len>0$.
+
+\ref*{prop: len ineq i j: exist} For $a$, $b$, $\len$, $i$, $j$ satisfying the two inequalities, a segment of length $\len$ can be found that has its endpoints in the interiors or on the outer edges of the two shaded tiles of a rectangle with normalized dimensions $i$ and $j$, as in Figure~\ref{fig: len ineq i j}, which is thus the discrete bounding rectangle of that segment.
 \begin{figure}
@@
 
-``There is a segment\ldots $\Rightarrow$ \eqref{eq: len > sqrt} '': Assume that a segment exists with length not exceeding $\len$ and with a discrete bounding rectangle with normalized dimensions $i$, $j$.
-From part~\ref{prop: len ineq i j: ineqs}, inequality \eqref{eq: len > sqrt} holds for that segment length, and thus for $\len$.
+``There is a segment\ldots $\Rightarrow$ \eqref{eq: len > sqrt} '': Assume that a segment exists with length $\len' \leq \len$ and with a discrete bounding rectangle of normalized dimensions $i$, $j$.
+From part~\ref{prop: len ineq i j: ineqs}, inequality \eqref{eq: len > sqrt} holds for the length $\len'$, and thus for $\len$.
 \end{proof}
 
-According to Proposition~\ref{prop: i+j-1}, in order to maximize the number of visited tiles for a given length, the position and orientation of the segment should be chosen to obtain $i+j-1$ as large as possible, where $i$ and $j$ are the normalized dimensions of its discrete bounding rectangle.
+Consider the problem of maximizing the number of visited tiles for a given length.
+According to Proposition~\ref{prop: i+j-1}, the position and orientation of the segment should be chosen to obtain $i+j-1$ as large as possible, where $i$ and $j$ are the normalized dimensions of its discrete bounding rectangle.
 On the other hand, Proposition~\ref{prop: len ineq i j} restricts the $i, j$ values that can be achieved with a given length.
-A relevant question is: are there any $(i,j)$ pairs that can be disregarded irrespective of the length $\len$?
+A relevant question is: are there any $(i,j)$ pairs that can be disregarded for all lengths $\len$?
 In other words, what is the ``smallest'' subset of $\mathbb N^2$ such that the $(i,j)$ pair that maximizes the number of tiles for any given length can always be found within that subset?
@@
 
-Consider $a=b=1$ first. This is illustrated in Figure~\ref{fig: ijLS_1}. Note that in this and in the next figures the axes represent $ia$ and $jb$, rather than
-% normalized coordinates
-$i$ and $j$. In this graph, each dashed diagonal line contains pairs $(ia,jb)$ with the same $i+j-1$; and the lower bound \eqref{eq: len > sqrt} corresponds to an arc centered at $(2a,2b)$.
+Consider $a=b=1$ first.
+This is illustrated in Figure~\ref{fig: ijLS_1}.
+Note that in this and in the next figures the axes represent $ia$ and $jb$ (not $i$ and $j$).
+In this graph, each dashed diagonal line contains points $(ia,jb)$ with the same $i+j-1$; and the lower bound \eqref{eq: len > sqrt} corresponds to an arc centered at $(2a,2b)$.
 
@@
 Suppose first that the grid is simplified to only horizontal lines with spacing $b$.
-This matches the setup of Buffon's original needle problem \cite{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines.
+This matches the set-up of Buffon's original needle problem \cite{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines.
 As shown in \cite{Ramaley69}, the expected number of lines crossed equals $2\len/(\pi b)$.

### 2021-11-25 paper.tex
Rewriting of Section 2; math expressions not checked

@@
 
-The optimal pairs in this specific case have $j=i$ or $j=i-1$, as seen in the figure.
+The optimal pairs in this specific case have the form $(i,i)$ or $(i,i-1)$, as seen in the figure.
 Due to symmetry, any pair $(i,i-1)$ could be replaced by $(i-1,i)$.
@@
 For each $\tiles$, the pair $(i_\tiles,j_\tiles)$ should be chosen as that on the line $i+j-1=\tiles$ which minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$.
-This allows maximizing the sum $i+j-1$, and thus the number of visited tiles, for a given length restriction (direct problem); or minimizing the required lengths for a specified number of visited files (inverse problem).
+This allows maximizing the sum $i+j-1$, and thus the number of visited tiles, for a given length restriction; or visiting a specified number of files with lengths as small as possible.
 
@@
 and $(i-2)^2 a^2 + (j-2)^2 b^2$ is expressed as $(x-2a)^2+(y-2b)^2$.
-The point minimizing this quadratic function along \eqref{eq: line 1} is the intersection of that line with the perpendicular line passing through $(2a,2b)$, that is,
+The point minimizing this quadratic function along the line \eqref{eq: line 1} is the intersection of that with the perpendicular line passing through $(2a,2b)$,
 \begin{equation}
@@
 \end{equation}
-This is depicted for $a=1.35$, $b=1$, $\tiles=7$ in Figure~\ref{fig: min suff set, form}, where \eqref{eq: line 2} is shown with solid line.
+An example with $a=1.35$, $b=1$, $\tiles=7$ is depicted in Figure~\ref{fig: min suff set, form}, where \eqref{eq: line 2} is shown with solid line.
 Solving the system of equations \eqref{eq: line 1} and \eqref{eq: line 2} gives
@@
 \end{align}
-In terms of the real-valued variables $i$, $j$, the solution $(\isolr, \jsolr)$ is
+In terms of the real-valued variables $i$, $j$, the solution $(\isolr, \jsolr)$ is thus
 \begin{align}
@@
 The pair $(i_\tiles, j_\tiles)$ that minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$ along the line $i+j-1=\tiles$ restricted to integer values is either $(\lceil \isolr \rceil, \lfloor \jsolr \rfloor)$ or $(\lfloor \isolr \rfloor, \lceil \jsolr \rceil)$, whichever gives $(i_\tiles a, j_\tiles b)$ closest to $(\isolr a, \jsolr b)$, as illustrated in Figure~\ref{fig: min suff set, form}.
-In case of a tie the first of the two pairs is chosen.
+In case of a tie the first of the two pairs is (arbitrarily) chosen.
 This can be expressed as
@@
 The pair $(i_\tiles, j_\tiles)$ is strictly below the line \eqref{eq: upper bound, line} because the used inequalities are strict.
-Similarly, the bounding line \eqref{eq: lower bound, line} results from the second inequality in \eqref{eq: prop: min suff set, form: isoli ineq} and the first in \eqref{eq: prop: min suff set, form: jsoli ineq}, and the fact that both inequalities are not strict implies that the bound \eqref{eq: lower bound, line} can actually be attained.
+Similarly, the bounding line \eqref{eq: lower bound, line} results from the second inequality in \eqref{eq: prop: min suff set, form: isoli ineq} and the first in \eqref{eq: prop: min suff set, form: jsoli ineq}, and the fact that those inequalities are not strict implies that the bound \eqref{eq: lower bound, line} can actually be attained.
 \end{proof}
@@
 This happens for instance in Figure~\ref{fig: pairs_bounds_1p35}, where the number of pairs for each $i$ equals either $2$ or $3$ without a clear pattern.\footnote{Strictly, there is a periodic pattern whenever $a^2/b^2$ is rational, which is the case in Figure~\ref{fig: pairs_bounds_1p35}.
-However, it is not easily discernible unless $a^2/b^2$ is a ratio of small numbers.}
+However, the pattern is not easily discernible unless $a^2/b^2$ is a ratio of small numbers.}
 On the other hand, a simple pattern arises when $a^2/b^2$ or $b^2/a^2$ is a natural number, as seen in Figures~\ref{fig: pairs_bounds_1} and \ref{fig: pairs_bounds_sqrt2}.

### 2021-11-25 paper.tex
Rewriting

@@
 The general case for rectangular grids with real-valued segment lengths is analyzed first, in \S\ref{part: max: arbitrary grid, real lengths}.
-The particular case of square grids is addressed in \S\ref{part: max: unit square grid, real lengths}, as it allows a specialized formula for the direct problem.
+The square grid with real-valued segment lengths is addressed in \S\ref{part: max: unit square grid, real lengths}, as it allows a specialized formula for the direct problem.
 Lastly, the analysis of a unit square grid with integer-valued lengths is presented in \S\ref{part: max: unit square grid, integer lengths}.
@@
 Note that $i$ and $j$ are not statistically independent.
+
+Rectangular grids with real-valued segment lengths are considered in \S\ref{part: probmax: arbitrary grid, real lengths}.
+The main result is the average number of visited tiles.
+The square grid with real-valued lengths is addressed in \S\ref{part: probmax: unit square grid, real lengths}.
+This more specialized setting allows computation of the probability of visiting the maximum possible number of tiles.
 

### 2021-11-25 paper.tex
Rewriting

@@
 
-The function $\funt$ can be obtained from $\funl$ by noting that the maximum number of tiles that can be visited by a segment of length $\len$ is the index of the largest term of the sequence $\funl(\tiles)$ that is less than $\len$:
+The function $\funt$ can be obtained from $\funl$ by noting that the maximum number of tiles that can be visited by a segment of length $\len$ is the index of the last term of the sequence $\funl(\tiles)$ that is less than $\len$:
 \begin{equation}
@@
 
-For arbitrary $a, b \in \mathbb R^+$, the functions $\funt$ and $\funl$ can be computed using an iterative procedure, which exploits the fact that the pairs $(i_3,j_3), (i_4,j_4), \ldots$ of $\mss$ are sorted by increasing $i+j-1$, and also by increasing $(i-2)^2 a^2 + (j-2)^2 b^2$.
+For arbitrary $a, b \in \mathbb R^+$, the functions $\funt$ and $\funl$ can be computed using an iterative procedure, which exploits the fact that the pairs $(i_3,j_3), (i_4,j_4), \ldots$ of the set $\mss$ are sorted by increasing $i+j-1$, and also by increasing $(i-2)^2 a^2 + (j-2)^2 b^2$.
 Namely, for $\funt$ the following procedure yields the solution: generate successive pairs to find the last one, $(i_\tiles,j_\tiles)$, that satisfies \eqref{eq: len > sqrt}; then $\funt(\len) = \tiles$.
@@
 Then, assuming that $i = \isoli$, the largest $j$ allowed by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, $\jsoli$, is obtained.
-As will be seen, in some cases the resulting $(\isoli,\jsoli)$ is in $\mss$ as defined by Proposition~\ref{prop: min suff set, form}, and in other cases it is not.
+As will be seen, in some cases the resulting $(\isoli,\jsoli)$ is in the set $\mss$ defined by Proposition~\ref{prop: min suff set, form}, and in other cases it is not.
 However, in either case $(\isoli,\jsoli)$ has the largest $i+j-1$ sum that can be achieved with segments of length up to $\len$.
@@
 \label{eq: quadratic, i}
-4a^2(\isolr-2)^2 - 4a^2(\isolr-2) + a^2+b^2-\frac{4\len^2 b^2}{a^2+b^2},
-\end{equation}
-which can have zero, one or two real-valued solutions (this respectively means that the line \eqref{eq: icont jcont 2} is exterior, tangent or secant to the circle \eqref{eq: icont jcont 1}).
+4a^2(\isolr-2)^2 - 4a^2(\isolr-2) + a^2+b^2-\frac{4\len^2 b^2}{a^2+b^2} = 0,
+\end{equation}
+which can have zero, one or two real-valued solutions.
+These cases respectively mean that the solid line in Figure~\ref{fig: ijLS_1p35_detail} is exterior, tangent or secant to the circle (the figure shows the latter case).
 Only the solution with $\isolr, \jsolr \geq 2$, if any, is of interest.
-This solution can only exist as the one with largest $\isolr$ value when there are real-valued solutions, and it is obtained as
+This solution can only exist as the one with largest $\isolr$ value when there are real-valued solutions, and it is obtained as %***Here rewriting
 \begin{equation}
@@
 
-If \eqref{eq: quadratic, i} has two real-valued solutions or one real-valued double solution, which is the case when $\len \geq \sqrt{a^2+b^2}/2$, the largest solution will have $\jsolr < 2$ if $\isolr < (b^2/a^2+5)/2$, as can be seen setting $\jsolr=2$ in \eqref{eq: icont jcont 2}; or equivalently if $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, as can be seen substituting $\jsolr < 2$ and $\isolr < (b^2/a^2+5)/2$ in \eqref{eq: icont jcont 1}.
+If \eqref{eq: quadratic, i} has two real-valued solutions or one real-valued double solution, which occurs when $\len \geq \sqrt{a^2+b^2}/2$, the largest solution will have $\jsolr < 2$ if $\isolr < (b^2/a^2+5)/2$, as can be seen setting $\jsolr=2$ in \eqref{eq: icont jcont 2}; or equivalently if $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, as can be seen substituting $\jsolr < 2$ and $\isolr < (b^2/a^2+5)/2$ in \eqref{eq: icont jcont 1}.
 Since $\isolr < (b^2/a^2+5)/2 < 3$ for $a \geq b$, only pairs with $i=2$ are achievable in this case again, and thus the solution $\isoli$ should also be $2$.
@@
 
-Although Theorem~\ref{theo: funt, form} is valid for $a \geq b$, the result could be applied for $a < b$ by simply swapping the values of $a$ and $b$.
-In other words, \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j}
+Although Theorem~\ref{theo: funt, form} requires $a \geq b$, the result could be applied for $a < b$ by simply swapping the values of $a$ and $b$.
+%In other words, \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j}
 % *!*
-can be used for any grid if $a$, $b$ are interpreted as the largest and smallest sides of a tile, respectively.
+%can be used for any grid if $a$, $b$ are interpreted as the largest and smallest sides of a tile, respectively.
 

### 2021-11-25 paper.tex
Rewriting

@@
 
-\affil[1]{\small Universidad Polit\'ecnica de Madrid. \texttt{luis.mendo@upm.es}}
+\affil[1]{\small Universidad Polit\'ecnica de Madrid. E-mail:  \texttt{luis.mendo@upm.es}}
 %Information Processing and Telecommunications Center, Universidad Polit\'ecnica de Madrid\\
 %Avenida Complutense, 30. 28040 Madrid, Spain.\\
-\affil[2]{\small \texttt{arkhipov@alum.mit.edu}}
+\affil[2]{\small E-mail: \texttt{arkhipov@alum.mit.edu}}
 
@@
 Only the solution with $\isolr, \jsolr \geq 2$, if any, is of interest.
-This solution can only exist as the one with largest $\isolr$ value when there are real-valued solutions, and it is obtained as %***Here rewriting
+This solution can only exist as the one with largest $\isolr$ value when there are real-valued solutions, and it is obtained as %***Here rewriting.
+Maybe rephrase the following: the expression is general (isn't it?), i.e.
+may give complex values; and the three cases (0, 1 or 2 real solutions are not clearly identified)
 \begin{equation}

### 2021-11-27 paper.tex
Tried to clarify proof of Theorem 1
The explanation of the three cases for the solution pair (i+,j+) were confusing. I think it's clearer now, but more work is still needed

--- 2021-11-25 paper.tex
@@
 \end{align}
-Substituting \eqref{eq: icont jcont 2} into \eqref{eq: icont jcont 1} yields a quadratic equation for $\isolr$,
+Substituting \eqref{eq: icont jcont 2} into \eqref{eq: icont jcont 1} gives a quadratic equation for $\isolr$,
 \begin{equation}
@@
 \end{equation}
-which can have zero, one or two real-valued solutions.
-These cases respectively mean that the solid line in Figure~\ref{fig: ijLS_1p35_detail} is exterior, tangent or secant to the circle (the figure shows the latter case).
-Only the solution with $\isolr, \jsolr \geq 2$, if any, is of interest.
-This solution can only exist as the one with largest $\isolr$ value when there are real-valued solutions, and it is obtained as %***Here rewriting
+which is solved as
 \begin{equation}
 \label{eq: icont jcont 3}
-\isolr = \frac{a + b \sqrt{4\len^2/(a^2+b^2)-1}}{2a} + 2.
-\end{equation}
-
-The case with no real-valued solutions to \eqref{eq: quadratic, i} corresponds to $\len < \sqrt{a^2+b^2}/2$.
-Since $a \geq b$, this implies that $\len < a$.
-Thus any achievable $(i,j)$ pair, if any, will have $i=2$.
-Therefore in this case $\isoli$ should be set to $2$; and then the maximum achievable $\jsoli$ will be obtained from \eqref{eq: icont jcont 2}.
-
-If \eqref{eq: quadratic, i} has two real-valued solutions or one real-valued double solution, which occurs when $\len \geq \sqrt{a^2+b^2}/2$, the largest solution will have $\jsolr < 2$ if $\isolr < (b^2/a^2+5)/2$, as can be seen setting $\jsolr=2$ in \eqref{eq: icont jcont 2}; or equivalently if $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, as can be seen substituting $\jsolr < 2$ and $\isolr < (b^2/a^2+5)/2$ in \eqref{eq: icont jcont 1}.
-Since $\isolr < (b^2/a^2+5)/2 < 3$ for $a \geq b$, only pairs with $i=2$ are achievable in this case again, and thus the solution $\isoli$ should also be $2$.
-
-Lastly, for $\len \geq (a^2+b^2) / (2a)$ the expressions \eqref{eq: icont jcont 2} and \eqref{eq: icont jcont 3} give $\isolr, \jsolr \geq 2$, and $\isoli$ should be taken as the greatest integer less than $\isolr$, i.e.~$\lceil \isolr \rceil-1$.
-
-The three cases are unified by taking the real part of \eqref{eq: icont jcont 3} and computing $\isoli = \lceil \isolr \rceil-1$, as is easily checked.
-This gives \eqref{eq: theo: funt, form; i}.
+\isolr = \frac{a \pm b \sqrt{4\len^2/(a^2+b^2)-1}}{2a} + 2,
+\end{equation}
+and then $\jsolr$ can be obtained from \eqref{eq: icont jcont 2}.
+This can produce zero, one or two real-valued solution pairs $(\isolr, \jsolr)$, which respectively corresponds to the solid line in Figure~\ref{fig: ijLS_1p35_detail} being exterior, tangent or secant to the circle (the figure shows the latter case).
+
+A solution $(\isolr, \jsolr)$ is of interest only if it is real-valued with $\isolr, \jsolr \geq 2$.
+This solution, if it exists, is always associated with the positive sign in \eqref{eq: icont jcont 3} (the negative sign either yields the same solution, in the tangent case, or yields a solution with $\jsolr < 2$, in the secant case).
+In this regard, three cases need to be distinguished: there are no real-valued solution pairs $(\isolr, \jsolr)$; there are one or two but do not satisfy $\isolr, \jsolr \geq 2$; there are one or two and and one of them satisfies that inequality.
+
+If $\len < \sqrt{a^2+b^2}/2$, \eqref{eq: icont jcont 3} yields no real-valued solutions for $\isolr$.
+Since $a \geq b$, $\len < \sqrt{a^2+b^2}/2$ implies that $\len < a$, and therefore any achievable $(i,j)$ pair, if any, will have $i=2$.
+Thus in this case $\isoli$ should be set to $2$; and then the maximum achievable $\jsoli$ will be obtained from \eqref{eq: icont jcont 2}.
+
+For $\len \geq \sqrt{a^2+b^2}/2$ there are either two real-valued solutions or one real-valued double solution for $\isolr$.
+These solutions will give $\jsolr < 2$ if $\isolr < (b^2/a^2+5)/2$, as can be seen setting $\jsolr=2$ in \eqref{eq: icont jcont 2}; or equivalently if $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, as can be seen substituting $\jsolr < 2$ and $\isolr < (b^2/a^2+5)/2$ in \eqref{eq: icont jcont 1}.
+Thus for $\len$ in this range, since $\isolr < (b^2/a^2+5)/2 < 3$ for $a \geq b$, only pairs with $i=2$ are achievable again, and therefore $\isoli$ must also be $2$.
+
+Lastly, for $\len \geq (a^2+b^2) / (2a)$ the expressions \eqref{eq: icont jcont 2} and \eqref{eq: icont jcont 3} with positive sign give $\isolr, \jsolr \geq 2$, and $\isoli$ should be taken as the greatest integer less than $\isolr$, i.e.~$\lceil \isolr \rceil-1$.
+
+The three cases are unified by taking the real part of \eqref{eq: icont jcont 3} and computing $\isoli = \lceil \isolr \rceil-1$, as can be easily checked.
+This yield \eqref{eq: theo: funt, form; i}.
 Once $\isoli$ is known, $\jsoli$ is computed, according to \eqref{eq: theo: funt, form; j}, as the greatest integer such that $(\isoli a,\jsoli b)$ is within the circle determined by $\len$.
@@
 
-There are two possibilities for the pair $(\isoli,\jsoli)$, as stated at the outset.
+There are two possibilities for the resulting pair $(\isoli,\jsoli)$, as stated at the outset.
 These are illustrated in Figures~\ref{fig: ijLS_1p35_detail_in} and \ref{fig: ijLS_1p35_detail_notin} respectively.

### 2021-11-28 paper.tex
... better now

@@
 \label{eq: theo: funt, form; i}
-\isoli &= \left\lceil \frac{a + b \Re \sqrt{4\len^2/(a^2+b^2)-1}}{2a} \right\rceil + 1, \\
+\isoli &= \left\lceil \frac 3 2 + \frac b a \Re \sqrt{\frac{\len^2}{a^2+b^2}-\frac 1 4} \right\rceil, \\
 \label{eq: theo: funt, form; j}
-\jsoli &= \left\lceil \frac{\sqrt{\len^2-(\isoli-2)^2a^2}}{b} \right\rceil + 1.
+\jsoli &= \left\lceil 1 + \frac{\sqrt{\len^2-(\isoli-2)^2a^2}}{b} \right\rceil.
 \end{align}
@@
 \end{align}
-Substituting \eqref{eq: icont jcont 2} into \eqref{eq: icont jcont 1} gives a quadratic equation for $\isolr$,
-\begin{equation}
-\label{eq: quadratic, i}
-4a^2(\isolr-2)^2 - 4a^2(\isolr-2) + a^2+b^2-\frac{4\len^2 b^2}{a^2+b^2} = 0,
-\end{equation}
-which is solved as
-\begin{equation}
-\label{eq: icont jcont 3}
-\isolr = \frac{a \pm b \sqrt{4\len^2/(a^2+b^2)-1}}{2a} + 2,
-\end{equation}
-and then $\jsolr$ can be obtained from \eqref{eq: icont jcont 2}.
-This can produce zero, one or two real-valued solution pairs $(\isolr, \jsolr)$, which respectively corresponds to the solid line in Figure~\ref{fig: ijLS_1p35_detail} being exterior, tangent or secant to the circle (the figure shows the latter case).
-
-A solution $(\isolr, \jsolr)$ is of interest only if it is real-valued with $\isolr, \jsolr \geq 2$.
-This solution, if it exists, is always associated with the positive sign in \eqref{eq: icont jcont 3} (the negative sign either yields the same solution, in the tangent case, or yields a solution with $\jsolr < 2$, in the secant case).
-In this regard, three cases need to be distinguished: there are no real-valued solution pairs $(\isolr, \jsolr)$; there are one or two but do not satisfy $\isolr, \jsolr \geq 2$; there are one or two and and one of them satisfies that inequality.
-
-If $\len < \sqrt{a^2+b^2}/2$, \eqref{eq: icont jcont 3} yields no real-valued solutions for $\isolr$.
-Since $a \geq b$, $\len < \sqrt{a^2+b^2}/2$ implies that $\len < a$, and therefore any achievable $(i,j)$ pair, if any, will have $i=2$.
+Expressing these equations in terms of new variables $\isolr-5/2$ and $\jsolr-3/2$, the solutions are easily found to be
+\begin{align}
+\label{eq: isolr sol}
+\isolr &= \frac 5 2 \pm \frac b a \sqrt{\frac{\len^2}{a^2+b^2} - \frac 1 4}, \\
+\label{eq: jsolr sol}
+\jsolr &= \frac 3 2 \pm \frac a b \sqrt{\frac{\len^2}{a^2+b^2} - \frac 1 4},
+\end{align}
+where the same sign (positive or negative) should be used in the two equations.
+This can yield zero, one or two real-valued solution pairs $(\isolr, \jsolr)$, which respectively corresponds to the solid line in Figure~\ref{fig: ijLS_1p35_detail} being exterior, tangent or secant to the circle (the figure depicts the latter case).
+
+A solution pair $(\isolr, \jsolr)$ from \eqref{eq: isolr sol} and \eqref{eq: jsolr sol} is meaningful only if it is real-valued with $\isolr, \jsolr \geq 2$.
+This solution, if it exists, is always associated with the positive sign in those equations.
+It is easily seen that $\jsolr \geq 2$ implies $\isolr \geq 2$, and thus it suffices the check the former.
+Three cases need to be distinguished: there are no real-valued solution pairs $(\isolr, \jsolr)$; there are one or two but none of them has $\jsolr \geq 2$; or there are one or two and and one of them satisfies that inequality.
+These cases correspond to different ranges of $\len$, as discussed next.
+
+For $\len < \sqrt{a^2+b^2}/2$, \eqref{eq: isolr sol} and \eqref{eq: jsolr sol} are not real-valued.
+Since $a \geq b$, from the inequality $\len < \sqrt{a^2+b^2}/2$ it follows that $\len < a$.
+This means that any achievable $(i,j)$ pair, if any, will have $i=2$.
 Thus in this case $\isoli$ should be set to $2$; and then the maximum achievable $\jsoli$ will be obtained from \eqref{eq: icont jcont 2}.
 
-For $\len \geq \sqrt{a^2+b^2}/2$ there are either two real-valued solutions or one real-valued double solution for $\isolr$.
-These solutions will give $\jsolr < 2$ if $\isolr < (b^2/a^2+5)/2$, as can be seen setting $\jsolr=2$ in \eqref{eq: icont jcont 2}; or equivalently if $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, as can be seen substituting $\jsolr < 2$ and $\isolr < (b^2/a^2+5)/2$ in \eqref{eq: icont jcont 1}.
-Thus for $\len$ in this range, since $\isolr < (b^2/a^2+5)/2 < 3$ for $a \geq b$, only pairs with $i=2$ are achievable again, and therefore $\isoli$ must also be $2$.
-
-Lastly, for $\len \geq (a^2+b^2) / (2a)$ the expressions \eqref{eq: icont jcont 2} and \eqref{eq: icont jcont 3} with positive sign give $\isolr, \jsolr \geq 2$, and $\isoli$ should be taken as the greatest integer less than $\isolr$, i.e.~$\lceil \isolr \rceil-1$.
-
-The three cases are unified by taking the real part of \eqref{eq: icont jcont 3} and computing $\isoli = \lceil \isolr \rceil-1$, as can be easily checked.
-This yield \eqref{eq: theo: funt, form; i}.
-Once $\isoli$ is known, $\jsoli$ is computed, according to \eqref{eq: theo: funt, form; j}, as the greatest integer such that $(\isoli a,\jsoli b)$ is within the circle determined by $\len$.
+For $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, \eqref{eq: isolr sol} and \eqref{eq: jsolr sol} give either two real-valued solutions or one real-valued double solution for $(\isolr, \jsolr)$; but $\jsolr < 2$.
+This in turn implies, according to \eqref{eq: icont jcont 2}, that $\isolr < (b^2/a^2+5)/2 < 3$ for $a \geq b$.
+Thus only pairs with $i=2$ are achievable again for $\len$ in this range, and therefore $\isoli$ must also be $2$.
+
+Lastly, for $\len \geq (a^2+b^2) / (2a)$ the expressions \eqref{eq: isolr sol} and \eqref{eq: jsolr sol} with positive sign give $\isolr, \jsolr \geq 2$, and $\isoli$ should be taken as the greatest integer less than $\isolr$, i.e.~$\lceil \isolr \rceil-1$.
+
+The three cases are unified by taking the real part of the positive-sign version of \eqref{eq: isolr sol} and computing $\isoli = \lceil \isolr \rceil-1$, as can be easily checked.
+This yields \eqref{eq: theo: funt, form; i}.
+Once $\isoli$ is known, \eqref{eq: theo: funt, form; j} computes $\jsoli$ as the greatest integer such that $(\isoli a,\jsoli b)$ is within the circle with center $(2a,2b)$ and radius $\len$.
 This ensures that $(\isoli,\jsoli)$ is achievable with lengths less than $\len$.

### 2021-11-28 paper.tex
More rewriting (proof of theorem 1)

@@
 This is always the case, because  $(\isoli,\jsoli+1) \in \mss$ (it is above or on the bounding line) and $(\isoli,\jsoli) \notin \mss$ (it is below the line), and due to how $\mss$ has been constructed, this implies that $(\isoli-1,\jsoli+1) \in \mss$ and $(\isoli-1,\jsoli+k) \notin \mss$ for $k =2, 3, \ldots$.
-Therefore, $(\isoli,\jsoli)$ is achievable and maximizes $i+j-1$, which implies that $\isoli+\jsoli-1$ is the desired solution.
+It follows that $(\isoli,\jsoli)$ is achievable and maximizes $i+j-1$, and thus $\isoli+\jsoli-1$ is the desired solution.
 
@@
 
-It should be noted that for the specific case that $a^2/b^2 = 2k-1$, $k \in \mathbb N$ the lower bounding line \eqref{eq: lower bound, line} becomes $j = i(2k-1)-5k+4$, which gives integer $j$ for integer $i$.
+Interestingly, for the specific case that $a^2/b^2 = 2k-1$, $k \in \mathbb N$ the lower bounding line \eqref{eq: lower bound, line} becomes $j = (2k-1)i - 5k + 4$, which gives integer $j$ for integer $i$.
 This means that the bound is actually achieved for all pairs in $\mss$ (see for example Figure~\ref{fig: pairs_bounds_1}), and the case $(\isoli,\jsoli) \notin \mss$ never occurs.
-Thus, when $a^2/b^2$ is an odd integer the obtained $(\isoli,\jsoli)$ is guaranteed to be in $\mss$, whereas for general $a^2/b^2$ either $(\isoli,\jsoli)$ or $(\isoli-1,\jsoli+1)$ are in in this set.
 % ^ If this is needed outside of the proof, it can be included at the end of the theorem statement
 
-As for the properties of $\funt$, it follows from \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j} that this function is piecewise constant and left-continuous.
+As for the properties of $\funt$, it stems from \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j} that this function is piecewise constant and left-continuous.
 From the procedure described in the previous paragraphs for obtaining $(\isoli, \jsoli)$ it is clear that $\isoli+\jsoli-1$ increases in steps of $1$ when $\len$ is increased continuously; that is, $\funt$ has jumps of unit height.
@@
 To see that $\funt$ has a jump at $\len$, assume for the sake of contradiction that $\funt$ is continuous, therefore constant, at that point.
-Then there is $\epsilon > 0$ for which $\funt(\len+\epsilon) = \tiles = \funt(\len-\epsilon)$.
+Then there is $\epsilon > 0$ for which $\funt(\len-\epsilon) = \tiles = \funt(\len+\epsilon)$.
 This means that there exists a segment with length $\len-\epsilon$ that visits $\tiles$ tiles, and thus $\funl(\tiles) \leq \len-\epsilon < \len$, in contradiction with \eqref{eq: theo funt, form: prop 1}.
 Therefore $\funt$ is discontinuous, from the right, at $\len$.
-By definition of $\funl$, \eqref{eq: theo funt, form: prop 1} implies that $\funt(\len) < \tiles$, and that there exists $\epsilon > 0$ such that $\funt(\len+h) = \tiles$ for $0 < h < \epsilon$.
-This means that
-\begin{align}
+By definition of $\funl$, from \eqref{eq: theo funt, form: prop 1} it follows that
+\begin{equation}
 \label{eq: theo funt, form: prop 2}
-\funt(\len) &< \tiles, \\
+\funt(\len) < \tiles
+\end{equation}
+and there exists $\epsilon > 0$ such that $\funt(\len+h) = \tiles$ for $0 < h < \epsilon$. This implies that
+\begin{equation}
 \label{eq: theo funt, form: prop 3}
-\lim_{h \rightarrow 0+} \funt(\len+h) &= \tiles,
-\end{align}
+\lim_{h \rightarrow 0+} \funt(\len+h) = \tiles,
+\end{equation}
 that is, $\funt$ has a jump at $\len$.
-In addition, since the jump has unit height, it follows from \eqref{eq: theo funt, form: prop 2} and \eqref{eq: theo funt, form: prop 3}  that $\funt(\len) = \tiles-1$.
-
-Conversely, assume that \eqref{eq: theo funt, form: prop 2} and \eqref{eq: theo funt, form: prop 3} hold for some $\len \in \mathbb R^+$, $\tiles \in \mathbb N$.
+In addition, since the jump has unit height, it stems from \eqref{eq: theo funt, form: prop 2} and \eqref{eq: theo funt, form: prop 3}  that $\funt(\len) = \tiles-1$.
+
+Conversely, assume that $\funt$ has a jump from $\tiles-1$ to $\tiles$ at some $\len \in \mathbb R^+$.
+This means that \eqref{eq: theo funt, form: prop 2} and \eqref{eq: theo funt, form: prop 3} hold for those $\tiles$ and $\len$.
 From \eqref{eq: theo funt, form: prop 2} it follows that  $\funl(\tiles) \geq \len$.
@@
 
-Although Theorem~\ref{theo: funt, form} requires $a \geq b$, the result could be applied for $a < b$ by simply swapping the values of $a$ and $b$.
+Although Theorem~\ref{theo: funt, form} requires $a \geq b$, the result could obviously be applied for $a < b$ by swapping the values of $a$ and $b$.
 %In other words, \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j}

### 2021-11-30 paper.tex
Rewriting

@@
 \newcommand{\st}{\mid}%{\,|\,}
-\newcommand{\cond}{\,|\,} % less spacing that \mid
+\newcommand{\cond}{\,|\,} % less spacing than \mid
 \newcommand{\diff}{\,\mathrm d}
@@
 \begin{proof}
-The $(i_\tiles,j_\tiles)$ pair in $\mss$ corresponds to a maximum of $\tiles$ visited tiles.
+The $(i_\tiles,j_\tiles)$ pair in set $\mss$ defined in Proposition~\ref{prop: min suff set, form} corresponds to a maximum of $\tiles$ visited tiles.
 By construction of this set, any segment that visits $\tiles$ tiles must have length greater than $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$.
-For $\tiles \geq 3$ the variables $\isoli$, $\jsoli$ in \eqref{eq: theo: funl, form; isoli} and \eqref{eq: theo: funl, form; jsoli} respectively coincide with $i_\tiles$, $j_\tiles$ as given by \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}.
-Therefore \eqref{eq: theo: funl, form; funl} gives the correct result.
+For $\tiles \geq 3$ the variables $\isoli$, $\jsoli$ computed in \eqref{eq: theo: funl, form; isolr}--\eqref{eq: theo: funl, form; jsoli}
+% -- *!*
+coincide with $i_\tiles$, $j_\tiles$ as given by \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}, and therefore \eqref{eq: theo: funl, form; funl} gives the correct result.
 For $\tiles \in \{1, 2\}$ both \eqref{eq: theo: funl, form; isoli} and \eqref{eq: theo: funl, form; jsoli} equal $2$, and \eqref{eq: theo: funl, form; funl} gives $0$, which is again the correct result.
@@
 \end{equation}
-The fact that both $(\isolr, \jsolr)$ and $(i_\tiles, j_\tiles)$ are on the line $i+j-1=\tiles$ implies that $\isolr+\jsolr = i_\tiles+ j_\tiles$, and thus $|i_\tiles-\isolr| = |j_\tiles-\jsolr| = \roundterm$.
+The fact that both $(\isolr, \jsolr)$ and $(i_\tiles, j_\tiles)$ are on the line $i+j-1=\tiles$ implies that $\isolr+\jsolr = i_\tiles+ j_\tiles$, and thus $|i_\tiles-\isolr| = |j_\tiles-\jsolr| = \roundterm$ with $\roundterm$ given by \eqref{eq: roundterm}.
 Consequently,
@@
 
-By Theorem~\ref{theo: funt, form}, $\funt$ has unit-height jumps at the values $\funl(\tiles)$, $\tiles \in \mathbb N$, $\tiles \geq 4$.
-This implies that $\funl$ is monotone increasing for $\tiles \geq 3$.
+The definition of $\funl$ implies that $\funl(\tiles) \geq \funl(\tiles-1)$ for any $\tiles \in \mathbb N$. On the other hand, 
+by Theorem~\ref{theo: funt, form}, $\funt$ is piecewise constant and has a unit-height jump from $\tiles-1$ to $\tiles$ at $\funl(\tiles)$, $\tiles \in \mathbb N$, $\tiles \geq 4$.
+This implies that $\funl(\tiles) > \funl(\tiles-1)$ for $\tiles \geq 4$.
 \end{proof}
 
-The expression \eqref{eq: theo: funl, form; funl 2} allows a neat interpretation of $\funl(\tiles)$ (as stems from the arguments used in the proof).
-Namely, $\funl^2(\tiles)$ is the sum of the two terms within the square root in that expression.
+The expression \eqref{eq: theo: funl, form; funl 2} allows a neat interpretation of $\funl(\tiles)$ (as stems from the arguments used in the proof of Theorem~\ref{theo: funl, form}).
+Namely, $\funl^2(\tiles)$ is the sum of the two terms that appear in that expression.
 The first term is the squared distance from $(2a, 2b)$ to the diagonal line defined by $i+j-1=t$; and the second term is additional squared distance incurred from rounding $i$, $j$ to integer values.
 
-Theorems~\ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the two problems stated in \S\ref{part: intro}; they also provide a way to actually position a segment of length slightly greater than $\len$ or $\funl(\tiles)$ so that it visits $\tiles$ or $\funt(\len)$ tiles.
-Namely, the segment should have its endpoints respectively in the interior of two tiles separated $\isoli-1$ steps horizontally and $\jsoli-1$ steps vertically, with its exact position and orientation  adjusted to avoid any grid points.
+Theorems~\ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the first two problems stated in \S\ref{part: intro}; they also provide a way to actually position a segment of length $\len$ or slightly greater than $\funl(\tiles)$, respectively, so that it visits $\funt(\len)$ or $\tiles$ tiles.
+Namely, for $\isoli$, $\jsoli$ computed as in the corresponding theorem, the segment should have its endpoints in the interiors of two tiles separated $\isoli-1$ steps horizontally and $\jsoli-1$ steps vertically, with its exact position and orientation  adjusted to avoid any grid points.
 

### 2021-11-30 paper.tex
Rewriting up to proof of Theorem 3, included

@@
 As for $a \gg b$, in this case the optimal discrete bounding rectangle has $\isoli = 2$, and $\jsoli$ as large as allowed by $\len$ (direct problem) or as required by $\tiles$ (inverse problem), corresponding to an almost vertical segment.
-Obviously, for $a \gg b$ the length of the segment is best invested in increasing the number of tiles traversed vertically (but the segment should be slightly tilted to cross a vertical edge), and the asymptotic value of \eqref{eq: asympt slope} is approximately $1/b$.
-
-For $a=b$, either from symmetry considerations or particularizing the formulas in the previous theorems it stems that the optimal orientation of the segment is close to $45^\circ$.
+In other words, for $a \gg b$ the length of the segment is best invested in increasing the number of tiles traversed vertically (but the segment should be slightly tilted to cross a vertical edge), and the asymptotic ratio \eqref{eq: asympt slope} is approximately $1/b$.
+
+For $a=b$, either from symmetry considerations or particularizing the formulas in the above theorems it stems that the optimal orientation of the segment is close to $45^\circ$.
 This case will be dealt with in \S\ref{part: max: unit square grid, real lengths}, as it lends itself to simplified formulas.
 
-Figure~\ref{fig: funt funl examples} shows $\funt$ and $\funl$ for several pairs of grid parameters $a$, $b$.
+Figure~\ref{fig: funt funl examples} shows the functions $\funt$ and $\funl$ for several pairs of grid parameters $a$, $b$.
 The graphs illustrate some of the observations of the previous paragraphs.
@@
 
-A square grid has $a=b$. For real-valued segment lengths it can be further assumed that $a=1$. For $a \neq 1$ the results to be obtained apply with $\len$ replaced by $\len/a$.
+A square grid has $a=b$.
+For real-valued segment lengths it can be further assumed that $a=1$.
+For $a \neq 1$ the expressions to be obtained are valid with $\len$ and $\funl(\tiles)$ replaced by $\len/a$ and $\funl(\tiles)/a$ respectively.
 
@@
 \label{eq: cor: funt, sq, form; i}
-\isoli &= \left\lceil \frac{1 + \Re \sqrt{2\len^2-1}}{2} \right\rceil + 1, \\
+\isoli &= \left\lceil \frac 3 2 + \Re\sqrt{\frac{\len^2}{2} - \frac 1 4} \right\rceil, \\
 \label{eq: cor: funt, sq, form; j}
-\jsoli &= \left\lceil \sqrt{\len^2-(\isoli-2)^2} \right\rceil + 1,
+\jsoli &= \left\lceil 1 + \sqrt{\len^2-(\isoli-2)^2} \right\rceil.
 \end{align}
@@
 For this new set, the lower bounding line \eqref{eq: lower bound, line} can be replaced by the simpler $j=i$.
-The same approach used in the proof of Theorem~\ref{theo: funt, form} can be followed, but using this line.
+The same approach followed in the proof of Theorem~\ref{theo: funt, form} can be applied here, but using this line.
 Thus $(\isolr, \jsolr)$ is obtained from
@@
 \end{equation}
-The pair $(\isoli,\jsoli)$ resulting from \eqref{eq: theo: funt, sq, form; proof 1} is in $\mss'$, and is given in \eqref{eq: theo: funt, sq, form; i} and \eqref{eq: theo: funt, sq, form; j}.
+The pair $(\isoli,\jsoli)$ resulting from \eqref{eq: theo: funt, sq, form; proof 1} is in $\mss'$, and is as given by \eqref{eq: theo: funt, sq, form; i} and \eqref{eq: theo: funt, sq, form; j}.
 This establishes \eqref{eq: theo: funt, sq, form; funt, with i, j}.
@@
 \end{equation}
-According to \eqref{eq: funt funl}, the function $\funt(\len)$ is given by the largest positive integer $t$ such that 
+According to \eqref{eq: funt funl}, $\funt(\len)$ is obtained by the largest positive integer $t$ such that 
 \begin{equation}
@@
 \end{equation}
-Finally, solving for $\tiles$ gives
+Solving for $\tiles$ gives
 \begin{equation}

### 2021-12-01 paper.tex
Rewriting some more

@@
 \label{eq: theo: funt, sq, form; j}
-\jsoli &= \left\lceil \sqrt{\len^2-(\isoli-2)^2} \right\rceil + 1.
-\end{align}
-Equivalently,
+\jsoli &= \left\lceil \sqrt{\len^2-(\isoli-2)^2} \right\rceil + 1,
+\end{align}
+where either $\jsoli = \isoli$ or $\jsoli = \isoli+1$. Equivalently,
 \begin{equation}
@@
 \begin{proof}
-The proof of \eqref{eq: theo: funt, sq, form; funt, with i, j} uses a variation of the set $\mss$ defined in Proposition~\ref{prop: min suff set, form} that is more suited to this situation.
+The proof of \eqref{eq: theo: funt, sq, form; funt, with i, j}--\eqref{eq: theo: funt, sq, form; j}
+% *!* --
+uses a variation of the set $\mss$ defined in Proposition~\ref{prop: min suff set, form} that is more suited to this situation.
 
 For $a=b=1$, the set $\mss$ consists of points of the form $(i,i)$ and $(i,i-1)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}.
-By symmetry, replacing each point $(i,i-1)$ by $(i-1,i)$ gives a set $\mss'$ which is also optimal.
+By symmetry, replacing each point $(i,i-1)$ by $(i-1,i)$ gives a set $\mss'$ that is also optimal.
 For this new set, the lower bounding line \eqref{eq: lower bound, line} can be replaced by the simpler $j=i$.
@@
 (\isolr-2)^2 + (\jsolr-2)^2  &= \len^2, \\
-\jsolr &= \isoli,
+\jsolr &= \isolr,
 \end{align}
@@
 \label{eq: theo: funt, sq, form; proof 1}
-\isolr = \len/\sqrt{2} + 2.
-\end{equation}
-The pair $(\isoli,\jsoli)$ resulting from \eqref{eq: theo: funt, sq, form; proof 1} is in $\mss'$, and is as given by \eqref{eq: theo: funt, sq, form; i} and \eqref{eq: theo: funt, sq, form; j}.
-This establishes \eqref{eq: theo: funt, sq, form; funt, with i, j}.
+\isolr = \jsolr = \len/\sqrt{2} + 2.
+\end{equation}
+As in Theorem~\ref{theo: funt, form}, $\isoli$ is obtained as $\lceil \isolr \rceil-1$; and then $\jsoli$ is chosen as the largest integer such that $(\isoli, \jsoli)$ is achievable, i.e.~its distance from $(2, 2)$ is less than $\len$.
+The resulting $\isoli$ and $\jsoli$ are given by \eqref{eq: theo: funt, sq, form; i} and \eqref{eq: theo: funt, sq, form; j}.
+
+The above procedure for choosing $\jsoli$ given $\isoli = \lceil \isolr \rceil-1$ always results in $\jsoli$ being either $\isoli$ or $\isoli+1$.
+This can be seen as follows.
+For $\jsoli = \isoli$, the point $(\isoli, \jsoli)$ is closer to $(2,2)$ than $(\isolr, \jsolr)$ is, and is therefore achievable.
+This implies that values of $\jsoli$ smaller than $\isoli$ will never be chosen.
+On the other hand, $\jsoli = \isoli+2$ or larger values are not achievable, because they would produce a sum $i+j-1$ greater than $\isolr+\jsolr$, which is impossible.
+
+The preceding analysis shows that the pair $(\isoli,\jsoli)$ is in $\mss'$ and maximizes $i+j-1$. Therefore \eqref{eq: theo: funt, sq, form; funt, with i, j} holds.
 
@@
 
-From Theorem~\ref{theo: funt, sq, form} it is easily seen that odd values of $\funt(\len)$ correspond to $\isoli = \jsoli$, whereas even values are achieved with $\jsoli = \isoli+1$.
-Given $\tiles \geq 3$ with $\tiles$ odd, $\funt(\len) = \tiles$ if and only
+From Theorem~\ref{theo: funt, sq, form} it stems that odd values of $\funt(\len)$ correspond to $\isoli = \jsoli$, whereas even values are achieved with $\jsoli = \isoli+1$.
+In addition, noting that $\funt(\len) = \tiles$ is equivalent to $\funl(\tiles) < \len \leq \funl(\tiles+1)$ and using Corollary~\ref{theo: funl, sq, form}, the following characterization of $\funt$ is obtained.
+For $\tiles \geq 3$ with $\tiles$ odd, $\funt(\len) = \tiles$ if and only if
 \begin{equation}
@@
 \end{equation}
+% *** The corollary says \sqrt{\frac{(\tiles-3)^2+1} {2}}, which comes from the corresponding theorem.
+This is equal to \frac{\sqrt{(\tiles-2)^2+(\tiles-4)^2}} {2}.
+Here we are using the second expression, which is probably needed for the random analysis.
+If so, the expression in the corollary should be changed from the first to the second.
+(But this change cannnot be made in the theorem it comes from.
+It's hopefully not a problem that the expression changes from the theorem to the corollary).
 

### 2021-12-01 paper.tex
Rewriting up to the end of Section 3

@@
 From Theorem~\ref{theo: funt, sq, form} it stems that odd values of $\funt(\len)$ correspond to $\isoli = \jsoli$, whereas even values are achieved with $\jsoli = \isoli+1$.
-In addition, noting that $\funt(\len) = \tiles$ is equivalent to $\funl(\tiles) < \len \leq \funl(\tiles+1)$ and using Corollary~\ref{theo: funl, sq, form}, the following characterization of $\funt$ is obtained.
+In addition, noting that $\funt(\len) = \tiles$ is equivalent to $\funl(\tiles) < \len \leq \funl(\tiles+1)$ and using Corollary~\ref{theo: funl, sq, form} the following characterization of $\funt$ is obtained.
 For $\tiles \geq 3$ with $\tiles$ odd, $\funt(\len) = \tiles$ if and only if
@@
 
-A natural variation of the first two problems introduced in \S\ref{part: intro} is to consider $a=b=1$ with the additional restriction that the segment length can only be a positive integer (equivalently, the square grid has spacing $a$ and the segment length can only be an integer multiple of $a$).
-
-The \emph{direct problem} in this setting corresponds to the restriction of the function $\funt$ to $\mathbb N$.
+A natural variation of the direct and inverse problems introduced in \S\ref{part: intro} is to consider $a=b=1$ with the additional restriction that the segment length can only be a positive integer (equivalently, the square grid has spacing $a$ and the segment lengths are restricted to integer multiples of $a$).
+
+The direct problem in this setting corresponds to the restriction of $\funt$ to $\mathbb N$.
 This will be denoted as a function $\funti: \mathbb N \to \mathbb N$ for greater clarity, although obviously $\funti(\leni) = \funt(\leni)$ for all $\leni \in \mathbb N$.
-The sequence $\funti(\leni)$, $\leni \in \mathbb N$, which takes values $3, 5, 7, 8, 9, \ldots$, is depicted in Figure~\ref{fig: funti}.
+The sequence $\funti(\leni)$, $\leni \in \mathbb N$ takes values $3, 5, 7, 8, 9, \ldots$, and is depicted in Figure~\ref{fig: funti}.
 This is A346232 in the On-Line Encyclopedia of Integer Sequences \cite{OEIS_unitsq_int_dir}.
-For this sequence, the expression in Theorem~\ref{theo: funt, sq, form} simplifies in the obvious way, and the following properties hold.
+For this sequence, the expression \eqref{eq: theo: funt, sq, form; funt, simpler} in Theorem~\ref{theo: funt, sq, form} simplifies in the obvious way, and the following properties hold.
 
@@
 
-The \emph{inverse problem} with integer-length segments can be formulated as follows: given $\tiles \in \mathbb N$, find the \emph{minimum} integer length that allows visiting at least $\tiles$ tiles.
-Observe that in this case, unlike with real-valued lengths, there is indeed a minimum length, because every subset of $\mathbb N$ has a minimum.
+The inverse problem with integer-length segments can be formulated as follows: given $\tiles \in \mathbb N$, find the \emph{minimum} integer length that allows visiting at least $\tiles$ tiles.
+Observe that in this case, unlike with real-valued lengths, there is indeed a minimum length, as every subset of $\mathbb N$ has a minimum.
 This can be expressed as a function $\funli: \mathbb N \to \mathbb N$:
@@
 \end{equation}
-which is related to the function $\funl$ corresponding to real-valued lengths by
+which is related to the function $\funl$ for real-valued lengths by
 \begin{equation}
@@
 
-The sequence $\funli(\tiles)$, $\tiles \in \mathbb N$ can be computed using \eqref{eq: theo: funt, sq, form; funl} and \eqref{eq: funli funl}, and has initial values $1, 1, 1, 2, 2, 3, 3, 4, 5 \ldots$, as shown in Figure~\ref{fig: funli}.
+The sequence $\funli(\tiles)$, $\tiles \in \mathbb N$ can be computed using \eqref{eq: theo: funt, sq, form; funl} and \eqref{eq: funli funl}.
+It has initial values $1, 1, 1, 2, 2, 3, 3, 4, 5 \ldots$, as seen in Figure~\ref{fig: funli}.
 This is A346693 in the On-Line Encyclopedia of Integer Sequences \cite{OEIS_unitsq_int_inv}.
@@
 \begin{equation}
-\leni \geq \left \lceil \sqrt{\frac{(\tiles-3)^2} 2 + 1} \ \right \rceil,
+\leni \geq \sqrt{\frac{(\tiles-3)^2} 2 + 1},
 \end{equation}

### 2021-12-02 paper.tex
Rewriting

@@
 Given $\len \in \mathbb R^+$, consider a segment of length $\len$ with uniformly random position and orientation.
-Specifically, the coordinates $x_1$, $y_1$ of the first endpoint are independent random variables uniformly distributed on $[0,a)$ and $[0,b)$ respectively.
+Specifically, the coordinates $x_1$, $y_1$ of the first endpoint are independent random variables uniformly distributed on $[0,a)$ and $[0,b)$ respectively, where $a$, $b$ are the grid parameters.
 The orientation $\theta$ of the segment is uniformly distributed on $[0,2\pi)$.
-The variables $x_1$, $y_1$ and $\theta$ determine the coordinates $x_2$, $y_2$ of the second endpoint.
+The coordinates $x_2$, $y_2$ of the second endpoint are determined by $x_1$, $y_1$ and $\theta$.
 
 Each realization of the random segment gives rise to a discrete bounding rectangle, whose normalized dimensions $i$ and $j$ are thus random variables, as is the number $\tiles$ of visited tiles.
-Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and $\tiles = i+j-1$ tiles.
+Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and $\tiles = i+j-1$.
 Note that $i$ and $j$ are not statistically independent.
 
-Rectangular grids with real-valued segment lengths are considered in \S\ref{part: probmax: arbitrary grid, real lengths}.
+This section deals with the two problems stated in \S\ref{part: intro} for random segments, namely obtaining the average number of visited tiles and the probability of visiting the maximum number of tiles.
+Segment lengths will be assumed to be real-valued.
+The results to be obtained are directly applicable for integer lengths as a particular case.
+
+Arbitrary grids are considered in \S\ref{part: probmax: arbitrary grid, real lengths}.
 The main result is the average number of visited tiles.
-The square grid with real-valued lengths is addressed in \S\ref{part: probmax: unit square grid, real lengths}.
+The square grid is addressed in \S\ref{part: probmax: unit square grid, real lengths}.
 This more specialized setting allows computation of the probability of visiting the maximum possible number of tiles.
@@
 
-Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles visited by a random segment of length $\len$, with the distributions specified in \S\ref{part: rand}.
+Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ is the average number of tiles visited by a random segment of length $\len$, with the distributions specified in \S\ref{part: rand}.
 This function can be easily computed using the fact that
@@
 \end{equation}
-and noting that events of probability $0$, such as $x_1=a$ or the segment passing through a grid point, can be disregarded.
+and noting that events of probability $0$, such as $x_1=0$ or the segment passing through a grid point, can be disregarded.
 
@@
 \label{theo: funta, form}
-Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above.
-The average number of tiles visited by the segment, $\funta(\len)$, is
+Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above. The average number of tiles visited by the segment is 
 \begin{equation}
@@
 \end{equation}
-As noted in Proposition \ref{prop: i+j-1}, the number of tiles visited by a segment is the count of its grid line crossings plus $1$, unless it exactly passes through any grid points, but that occurs with probability $0$ and therefore does not affect the expectation.
-Thus, $\funta(\len)$ is given by the right-hand side of \eqref{eq: funca} plus $1$.
+As noted in Proposition \ref{prop: i+j-1}, the number of tiles visited by a segment is the count of its grid line crossings plus $1$, unless it exactly passes through any grid points, but that occurs with probability $0$ and therefore does not affect the expected value.
+Thus, $\funta(\len)$ is obtained by adding $1$ to \eqref{eq: funca}, which gives \eqref{eq: funta, form}.
 \end{proof}
@@
 
-The results in Theorem~\ref{theo: funta, form} and Proposition~\ref{prop: Pr i} make clear the relationship between the problem considered in this section here and Buffon's needle problem, i.e.~the probability that a random segment of fixed length crosses a line in a regular structure of parallel strips \cite[section 1.1]{Mathai99}.
+The results in Theorem~\ref{theo: funta, form} and Proposition~\ref{prop: Pr i} make clear the relationship between the problem considered here and Buffon's needle problem, i.e.~the probability that a random segment of fixed length crosses a line in a regular structure of parallel strips \cite[section 1.1]{Mathai99}.
 Firstly, since the number $\pi$ is involved in \eqref{eq: funta, form}, it is possible to design a simple probabilistic experiment to estimate the value of $\pi$, as in Buffon's original experiment.
@@
 
-The probability of visiting the maximum number of tiles for a rectangular grid is difficult to compute, due to the irregularity of the optimal $(i,j)$ pairs (see \S\ref{part: fund results}).
-In the square case, however, the simplicity of the set makes the problem more tractable.
+The probability of visiting the maximum number of tiles for a rectangular grid is difficult to compute, due to the irregularity of the optimal $(i,j)$ pairs (as was discussed in \S\ref{part: fund results}; in particular see Figure~\ref{fig: pairs_bounds}).
+In the square case, however, the simplicity of the set of optimal pairs makes the problem more tractable.
 

### 2021-12-02 paper.tex
Content changes in relationship with Buffon's needle problem, third item (Pr[i>=n])
Also some rewriting

@@
 \newcommand{\funta}{\varphi} % function that gives average number of touched tiles for real-valued length
-\newcommand{\funca}{\psi} % function that gives average number of grid line crossings for real-valued length
+%\newcommand{\funca}{\psi} % function that gives average number of grid line crossings for real-valued length
 \newcommand{\probmax}{\rho} % function that gives the probability of achieving the maximum number of touched tiles in the square case for real-valued length
@@
 
-The Buffon-Laplace needle problem \cite{Mathai99} considers a needle tossed onto a grid of rectangular tiles of width $a$ and height $b$.
+The Buffon-Laplace needle problem \cite[section 1.1]{Mathai99} considers a needle tossed onto a grid of rectangular tiles of width $a$ and height $b$.
 The number of visited tiles equals one plus the number of crossings almost surely.
@@
 Given $\len \in \mathbb R^+$, consider a segment of length $\len$ with uniformly random position and orientation.
-Specifically, the coordinates $x_1$, $y_1$ of the first endpoint are independent random variables uniformly distributed on $[0,a)$ and $[0,b)$ respectively, where $a$, $b$ are the grid parameters.
+Specifically, the coordinates $x_1, y_1$ of the first endpoint are independent random variables uniformly distributed on $[0,a)$ and $[0,b)$ respectively, where $a$, $b$ are the grid parameters.
 The orientation $\theta$ of the segment is uniformly distributed on $[0,2\pi)$.
-The coordinates $x_2$, $y_2$ of the second endpoint are determined by $x_1$, $y_1$ and $\theta$.
+The variables $x_1$, $y_1$ and $\theta$ determine the coordinates $x_2, y_2$ of the second endpoint.
 
@@
 
-This section deals with the two problems stated in \S\ref{part: intro} for random segments, namely obtaining the average number of visited tiles and the probability of visiting the maximum number of tiles.
+This section deals with the two problems stated in \S\ref{part: intro} for random segments, namely obtaining the average number of visited tiles and the probability of visiting the maximum possible number of tiles.
 Segment lengths will be assumed to be real-valued.
@@
 The square grid is addressed in \S\ref{part: probmax: unit square grid, real lengths}.
-This more specialized setting allows computation of the probability of visiting the maximum possible number of tiles.
+This more specialized setting allows computation of the probability of visiting the maximum number of tiles.
 
@@
 Suppose first that the grid is simplified to only horizontal lines with spacing $b$.
-This matches the set-up of Buffon's original needle problem \cite{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines.
+This matches the set-up of Buffon's original needle problem \cite[section 1.1]{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines.
 As shown in \cite{Ramaley69}, the expected number of lines crossed equals $2\len/(\pi b)$.
@@
 The grid crossings decompose into crossings of horizontal and vertical grid lines.
-By linearity of expectation, the expected number of crossings $\funca(\len)$ is the sum of the expectations for parallel lines with spacing $a$ and $b$ respectively, which gives
-\begin{equation}
-\label{eq: funca}
-\funca(\len) = \frac{2\len}{\pi}\left(\frac 1 a + \frac 1 b\right).
-\end{equation}
+By linearity of expectation, the expected number of crossings is the sum of the expectations for parallel lines with spacing $a$ and $b$ respectively, which gives
+\[
+\frac{2\len}{\pi}\left(\frac 1 a + \frac 1 b\right).
+\]
 As noted in Proposition \ref{prop: i+j-1}, the number of tiles visited by a segment is the count of its grid line crossings plus $1$, unless it exactly passes through any grid points, but that occurs with probability $0$ and therefore does not affect the expected value.
-Thus, $\funta(\len)$ is obtained by adding $1$ to \eqref{eq: funca}, which gives \eqref{eq: funta, form}.
+Thus $\funta(\len)$ is obtained by adding $1$ to the above expression, which gives \eqref{eq: funta, form}.
 \end{proof}
@@
 
-Given the coordinates $(x_1, y_1)$ of the first endpoint of the segment, with $0 \leq x_1 \leq a$, $0 \leq y_1 \leq b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure~\ref{fig: Pr_i}.
+Given the coordinates $(x_1, y_1)$ of the first endpoint of the segment, with $0 \leq x_1 < a$, $0 \leq y_1 < b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure~\ref{fig: Pr_i}.
 The segment orientation is a random angle $\theta$ uniformly distributed on $[0,2\pi)$.
@@
 \label{eq: Pr i = 2 Pr}
-\Pr[i \geq n] = 2 \Pr[x_2 \geq a(n-1)] = 2 \E[\Pr[x_2 \geq a(n-1) \cond x_1]].
+\Pr[i \geq n] = 2 \Pr[x_2 \geq a(n-1)].
 \end{equation}
@@
 If $a(n-1) < \len$, regardless of $x_1$ the length $\len$ is enough for $x_2$ to exceed $a(n-1)$ for some angles $\theta$.
-This is depicted in Figure~\ref{fig: Pr_i_full}, where the section of the arc with solid line represents those angles for which $x_2 > a(n-1)$.
-If $a(n-2) < \len \leq  a(n-1)$, the length will be enough provided that $x_1 \geq a(n-1)-\len$, which corresponds to the shaded region in Figure~\ref{fig: Pr_i_notfull}.
-Lastly, if $\len \leq  a(n-2)$ it is not possible for $x_2$ to exceed $a(n-1)$.
-Note that the coordinate $y_1$ is irrelevant to this.
+This is depicted in Figure~\ref{fig: Pr_i_full}, where the section of the arc with solid line represents, for a given $x_1$, those angles for which $x_2 > a(n-1)$.
+If $a(n-2) < \len \leq  a(n-1)$, the length will be enough provided that $x_1 \geq a(n-1)-\len$, and then only for certain angles.
+This restriction on $x_1$ corresponds to the shaded region in Figure~\ref{fig: Pr_i_notfull}.
+Lastly, if $\len \leq  a(n-2)$ it is not possible for $x_2$ to exceed $a(n-1)$, regardless of $x_1$ or $\theta$.
+The figure makes it clear that the coordinate $y_1$ is irrelevant to this.
 
@@
 \end{equation}
-In the first case $x_1$ has a uniform distribution on $(0,a)$, and $\Pr[x_2 \geq a(n-1)]$ is easily obtained from \eqref{eq: Pr x_2 geq a(n-1) cond} as
+In the first case $x_1$ has a uniform distribution on $[0,a)$, and $\Pr[x_2 \geq a(n-1)]$ is easily obtained from \eqref{eq: Pr x_2 geq a(n-1) cond} as
 \begin{equation}
@@
 \begin{split}
-\Pr[x_2 \geq a(n-1)] &= \frac 1 {\pi a} \int_0^a\arccos \frac{a(n-1)-x_1}{\len} \diff x_1 \\
+\Pr[x_2 \geq a(n-1)] &= \E[\Pr[x_2 \geq a(n-1) \cond x_1]] \\
+&= \frac 1 {\pi a} \int_0^a\arccos \frac{a(n-1)-x_1}{\len} \diff x_1 \\
 &= \frac{\len}{\pi a}\,\, \left( f\left( \frac{a(n-1)}{\len} \right) - f\left( \frac{a(n-2)}{\len} \right) \right),
@@
 
-The above arguments can be applied to $\Pr[j \geq n]$ if the $x$ and $y$ axes are interchanged. Thus the result is the same with $a$ replaced by $b$.
+The above arguments can be applied to $\Pr[j \geq n]$ if the $x$ and $y$ axes are interchanged. Thus the formulas are the same with $a$ replaced by $b$.
 \end{proof}
@@
 gives the average number of lines crossed in Buffon's experiment.
-For $\len \leq a$ the segment can cross at most one line, and \eqref{eq: number of lines crossed, Buffon} coincides with the probability of crossing \cite[section 1.1]{Mathai99}.
-
-Lastly, $\Pr[i \geq n]$ as computed in Proposition~\ref{prop: Pr i} for $n \geq 2$ can be seen as a generalization of the probability of crossing at least one line in Buffon's experiment, when a needle endpoint is only allowed to move in a narrower region of width $n-1$ times smaller than the width of the strips (instead of across the full strip as in Buffon's setting).
+For $\len \leq a$ the segment can cross at most one line, and \eqref{eq: number of lines crossed, Buffon} coincides with the probability of crossing \cite[theorem 1.1.1]{Mathai99}.
+
+% ***Alex: please check this, as I've done some content changes
+Lastly, $\Pr[i \geq 2]$ as computed in Proposition~\ref{prop: Pr i} can be interpreted as the probability of crossing at least one line in Buffon's experiment.
+Indeed, for $n=2$ the third line in \eqref{eq: Pr i geq n} reduces again to $2\len/(\pi a)$ for $\len \leq a$, whereas the second line reduces to $2\len f(a/\len)/(\pi a)$ for $\len > a$.
+Using \eqref{eq: f} this is expressed as
+\[
+\frac 2 \pi \arccos\frac a \len + \frac{2\len}{\pi a} \left(1 - \sqrt{1-\frac{a^2}{\len^2}} \right),
+\]
+in agreement with the known result for Buffon's needle problem with $\len > a$ \cite[theorem 1.1.2]{Mathai99}.
+On the other hand, $\Pr[i \geq n]$ for $n \geq 3$ can be seen as the probability of crossing at least one line in a modified version of Buffon's setting in which a needle endpoint is only allowed to move in a narrower region in the middle of a strip, of width $2n-3$ times smaller than the strip width.
 
@@
 
+% ***Change vertical axis scale: [0 1]? [0.5 1]?
 \begin{figure}

### 2021-12-03 paper.tex
Rewriting and corrections

@@
 
-For $\len \rightarrow \infty$, the average number of visited tiles $\funta$ has the following asymptotic slope, as stems from \eqref{eq: funta, form}:
+For $\len \rightarrow \infty$, the average number of visited tiles $\funta(\len)$ has the following asymptotic slope, as stems from \eqref{eq: funta, form}:
 \begin{equation}
@@
 \end{equation}
-On the other hand, the maximum number of visited tiles $\funt$ has an asymptotic slope given by \eqref{eq: asympt slope}.
+On the other hand, the maximum number of visited tiles $\funt(\len)$ has an asymptotic slope given by \eqref{eq: asympt slope}.
 It is interesting to consider the ratio of these values, which also gives the asymptotic ratio of average and maximum numbers of visited tiles.
-This depends only on $a/b$:
+This depends only on $a/b$, and is thus expressed as a function $\ras: \mathbb R^+ \to (0,1)$:
 \begin{equation}
@@
 = \frac
-{ \lim_{\len \rightarrow \infty} \funta(\len)/\len }
-{ \lim_{\len \rightarrow \infty} \funt(\len)/\len }
+{ \lim_{\len \rightarrow \infty} (\funta(\len)/\len) }
+{ \lim_{\len \rightarrow \infty} (\funt(\len)/\len) }
 = \frac{2\left(1 + a/b\right)}{\pi\sqrt{1 + \left(a/b\right)^2}}.
 \end{equation}
-The function $\ras$ satisfies $\ras(x) = \ras(1/x)$, and is therefore symmetric in logarithmic scale, as can be seen in Figure~\ref{fig: ratio asympt slopes}.
+The function $\ras$ satisfies $\ras(x) = \ras(1/x)$, and is therefore symmetric in logarithmic scale, as can be observed in Figure~\ref{fig: ratio asympt slopes}.
 It is unimodal with maximum value $2\sqrt 2/\pi \approx 0.9003$ for $a/b=1$, and its limit when $a/b$ tends to $0$ or $\infty$ is $2/\pi \approx 0.6366$.
@@
 \label{eq: g def}
-g(\genvar, u, v) = \int_{\arcsin \frac u {2r}}^{\arccos \frac u {2r}} \left( \genvar \sin \theta - \frac u 2 \right) \left( \genvar \cos \theta - \frac v 2 \right) \diff \theta.
-\end{equation}
-Using the identities $\cos^2 \alpha = 1-\sin^2 \alpha$ and $\cos(2\alpha) = 2\cos^2\alpha-1 = 1 - 2\sin^2\alpha$, the integral in \eqref{eq: g def} is computed as
+g(\genvar, u, v) = \frac 2 \pi \int_{\arcsin \frac v {2\genvar}}^{\arccos \frac u {2\genvar}} \left( \genvar \sin \theta - \frac u 2 \right) \left( \genvar \cos \theta - \frac v 2 \right) \diff \theta.
+\end{equation}
+Computing the integral in \eqref{eq: g def} and using the identities $\cos^2 \alpha = 1-\sin^2 \alpha$ and $\cos(2\alpha) = 2\cos^2\alpha-1 = 1 - 2\sin^2\alpha$,
 \begin{equation}

### 2021-12-03 paper.tex
Rewriting

@@
 Given $\len \in \mathbb R^+$, consider a unit square grid and a uniformly random segment of length $\len$, as previously defined.
-The probability $\probmax(\len)$ that the segment visits the maximum number of tiles $\funt(\len)$ is
-\begin{equation}
-\label{eq: probmax}
+The probability $\probmax(\len)$ that the segment visits the maximum number of tiles $\tiles = \funt(\len)$ is
+\begin{equation}
+\label{eq: probmax} %***Change order of the two summands under each square root? here and in the rest of the proof, in accordance with \eqref{eq: square, odd max tiles, lengths}. Or change there
 \begin{split}
@@
 \begin{proof}
-The approach is similar to that used in the proof of Proposition~\ref{prop: Pr i}, but conditioning on the segment orientation $\theta$ instead of on the location of its first endpoint.
-
-The first endpoint of the segment is assumed to be in the reference tile with lower-left corner $(0,0)$.
-For an arbitrary value $\tiles$, the segment visits $\tiles$ tiles almost surely if and only if its second endpoint is in a tile with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, $i+j-1=\tiles$, or in a tile symmetrical to this with respect to the line $x=1/2$ (with $i \leq 0$), or $y=1/2$ (with $j \leq 0$), or both.
-
-Consider $i,j\geq 1$ for the moment, and let $\tiles = \funt(\len)$.
-For a given orientation $\theta$, the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ if and only if shifting the segment so that its second endpoint has coordinates $(i-1,j-1)$ results in the first endpoint being in the reference tile.
-It cannot be the case that the segment ``overshoots'' past this tile, because that would result in a number of visited tiles exceeding the assumed maximum, $\tiles$.
+The approach is analogous to that used in the proof of Proposition~\ref{prop: Pr i}, but conditioning on the segment orientation $\theta$ instead of on the location of its first endpoint.
+
+For $i,j \in \mathbb Z$, the (square) tile with lower-left coordinate $(i,j)$ is defined as $[i,i+1) \times [j,j+1)$, i.e.~the tile includes its lower and left edges, and tiles are disjoint.
+The first endpoint of the segment is contained in the tile with lower-left corner $(0,0)$, which will be referred to as reference tile.
+For an arbitrary value $\tiles$, the segment visits $\tiles$ tiles almost surely if and only if its second endpoint is in a tile with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, $i+j-1=\tiles$, or in a tile symmetrical to this about the line $x=1/2$ (with $i \leq 0$), or about $y=1/2$ (with $j \leq 0$), or about both.
+
+Given $\len \in \mathbb R^+$, let $\tiles = \funt(\len)$, and consider $i,j\geq 1$ such that $i+j-1=\tiles$.
+For a fixed orientation $\theta$, the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ if and only if shifting the segment so that its second endpoint has coordinates $(i-1,j-1)$ results in the first endpoint being in the reference tile.
+It cannot be the case that the segment ``overshoots'' past this tile, because that would result in a number of visited tiles exceeding $\tiles$, which is the maximum.
 
@@
 For convenience, the angle $\theta$ is defined clockwise from the downward direction.
-For each $\theta$ in the allowed range, the valid positions for the first endpoint of the (shifted) segment are in the shaded rectangle, whose dimensions depend on $\theta$.
+For each $\theta$ in the allowed range, the valid positions for the first endpoint of the (shifted) segment are in a rectangle contained in the reference tile.
+The rectangle associated with an example $\theta$ is shown in the figure, shaded.
 The area occupied by this rectangle is the probability $\probmax_{i-1,j-1}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$.
-The desired probability, $\probmax(\len)$, will be obtained by averaging $\probmax_{i-1,j-1}(\len, \theta)$ over $\theta$, summing over all $i, j \geq 1$ such that the tile with lower-left corner $(i-1, j-1)$ can be reached from the reference tile, and then using symmetry to take into account the tiles with $i \leq 0$ or $j \leq 0$.
-
-The situation is different depending on whether $\tiles$ is odd or even.
+The desired probability $\probmax(\len)$ will be obtained by averaging $\probmax_{i-1,j-1}(\len, \theta)$ over $\theta$, summing over all $i, j \geq 1$ such that the tile with lower-left corner $(i-1, j-1)$ can be reached from the reference tile, and then using symmetry to take into account the tiles with $i \leq 0$ or $j \leq 0$.
+
+The situation is different depending on whether $\tiles = \funt(\len)$ is odd or even.
 Consider $\tiles$ odd first.
-Under this assumption, the range of lengths corresponding to the condition $\tiles = \funt(\len)$ is given by \eqref{eq: square, odd max tiles, lengths}.
+Under this assumption, the range of lengths for which $\funt(\len)$ equals $\tiles$ is given by \eqref{eq: square, odd max tiles, lengths}.
 For any length in this interval, there are some positions within the reference tile from which the tile with lower-left coordinate $((\tiles-1)/2, (\tiles-1)/2)$ can be reached.
@@
 In this case, by symmetry, each of those two tiles contributes the same probability.
-Other tiles with $i,j\geq 1$ corresponding to the same $\tiles$ do not need to be considered, because they cannot be reached with lengths satisfying \eqref{eq: square, odd max tiles, lengths}.
+Other tiles with $i,j\geq 1$, $i+j-1 = \tiles$ do not need to be considered, because they cannot be reached with lengths satisfying \eqref{eq: square, odd max tiles, lengths}.
 
@@
 This can only happen for $\tiles \geq 8$.
-No other tiles with $i,j\geq 1$ can be reached from the reference tile.
+No other tiles with $i,j\geq 1$ can be reached from the reference tile $i+j-1 = \tiles$.
 
@@
 
-According to the preceding analysis, two cases need to be distinguished for $\tiles$ odd, depending on whether \eqref{eq: probmax len odd2} is satisfied or not; and similarly two cases exist for $\tiles$ even, depending on whether \eqref{eq: probmax len even2} holds or not.
+According to the preceding analysis, two cases need to be distinguished for $\tiles = \funt(\len)$ odd, depending on whether \eqref{eq: probmax len odd2} is satisfied or not; and similarly two cases exist for $\tiles = \funt(\len)$ even, depending on whether \eqref{eq: probmax len even2} holds or not.
 These are, respectively, the four cases in \eqref{eq: probmax}.
 
-In the first case indicated in \eqref{eq: probmax}, as argued, $\tiles$ is odd and only the tile with lower-left coordinates $((\tiles-1)/2, (\tiles-1)/2)$ is reachable.
+In the first case indicated in \eqref{eq: probmax}, as previously argued, $\tiles$ is odd and only the tile with lower-left coordinates $((\tiles-1)/2, (\tiles-1)/2)$ is reachable.
 For a given $\theta$, the conditional probability $\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)$ is the area of the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_diag}:
@@
 \end{align}
-Computing the average of \eqref{eq: probmax: probmax theta odd} over $\theta$, multiplying by $4$ to include all quadrants and comparing with \eqref{eq: g def} gives $\probmax(\len)$ as
+Averaging \eqref{eq: probmax: probmax theta odd} over $\theta$, multiplying by $4$ to include the symmetrical tiles in other quadrants and using \eqref{eq: g def} gives $\probmax(\len)$ as
 \begin{equation}

### 2021-12-04 paper.tex
Rewriting

@@
 \label{eq: g def}
-g(\genvar, u, v) = \frac 2 \pi \int_{\arcsin \frac v {2\genvar}}^{\arccos \frac u {2\genvar}} \left( \genvar \sin \theta - \frac u 2 \right) \left( \genvar \cos \theta - \frac v 2 \right) \diff \theta.
+g(\genvar, u, v) = \frac 2 \pi \int_{\arcsin \frac u {2\genvar}}^{\arccos \frac v {2\genvar}} \left( \genvar \sin \theta - \frac u 2 \right) \left( \genvar \cos \theta - \frac v 2 \right) \diff \theta.
 \end{equation}
@@
 g(\genvar, u, v) &= \frac 1 {2\pi} \biggl(
-\left(\arccos\frac{u}{2\genvar}-\arcsin\frac{v}{2\genvar}\right) u v + 2\genvar^2 + \frac{u^2} 2 + \frac{v^2} 2 \\
+\left(\arccos\frac{v}{2\genvar}-\arcsin\frac{u}{2\genvar}\right) u v + 2\genvar^2 + \frac{u^2} 2 + \frac{v^2} 2 \\
 & \quad  - u \sqrt{4\genvar^2-v^2} - v \sqrt{4\genvar^2-u^2} \biggr).
@@
 \begin{equation}
-\label{eq: probmax} %***Change order of the two summands under each square root? here and in the rest of the proof, in accordance with \eqref{eq: square, odd max tiles, lengths}. Or change there
+\label{eq: probmax} %***Change order of the two summands under each square root here and in the rest of the proof, in accordance with \eqref{eq: square, odd max tiles, lengths}.
+(Or change there, but I think that is less preferable because it would correspond to using the y coordinate, then the x coordinate from the figure)
 \begin{split}
@@
 
-For $i,j \in \mathbb Z$, the (square) tile with lower-left coordinate $(i,j)$ is defined as $[i,i+1) \times [j,j+1)$, i.e.~the tile includes its lower and left edges, and tiles are disjoint.
+For $i,j \in \mathbb Z$, the (square) tile with lower-left coordinate $(i,j)$ is defined as $[i,i+1) \times [j,j+1)$, i.e.~including its lower and left edges.
 The first endpoint of the segment is contained in the tile with lower-left corner $(0,0)$, which will be referred to as reference tile.
-For an arbitrary value $\tiles$, the segment visits $\tiles$ tiles almost surely if and only if its second endpoint is in a tile with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, $i+j-1=\tiles$, or in a tile symmetrical to this about the line $x=1/2$ (with $i \leq 0$), or about $y=1/2$ (with $j \leq 0$), or about both.
+For an arbitrary value $\tiles$, the segment visits $\tiles$ tiles almost surely if its second endpoint is in a tile with lower-left corner $(i-1,j-1)$ such that $i,j\geq 1$ and $i+j-1=\tiles$, or in a tile symmetrical to this about the line $x=1/2$ (with $i \leq 0$), or about $y=1/2$ (with $j \leq 0$), or about both.
 
@@
 }%
-\caption{Segment orientations and positions of the first endpoint within the reference tile for a given number $\tiles$ of visited tiles. Example for odd $\tiles$
+\caption{Segment orientations and positions of the first endpoint within the reference tile for a given $\tiles = \funt(\len)$, with $\tiles$ odd. Example for $\tiles = 7$
 }%
@@
 The rectangle associated with an example $\theta$ is shown in the figure, shaded.
-The area occupied by this rectangle is the probability $\probmax_{i-1,j-1}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$.
+The area of this rectangle is the probability $\probmax_{i-1,j-1}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$.
 The desired probability $\probmax(\len)$ will be obtained by averaging $\probmax_{i-1,j-1}(\len, \theta)$ over $\theta$, summing over all $i, j \geq 1$ such that the tile with lower-left corner $(i-1, j-1)$ can be reached from the reference tile, and then using symmetry to take into account the tiles with $i \leq 0$ or $j \leq 0$.
@@
 This can only happen for $\tiles \geq 5$ (for $\tiles=3$ the inequality \eqref{eq: probmax len odd2} cannot be satisfied for lengths in the range given by \eqref{eq: square, odd max tiles, lengths}).
-In this case, by symmetry, each of those two tiles contributes the same probability.
+By symmetry, each of those two tiles contributes the same probability.
 Other tiles with $i,j\geq 1$, $i+j-1 = \tiles$ do not need to be considered, because they cannot be reached with lengths satisfying \eqref{eq: square, odd max tiles, lengths}.
@@
 
-The above two paragraphs show that, among all possible tiles with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, those whose centers lie on the lines $x=1/2$ or $y=1/2$ (i.e.~those with $i=1$ of $j=1$) are never reachable from the reference tile, in the sense that a segment that touches the maximum number of tiles cannot have its second endpoint in those tiles.
-This implies that $\probmax(\len)$ can be computed by considering only the tiles with $i,j\geq 1$ (actually $i,j \geq 2$), and then multiplying by $4$ to include the symmetrical tiles with respect to $x=1/2$ and $y=1/2$.
+From the above two paragraphs it stems that, among all possible tiles with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, those whose centers lie on the lines $x=1/2$ or $y=1/2$ (i.e.~those with $i=1$ of $j=1$) are never reachable from the reference tile, in the sense that a segment that touches the maximum number of tiles cannot have its second endpoint in those tiles.
+This implies that $\probmax(\len)$ can be computed by considering only the tiles with $i,j\geq 1$ (actually $i,j \geq 2$), and then multiplying by $4$ to include the symmetrical tiles with respect to the lines $x=1/2$ and $y=1/2$.
 As an example, Figure~\ref{fig: probmax_possible_tiles} shows the reference tile (solid line) and the tiles that can contain the second endpoint (dotted lines) for $\funt(\len) = 5$.
@@
 \includegraphics[width=.52\textwidth]{probmax_possible_tiles}%
-\caption{Tiles that can contain the second endpoint of a segment that visits the maximum number of tiles $\funt(\len)$ in a square grid (example for $\funt(\len)=5$)%
+\caption{Tiles that can contain the second endpoint of a segment that visits the maximum number of tiles $\tiles= \funt(\len)$, with $\tiles$ odd. Example for $\tiles = 5$%
 }%
@@
 According to the preceding analysis, two cases need to be distinguished for $\tiles = \funt(\len)$ odd, depending on whether \eqref{eq: probmax len odd2} is satisfied or not; and similarly two cases exist for $\tiles = \funt(\len)$ even, depending on whether \eqref{eq: probmax len even2} holds or not.
-These are, respectively, the four cases in \eqref{eq: probmax}.
+These are the four cases in \eqref{eq: probmax}.
 
@@
 \end{equation}
-The range of allowed values for $\theta$, as can be seen from the figure, is $(\theta_0, \theta_1)$ with
+The range of allowed values for $\theta$, as can be deduced from the figure, is $(\theta_0, \theta_1)$ with
 \begin{align}
@@
 \end{equation}
-in accordance with \eqref{eq: probmax}.
+in accordance with the first expression in \eqref{eq: probmax}.
 

### 2021-12-04 figures/figures.m

@@ -1771,11 +1771,12 @@ limit(G*t, t, inf)
 syms t, assume(t, 'real'), pretty(simplify(diff((acos((t-3)/hypot(t-3, t-1)) - asin((t-3)/hypot(t-1, t-3))), t)))
 
 
-%% #67. Figure
+%% #67. Figure: probmax
 
 clear, clc
 a = 1;
-g = @(r,u,v) ((acos(u/2./r)-asin(v/2./r)).*u.*v + 2*r.^2 + u.^2/2 + v.^2/2 - u.*sqrt(4*r.^2-v.^2) - v.*sqrt(4*r.^2-u.^2))/2/pi;
+%g= @(r,u,v) ((acos(u/2./r)-asin(v/2./r)).*u.*v + 2*r.^2 + u.^2/2 + v.^2/2 - u.*sqrt(4*r.^2-v.^2) - v.*sqrt(4*r.^2-u.^2))/2/pi; % #65
+g = @(r,u,v) ((acos(v/2./r)-asin(u/2./r)).*u.*v + 2*r.^2 + u.^2/2 + v.^2/2 - u.*sqrt(4*r.^2-v.^2) - v.*sqrt(4*r.^2-u.^2))/2/pi; % #74
 L_fine = .01:.01:16;
 ii = ceil(L_fine/sqrt(2)) + 1;
 jj = ceil(sqrt(L_fine.^2 - (ii-2).^2 )) + 1;
@@ -1929,3 +1930,38 @@ as_ratio_slopes = 2/pi*(1+a/b)./hypot(1, a/b)
 plot(funta./funt), hold on, plot(xlim, repmat(as_ratio_slopes, 1, 2))
 
 
+%% #74. Like #65 but defining the g function differently (gives the same result)
+% Checked
+
+clear, clc
+a = 1;
+L = .1:.1:10; %3.8:.1:4.2;%3:.1:3.5;
+N = 3e5;
+ii = ceil(L/sqrt(2)) + 1;
+jj = ceil(sqrt(L.^2 - (ii-2).^2 )) + 1;
+funt = ii+jj-1 % formula valid for real-valued lengths
+z0 = rand(N, numel(L)) + 1j*rand(N, numel(L));
+z1 = z0 + L.*exp(1j*2*pi*rand(N, numel(L)));
+ii = ceil(abs(real(z1))) + (real(z1)<0); % add 1 for negative
[…]

### 2021-12-04 paper.tex
The function g is symmetric in its second and third arguments

@@
 
-% The rounding down (floor) and rounding up (ceiling) operations will be denoted as $\lfloor \cdot \rfloor$ and $\lceil \cdot \rceil$.
-The notation $[\mathsf S]$ will be used for the Iverson function, which equals $1$ if statement $\mathsf S$ is true and $0$ otherwise.
+The symbols $\lfloor x \rfloor$ and $\lceil x \rceil$ respectively denote the floor (rounding down) and ceiling (rounding up) operations.
+The functions $\arcsin x$ and $\arccos x$ are defined as their principal branches in the usual way.
+In particular, the output angles are in $[0, \pi/2]$ for $x \in [0,1]$.
+% The notation $[\mathsf S]$ will be used for the Iverson function, which equals $1$ if statement $\mathsf S$ is true and $0$ otherwise.
 
@@
 \end{equation}
+Taking into account that $\arcsin x = \pi/2 - \arccos x$, it follows from \eqref{eq: g fin} that $g(\genvar, u, v) = g(\genvar, v, u)$.
 

### 2021-12-05 paper.tex
Rewriting

@@
 Namely, $\funl^2(\tiles)$ is the sum of the two terms that appear in that expression.
-The first term is the squared distance from $(2a, 2b)$ to the diagonal line defined by $i+j-1=t$; and the second term is additional squared distance incurred from rounding $i$, $j$ to integer values.
+The first term is the squared distance from $(2a, 2b)$ to the diagonal line defined by $i+j-1 = \tiles$; and the second term is additional squared distance incurred from rounding $i$, $j$ to integer values.
 
@@
 This is equal to \frac{\sqrt{(\tiles-2)^2+(\tiles-4)^2}} {2}.
-Here we are using the second expression, which is probably needed for the random analysis.
+Here we are using the latter expression, which is probably needed for the random analysis.
 If so, the expression in the corollary should be changed from the first to the second.
@@
 \probmax(\len) & = g\left(\len, 
-\tiles-3, \tiles-3 \right) \quad\ \text{for } \tiles \text{ odd, } \textstyle \len \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2 \right]; \\
+\tiles-3, \tiles-3 \right) \quad\ \text{for } \tiles \text{ odd, } \textstyle \len \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2 \right] \\
 \probmax(\len) & = g\left(\len, \tiles-3, \tiles-3 \right) + 2 g\left(\len, \tiles-5, \tiles-1 \right) \\
-& \text{for } \tiles \text{ odd, } \textstyle \len \in \left (\frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2,  \frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2 \right]; \\
-\probmax(\len) & = 2 g\left(\len, \tiles-4, \tiles-2 \right) \quad \text{for } \tiles \text{ even, } \textstyle \len \in \left( \frac{\sqrt{(\tiles-4)^2+(\tiles-2)^2}} 2, \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2 \right]; \\
+& \text{for } \tiles \text{ odd, } \textstyle \len \in \left (\frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2,  \frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2 \right] \\
+\probmax(\len) & = 2 g\left(\len, \tiles-4, \tiles-2 \right) \quad \text{for } \tiles \text{ even, } \textstyle \len \in \left( \frac{\sqrt{(\tiles-4)^2+(\tiles-2)^2}} 2, \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2 \right] \\
 \probmax(\len) & = 2 g\left(\len, \tiles-4, \tiles-2 \right) + 2 g\left(\len, \tiles-6, \tiles \right) \\
@@
 }%
-\caption{Segment orientations and positions of the first endpoint within the reference tile for a given $\tiles = \funt(\len)$, with $\tiles$ odd. Example for $\tiles = 7$
+\caption{Segment orientations and positions for a given $\tiles = \funt(\len)$ on a unit square grid, with $\tiles$ odd. Example for $\tiles = 7$
 }%
@@
 \includegraphics[width=.52\textwidth]{probmax_possible_tiles}%
-\caption{Tiles that can contain the second endpoint of a segment that visits the maximum number of tiles $\tiles= \funt(\len)$, with $\tiles$ odd. Example for $\tiles = 5$%
+\caption{Possible tiles for the endpoints of a segment that visits the maximum number of tiles $\tiles= \funt(\len)$ on a unit square grid, with $\tiles$ odd. Example for $\tiles = 5$%
 }%
@@
 The technique used in the proof of Theorem~\ref{theo: probmax, sq} could in principle be used for obtaining the probability that the number of tiles visited by the segment equals or exceeds any given value $t < \funt(\len)$.
-The resulting expressions, however, are cumbersome, because more cases (more subintervals for the length) need to be distinguished depending on how many tiles with lower-left corner $(i-1,j-1)$ such that $i+j-1=t$ can be reached from the reference tile.
+The resulting expressions, however, are cumbersome, because more cases (more subintervals for the length) need to be distinguished depending on how many tiles with lower-left corner $(i-1,j-1)$ such that $i+j-1=\tiles$ can be reached from the reference tile.
 
@@
 \includegraphics[width=.75\textwidth]{probmax_len}%
-\caption{Probability $\probmax(\len)$ that a random segment of length $\len$ visits the maximum number of tiles%
+\caption{Probability $\probmax(\len)$ that a random segment of length $\len$ visits the maximum number of tiles on a unit square grid%
 }%
@@
 Figure~\ref{fig: probmax_len} shows the probability $\probmax(\len)$ of visiting the maximum number of tiles on a unit square grid, as given by \eqref{eq: probmax}.
-As $\len$ grows, $\probmax(\len)$ has a jump when the maximum number of tiles that can be visited increases by $1$ (and there the function takes the value approached from the left).
-This happens when $\len$ equals the right endpoint of the interval \eqref{eq: square, odd max tiles, lengths} for $\tiles$ odd or of \eqref{eq: square, even max tiles, lengths} for $\tiles$ even.
-Let these length values be denoted as $\len_\tiles$:
+As $\len$ grows, $\probmax(\len)$ has a jump when $\funt(\len)$ increases by $1$ (and there $\probmax(\len)$ takes the value approached from the left).
+This happens when $\len$ equals the right endpoint of the interval \eqref{eq: square, odd max tiles, lengths} for $\tiles$ odd or of the interval \eqref{eq: square, even max tiles, lengths} for $\tiles$ even.
+These length values will be denoted as $\len_\tiles$:
 \begin{equation}
@@
 \displaystyle
-\frac{\tiles-2}{\sqrt{2}} & \text{ for $\tiles$ odd, $\tiles \geq 4$}.
+\frac{\tiles-2}{\sqrt{2}} & \text{ for $\tiles$ odd, $\tiles \geq 4$},
 \end{cases}
 \end{equation}
-For example, the first continuous section seen in Figure~\ref{fig: probmax_len} corresponds to a maximum number of tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, where $\len_3= 1$.
+and satisfy $\funt(\len_\tiles) = \tiles$.
+
+The first continuous section seen in Figure~\ref{fig: probmax_len} corresponds to a maximum number of visited tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, where $\len_3= 1$.
 The second corresponds to $\tiles=4$, for lengths in $(\len_3, \len_4]$, where $\len_4= \sqrt{2}$.
@@
 \label{eq: probmax liminf}
-\liminf_{\len \rightarrow \infty}\, \probmax(\len) & = 0 \\ 
+\liminf_{\len \rightarrow \infty}\, \probmax(\len) & = 0, \\ 
 \label{eq: probmax lim odd}
 \lim_{\substack{\tiles \rightarrow \infty \\
-\tiles \ \mathrm{odd}}}\, \tiles \probmax(\len_\tiles) & = \frac 2 \pi \\
+\tiles \ \mathrm{odd}}}\, \tiles \probmax(\len_\tiles) & = \frac 2 \pi, \\
 \label{eq: probmax lim even}
-\lim_{\substack{\tiles \rightarrow \infty \\ \tiles \ \mathrm{even}}}\, \tiles \probmax(\len_\tiles) & = \frac 8 {3\pi} \\
+\lim_{\substack{\tiles \rightarrow \infty \\ \tiles \ \mathrm{even}}}\, \tiles \probmax(\len_\tiles) & = \frac 8 {3\pi}, \\
 \label{eq: probmax limsup}

### 2021-12-05 figures/figures.m

@@ -1795,7 +1795,7 @@ plot(L_fine.*(ind_odd|ind_odd2)./(ind_odd|ind_odd2), result_computed.*(ind_odd|i
 set(gca, 'colororderindex', 1)
 plot(L_fine.*(ind_even|ind_even2)./(ind_even|ind_even2), result_computed.*(ind_even|ind_even2)./(ind_even|ind_even2), '-')
 %plot(L_fine, result_computed, '-')
-legend({'Odd maximum number of squares' 'Even maximum number of squares'})
+legend({'Odd maximum number of tiles' 'Even maximum number of tiles'})
 xlabel(char(hex2dec('2113'))), ylabel([char(hex2dec('03C1')) '(' char(hex2dec('2113')) ')'])
 
 

### 2021-12-05 paper.tex
Rewriting

@@
 
-Theorems~\ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the first two problems stated in \S\ref{part: intro}; they also provide a way to actually position a segment of length $\len$ or slightly greater than $\funl(\tiles)$, respectively, so that it visits $\funt(\len)$ or $\tiles$ tiles.
+Theorems \ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the first two problems stated in \S\ref{part: intro}; they also provide a way to actually position a segment of length $\len$ or slightly greater than $\funl(\tiles)$, respectively, so that it visits $\funt(\len)$ or $\tiles$ tiles.
 Namely, for $\isoli$, $\jsoli$ computed as in the corresponding theorem, the segment should have its endpoints in the interiors of two tiles separated $\isoli-1$ steps horizontally and $\jsoli-1$ steps vertically, with its exact position and orientation  adjusted to avoid any grid points.
@@
 \displaystyle
-\sqrt{\frac{(\tiles-3)^2+1} {2}} & \text{for } \tiles \text{ even, } \tiles \geq 4,
+\frac{\sqrt{(\tiles-4)^2+(\tiles-2)^2}} {2} & \text{for } \tiles \text{ even, } \tiles \geq 4,
 \end{cases}
@@
 \end{corollary}
+% Theorem~\ref{theo: funl, form} for $a=b=1$ gives $\sqrt{\frac{(\tiles-3)^2+1} {2}}$.
+Instead, this corollary uses the equivalent expression $\frac{\sqrt{(\tiles-4)^2+(\tiles-2)^2}} {2}$, which is better for subsequent analysis.
 
@@
 \label{eq: square, odd max tiles, lengths}
-\len \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{ \sqrt{(\tiles-1)^2+(\tiles-3)^2}}{2} \right].
+\len \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{ \sqrt{(\tiles-3)^2+(\tiles-1)^2}}{2} \right].
 \end{equation}
@@
 \label{eq: square, even max tiles, lengths}
-\len \in \left(\frac{ \sqrt{(\tiles-2)^2+(\tiles-4)^2}}{2}, \frac{\tiles-2}{\sqrt{2}} \right].
-\end{equation}
-% *** The corollary says \sqrt{\frac{(\tiles-3)^2+1} {2}}, which comes from the corresponding theorem.
-This is equal to \frac{\sqrt{(\tiles-2)^2+(\tiles-4)^2}} {2}.
-Here we are using the latter expression, which is probably needed for the random analysis.
-If so, the expression in the corollary should be changed from the first to the second.
-(But this change cannnot be made in the theorem it comes from.
-It's hopefully not a problem that the expression changes from the theorem to the corollary).
+\len \in \left(\frac{ \sqrt{(\tiles-4)^2+(\tiles-2)^2}}{2}, \frac{\tiles-2}{\sqrt{2}} \right].
+\end{equation}
 
@@
 \]
-As noted in Proposition \ref{prop: i+j-1}, the number of tiles visited by a segment is the count of its grid line crossings plus $1$, unless it exactly passes through any grid points, but that occurs with probability $0$ and therefore does not affect the expected value.
+As noted in Proposition~\ref{prop: i+j-1}, the number of tiles visited by a segment is the count of its grid line crossings plus $1$, unless it exactly passes through any grid points, but that occurs with probability $0$ and therefore does not affect the expected value.
 Thus $\funta(\len)$ is obtained by adding $1$ to the above expression, which gives \eqref{eq: funta, form}.
@@
 \begin{proof}
-The probability $\probmax(\len)$ becomes $0$ whenever the length causes the maximum number of visited tiles to increase by $1$. This establishes \eqref{eq: probmax liminf}.
+The probability $\probmax(\len)$ becomes $0$ whenever the length causes the maximum number of visited tiles to increase by $1$. This proves \eqref{eq: probmax liminf}.
 
 For \eqref{eq: probmax lim odd} it suffices to note that
-\begin{align}
-\label{eq: tedious limit, odd 1}
-\lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2, \tiles-3, \tiles-3 \right) &= \frac 2 {3\pi} \\
-\label{eq: tedious limit, odd 2}
-\lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2, \tiles-5, \tiles-1 \right)
-& = \frac 2 {3\pi},
-\end{align}
+\begin{equation}
+\begin{split}
+\label{eq: tedious limit, odd}
+&\lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2, \tiles-3, \tiles-3 \right) \\
+&\quad = \lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2, \tiles-5, \tiles-1 \right) = \frac 2 {3\pi},
+\end{split}
+\end{equation}
 as can be (somewhat tediously) checked using L'H\^opital's rule \cite[section~5.3]{Abbott15}.
@@
 Similarly,
-\begin{align}
-\label{eq: tedious limit, even 1}
-\lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\tiles-2}{\sqrt 2}, \tiles-4, \tiles-2 \right) &= \frac 2 {3\pi} \\
-\label{eq: tedious limit, even 2}
+\begin{equation}
+\label{eq: tedious limit, even}
+\lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\tiles-2}{\sqrt 2}, \tiles-4, \tiles-2 \right) =
 \lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\tiles-2}{\sqrt 2}, \tiles-6, \tiles \right)
-& = \frac 2 {3\pi},
-\end{align}
+= \frac 2 {3\pi},
+\end{equation}
 and combining with the fourth equality in \eqref{eq: probmax} gives \eqref{eq: probmax lim even}.

### 2021-12-05 paper.tex
Abstract

@@
 \begin{abstract}
-***
-
-\emph{Keywords:} ***.
-
-%\emph{MSC2010:} 65C10, 65C50.
+The relationship between the length of a line segment and the number of tiles it visits on a rectangular grid is addressed in this paper.
+For a segment with deterministic but arbitrary position and orientation, the maximum number of visited tiles given its length and the infimum length for a desired number of visited tiles are characterized.
+For a segment with random position and orientation, the average number of visited tiles and the probability of visiting the maximum number of tiles are studied.
+These questions are related to Buffon's needle problem and its extension by Laplace.
+
+\emph{Keywords:} Discrete geometry, Geometric probability, Rectangular lattice, Buffon's needle problem.
+
+\emph{MSC2020:} 52C99, 60D05. % ***Alex: check if there's a better one
+% 52C99: Discrete geometry: None of the above, but in this section
+% 60D05: Geometric probability and stochastic geometry
 \end{abstract}
@@
 \begin{equation}
-\label{eq: probmax} %***Change order of the two summands under each square root here and in the rest of the proof, in accordance with \eqref{eq: square, odd max tiles, lengths}.
-(Or change there, but I think that is less preferable because it would correspond to using the y coordinate, then the x coordinate from the figure)
+\label{eq: probmax}
 \begin{split}

### 2021-12-06 paper.tex
Changed abstract

@@
 
-\begin{abstract}
+\begin{abstract} % ***Alex: please check / revise 
 The relationship between the length of a line segment and the number of tiles it visits on a rectangular grid is addressed in this paper.
-For a segment with deterministic but arbitrary position and orientation, the maximum number of visited tiles given its length and the infimum length for a desired number of visited tiles are characterized.
-For a segment with random position and orientation, the average number of visited tiles and the probability of visiting the maximum number of tiles are studied.
+The square grid is also considered explicitly, as some of the specific problems studied are more tractable in that particular case.
+The segment position and orientation are modelled as either deterministic or random.
+In the former setting, the maximum number of visited tiles given the segment length and the infimum length for a desired number of visited tiles are characterized.
+For a segment of fixed length with random position and orientation, the average number of visited tiles and the probability of visiting the maximum number of tiles on a square grid are studied.
 These questions are related to Buffon's needle problem and its extension by Laplace.
@@
 
-\emph{MSC2020:} 52C99, 60D05. % ***Alex: check if there's a better one
+\emph{MSC2020:} 52C99, 60D05. % ***Alex: is there a better one?
 % 52C99: Discrete geometry: None of the above, but in this section

### 2021-12-08 paper.tex
Rewriting

@@
 
-\addtolength{\textheight}{28mm}
-\addtolength{\voffset}{-17mm}
+\addtolength{\textheight}{34mm}
+\addtolength{\voffset}{-20mm}
 \addtolength{\textwidth}{-3mm}
@@
 
-As an example of the first question \ref{prob: determ dir}, consider $a=1.35$, $b=1$.
+As an example of question \ref{prob: determ dir}, consider $a=1.35$, $b=1$.
 A segment of unit length can be placed as shown in Figure~\ref{fig: examples} (left) to make it visit $3$ tiles.
@@
 The segment orientation has a \emph{uniform} distribution on $[0,2\pi)$, and is \emph{independent} of the endpoint position.
-Solving the problem of how many tiles the segment visits on average \ref{prob: rand ave} also answers, as will be seen, the inverse question (segment length to visit a given number of tiles on average).
-A natural, related question \ref{prob: rand prob} is with what probability the segment visits the maximum number of tiles.
-
-% ***Do we need to discuss potential applications of this (if we can find any)? This probably depends on the journal
-% ***Do we need a section with conclusions? Probably not; depends on the journal
+Solving the problem \ref{prob: rand ave} of how many tiles the segment visits on average also answers, as will be seen, the inverse question (segment length to visit a given number of tiles on average).
+A natural, related problem \ref{prob: rand prob} is with what probability the segment visits the maximum number of tiles.
+
+% Do we need to discuss potential applications of this (if we can find any)? This probably depends on the journal
+% Do we need a section with conclusions? Probably not; depends on the journal
 
@@
 and $(i-2)^2 a^2 + (j-2)^2 b^2$ is expressed as $(x-2a)^2+(y-2b)^2$.
-The point minimizing this quadratic function along the line \eqref{eq: line 1} is the intersection of that with the perpendicular line passing through $(2a,2b)$,
+The point minimizing this quadratic function along the line \eqref{eq: line 1} is the intersection of the latter with the perpendicular line passing through $(2a,2b)$,
 \begin{equation}
@@
 \end{equation}
-An example with $a=1.35$, $b=1$, $\tiles=7$ is depicted in Figure~\ref{fig: min suff set, form}, where \eqref{eq: line 2} is shown with solid line.
+An example with $a=1.35$, $b=1$, $\tiles=7$ is shown in Figure~\ref{fig: min suff set, form}, where \eqref{eq: line 2} is depicted as a solid line.
 Solving the system of equations \eqref{eq: line 1} and \eqref{eq: line 2} gives
 \begin{align}
-\frac x a - 2 &= \frac{(t-3)b^2}{a^2+b^2}, \\
-\frac y b - 2 &= \frac{(t-3)a^2}{a^2+b^2}.
+\frac x a - 2 &= \frac{(\tiles-3)b^2}{a^2+b^2}, \\
+\frac y b - 2 &= \frac{(\tiles-3)a^2}{a^2+b^2}.
 \end{align}
@@
 \label{eq: prop: min suff set, form: isolr}
-\isolr &= \frac{(t-3)b^2}{a^2+b^2} + 2, \\
+\isolr &= \frac{(\tiles-3)b^2}{a^2+b^2} + 2, \\
 \label{eq: prop: min suff set, form: jsolr}
-\jsolr &= \frac{(t-3)a^2}{a^2+b^2} + 2.
+\jsolr &= \frac{(\tiles-3)a^2}{a^2+b^2} + 2.
 \end{align}
@@
 The variables $i$, $j$ are actually limited to integer values.
-The pair $(i_\tiles, j_\tiles)$ that minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$ along the line $i+j-1=\tiles$ restricted to integer values is either $(\lceil \isolr \rceil, \lfloor \jsolr \rfloor)$ or $(\lfloor \isolr \rfloor, \lceil \jsolr \rceil)$, whichever gives $(i_\tiles a, j_\tiles b)$ closest to $(\isolr a, \jsolr b)$, as illustrated in Figure~\ref{fig: min suff set, form}.
+The pair $(i_\tiles, j_\tiles) \in \mathbb Z^2$ that minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$ along the line $i+j-1=\tiles$ is either $(\lceil \isolr \rceil, \lfloor \jsolr \rfloor)$ or $(\lfloor \isolr \rfloor, \lceil \jsolr \rceil)$, whichever gives $(i_\tiles a, j_\tiles b)$ closest to $(\isolr a, \jsolr b)$, as illustrated in Figure~\ref{fig: min suff set, form}.
 In case of a tie the first of the two pairs is (arbitrarily) chosen.
@@
 \label{eq: prop: min suff set, form: isoli ineq}
-\frac{(t-3)b^2}{a^2+b^2} + \frac 3 2 < i_\tiles \leq \frac{(t-3)b^2}{a^2+b^2} + \frac 5 2,
+\frac{(\tiles-3)b^2}{a^2+b^2} + \frac 3 2 < i_\tiles \leq \frac{(\tiles-3)b^2}{a^2+b^2} + \frac 5 2,
 \end{equation}
@@
 \label{eq: prop: min suff set, form: jsoli ineq}
-\frac{(t-3)a^2}{a^2+b^2} + \frac 3 2 \leq j_\tiles < \frac{(t-3)a^2}{a^2+b^2} + \frac 5 2.
+\frac{(\tiles-3)a^2}{a^2+b^2} + \frac 3 2 \leq j_\tiles < \frac{(\tiles-3)a^2}{a^2+b^2} + \frac 5 2.
 \end{equation}
@@
 The pair $(i_\tiles, j_\tiles)$ is strictly below the line \eqref{eq: upper bound, line} because the used inequalities are strict.
-Similarly, the bounding line \eqref{eq: lower bound, line} results from the second inequality in \eqref{eq: prop: min suff set, form: isoli ineq} and the first in \eqref{eq: prop: min suff set, form: jsoli ineq}, and the fact that those inequalities are not strict implies that the bound \eqref{eq: lower bound, line} can actually be attained.
+Similarly, \eqref{eq: lower bound, line} results from the second inequality in \eqref{eq: prop: min suff set, form: isoli ineq} and the first in \eqref{eq: prop: min suff set, form: jsoli ineq}, and the fact that those inequalities are not strict implies that the bound \eqref{eq: lower bound, line} can actually be attained.
 \end{proof}
@@
 \label{fig: pairs_bounds_1}%
-\includegraphics[height=.24\textheight]{pairs_bounds_1}% *!*
+\includegraphics[height=.235\textheight]{pairs_bounds_1}% *!*
 }\hfill%
@@
 \label{fig: pairs_bounds_1p35}%
-\includegraphics[height=.24\textheight]{pairs_bounds_1p35}% *!*
+\includegraphics[height=.235\textheight]{pairs_bounds_1p35}% *!*
 }\hfill%
@@
 \label{fig: pairs_bounds_sqrt2}%
-\includegraphics[height=.24\textheight]{pairs_bounds_sqrt2}% *!*
+\includegraphics[height=.235\textheight]{pairs_bounds_sqrt2}% *!*
 }%
@@
 \label{eq: theo: funl, form; funl 2}
-\funl(\tiles) = \sqrt{\frac{(t-3)^2 a^2 b^2}{a^2+b^2} + \roundterm^2(a^2+b^2)}
+\funl(\tiles) = \sqrt{\frac{(\tiles-3)^2 a^2 b^2}{a^2+b^2} + \roundterm^2(a^2+b^2)}
 \end{equation}
@@
 \end{equation}
-According to \eqref{eq: funt funl}, $\funt(\len)$ is obtained by the largest positive integer $t$ such that 
+According to \eqref{eq: funt funl}, $\funt(\len)$ is obtained by the largest positive integer $\tiles$ such that 
 \begin{equation}
@@
 
-The technique used in the proof of Theorem~\ref{theo: probmax, sq} could in principle be used for obtaining the probability that the number of tiles visited by the segment equals or exceeds any given value $t < \funt(\len)$.
+The technique used in the proof of Theorem~\ref{theo: probmax, sq} could in principle be used for obtaining the probability that the number of tiles visited by the segment equals or exceeds any given value $\tiles < \funt(\len)$.
 The resulting expressions, however, are cumbersome, because more cases (more subintervals for the length) need to be distinguished depending on how many tiles with lower-left corner $(i-1,j-1)$ such that $i+j-1=\tiles$ can be reached from the reference tile.

### 2021-12-30 paper.tex
Scale in Figure ratio_asympt_slopes
Wording ("line" could be confusing)

@@
 
-\emph{MSC2020:} 52C99, 60D05. % ***Alex: is there a better one?
+\emph{MSC2020:} 52C99, 60D05.
 % 52C99: Discrete geometry: None of the above, but in this section
@@
 Lastly, $\Pr[i \geq 2]$ as computed in Proposition~\ref{prop: Pr i} can be interpreted as the probability of crossing at least one line in Buffon's experiment.
-Indeed, for $n=2$ the third line in \eqref{eq: Pr i geq n} reduces again to $2\len/(\pi a)$ for $\len \leq a$, whereas the second line reduces to $2\len f(a/\len)/(\pi a)$ for $\len > a$.
+Indeed, for $n=2$ the third expression in \eqref{eq: Pr i geq n} reduces again to $2\len/(\pi a)$ for $\len \leq a$, whereas the second reduces to $2\len f(a/\len)/(\pi a)$ for $\len > a$.
 Using \eqref{eq: f} this is expressed as
@@
 
-% ***Change vertical axis scale: [0 1]? [0.5 1]?
 \begin{figure}

### 2022-01-02 paper.tex
Revised abstract

@@
 
-\begin{abstract} % ***Alex: please check / revise 
-The relationship between the length of a line segment and the number of tiles it visits on a rectangular grid is addressed in this paper.
+\begin{abstract}
+Consider a line segment placed on a two-dimensional grid of rectangular tiles.
+This paper addresses the relationship between the length of the segment and the number of tiles it visits (i.e.
+has intersection with).
 The square grid is also considered explicitly, as some of the specific problems studied are more tractable in that particular case.
-The segment position and orientation are modelled as either deterministic or random.
-In the former setting, the maximum number of visited tiles given the segment length and the infimum length for a desired number of visited tiles are characterized.
-For a segment of fixed length with random position and orientation, the average number of visited tiles and the probability of visiting the maximum number of tiles on a square grid are studied.
+The segment position and orientation can be modelled as either deterministic or random.
+In the deterministic setting, the maximum possible number of visited tiles is characterized for a given length, and conversely, the infimum segment length needed to visit a desired number of tiles is analyzed.
+The probabilistic setting regards the segment position and orientation as random.
+The average number of visited tiles and the probability of visiting the maximum number of tiles on a square grid are studied as a function of segment length.
 These questions are related to Buffon's needle problem and its extension by Laplace.

### 2022-01-04 paper.tex
Clarification about Pr[i>=n] generalizing Buffon's needle setting

@@
 in agreement with the known result for Buffon's needle problem with $\len > a$ \cite[theorem 1.1.2]{Mathai99}.
-On the other hand, $\Pr[i \geq n]$ for $n \geq 3$ can be seen as the probability of crossing at least one line in a modified version of Buffon's setting in which a needle endpoint is only allowed to move in a narrower region in the middle of a strip, of width $2n-3$ times smaller than the strip width.
+On the other hand, $\Pr[i \geq n]$ for $n \geq 3$ can be seen as the probability of crossing at least one line in a modified version of Buffon's setting in which a needle endpoint is only allowed to move in a region of width $a$ located in the middle of a strip of width $(2n-3)a$.
 

### 2022-01-09 paper.tex
Unifying use of $a$ or $b$ in Buffon's problem: always use $a·

@@
 The questions studied in this paper are related to Buffon's needle problem and Laplace's extension of it.
-Buffon's original problem considers tossing a needle of length $\len$ onto ruled paper with parallel lines a distance $b$ apart, and asks what is the probability of the needle crossing a line if it lands in a uniformly random position and orientation.
-For a needle of length $\len = b$, this is $2/\pi$ (and hence repeated trials of this experiment can be used to estimate $\pi$).
-This is generalized in \cite{Ramaley69} to $2\len /(\pi b)$ for the expected number of crossings of a needle with arbitrary length.
-
-The Buffon-Laplace needle problem \cite[section 1.1]{Mathai99} considers a needle tossed onto a grid of rectangular tiles of width $a$ and height $b$.
+Buffon's original problem considers a plane with vertical lines a distance $a$ apart.
+A needle is placed on the plane with uniformly random position\footnote{It suffices to define position using a horizontal coordinate modulo $a$, for which a uniform distribution can be defined.} and orientation, and the probability of the needle crossing a line is studied.
+For a needle of length $\len = a$, this probability is $2/\pi$ (and hence repeated trials of this experiment can be used to estimate $\pi$).
+This is generalized in \cite{Ramaley69} to $2\len /(\pi a)$ for the expected number of crossings of a needle with arbitrary length.
+
+The Buffon-Laplace needle problem \cite[section 1.1]{Mathai99} considers a needle randomly placed on a grid of rectangular tiles of width $a$ and height $b$.
 The number of visited tiles equals one plus the number of crossings almost surely.
-The probability of the needle staying within a single tile is computed in \cite{Arnow94} for the case where $\len < \min(a,b)$.
+The probability of the needle staying within a single tile is computed in \cite{Arnow94} for the case where $\len < \min\{a,b\}$.
 One of the problems considered in this paper, as mentioned above, is the complementary question \ref{prob: rand prob} of the probability that the needle visits the maximum number of tiles possible for its length.
@@
 \begin{proof}
-Suppose first that the grid is simplified to only horizontal lines with spacing $b$.
-This matches the set-up of Buffon's original needle problem \cite[section 1.1]{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines.
-As shown in \cite{Ramaley69}, the expected number of lines crossed equals $2\len/(\pi b)$.
-% When $\len<b$ this expectation matches the probability of the needle crossing a grid line, since it cannot cross more than one.
+Suppose first that the grid is simplified to only vertical lines with spacing $a$.
+This matches the set-up of Buffon's original needle problem \cite[section 1.1]{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $a$, allowing it to cross multiple lines.
+As shown in \cite{Ramaley69}, the expected number of lines crossed equals $2\len/(\pi a)$.
+% When $\len<a$ this expectation matches the probability of the needle crossing a grid line, since it cannot cross more than one.
 
@@
 
-% ***Alex: please check this, as I've done some content changes
 Lastly, $\Pr[i \geq 2]$ as computed in Proposition~\ref{prop: Pr i} can be interpreted as the probability of crossing at least one line in Buffon's experiment.

### 2022-01-19 paper.tex
Errata / wording

@@
 
-An equivalent formulation of the problem is obtained allowing segments of length $\len$ \emph{or smaller}.
+An equivalent formulation of the problem is obtained by allowing segments of length $\len$ \emph{or smaller}.
 The equivalence is clear from the fact that reducing the length cannot increase the number of visited tiles.
@@
 The \emph{inverse} problem \ref{prob: determ inv} is, given $\tiles \in \mathbb N$, to determine the infimum length of all segments that visit at least $\tiles$ tiles.
-If the length can take any real value the infimum is not a minimum, because given any segment can be shortened by some small amount without changing the number of visited tiles.
+If the length can take any real value the infimum is not a minimum, because given any segment it can be shortened by some small amount without changing the number of visited tiles.
 This is a consequence of the interior of each tile being an open set.

### 2022-03-02 paper.tex
Small corrections for (first) submission

@@
 
-\addtolength{\textheight}{34mm}
-\addtolength{\voffset}{-20mm}
+\addtolength{\textheight}{38mm}
+\addtolength{\voffset}{-22mm}
 \addtolength{\textwidth}{-3mm}
@@
 In the deterministic setting, the maximum possible number of visited tiles is characterized for a given length, and conversely, the infimum segment length needed to visit a desired number of tiles is analyzed.
-The probabilistic setting regards the segment position and orientation as random.
-The average number of visited tiles and the probability of visiting the maximum number of tiles on a square grid are studied as a function of segment length.
+In the random setting, the average number of visited tiles and the probability of visiting the maximum number of tiles on a square grid are studied as a function of segment length.
 These questions are related to Buffon's needle problem and its extension by Laplace.
@@
 
-The Buffon-Laplace needle problem \cite[section 1.1]{Mathai99} considers a needle randomly placed on a grid of rectangular tiles of width $a$ and height $b$.
+The Buffon-Laplace needle problem \cite[section~1.1]{Mathai99} considers a needle randomly placed on a grid of rectangular tiles of width $a$ and height $b$.
 The number of visited tiles equals one plus the number of crossings almost surely.
@@
 The rest of the paper is organized as follows.
-Fundamental results are presented in \S\ref{part: fund results}, which form the basis of the subsequent analysis.
+Fundamental results are presented in \S\ref{part: fund results}, which form the basis of both the deterministic and probabilistic analyses.
 The direct and inverse problems for a deterministic segment are considered in \S\ref{part: max}, first for arbitrary grids and then for a square grid.
@@
 \end{align*}
-The dimensions of the discrete bounding rectangle, normalized to the tile width and height respectively, are two integer numbers $i$, $j$.
+The dimensions of the discrete bounding rectangle, \emph{normalized} to the tile width and height respectively, are two integer numbers $i$, $j$.
 Two examples are illustrated in Figure~\ref{fig: discrete bounding rectangle and touched tiles}, both with $i=5$, $j=4$.
@@
 \ref*{prop: len ineq i j: ineqs} The inequalities follow from the fact that the segment endpoints lie in the interiors or on the outer edges of two tiles in opposite corners of the discrete bounding rectangle.
-This is illustrated in Figures~\ref{fig: len ineq 4 3} and \ref{fig: len ineq 4 2} for two specific $(i,j)$ pairs respectively.
+This is illustrated in Figure~\ref{fig: len ineq i j} for two specific $(i,j)$ pairs.
 For each $(i,j)$, segments are shown with lengths close to either of the two bounds.
@@
 
-\ref*{prop: len ineq i j: exist} For $a$, $b$, $\len$, $i$, $j$ satisfying the two inequalities, a segment of length $\len$ can be found that has its endpoints in the interiors or on the outer edges of the two shaded tiles of a rectangle with normalized dimensions $i$ and $j$, as in Figure~\ref{fig: len ineq i j}, which is thus the discrete bounding rectangle of that segment.
+\ref*{prop: len ineq i j: exist} For $a$, $b$, $\len$, $i$, $j$ satisfying the two inequalities, a segment of length $\len$ can be found that has its endpoints in the interiors or on the outer edges of the two shaded tiles of a rectangle with normalized dimensions $i$ and $j$ (see Figure~\ref{fig: len ineq i j}), which is thus the discrete bounding rectangle of that segment.
 \begin{figure}
@@
 
-``There is a segment\ldots $\Rightarrow$ \eqref{eq: len > sqrt} '': Assume that a segment exists with length $\len' \leq \len$ and with a discrete bounding rectangle of normalized dimensions $i$, $j$.
-From part~\ref{prop: len ineq i j: ineqs}, inequality \eqref{eq: len > sqrt} holds for the length $\len'$, and thus for $\len$.
+``There is a segment\ldots $\Rightarrow$ \eqref{eq: len > sqrt}'': Assume that a segment exists with length $\len' \leq \len$ and with a discrete bounding rectangle of normalized dimensions $i$, $j$.
+From part~\ref{prop: len ineq i j: ineqs} it follows that inequality \eqref{eq: len > sqrt} holds for the length $\len'$, and thus for $\len$.
 \end{proof}
@@
 On the other hand, Proposition~\ref{prop: len ineq i j} restricts the $i, j$ values that can be achieved with a given length.
-A relevant question is: are there any $(i,j)$ pairs that can be disregarded for all lengths $\len$?
+A relevant question is: are there any $(i,j)$ pairs that can be disregarded irrespective of the length $\len$?
 In other words, what is the ``smallest'' subset of $\mathbb N^2$ such that the $(i,j)$ pair that maximizes the number of tiles for any given length can always be found within that subset?
@@
 For a given $\len \in \mathbb R^+$, the $(i,j)$ pairs that can be achieved with segments of length not exceeding $\len$ are, by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, those that satisfy \eqref{eq: len > sqrt}.
-Thus for a given value of $i+j-1$ the optimal $(i,j)$ pair (the one that can be achieved with the smallest possible length) is determined by the condition that the point $(ia,jb)$ minimizes the distance to $(2a,2b)$.
-Such pairs are depicted as filled circles in the figure.
-The displayed arcs represent the inequality \eqref{eq: len > sqrt} for each of the optimal pairs.
-Conversely, given a length $\len$, the maximum number of visited tiles will be achieved with one of these pairs, namely the pair $(i,j)$ such that $(ia,jb)$ is on the highest (or rightmost) diagonal line while still being contained in the circle of radius $\len$ centered at $(2a,2b)$.
+Thus for a given value of $i+j-1$ the optimal $(i,j)$ pair (the one that can be achieved with the smallest length) is determined by the condition that the point $(ia,jb)$ minimizes the distance to $(2a,2b)$.
+Such pairs are depicted as filled circles in the figure, and the arcs represent the inequality \eqref{eq: len > sqrt} for each of the optimal pairs.
+
+Conversely, given a length $\len$, the maximum number of visited tiles will be achieved with one of these pairs, namely the pair $(i,j)$ such that $(ia,jb)$ is on the uppermost (or rightmost) diagonal line while still being contained in the circle of radius $\len$ centered at $(2a,2b)$.
 
@@
 \label{prop: min suff set, form}
-Given $a, b \in \mathbb R^+$ and $\tiles \in \mathbb N$, $\tiles \geq 3$, a set of optimal pairs $\mss = \{(i_3,j_3), (i_4.j_4), \ldots\}$ can be obtained as
+Given $a, b \in \mathbb R^+$, a set of optimal pairs $\mss = \{(i_3,j_3), (i_4.j_4), \ldots\}$ can be obtained as
 \begin{align}
@@
 \begin{proof}
-For each $\tiles$, the pair $(i_\tiles,j_\tiles)$ should be chosen as that on the line $i+j-1=\tiles$ which minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$.
+For each $\tiles \geq 3$, the pair $(i_\tiles,j_\tiles)$ should be chosen as that on the line $i+j-1=\tiles$ which minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$.
 This allows maximizing the sum $i+j-1$, and thus the number of visited tiles, for a given length restriction; or visiting a specified number of files with lengths as small as possible.
@@
 The pair $(i_\tiles, j_\tiles) \in \mathbb Z^2$ that minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$ along the line $i+j-1=\tiles$ is either $(\lceil \isolr \rceil, \lfloor \jsolr \rfloor)$ or $(\lfloor \isolr \rfloor, \lceil \jsolr \rceil)$, whichever gives $(i_\tiles a, j_\tiles b)$ closest to $(\isolr a, \jsolr b)$, as illustrated in Figure~\ref{fig: min suff set, form}.
-In case of a tie the first of the two pairs is (arbitrarily) chosen.
+In case of a tie the first of the two options is (arbitrarily) chosen.
 This can be expressed as
@@
 \label{fig: pairs_bounds_1}%
-\includegraphics[height=.235\textheight]{pairs_bounds_1}% *!*
+\includegraphics[height=.23\textheight]{pairs_bounds_1}% *!*
 }\hfill%
@@
 \label{fig: pairs_bounds_1p35}%
-\includegraphics[height=.235\textheight]{pairs_bounds_1p35}% *!*
+\includegraphics[height=.23\textheight]{pairs_bounds_1p35}% *!*
 }\hfill%
@@
 \label{fig: pairs_bounds_sqrt2}%
-\includegraphics[height=.235\textheight]{pairs_bounds_sqrt2}% *!*
+\includegraphics[height=.23\textheight]{pairs_bounds_sqrt2}% *!*
 }%
@@
 
-For $a^2/b^2$ arbitrary, the number of pairs in $\mss$ with the same $i$ coordinate, or with the same $j$, is in general irregular, because the lines \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} do not follow a ``natural'' direction of the grid.
+For $a^2/b^2$ arbitrary, the number of pairs in $\mss$ with the same $i$, or with the same $j$, is in general irregular, because the lines \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} do not follow a ``natural'' direction of the grid.
 This happens for instance in Figure~\ref{fig: pairs_bounds_1p35}, where the number of pairs for each $i$ equals either $2$ or $3$ without a clear pattern.\footnote{Strictly, there is a periodic pattern whenever $a^2/b^2$ is rational, which is the case in Figure~\ref{fig: pairs_bounds_1p35}.
@@
 A segment whose discrete bounding rectangle has normalized width $i$ and height $j$ is oriented with approximate slope $jb/(ia)$ with respect to the $x$ axis (see Figure~\ref{fig: len ineq i j}); and this approximation becomes better for greater segment lengths.
-From \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} it can be seen that the positions of the pairs $(i,j) \in \mss$ have $j/i \approx a^2/b^2$ for large $i, j$.
+From \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} it can be seen that the pairs $(i,j) \in \mss$ have $j/i \approx a^2/b^2$ for large $i, j$.
 Therefore the optimal slope for long segments is approximately $a/b$.
-This formalizes the intuition that to maximize the number of visited tiles, the segment should follow a direction along which the perceived ``length'' of the tile is smaller.
-% but not ``smallest'', if that were the case it would always be either vertical or horizontal
+This substantiates the intuition that to maximize the number of visited tiles, the segment direction should strike a balance between achieving a small perceived ``length'' of the tile on one hand, and crossing both horizontal and vertical grid lines on the other hand.
 
@@
 
-For a grid with parameters $a, b \in \mathbb R^+$, the maximum number $\tiles$ of visited tiles with a given real-valued length $\len$ can be represented by a function $\funt: \mathbb R^+ \to \mathbb N$ such that $\tiles = \funt(\len)$.
+Given a grid with parameters $a, b \in \mathbb R^+$, the maximum number $\tiles$ of visited tiles for an arbitrary real-valued length $\len$ can be represented by a function $\funt: \mathbb R^+ \to \mathbb N$ such that $\tiles = \funt(\len)$.
 Similarly, for the inverse problem a function $\funl: \mathbb N \to \mathbb R^+$ can be defined such that $\funl(\tiles)$ gives the infimum length of all segments that visit at least $\tiles$ tiles.
-
-The function $\funt$ can be obtained from $\funl$ by noting that the maximum number of tiles that can be visited by a segment of length $\len$ is the index of the last term of the sequence $\funl(\tiles)$ that is less than $\len$:
-\begin{equation}
+Clearly, these two functions are related as
+\begin{align}
 \label{eq: funt funl}
-\funt(\len) = \max \{\tiles \in \mathbb N \st \funl(\tiles)<\len\}.
-\end{equation}
-Conversely, $\funl$ can be obtained from $\funt$ as
-\begin{equation}
+\funt(\len) &= \max \{\tiles \in \mathbb N \st \funl(\tiles)<\len\},\\
 \label{eq: funl funt}
-\funl(\tiles) = \inf\{\len \in \mathbb R^+ \st \funt(\len) \geq \tiles\}.
-\end{equation}
+\funl(\tiles) &= \inf\{\len \in \mathbb R^+ \st \funt(\len) \geq \tiles\}.
+\end{align}
 
 For arbitrary $a, b \in \mathbb R^+$, the functions $\funt$ and $\funl$ can be computed using an iterative procedure, which exploits the fact that the pairs $(i_3,j_3), (i_4,j_4), \ldots$ of the set $\mss$ are sorted by increasing $i+j-1$, and also by increasing $(i-2)^2 a^2 + (j-2)^2 b^2$.
-Namely, for $\funt$ the following procedure yields the solution: generate successive pairs to find the last one, $(i_\tiles,j_\tiles)$, that satisfies \eqref{eq: len > sqrt}; then $\funt(\len) = \tiles$.
+Namely, for $\funt$ the following method yields the solution: generate successive pairs to find the last one, $(i_\tiles,j_\tiles)$, that satisfies \eqref{eq: len > sqrt}; then $\funt(\len) = \tiles$.
 For $\funl$ the analogous method gives a direct formula.
@@
 \begin{proof}
-The approach is similar to that used in Proposition~\ref{prop: min suff set, form}.
-Namely, for $i, j \geq 2$ the intersection $(\isolr a, \jsolr b)$ of the line defined by \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ is found, and from that the actual integer solution $(\isoli, \jsoli)$ is computed.
-
-More specifically, assuming that \eqref{eq: lower bound, line} holds, the largest $i$ that can be achieved with lengths not exceeding $\len$ is obtained, as given by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}.
-This value, $\isoli$, is computed by rounding down $\isolr$.
-Then, assuming that $i = \isoli$, the largest $j$ allowed by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, $\jsoli$, is obtained.
-As will be seen, in some cases the resulting $(\isoli,\jsoli)$ is in the set $\mss$ defined by Proposition~\ref{prop: min suff set, form}, and in other cases it is not.
-However, in either case $(\isoli,\jsoli)$ has the largest $i+j-1$ sum that can be achieved with segments of length up to $\len$.
-Thus the desired result is $\isoli+\jsoli-1$.
-
-The two possibilities are illustrated in Figure~\ref{fig: ijLS_1p35_detail} for $a=1.35$, $b=1$.
-In each case, the arc displayed in the graph is centered at $(2a,2b)$ and has radius $\len$.
+The approach is similar to that used in the proof of Proposition~\ref{prop: min suff set, form}.
+First, the intersection point $(\isolr a, \jsolr b)$, $\isolr, \jsolr \geq 2$ between the line defined by \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ is computed, if it exists.
+Then, based on either the values $\isolr, \jsolr$ or the non-existence of the intersection point, a pair of integer values $(\isoli, \jsoli)$ is obtained that maximizes the $i+j-1$ sum that can be achieved with segments of length up to $\len$.
+
+As will be seen, the obtained pair $(\isoli,\jsoli)$ may belong to the set $\mss$ defined by Proposition~\ref{prop: min suff set, form} or not.
+However, in either case $\funt(\len)$ is given by $\isoli+\jsoli-1$.
+The two possibilities are respectively illustrated in Figures~\ref{fig: ijLS_1p35_detail_in} and \ref{fig: ijLS_1p35_detail_notin} for $a=1.35$, $b=1$.
+In each case, the displayed arc is centered at $(2a,2b)$ and has radius $\len$.
 The inner region defined by the arc contains all $(ia,jb)$ points such that the pair $(i,j)$ is achievable according to Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}.
@@
 
-The pair $(\isolr, \jsolr)$ can be obtained as a solution of the equation system
+The pair $(\isolr, \jsolr)$ results from solving the equation system
 \begin{align}
@@
 where the same sign (positive or negative) should be used in the two equations.
-This can yield zero, one or two real-valued solution pairs $(\isolr, \jsolr)$, which respectively corresponds to the solid line in Figure~\ref{fig: ijLS_1p35_detail} being exterior, tangent or secant to the circle (the figure depicts the latter case).
-
-A solution pair $(\isolr, \jsolr)$ from \eqref{eq: isolr sol} and \eqref{eq: jsolr sol} is meaningful only if it is real-valued with $\isolr, \jsolr \geq 2$.
-This solution, if it exists, is always associated with the positive sign in those equations.
-It is easily seen that $\jsolr \geq 2$ implies $\isolr \geq 2$, and thus it suffices the check the former.
+This can yield zero, one or two real-valued solution pairs $(\isolr, \jsolr)$, which respectively corresponds to the solid line in Figure~\ref{fig: ijLS_1p35_detail} being exterior, tangent or secant to the circle (the figure depicts the latter situation).
+
+A solution pair $(\isolr, \jsolr)$ given by \eqref{eq: isolr sol} and \eqref{eq: jsolr sol} is meaningful only if it is real-valued with $\isolr, \jsolr \geq 2$.
+This solution, if it exists, is always associated with the positive sign in those expressions.
+Since $a \geq b$, it is easily seen that $\jsolr \geq 2$ implies $\isolr \geq 2$, and thus it suffices to check the former condition.
 Three cases need to be distinguished: there are no real-valued solution pairs $(\isolr, \jsolr)$; there are one or two but none of them has $\jsolr \geq 2$; or there are one or two and and one of them satisfies that inequality.
-These cases correspond to different ranges of $\len$, as discussed next.
+These cases correspond to different ranges of $\len$, as seen next.
 
 For $\len < \sqrt{a^2+b^2}/2$, \eqref{eq: isolr sol} and \eqref{eq: jsolr sol} are not real-valued.
-Since $a \geq b$, from the inequality $\len < \sqrt{a^2+b^2}/2$ it follows that $\len < a$.
+Since $a \geq b$, from the inequality $\len < \sqrt{a^2+b^2}/2$ it follows that $\len < a/\sqrt{2} < a$.
 This means that any achievable $(i,j)$ pair, if any, will have $i=2$.
-Thus in this case $\isoli$ should be set to $2$; and then the maximum achievable $\jsoli$ will be obtained from \eqref{eq: icont jcont 2}.
-
-For $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, \eqref{eq: isolr sol} and \eqref{eq: jsolr sol} give either two real-valued solutions or one real-valued double solution for $(\isolr, \jsolr)$; but $\jsolr < 2$.
-This in turn implies, according to \eqref{eq: icont jcont 2}, that $\isolr < (b^2/a^2+5)/2 < 3$ for $a \geq b$.
-Thus only pairs with $i=2$ are achievable again for $\len$ in this range, and therefore $\isoli$ must also be $2$.
+Thus in this case $\isoli$ should be set to $2$.
+
+For $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, \eqref{eq: isolr sol} and \eqref{eq: jsolr sol} give either two real-valued solutions or one real-valued double solution for $(\isolr, \jsolr)$, with $\jsolr < 2$.
+This in turn implies, according to \eqref{eq: icont jcont 2}, that $\isolr < (b^2/a^2+5)/2 \leq 3$ for $a \geq b$.
+Thus only pairs with $i=2$ are achievable again for $\len$ in this range, and therefore $\isoli$ must be $2$.
 
@@
 
-The three cases are unified by taking the real part of the positive-sign version of \eqref{eq: isolr sol} and computing $\isoli = \lceil \isolr \rceil-1$, as can be easily checked.
+The three cases are unified, as can be easily checked, by taking the real part of the positive-sign version of \eqref{eq: isolr sol} and computing $\isoli = \lceil \isolr \rceil-1$.
 This yields \eqref{eq: theo: funt, form; i}.
@@
 
-There are two possibilities for the resulting pair $(\isoli,\jsoli)$, as stated at the outset.
-These are illustrated in Figures~\ref{fig: ijLS_1p35_detail_in} and \ref{fig: ijLS_1p35_detail_notin} respectively.
-The first possibility is that $(\isoli,\jsoli) \in \mss$ (Figure~\ref{fig: ijLS_1p35_detail_in}).
+To see that $\funt(\len) = \isoli+\jsoli-1$, the two situations stated at the outset need to be considered separately.
+The first possibility is that $(\isoli,\jsoli) \in \mss$ (upper part of Figure~\ref{fig: ijLS_1p35_detail}).
 Then, by construction  $(\isoli,\jsoli)$ maximizes $i+j-1$ among all achievable pairs of $\mss$, and is therefore optimal.
@@
 
-Therefore, regardless of whether $(\isoli,\jsoli)$ is in $\mss$ or not, \eqref{eq: theo: funt, form; i} and \eqref{eq: theo: funt, form; j} give $\isoli+\jsoli-1$ equal to $\funt(\len)$, as claimed.
+Therefore, regardless of whether $(\isoli,\jsoli)$ is in $\mss$ or not, \eqref{eq: theo: funt, form; i} and \eqref{eq: theo: funt, form; j} give $\isoli+\jsoli-1$ equal to $\funt(\len)$.
+This establishes \eqref{eq: theo: funt, form; funt}.
 
 Interestingly, for the specific case that $a^2/b^2 = 2k-1$, $k \in \mathbb N$ the lower bounding line \eqref{eq: lower bound, line} becomes $j = (2k-1)i - 5k + 4$, which gives integer $j$ for integer $i$.
-This means that the bound is actually achieved for all pairs in $\mss$ (see for example Figure~\ref{fig: pairs_bounds_1}), and the case $(\isoli,\jsoli) \notin \mss$ never occurs.
+This means that for each $i$ there is a pair $(i,j) \in \mss$ that is on that line (see for example Figure~\ref{fig: pairs_bounds_1}), and the case $(\isoli,\jsoli) \notin \mss$ never occurs.
 % ^ If this is needed outside of the proof, it can be included at the end of the theorem statement
@@
 \end{equation}
-To see that $\funt$ has a jump at $\len$, assume for the sake of contradiction that $\funt$ is continuous, therefore constant, at that point.
-Then there is $\epsilon > 0$ for which $\funt(\len-\epsilon) = \tiles = \funt(\len+\epsilon)$.
+To see that $\funt$ has a jump at $\len$, assume for the sake of contradiction that $\funt$ is continuous at that point.
+Therefore $\funt$ is constant on an interval containing that point, which implies that $\funt(\len-\epsilon) = \tiles = \funt(\len+\epsilon)$ for some $\epsilon > 0$.
 This means that there exists a segment with length $\len-\epsilon$ that visits $\tiles$ tiles, and thus $\funl(\tiles) \leq \len-\epsilon < \len$, in contradiction with \eqref{eq: theo funt, form: prop 1}.
-Therefore $\funt$ is discontinuous, from the right, at $\len$.
+Therefore $\funt$ is discontinuous (from the right) at $\len$.
 By definition of $\funl$, from \eqref{eq: theo funt, form: prop 1} it follows that
@@
 \begin{proof}
-The $(i_\tiles,j_\tiles)$ pair in set $\mss$ defined in Proposition~\ref{prop: min suff set, form} corresponds to a maximum of $\tiles$ visited tiles.
+The $(i_\tiles,j_\tiles)$ pair in set $\mss$ defined in Proposition~\ref{prop: min suff set, form} corresponds to at most $\tiles$ visited tiles.
 By construction of this set, any segment that visits $\tiles$ tiles must have length greater than $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$.
@@
 (\isolr-2)^2a^2 + (\jsolr-2)^2b^2 =
-\frac{(\tiles-3)^2 (b^4a^2+a^2b^4)}{(a^2+b^2)^2} =
+\frac{(\tiles-3)^2 (a^2b^4+a^4b^2)}{(a^2+b^2)^2} =
 \frac{(\tiles-3)^2 a^2b^2}{a^2+b^2}.
 \end{equation}
-The fact that both $(\isolr, \jsolr)$ and $(i_\tiles, j_\tiles)$ are on the line $i+j-1=\tiles$ implies that $\isolr+\jsolr = i_\tiles+ j_\tiles$, and thus $|i_\tiles-\isolr| = |j_\tiles-\jsolr| = \roundterm$ with $\roundterm$ given by \eqref{eq: roundterm}.
+The fact that both $(\isolr, \jsolr)$ and $(i_\tiles, j_\tiles)$ are on the line $i+j-1=\tiles$ implies that $\isolr+\jsolr = i_\tiles+ j_\tiles$.
+Taking into account that $i_\tiles=\isoli$ and $j_\tiles=\jsoli$, it stems that $\vert i_\tiles-\isolr \vert = \vert j_\tiles-\jsolr \vert = \roundterm$ with $\roundterm$ given by \eqref{eq: roundterm}.
 Consequently,
@@
 \end{equation}
-Substituting \eqref{eq: distance as hypotenuse, first leg} and \eqref{eq: distance as hypotenuse, second leg} into \eqref{eq: distance as hypotenuse} yields \eqref{eq: theo: funl, form; funl 2}.
+Substituting \eqref{eq: distance as hypotenuse, first leg} and \eqref{eq: distance as hypotenuse, second leg} into \eqref{eq: distance as hypotenuse} and using \eqref{eq: theo: funl, form; funl} yields \eqref{eq: theo: funl, form; funl 2}.
 
@@
 
-Theorems \ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the first two problems stated in \S\ref{part: intro}; they also provide a way to actually position a segment of length $\len$ or slightly greater than $\funl(\tiles)$, respectively, so that it visits $\funt(\len)$ or $\tiles$ tiles.
-Namely, for $\isoli$, $\jsoli$ computed as in the corresponding theorem, the segment should have its endpoints in the interiors of two tiles separated $\isoli-1$ steps horizontally and $\jsoli-1$ steps vertically, with its exact position and orientation  adjusted to avoid any grid points.
+Theorems \ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the first two questions posed in \S\ref{part: intro}; they also provide a way to actually position a segment of length $\len$ or slightly greater than $\funl(\tiles)$, respectively, so that it visits $\funt(\len)$ or $\tiles$ tiles.
+Namely, for $\isoli$, $\jsoli$ computed as in the corresponding theorem, the segment should have its endpoints in the interiors of two tiles shifted $\isoli-1$ steps horizontally and $\jsoli-1$ steps vertically with respect to each other, with the exact position and orientation of the segment adjusted to avoid any grid points.
 
@@
 
-For $a=b$, either from symmetry considerations or particularizing the formulas in the above theorems it stems that the optimal orientation of the segment is close to $45^\circ$.
+For $a=b$, either from symmetry considerations or particularizing the results in the above theorems it stems that the optimal orientation of the segment is close to $45^\circ$.
 This case will be dealt with in \S\ref{part: max: unit square grid, real lengths}, as it lends itself to simplified formulas.
@@
 The graphs illustrate some of the observations of the previous paragraphs.
-Indeed, the asymptotic slope in Figure~\ref{fig: funt examples}, or the inverse of the asymptotic slope in Figure~\ref{fig: funl examples}, is approximately $\sqrt{2}$ for $a, b=1$; and it is $1/b$ for the case $a=5,b=1$, or even for $a=5,b=1.5$ or $a=10,b=3$.
+Indeed, the asymptotic slope in Figure~\ref{fig: funt examples}, or the inverse of the asymptotic slope in Figure~\ref{fig: funl examples}, is approximately $\sqrt{2}$ for $a=b=1$; and it is roughly $1/b$ for the case $a=5,b=1$, or even for $a=5,b=1.5$ or $a=10,b=3$.
 Comparing the latter two cases it is also seen that scaling $a$, $b$ and $\len$ by the same factor does not alter $\funt(\len)$, and results in $\funl(\tiles)$ being scaled by that factor.
@@
 }%
-\caption{Functions $\funt$ and $\funl$ for several pairs $a$, $b$
+\caption{Functions $\funt$ and $\funl$ for several pairs $a, b$
 }%
@@
 A square grid has $a=b$.
-For real-valued segment lengths it can be further assumed that $a=1$.
+For real-valued segment lengths it can be further assumed that $a=1$ (unit square grid).
 For $a \neq 1$ the expressions to be obtained are valid with $\len$ and $\funl(\tiles)$ replaced by $\len/a$ and $\funl(\tiles)/a$ respectively.
@@
 \end{align}
-where either $\jsoli = \isoli$ or $\jsoli = \isoli+1$. Equivalently,
+which gives either $\jsoli = \isoli$ or $\jsoli = \isoli+1$. Equivalently,
 \begin{equation}
@@
 This can be seen as follows.
-For $\jsoli = \isoli$, the point $(\isoli, \jsoli)$ is closer to $(2,2)$ than $(\isolr, \jsolr)$ is, and is therefore achievable.
+If $\jsoli = \isoli$, the point $(\isoli, \jsoli)$ is closer to $(2,2)$ than $(\isolr, \jsolr)$ is, and is therefore achievable.
 This implies that values of $\jsoli$ smaller than $\isoli$ will never be chosen.
@@
 }%
-\label{fig: funti funti}%
+\label{fig: funti funli}%
 \end{figure}%
@@
 Therefore at least one of the three increments from $\funti(\leni)$ to $\funti(\leni+3)$ is $2$.
-Since $\funti(2)-\funti(1) = \funti(3)-\funti(2) = 2$, the result holds for all $\leni \in \mathbb N$.
+Since $\funti(2)-\funti(1) = \funti(3)-\funti(2) = 2$, this result holds for all $\leni \in \mathbb N$.
 
@@
 This is A346693 in the On-Line Encyclopedia of Integer Sequences \cite{OEIS_unitsq_int_inv}.
-However, a slightly simpler expression can be obtained from \eqref{eq: funti leni} and \eqref{eq: funli funti}.
+Moreover, a slightly simpler expression can be obtained from \eqref{eq: funti leni} and \eqref{eq: funli funti}.
 This is established by the next theorem, which also states some properties of $\funli$, parallel to those of $\funti$.
@@
 \item
-There can be no more than $3$ different consecutive terms.
+There can be no more than $3$ consecutive terms that are different.
 \item
@@
 \end{equation}
-For $\tiles \geq 4$, taking squares and rearranging gives
+Assuming $\tiles \geq 4$, taking squares and rearranging gives
 \begin{equation}
@@
 which combined with \eqref{eq: funli funti} yields the second part of \eqref{eq: funli}.
-The first part results from noting that $\tiles-3 \leq 0$ for $\tiles \leq 3$, and thus $\leni=1$ satisfies \eqref{eq: theo funli ineq}.
+The first part results from noting that for $\tiles \leq 3$ the value $\leni=1$ satisfies \eqref{eq: theo funli ineq}.
 
@@
 Each realization of the random segment gives rise to a discrete bounding rectangle, whose normalized dimensions $i$ and $j$ are thus random variables, as is the number $\tiles$ of visited tiles.
-Except for a set of realizations with zero probability, $i$ and $j$ are at least $1$, and $\tiles = i+j-1$.
+Except for a set of realizations with probability $0$, $i$ and $j$ are at least $1$, and $\tiles = i+j-1$.
 Note that $i$ and $j$ are not statistically independent.
@@
 The square grid is addressed in \S\ref{part: probmax: unit square grid, real lengths}.
-This more specialized setting allows computation of the probability of visiting the maximum number of tiles.
+This more specialized setting allows computation of the probability of visiting the maximum number of tiles, which would be difficult in the general case.
 
@@
 
-Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ is the average number of tiles visited by a random segment of length $\len$, with the distributions specified in \S\ref{part: rand}.
-This function can be easily computed using the fact that
-\begin{equation}
-\label{eq: funta E[i] E[j]}
-\funta(\len) = \E[\tiles] = \E[i]+\E[j]-1,
-\end{equation}
-and noting that events of probability $0$, such as $x_1=0$ or the segment passing through a grid point, can be disregarded.
+Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ is the average number of tiles visited by a random segment of length $\len$, with the distributions specified in the preceding.
 
@@
 Suppose first that the grid is simplified to only vertical lines with spacing $a$.
-This matches the set-up of Buffon's original needle problem \cite[section 1.1]{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $a$, allowing it to cross multiple lines.
+This matches the set-up of Buffon's original needle problem \cite[section~1.1]{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $a$, allowing it to cross multiple lines.
 As shown in \cite{Ramaley69}, the expected number of lines crossed equals $2\len/(\pi a)$.
@@
 The segment orientation is a random angle $\theta$ uniformly distributed on $[0,2\pi)$.
-It is clear from the figure that $i \geq n$ if and only if $x_2 \geq a(n-1)$ or $x_2 \leq -a(n-2)$; and these events are exclusive for $n \geq 2$.
+It is clear from the figure that $i \geq n$ if and only if $x_2 > a(n-1)$ or $x_2 < -a(n-2)$; and for $n \geq 2$ these events are exclusive.
 Thus
@@
 \label{eq: Pr i cond}
-\Pr[i \geq n \cond x_1, y_1] = \Pr[x_2 \geq a(n-1) \cond x_1, y_1] + \Pr[x_2 \leq -a(n-2) \cond x_1, y_1].
+\Pr[i \geq n \cond x_1, y_1] = \Pr[x_2 > a(n-1) \cond x_1, y_1] + \Pr[x_2 < -a(n-2) \cond x_1, y_1].
 \end{equation}
 The two conditional probabilities on the right-hand side of \eqref{eq: Pr i cond} are different in general.
-However, averaging over $x_1, y_1$ gives, by symmetry, $\Pr[x_2 \geq a(n-1)] = \Pr[x_2 \leq -a(n-2)]$.
+However, averaging over $x_1, y_1$ gives, by symmetry, $\Pr[x_2 > a(n-1)] = \Pr[x_2 < -a(n-2)]$.
 In addition, the coordinate $y_1$ does not have any influence on these events, and therefore conditioning on $x_1, y_1$ is the same as conditioning on $x_1$.
@@
 \label{eq: Pr i = 2 Pr}
-\Pr[i \geq n] = 2 \Pr[x_2 \geq a(n-1)].
+\Pr[i \geq n] = 2 \Pr[x_2 > a(n-1)].
 \end{equation}
@@
 
-Consider the event $x_2 \geq a(n-1)$ conditioned on $x_1$, with $n \geq 2$.
+Consider the event $x_2 > a(n-1)$ conditioned on $x_1$, with $n \geq 2$.
 There are three possibilities depending on $x_1$, $n$ and $\len$.
@@
 This is depicted in Figure~\ref{fig: Pr_i_full}, where the section of the arc with solid line represents, for a given $x_1$, those angles for which $x_2 > a(n-1)$.
-If $a(n-2) < \len \leq  a(n-1)$, the length will be enough provided that $x_1 \geq a(n-1)-\len$, and then only for certain angles.
+If $a(n-2) < \len \leq  a(n-1)$, the length will be enough provided that $x_1 > a(n-1)-\len$, and then only for certain angles.
 This restriction on $x_1$ corresponds to the shaded region in Figure~\ref{fig: Pr_i_notfull}.
@@
 
-In the first two cases above, the probability that $x_2 \geq a(n-1)$, conditioned on $x_1$, is the length of the arc to the right of the line $x=a(n-1)$ divided by $2\pi\len$, that is,
-\begin{equation}
-\label{eq: Pr x_2 geq a(n-1) cond}
-\Pr[x_2 \geq a(n-1) \cond x_1] = \frac 1 \pi \arccos \frac{a(n-1)-x_1}{\len}.
-\end{equation}
-In the first case $x_1$ has a uniform distribution on $[0,a)$, and $\Pr[x_2 \geq a(n-1)]$ is easily obtained from \eqref{eq: Pr x_2 geq a(n-1) cond} as
-\begin{equation}
-\label{eq: Pr x_2 geq a(n-1), first case}
+In the first two cases above, the probability that $x_2 > a(n-1)$, conditioned on $x_1$, is the length of the arc to the right of the line $x=a(n-1)$ divided by $2\pi\len$, that is,
+\begin{equation}
+\label{eq: Pr x_2 > a(n-1) cond}
+\Pr[x_2 > a(n-1) \cond x_1] = \frac 1 \pi \arccos \frac{a(n-1)-x_1}{\len}.
+\end{equation}
+In the first case $x_1$ has a uniform distribution on $[0,a)$, and $\Pr[x_2 > a(n-1)]$ is easily obtained from \eqref{eq: Pr x_2 > a(n-1) cond} as
+\begin{equation}
+\label{eq: Pr x_2 > a(n-1), first case}
 \begin{split}
-\Pr[x_2 \geq a(n-1)] &= \E[\Pr[x_2 \geq a(n-1) \cond x_1]] \\
+\Pr[x_2 > a(n-1)] &= \E[\Pr[x_2 > a(n-1) \cond x_1]] \\
 &= \frac 1 {\pi a} \int_0^a\arccos \frac{a(n-1)-x_1}{\len} \diff x_1 \\
@@
 \begin{equation}
-\label{eq: Pr x_2 geq a(n-1), second case}
-\Pr[x_2 \geq a(n-1)] = \frac{\len}{\pi a}\,\, \left( 1 - f\left( \frac{a(n-2)}{\len} \right) \right),
+\label{eq: Pr x_2 > a(n-1), second case}
+\Pr[x_2 > a(n-1)] = \frac{\len}{\pi a}\,\, \left( 1 - f\left( \frac{a(n-2)}{\len} \right) \right),
 \end{equation}
@@
 
-The results in Theorem~\ref{theo: funta, form} and Proposition~\ref{prop: Pr i} make clear the relationship between the problem considered here and Buffon's needle problem, i.e.~the probability that a random segment of fixed length crosses a line in a regular structure of parallel strips \cite[section 1.1]{Mathai99}.
-Firstly, since the number $\pi$ is involved in \eqref{eq: funta, form}, it is possible to design a simple probabilistic experiment to estimate the value of $\pi$, as in Buffon's original experiment.
-For example, choosing $a=b=\len=1$ gives $\funta(\len) = 1+1/\pi$.
+The results in Theorem~\ref{theo: funta, form} and Proposition~\ref{prop: Pr i} make clear the relationship between the problem considered here and Buffon's needle experiment, which considers a random segment of fixed length in a regular structure of parallel strips \cite[section~1.1]{Mathai99}.
+Firstly, since the number $\pi$ is involved in \eqref{eq: funta, form}, it is possible to design a simple probabilistic experiment to estimate the value of $\pi$, as in Buffon's original problem.
+For example, choosing $a=b=\len=1$ gives $\funta(\len) = 1/\pi+1$.
 
@@
 
-Lastly, $\Pr[i \geq 2]$ as computed in Proposition~\ref{prop: Pr i} can be interpreted as the probability of crossing at least one line in Buffon's experiment.
+Lastly, $\Pr[i \geq 2]$ as computed in Proposition~\ref{prop: Pr i} can be interpreted as the probability of crossing at least one line in Buffon's experiment with arbitrary needle length.
 Indeed, for $n=2$ the third expression in \eqref{eq: Pr i geq n} reduces again to $2\len/(\pi a)$ for $\len \leq a$, whereas the second reduces to $2\len f(a/\len)/(\pi a)$ for $\len > a$.
@@
 in agreement with the known result for Buffon's needle problem with $\len > a$ \cite[theorem 1.1.2]{Mathai99}.
-On the other hand, $\Pr[i \geq n]$ for $n \geq 3$ can be seen as the probability of crossing at least one line in a modified version of Buffon's setting in which a needle endpoint is only allowed to move in a region of width $a$ located in the middle of a strip of width $(2n-3)a$.
+On the other hand, $\Pr[i \geq n]$ for $n \geq 3$ can be seen as the probability of crossing at least one line in a modified version of Buffon's set-up in which a needle endpoint is only allowed to move in a region of width $a$ located in the middle of a strip of width $(2n-3)a$.
 
@@
 On the other hand, the maximum number of visited tiles $\funt(\len)$ has an asymptotic slope given by \eqref{eq: asympt slope}.
-It is interesting to consider the ratio of these values, which also gives the asymptotic ratio of average and maximum numbers of visited tiles.
+It is interesting to consider the ratio of these values, which also gives the asymptotic ratio between the average and the maximum numbers of visited tiles.
 This depends only on $a/b$, and is thus expressed as a function $\ras: \mathbb R^+ \to (0,1)$:
@@
 \end{equation}
-The function $\ras$ satisfies $\ras(x) = \ras(1/x)$, and is therefore symmetric in logarithmic scale, as can be observed in Figure~\ref{fig: ratio asympt slopes}.
-It is unimodal with maximum value $2\sqrt 2/\pi \approx 0.9003$ for $a/b=1$, and its limit when $a/b$ tends to $0$ or $\infty$ is $2/\pi \approx 0.6366$.
-This implies that for $\len$ large the average number of tiles $\funta(\len)$ cannot very small compared with the maximum number of tiles $\funt(\len)$.
-For example, $\funta(\len)$ asymptotically exceeds $0.8\funt(\len)$ for $1/3 \leq a/b \leq 3$.
+This function is represented in Figure~\ref{fig: ratio asympt slopes}.
+It satisfies $\ras(x) = \ras(1/x)$, and is therefore symmetric in logarithmic scale, as seen in the figure.
+The function $\ras$ is unimodal with maximum value $2\sqrt 2/\pi \approx 0.9003$ for $a/b=1$, and its limit when $a/b$ tends to $0$ or $\infty$ is $2/\pi \approx 0.6366$.
+This implies that for $\len$ large the average number of tiles $\funta(\len)$ cannot be very small compared with the maximum number of tiles $\funt(\len)$.
+For example, $\funta(\len)$ exceeds $0.8\funt(\len)$ asymptotically when $\len \rightarrow \infty$ for $1/3 \leq a/b \leq 3$.
 
@@
 
-The probability of visiting the maximum number of tiles for a rectangular grid is difficult to compute, due to the irregularity of the optimal $(i,j)$ pairs (as was discussed in \S\ref{part: fund results}; in particular see Figure~\ref{fig: pairs_bounds}).
-In the square case, however, the simplicity of the set of optimal pairs makes the problem more tractable.
+The probability of visiting the maximum number of tiles for a rectangular grid is difficult to compute, due to the irregularity of the optimal $(i,j)$ pairs (which was discussed in \S\ref{part: fund results}; in particular see Figure~\ref{fig: pairs_bounds}).
+In a square grid, however, the simplicity of the set of optimal pairs makes the problem more tractable.
 
@@
 \end{equation}
-Computing the integral in \eqref{eq: g def} and using the identities $\cos^2 \alpha = 1-\sin^2 \alpha$ and $\cos(2\alpha) = 2\cos^2\alpha-1 = 1 - 2\sin^2\alpha$,
+Computing this integral and using the identities $\cos^2 \alpha = 1-\sin^2 \alpha$ and $\cos(2\alpha) = 2\cos^2\alpha-1 = 1 - 2\sin^2\alpha$,
 \begin{equation}
@@
 Given $\len \in \mathbb R^+$, consider a unit square grid and a uniformly random segment of length $\len$, as previously defined.
-The probability $\probmax(\len)$ that the segment visits the maximum number of tiles $\tiles = \funt(\len)$ is
+The probability $\probmax(\len)$ that the segment visits the maximum number of tiles is
 \begin{equation}
@@
 
-For $i,j \in \mathbb Z$, the (square) tile with lower-left coordinate $(i,j)$ is defined as $[i,i+1) \times [j,j+1)$, i.e.~including its lower and left edges.
+For $i,j \in \mathbb Z$, let the (square) tile with lower-left coordinates $(i,j)$ be defined as $[i,i+1) \times [j,j+1)$, i.e.~including its lower and left edges.
 The first endpoint of the segment is contained in the tile with lower-left corner $(0,0)$, which will be referred to as reference tile.
@@
 The area of this rectangle is the probability $\probmax_{i-1,j-1}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$.
-The desired probability $\probmax(\len)$ will be obtained by averaging $\probmax_{i-1,j-1}(\len, \theta)$ over $\theta$, summing over all $i, j \geq 1$ such that the tile with lower-left corner $(i-1, j-1)$ can be reached from the reference tile, and then using symmetry to take into account the tiles with $i \leq 0$ or $j \leq 0$.
+The desired probability $\probmax(\len)$ will be obtained by averaging $\probmax_{i-1,j-1}(\len, \theta)$ with respect $\theta$, summing over all $i, j \geq 1$ such that the tile with lower-left corner $(i-1, j-1)$ can be reached from the reference tile, and then using symmetry to take into account the tiles with $i \leq 0$ or $j \leq 0$.
 
@@
 Under this assumption, the range of lengths for which $\funt(\len)$ equals $\tiles$ is given by \eqref{eq: square, odd max tiles, lengths}.
-For any length in this interval, there are some positions within the reference tile from which the tile with lower-left coordinate $((\tiles-1)/2, (\tiles-1)/2)$ can be reached.
+For any length in this interval, there are some positions within the reference tile from which the tile with lower-left coordinates $((\tiles-1)/2, (\tiles-1)/2)$ can be reached.
 Equivalently, the shifted segment with one endpoint at $((\tiles-1)/2, (\tiles-1)/2)$ can reach the reference tile.
@@
 By symmetry, each of those two tiles contributes the same probability.
-Other tiles with $i,j\geq 1$, $i+j-1 = \tiles$ do not need to be considered, because they cannot be reached with lengths satisfying \eqref{eq: square, odd max tiles, lengths}.
+Other tiles with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, $i,j\geq 1$, $i+j-1 = \tiles$ do not need to be considered, because they cannot be reached with lengths satisfying \eqref{eq: square, odd max tiles, lengths}.
 
@@
 This can only happen for $\tiles \geq 8$.
-No other tiles with $i,j\geq 1$ can be reached from the reference tile $i+j-1 = \tiles$.
+No other tiles with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, $i+j-1 = \tiles$ can be reached from the reference tile.
 
@@
 
-According to the preceding analysis, two cases need to be distinguished for $\tiles = \funt(\len)$ odd, depending on whether \eqref{eq: probmax len odd2} is satisfied or not; and similarly two cases exist for $\tiles = \funt(\len)$ even, depending on whether \eqref{eq: probmax len even2} holds or not.
+Based on the preceding analysis, two cases need to be distinguished for $\tiles = \funt(\len)$ odd, depending on whether \eqref{eq: probmax len odd2} is satisfied or not; and similarly two cases exist for $\tiles = \funt(\len)$ even, depending on whether \eqref{eq: probmax len even2} is satisfied or not.
 These are the four cases in \eqref{eq: probmax}.
 
-In the first case indicated in \eqref{eq: probmax}, as previously argued, $\tiles$ is odd and only the tile with lower-left coordinates $((\tiles-1)/2, (\tiles-1)/2)$ is reachable.
-For a given $\theta$, the conditional probability $\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)$ is the area of the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_diag}:
+In the first case indicated in \eqref{eq: probmax}, $\tiles$ is odd and only the tile with lower-left coordinates $((\tiles-1)/2, (\tiles-1)/2)$ is reachable.
+For a given $\theta$, the conditional probability $\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)$ is the area of the shaded rectangle shown in Figure~\ref{fig: probmax_funt_odd_diag}:
 \begin{equation}
@@
 \end{align}
-Averaging \eqref{eq: probmax: probmax theta odd} over $\theta$, multiplying by $4$ to include the symmetrical tiles in other quadrants and using \eqref{eq: g def} gives $\probmax(\len)$ as
+For $\theta$ outside of this range the conditional probability $\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)$ is $0$.
+Averaging $\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)$ over $\theta$, multiplying by $4$ to include the symmetrical tiles in other quadrants and using \eqref{eq: g def} gives $\probmax(\len)$ as
 \begin{equation}
@@
 
-For the second case in \eqref{eq: probmax}, $\probmax(\len)$ is obtained as above plus an additional term corresponding to the tile with lower-left corner $((\tiles-3)/2, (\tiles+1)/2)$, multiplied by $2$ to account for its symmetrical tile about the line $y=x$, namely that with lower-left corner $((\tiles+1)/2, (\tiles-3)/2)$, and by $4$ to include all quadrants.
+For the second case in \eqref{eq: probmax}, $\probmax(\len)$ is obtained as above plus an additional term corresponding to the tile with lower-left corner $((\tiles-3)/2, (\tiles+1)/2)$, multiplied by $2$ to account for its symmetrical tile about the line $y=x$, which has lower-left corner $((\tiles+1)/2, (\tiles-3)/2)$, and by $4$ to include all quadrants.
 Observing the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_offdiag},
@@
 \end{align}
-The additional contribution to $\probmax(\len)$, given by
+The additional contribution to $\probmax(\len)$ is thus
 \begin{multline}
@@
 \end{multline}
-is recognized to be $2 g(\len, \tiles-5, \tiles-1)$. This proves the second equality in \eqref{eq: probmax}.
+which is $2 g(\len, \tiles-5, \tiles-1)$. This proves the second equality in \eqref{eq: probmax}.
 

### 2023-03-28 paper.tex
Small changes in paper text

@@
 
-This paper deals with the following problems, which relate the length of the segment with the number of visited tiles.
-Two settings are considered, which correspond respectively to the segment position and orientation being deterministic or random.
+This paper studies the relationship between the length of the segment and the number of visited tiles.
+The motivation comes from the classical Buffon-Laplace needle problem (i.e.~a segment with random position and orientation on a rectangular grid), of which a modified version is considered, wherein the segment position and orientation are parameters that can be chosen to maximize the number of visited tiles.
+In addition, the probability of visiting that maximum number of tiles in the classical (random) setting is studied.
+
+Specifically, two different settings are considered, which correspond to the segment position and orientation being deterministic or random, respectively.
 In the \emph{deterministic} case, the relevant questions are:
@@
 
-The questions studied in this paper are related to Buffon's needle problem and Laplace's extension of it.
+The questions studied in this paper are related to Buffon's needle problem and Laplace's extension of it, as stated at the outset.
 Buffon's original problem considers a plane with vertical lines a distance $a$ apart.

### 2023-03-28 figures/figures.m
A reviewer has found a mistake in Theorem 7

@@ -1965,3 +1965,52 @@ result_computed(ind_even2) = 2*g(L_fine(ind_even2)/a, funt_fine(ind_even2)-4, fu
 plot(L_fine.*(ind_odd|ind_odd2)./(ind_odd|ind_odd2), result_computed.*(ind_odd|ind_odd2)./(ind_odd|ind_odd2), '-', 'linewidth', .7), hold on, grid on
 plot(L_fine.*(ind_even|ind_even2)./(ind_even|ind_even2), result_computed.*(ind_even|ind_even2)./(ind_even|ind_even2), '-', 'linewidth', .7), hold on, grid on
 legend({'Simulated' 'Computed, odd max number of squares' 'Computed, even max number of squares'})
+
+
+%% #75. 27iii23: a reviewer finds a mistake in the probability of visiting the maximum number of tiles
+
+% From #65
+
+% Simulation
+clear, clc
+a = 1;
+L = 6.35; %3.8:.1:4.2;%3:.1:3.5;
+N = 2e7;
+ii = ceil(L/sqrt(2)) + 1;
+jj = ceil(sqrt(L.^2 - (ii-2).^2 )) + 1;
+funt = ii+jj-1 % formula valid for real-valued lengths
+z0 = rand(N, numel(L)) + 1j*rand(N, numel(L));
+z1 = z0 + L.*exp(1j*2*pi*rand(N, numel(L)));
+ii = ceil(abs(real(z1))) + (real(z1)<0); % add 1 for negative
+jj = ceil(abs(imag(z1))) + (imag(z1)<0); % add 1 for negative
+result = mean(ii+jj-1 == funt , 1)
+
+% Theoretical
+g = @(r,u,v) ((acos(u/2./r)-asin(v/2./r)).*u.*v + 2*r.^2 + u.^2/2 + v.^2/2 - u.*sqrt(4*r.^2-v.^2) - v.*sqrt(4*r.^2-u.^2))/2/pi;
+L_fine = L;
+ii_theo = ceil(L_fine/sqrt(2)) + 1;
+jj_theo = ceil(sqrt(L_fine.^2 - (ii_theo-2).^2 )) + 1;
+funt_fine = ii_theo+jj_theo-1; % formula valid for real-valued lengths
+result_computed = NaN(size(funt_fine));
+ind_odd = (mod(funt_fine,2)==1) & (L_fine > (funt_fine-3)/sqrt(2)) & (L_fine <= sqrt((funt_fine-5).^2+(funt_fine-1).^2)/2);
[…]

### 2023-03-28 figures/figures.m
Working in fixing the mistake
Simulations match the new computation, not the old

According to the new computation, the probability maxima vary asymptotically as expected

@@ -2014,3 +2014,113 @@ histogram2(ii(ind), jj(ind), 'binmethod', 'integer', 'DisplayStyle' ,'tile', 'Sh
 % That's why I didn't catch this mistake with the simulations.
 
 
+%% # 76. Like #65 but correcting the mistake by including more terms
+
+% The computations here seem to be correct. For L around 10 the difference between the old (wrong) and new
+% computation is already noticeable, and the simulation results match the new, not the old.
+%  Computing for L_fine large enough (around 1000, with fine enough sampling) and using log-log
+% scale the 1/sqrt(L) asymptotic variation of the probability maxima is clearly visible.
+
+clear, clc
+L = .1:.1:10; %3.8:.1:4.2;%3:.1:3.5;
+N = 3e5;
+ii = ceil(L/sqrt(2)) + 1;
+jj = ceil(sqrt(L.^2 - (ii-2).^2 )) + 1;
+funt = ii+jj-1 % formula valid for real-valued lengths
+z0 = rand(N, numel(L)) + 1j*rand(N, numel(L));
+z1 = z0 + L.*exp(1j*2*pi*rand(N, numel(L)));
+ii = ceil(abs(real(z1))) + (real(z1)<0); % add 1 for negative
+jj = ceil(abs(imag(z1))) + (imag(z1)<0); % add 1 for negative
+result = mean(ii+jj-1 == funt , 1);
+plot(L, result, 'o', 'markersize', 4), hold on, grid on
+
+g = @(r,u,v) ((acos(u/2./r)-asin(v/2./r)).*u.*v + 2*r.^2 + u.^2/2 + v.^2/2 - u.*sqrt(4*r.^2-v.^2) - v.*sqrt(4*r.^2-u.^2))/2/pi;
+L_fine = linspace(.01,10,1e4);
+ii = ceil(L_fine/sqrt(2)) + 1;
+jj = ceil(sqrt(L_fine.^2 - (ii-2).^2 )) + 1;
+funt_fine = ii+jj-1; % formula valid for real-valued lengths
+result_computed = NaN(size(funt_fine));
+num_k =  NaN(size(funt_fine));
+for ind_fine = 1:numel(funt_fine)
+    t = funt_fine(ind_fine);
+    L = L_fine(ind_fine);
+    if mod(t,2) % odd
+        res = g(L, t-3, t-3); % this term is always present
+        potential_k = 1:(t-3)/2;
[…]

### 2023-03-31 paper.tex
Issue found by the reviewer fixed
in Matlab. The paper text had not been modified yet regarding this

@@
 For a fixed orientation $\theta$, the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ if and only if shifting the segment so that its second endpoint has coordinates $(i-1,j-1)$ results in the first endpoint being in the reference tile.
-It cannot be the case that the segment ``overshoots'' past this tile, because that would result in a number of visited tiles exceeding $\tiles$, which is the maximum.
+It cannot be the case that the segment ``overshoots'' past this tile, because that would result in a number of visited tiles exceeding $\tiles$, which is the maximum by hypothesis.
 
@@
 The area of this rectangle is the probability $\probmax_{i-1,j-1}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$.
-The desired probability $\probmax(\len)$ will be obtained by averaging $\probmax_{i-1,j-1}(\len, \theta)$ with respect $\theta$, summing over all $i, j \geq 1$ such that the tile with lower-left corner $(i-1, j-1)$ can be reached from the reference tile, and then using symmetry to take into account the tiles with $i \leq 0$ or $j \leq 0$.
+The desired probability $\probmax(\len)$ will be obtained by averaging $\probmax_{i-1,j-1}(\len, \theta)$ with respect to $\theta$, summing over all $i, j \geq 1$ such that the tile with lower-left corner $(i-1, j-1)$ can be reached from the reference tile, and then using symmetry to take into account the tiles with $i \leq 0$ or $j \leq 0$.
 

### 2023-04-10 paper.tex
Redefined g
g_new(s,u,v) equals g_old(s,2*u, 2*v)
Fixing issue found by reviewer

--- 2023-03-28 paper.tex
@@
 \newcommand{\cond}{\,|\,} % less spacing than \mid
-\newcommand{\diff}{\,\mathrm d}
+\newcommand{\diff}{\mathrm d}
 \let\Re\relax % remove the command...
@@
 \newcommand{\mss}{M}
+\newcommand{\nextrat}{K}
+\newcommand{\funcirc}{\Omega}
+\newcommand{\funcircesc}{\bar \Omega}
 
@@
 Consider a line segment placed on a two-dimensional grid of rectangular tiles.
-This paper addresses the relationship between the length of the segment and the number of tiles it visits (i.e.
-has intersection with).
+This paper addresses the relationship between the length of the segment and the number of tiles it visits (i.e.~has intersection with).
 The square grid is also considered explicitly, as some of the specific problems studied are more tractable in that particular case.
@@
 
-As an example of question \ref{prob: determ dir}, consider $a=1.35$, $b=1$.
+As an example of question~\ref{prob: determ dir}, consider $a=1.35$, $b=1$.
 A segment of unit length can be placed as shown in Figure~\ref{fig: examples} (left) to make it visit $3$ tiles.
@@
 
-The \emph{inverse} problem \ref{prob: determ inv} is, given $\tiles \in \mathbb N$, to determine the infimum length of all segments that visit at least $\tiles$ tiles.
+The \emph{inverse} problem~\ref{prob: determ inv} is, given $\tiles \in \mathbb N$, to determine the infimum length of all segments that visit at least $\tiles$ tiles.
 If the length can take any real value the infimum is not a minimum, because given any segment it can be shortened by some small amount without changing the number of visited tiles.
@@
 The segment orientation has a \emph{uniform} distribution on $[0,2\pi)$, and is \emph{independent} of the endpoint position.
-Solving the problem \ref{prob: rand ave} of how many tiles the segment visits on average also answers, as will be seen, the inverse question (segment length to visit a given number of tiles on average).
-A natural, related problem \ref{prob: rand prob} is with what probability the segment visits the maximum number of tiles.
+Solving the problem~\ref{prob: rand ave} of how many tiles the segment visits on average also answers, as will be seen, the inverse question (segment length to visit a given number of tiles on average).
+A natural, related problem~\ref{prob: rand prob} is with what probability the segment visits the maximum number of tiles.
 
@@
 The probability of the needle staying within a single tile is computed in \cite{Arnow94} for the case where $\len < \min\{a,b\}$.
-One of the problems considered in this paper, as mentioned above, is the complementary question \ref{prob: rand prob} of the probability that the needle visits the maximum number of tiles possible for its length.
+One of the problems considered in this paper, as mentioned above, is the complementary question~\ref{prob: rand prob} of the probability that the needle visits the maximum number of tiles possible for its length.
 
@@
 In addition, it is possible to obtain a direct formula also for $\funt$ using a different approach.
-These formulas are given in Theorems \ref{theo: funt, form} and \ref{theo: funl, form}.
+These formulas are given in Theorems~\ref{theo: funt, form} and \ref{theo: funl, form}.
 
@@
 
-The definition of $\funl$ implies that $\funl(\tiles) \geq \funl(\tiles-1)$ for any $\tiles \in \mathbb N$. On the other hand, 
-by Theorem~\ref{theo: funt, form}, $\funt$ is piecewise constant and has a unit-height jump from $\tiles-1$ to $\tiles$ at $\funl(\tiles)$, $\tiles \in \mathbb N$, $\tiles \geq 4$.
+The definition of $\funl$ implies that $\funl(\tiles) \geq \funl(\tiles-1)$ for any $\tiles \in \mathbb N$.
+On the other hand, by Theorem~\ref{theo: funt, form}, $\funt$ is piecewise constant and has a unit-height jump from $\tiles-1$ to $\tiles$ at $\funl(\tiles)$, $\tiles \in \mathbb N$, $\tiles \geq 4$.
 This implies that $\funl(\tiles) > \funl(\tiles-1)$ for $\tiles \geq 4$.
@@
 
-Theorems \ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the first two questions posed in \S\ref{part: intro}; they also provide a way to actually position a segment of length $\len$ or slightly greater than $\funl(\tiles)$, respectively, so that it visits $\funt(\len)$ or $\tiles$ tiles.
+Theorems~\ref{theo: funt, form} and \ref{theo: funl, form} not only give the solutions $\funt(\len)$ and $\funl(\tiles)$ to the first two questions posed in \S\ref{part: intro}; they also provide a way to actually position a segment of length $\len$ or slightly greater than $\funl(\tiles)$, respectively, so that it visits $\funt(\len)$ or $\tiles$ tiles.
 Namely, for $\isoli$, $\jsoli$ computed as in the corresponding theorem, the segment should have its endpoints in the interiors of two tiles shifted $\isoli-1$ steps horizontally and $\jsoli-1$ steps vertically with respect to each other, with the exact position and orientation of the segment adjusted to avoid any grid points.
@@
 \label{eq: f}
-f(\genvar) = \int_0^\genvar \arccos z \diff z = \genvar \arccos \genvar - \sqrt{1-\genvar^2} + 1.
+f(\genvar) = \int_0^\genvar \arccos z \, \diff z = \genvar \arccos \genvar - \sqrt{1-\genvar^2} + 1.
 \end{equation}
@@
 \Pr[x_2 > a(n-1)] &= \E[\Pr[x_2 > a(n-1) \cond x_1]] \\
-&= \frac 1 {\pi a} \int_0^a\arccos \frac{a(n-1)-x_1}{\len} \diff x_1 \\
+&= \frac 1 {\pi a} \int_0^a\arccos \frac{a(n-1)-x_1}{\len} \, \diff x_1 \\
 &= \frac{\len}{\pi a}\,\, \left( f\left( \frac{a(n-1)}{\len} \right) - f\left( \frac{a(n-2)}{\len} \right) \right),
@@
 \label{eq: g def}
-g(\genvar, u, v) = \frac 2 \pi \int_{\arcsin \frac u {2\genvar}}^{\arccos \frac v {2\genvar}} \left( \genvar \sin \theta - \frac u 2 \right) \left( \genvar \cos \theta - \frac v 2 \right) \diff \theta.
+g(\genvar, u, v) = \frac 2 \pi \int_{\arcsin \frac v {\genvar}}^{\arccos \frac u {\genvar}} \left( \genvar \cos \theta - u \right) \left( \genvar \sin \theta - v \right) \, \diff \theta.
 \end{equation}
@@
 \begin{split}
-g(\genvar, u, v) &= \frac 1 {2\pi} \biggl(
-\left(\arccos\frac{v}{2\genvar}-\arcsin\frac{u}{2\genvar}\right) u v + 2\genvar^2 + \frac{u^2} 2 + \frac{v^2} 2 \\
-& \quad  - u \sqrt{4\genvar^2-v^2} - v \sqrt{4\genvar^2-u^2} \biggr).
+g(\genvar, u, v) &= \frac 1 {\pi} \biggl(
+2\left(\arccos\frac{u}{\genvar}-\arcsin\frac{v}{\genvar}\right) u v + \genvar^2 + u^2 + v^2 \\
+& \quad - 2 u \sqrt{\genvar^2-v^2} - 2 v \sqrt{\genvar^2-u^2} \biggr).
 \end{split}
@@
 \begin{equation}
-\label{eq: probmax}
-\begin{split}
-\probmax(\len) & = g\left(\len, 
-\tiles-3, \tiles-3 \right) \quad\ \text{for } \tiles \text{ odd, } \textstyle \len \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2 \right] \\
-\probmax(\len) & = g\left(\len, \tiles-3, \tiles-3 \right) + 2 g\left(\len, \tiles-5, \tiles-1 \right) \\
-& \text{for } \tiles \text{ odd, } \textstyle \len \in \left (\frac{\sqrt{(\tiles-5)^2+(\tiles-1)^2}} 2,  \frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2 \right] \\
-\probmax(\len) & = 2 g\left(\len, \tiles-4, \tiles-2 \right) \quad \text{for } \tiles \text{ even, } \textstyle \len \in \left( \frac{\sqrt{(\tiles-4)^2+(\tiles-2)^2}} 2, \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2 \right] \\
-\probmax(\len) & = 2 g\left(\len, \tiles-4, \tiles-2 \right) + 2 g\left(\len, \tiles-6, \tiles \right) \\
-& \text{for } \tiles \text{ even, } \textstyle \len \in \left( \frac{\sqrt{(\tiles-6)^2+\tiles^2}} 2, \frac{\tiles-2}{\sqrt{2}} \right].
-\end{split}
+\label{eq: probmax impl}
+\probmax(\len) = \sum_{\substack{i,j \geq 2,\ % 
+i+j-1=\tiles,\\
+(i-2)^2+(j-2)^2<\len^2
+}} g\left(\len, i-2, j-2 \right),
+\end{equation}
+or equivalently, defining $\nextrat = \sqrt{\len^2/2-(\tiles-3)^2/4}$,
+\begin{equation}
+\label{eq: probmax expl}
+\probmax(\len) =
+\begin{cases}
+\displaystyle
+g\left(\len, \frac{\tiles-3} 2, \frac{\tiles-3} 2 \right) +
+2 \sum_{k=1}^{\lceil\nextrat-1 \rceil} g\left(\len, \frac{\tiles-3} 2 - k, \frac{\tiles-3} 2 + k \right) &\text{for $\tiles$ odd} \\[6mm] % *!*
+\displaystyle
+2 g\left(\len, \frac{\tiles-4} 2, \frac{\tiles-2} 2 \right) +
+2 \sum_{k=1}^{\lceil\nextrat-3/2 \rceil} g\left(\len, \frac{\tiles-4} 2 - k, \frac{\tiles-2} 2 + k \right) &\text{for $\tiles$ even}.
+\end{cases}
 \end{equation}
@@
 
-For $i,j \in \mathbb Z$, let the (square) tile with lower-left coordinates $(i,j)$ be defined as $[i,i+1) \times [j,j+1)$, i.e.~including its lower and left edges.
-The first endpoint of the segment is contained in the tile with lower-left corner $(0,0)$, which will be referred to as reference tile.
-For an arbitrary value $\tiles$, the segment visits $\tiles$ tiles almost surely if its second endpoint is in a tile with lower-left corner $(i-1,j-1)$ such that $i,j\geq 1$ and $i+j-1=\tiles$, or in a tile symmetrical to this about the line $x=1/2$ (with $i \leq 0$), or about $y=1/2$ (with $j \leq 0$), or about both.
-
-Given $\len \in \mathbb R^+$, let $\tiles = \funt(\len)$, and consider $i,j\geq 1$ such that $i+j-1=\tiles$.
-For a fixed orientation $\theta$, the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ if and only if shifting the segment so that its second endpoint has coordinates $(i-1,j-1)$ results in the first endpoint being in the reference tile.
-It cannot be the case that the segment ``overshoots'' past this tile, because that would result in a number of visited tiles exceeding $\tiles$, which is the maximum.
+% For $i,j \in \mathbb Z$, let the (square) tile with lower-left coordinates $(i,j)$ be defined as $[i,i+1) \times [j,j+1)$, i.e.~including its lower and left edges.
+
+Without loss of generality, the first endpoint of the segment can be assumed to be contained in the tile with lower-left corner $(0,0)$, which will be called the reference tile.
+
+Consider $\len \in \mathbb R^+$, and let $\tiles = \funt(\len)$.
+By Proposition~\ref{prop: i+j-1}, the segment visits $\tiles$ tiles almost surely if the dimensions $i$, $j$ of its discrete bounding rectangle satisfy $i+j-1=\tiles$.
+In addition, a segment that visits the maximum number of tiles always has $i, j \geq 2$, because if $i=1$ the segment can always be shifted and possibly tilted to cross a vertical grid line, thus increasing $i$ to $2$, without reducing $j$.
+The reasoning for $j=1$ is analogous.
+Thus for $i, j \geq 2$, $i+j-1=\tiles$ let $\probmax_{i,j}(\len)$ denote the probability that the segment has a discrete bounding rectangle with dimensions $i, j$.
+Then
+\begin{equation}
+\label{eq: probmax len probmax i,j len}
+\probmax(\len) =  \sum_{i,j \geq 2,\ i+j-1=\tiles} \probmax_{i,j}(\len).
+\end{equation}
+
+For any $i, j \geq 2$ with $i+j-1=\tiles$, the segment has a discrete bounding rectangle with dimensions $i, j$ if and only if its second endpoint is in the tile with lower-left corner $(i-1,j-1)$ or in one of the other three symmetrical tiles with respect to the lines $x=1/2$ and $y=1/2$, i.e.~those with lower-left corners $(i-1,1-j)$, $(1-i,1-j)$ and $(1-i,1-j)$.
+An example is shown in Figure~\ref{fig: probmax_symmetric_tiles}.
+Consequently, $\probmax_{i,j}(\len)$ can be obtained by computing the probability that the second endpoint is in the tile with lower-left corner $(i-1,j-1)$ and then multiplying by $4$.
 
@@
 \centering%
-\subfigure[The second endpoint is in the tile with lower-left corner $((\tiles-1)/2,(\tiles-1)/2)$]{%
-\label{fig: probmax_funt_odd_diag}%
-\includegraphics[width=.56\textwidth]{probmax_funt_odd_diag}%
-}\\%\hfill%
-\subfigure[The second endpoint is in the tile with lower-left corner $((\tiles-3)/2,(\tiles+1)/2)$]{%
-\label{fig: probmax_funt_odd_offdiag}%
-\includegraphics[width=.56\textwidth]{probmax_funt_odd_offdiag}%
-}%
-\caption{Segment orientations and positions for a given $\tiles = \funt(\len)$ on a unit square grid, with $\tiles$ odd. Example for $\tiles = 7$
+\includegraphics[width=.5\textwidth]{probmax_symmetric_tiles_}%
+\caption{Symmetric tiles for the computation of $\probmax_{i,j}(\len)$ on a unit square grid. Example for $i=3$, $j=4$%
+}%
+\label{fig: probmax_symmetric_tiles}%
+\end{figure}%
+
+By Proposition~\ref{prop: len ineq i j}, the tile with lower-left corner $(i-1,j-1)$ can be visited only if
+\begin{equation}
+\label{eq: i-2 j-2 len}
+(i-2)^2+(j-2)<\len^2.
+\end{equation}
+Thus $\probmax_{i,j}(\len)=0$ if this restriction does not hold. In the following it will be assumed that it does.
+
+For a fixed orientation $\theta$, the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ if and only if shifting the segment up and to the right so that its first endpoint is at $(1,1)$ results in the second endpoint still being in the tile with lower-left corner $(i-1,j-1)$, as shown in Figure~\ref{fig: probmax_funt_odd}.
+It cannot be the case that the segment ``overshoots'' past this tile, because that would yield a number of visited tiles exceeding $\tiles$, which is the maximum.
+In other words, only one corner of the tile can be contained within the arc in the figure, namely its lower-left corner $(i-1,j-1)$.
+
+\begin{figure}
+\centering
+\includegraphics[width=.56\textwidth]{probmax_funt_odd}%
+\caption{Reachability from the reference tile for a given $\tiles = \funt(\len)$ on a unit square grid. Example for $\tiles = 7$
 }%
@@
 
-There is a range of possible values of $\theta$ for which the above condition is satisfied.
+There is a range of values of $\theta$ for which the above condition is satisfied.
 This range corresponds to the part of the arc with solid line in Figure~\ref{fig: probmax_funt_odd}.
-For convenience, the angle $\theta$ is defined clockwise from the downward direction.
-For each $\theta$ in the allowed range, the valid positions for the first endpoint of the (shifted) segment are in a rectangle contained in the reference tile.
-The rectangle associated with an example $\theta$ is shown in the figure, shaded.
-The area of this rectangle is the probability $\probmax_{i-1,j-1}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$.
-The desired probability $\probmax(\len)$ will be obtained by averaging $\probmax_{i-1,j-1}(\len, \theta)$ with respect $\theta$, summing over all $i, j \geq 1$ such that the tile with lower-left corner $(i-1, j-1)$ can be reached from the reference tile, and then using symmetry to take into account the tiles with $i \leq 0$ or $j \leq 0$.
-
-The situation is different depending on whether $\tiles = \funt(\len)$ is odd or even.
-Consider $\tiles$ odd first.
-Under this assumption, the range of lengths for which $\funt(\len)$ equals $\tiles$ is given by \eqref{eq: square, odd max tiles, lengths}.
-For any length in this interval, there are some positions within the reference tile from which the tile with lower-left coordinates $((\tiles-1)/2, (\tiles-1)/2)$ can be reached.
-Equivalently, the shifted segment with one endpoint at $((\tiles-1)/2, (\tiles-1)/2)$ can reach the reference tile.
-This is illustrated in Figure~\ref{fig: probmax_funt_odd_diag}.
-In addition, if
-\begin{equation}
-\label{eq: probmax len odd2}
-\len > \frac{\sqrt{(\tiles-5)^2 + (\tiles-1)^2}}{2}
-\end{equation}
-the two tiles with lower-left coordinates $((\tiles-3)/2, (\tiles+1)/2)$ and $((\tiles+1)/2, (\tiles-3)/2)$ are also reachable from the reference tile; or equivalently the shifted segment can reach the reference tile, as shown in Figure~\ref{fig: probmax_funt_odd_offdiag}.
-This can only happen for $\tiles \geq 5$ (for $\tiles=3$ the inequality \eqref{eq: probmax len odd2} cannot be satisfied for lengths in the range given by \eqref{eq: square, odd max tiles, lengths}).
-By symmetry, each of those two tiles contributes the same probability.
-Other tiles with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, $i,j\geq 1$, $i+j-1 = \tiles$ do not need to be considered, because they cannot be reached with lengths satisfying \eqref{eq: square, odd max tiles, lengths}.
-
-For $\tiles$ even, which corresponds to lengths in the interval \eqref{eq: square, even max tiles, lengths}, the two tiles with lower-left coordinates $(\tiles/2-1, \tiles/2)$ and $(\tiles/2, \tiles/2-1)$ can always be reached from some positions within the reference tile.
-In addition, if
-\begin{equation}
-\label{eq: probmax len even2}
-\len > \frac{\sqrt{((\tiles-6)^2 + \tiles^2}}{2}
-\end{equation}
-the two tiles with lower-left coordinates $(\tiles/2-2, \tiles/2+1)$ and $(\tiles/2+1, \tiles/2-2)$ have to be considered too.
-This can only happen for $\tiles \geq 8$.
-No other tiles with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, $i+j-1 = \tiles$ can be reached from the reference tile.
-
-From the above two paragraphs it stems that, among all possible tiles with lower-left corner $(i-1,j-1)$, $i,j\geq 1$, those whose centers lie on the lines $x=1/2$ or $y=1/2$ (i.e.~those with $i=1$ of $j=1$) are never reachable from the reference tile, in the sense that a segment that touches the maximum number of tiles cannot have its second endpoint in those tiles.
-This implies that $\probmax(\len)$ can be computed by considering only the tiles with $i,j\geq 1$ (actually $i,j \geq 2$), and then multiplying by $4$ to include the symmetrical tiles with respect to the lines $x=1/2$ and $y=1/2$.
-As an example, Figure~\ref{fig: probmax_possible_tiles} shows the reference tile (solid line) and the tiles that can contain the second endpoint (dotted lines) for $\funt(\len) = 5$.
-
+For each $\theta$ in the allowed range, the valid positions for the second endpoint of the segment are in a rectangle contained in the tile with lower-left corner $(i-1,j-1)$.
+These are the positions that cause the first endpoint to be in the reference tile.
+The rectangle associated with an example $\theta$ is shown in Figure~\ref{fig: probmax_funt_odd}, shaded.
+The area of this rectangle is the probability $\probmax_{i, j}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$.
+As seen in the figure, this area is
+\begin{equation}
+\label{eq: probmax: probmax theta odd}
+\probmax_{i, j}(\len, \theta) = \left(\len\cos\theta - i+2\right) \left(\len\sin\theta - j + 2\right).
+\end{equation}
+The range of allowed values for $\theta$, as can also be deduced from the figure, is $(\theta_0, \theta_1)$ with
+\begin{align}
+\label{eq: odd theta 0}
+\theta_0 & = \arcsin\frac{j-2}{\len}, \\
+\label{eq: odd theta 1}
+\theta_1 & = \arccos\frac{i-2}{\len}.
+\end{align}
+For $\theta$ outside of this range the conditional probability $\probmax_{i, j}(\len, \theta)$ is $0$.
+Averaging $\probmax_{i, j}(\len, \theta)$ over $\theta$ gives the probability that the segment visits this tile, and then multiplying by $4$ yields $\probmax_{i.j}(\len)$:
+\begin{equation}
+\begin{split}
+\label{eq: probmax: cond on theta}
+\probmax_{i,j}(\len) &= 4 \int_{\theta_0}^{\theta_1} \frac{\probmax_{i, j}(\len, \theta)}{2\pi} \, \diff \theta \\
+&= \frac 2 {\pi} \int_{\arcsin\frac{j-2}{\len}}^{\arccos\frac{i-2}{\len}} \left( \len\cos\theta - i + 2 \right)  \left( \len\sin\theta - j + 2 \right) \, \diff \theta
+= g( \len, i-2, j-2).
+\end{split}
+\end{equation}
+
+The above arguments show that $\probmax_{i,j}(\len)$ is given by \eqref{eq: probmax: cond on theta} if the condition \eqref{eq: i-2 j-2 len} holds, and equals $0$ otherwise.
+Using this into \eqref{eq: probmax len probmax i,j len} yields \eqref{eq: probmax impl} .
+
+Given $\tiles$, the range of lengths $\len$ such that $\tiles = \funt(\len)$ is either \eqref{eq: square, odd max tiles, lengths} or \eqref{eq: square, even max tiles, lengths}.
+For $\len$ in this range, the number of tiles with lower-left corners $(i-1,j-1)$, $i,j \geq 2$, $i+j-1=\tiles$ that can be reached, i.e.~the number of terms in \eqref{eq: probmax impl}, depends on $\len$.
+If $\tiles$ is odd the minimum number is $1$, corresponding to a tile with lower-left corner on the line $x=y$, which gives the first term in the first expression in \eqref{eq: probmax expl}.
+For $\tiles$ even the minimum number is $2$, corresponding to two symmetrical tiles whose $x$ and $y$ coordinates differ by $1$, which give the first term in the second expression.
+Either for $\tiles$ odd or even, additional tiles may become reachable as $\len$ grows (within the range allowed by $\tiles$), and these always occur in symmetrical pairs.
+These tiles (if any) correspond to the terms in the (possibly empty) sum indexed by $k$ in either expression in \eqref{eq: probmax expl}.
+
+Figure~\ref{fig: probmax_terms} contains examples for several values of $\tiles$, with a few values of $\len$ for each $\tiles$.
+Each length is represented by an arc with radius $\len$ centered at $(1,1)$.
+For a given $\tiles$, the filled circular markers are the lower-left corners of the tiles that are always reachable, whereas the empty circles correspond to tiles whose reachability depends on $\len$.
+% For example, with $\tiles=11$ and the greatest $\len$ shown in the figure the sum over $k$ in \eqref{eq: probmax expl} has $2$ terms, corresponding to the $2$ pairs of empty markers on the diagonal line that are enclosed by the arc.
+
+\begin{figure}
+\centering
+\includegraphics[width=.7\textwidth]{probmax_terms}%
+\caption{Tiles with lower-left corners $(i-1,j-1)$, $i+j-1=\tiles=\funt(\len)$ that can be reached from the reference tile on a unit square grid. Examples for several values of $\tiles$ and $\len$.
+}%
+\label{fig: probmax_terms}%
+\end{figure}%
+
+As can be seen in Figure~\ref{fig: probmax_terms}, for each $\tiles$ there is a maximum number of tiles that can be reached, beyond which incrementing $\len$ only results in $\tiles$ increasing by $1$.
+For a given $\len$, the number of reachable tiles can be determined by counting how many markers are covered by the length $w$ defined in Figure~\ref{fig: probmax_nterms}.
+This length satisfies the relationship $w^2 = \len^2 - (\tiles-3)^2/2$, and thus $K$ as defined in the theorem equals $w/\sqrt{2}$.
+From the figure, the number of terms in the sum over $k$ (empty circular markers) is $\lceil w/\sqrt 2 - 1 \rceil$ or $\lceil w/\sqrt 2 - 3/2 \rceil$ for $\tiles$ odd or even, i.e.~$\lceil \nextrat - 1 \rceil$ or $\lceil \nextrat - 3/2 \rceil$ respectively.
+This establishes \eqref{eq: probmax expl}.
 \begin{figure}
 \centering%
-\includegraphics[width=.52\textwidth]{probmax_possible_tiles}%
-\caption{Possible tiles for the endpoints of a segment that visits the maximum number of tiles $\tiles= \funt(\len)$ on a unit square grid, with $\tiles$ odd. Example for $\tiles = 5$%
-}%
-\label{fig: probmax_possible_tiles}%
+\subfigure[Odd $\tiles$]{%
+\label{fig: probmax_nterms_odd}%
+\includegraphics[width=.65\textwidth]{probmax_nterms_odd}%
+}\\%
+\subfigure[Even $\tiles$]{%
+\label{fig: probmax_nterms_even}%
+\includegraphics[width=.65\textwidth]{probmax_nterms_even}%
+}%
+\caption{Computation of the number of reachable tiles with lower-left corners $(i-1,j-1)$, $i+j-1=\tiles=\funt(\len)$ on a unit square grid}%
+\label{fig: probmax_nterms}%
 \end{figure}%
-
-Based on the preceding analysis, two cases need to be distinguished for $\tiles = \funt(\len)$ odd, depending on whether \eqref{eq: probmax len odd2} is satisfied or not; and similarly two cases exist for $\tiles = \funt(\len)$ even, depending on whether \eqref{eq: probmax len even2} is satisfied or not.
-These are the four cases in \eqref{eq: probmax}.
-
-In the first case indicated in \eqref{eq: probmax}, $\tiles$ is odd and only the tile with lower-left coordinates $((\tiles-1)/2, (\tiles-1)/2)$ is reachable.
-For a given $\theta$, the conditional probability $\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)$ is the area of the shaded rectangle shown in Figure~\ref{fig: probmax_funt_odd_diag}:
-\begin{equation}
-\label{eq: probmax: probmax theta odd}
-\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta) = \left(\len\sin\theta - \frac{\tiles-3}2\right) \left(\len\cos\theta - \frac{\tiles-3}2\right).
-\end{equation}
-The range of allowed values for $\theta$, as can be deduced from the figure, is $(\theta_0, \theta_1)$ with
-\begin{align}
-\label{eq: odd theta 0}
-\theta_0 & = \arcsin\frac{\tiles-3}{2\len}, \\
-\label{eq: odd theta 1}
-\theta_1 & = \arccos\frac{\tiles-3}{2\len}.
-\end{align}
-For $\theta$ outside of this range the conditional probability $\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)$ is $0$.
-Averaging $\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)$ over $\theta$, multiplying by $4$ to include the symmetrical tiles in other quadrants and using \eqref{eq: g def} gives $\probmax(\len)$ as
-\begin{equation}
-\label{eq: probmax: probmax odd}
-\begin{split}
-\probmax(\len) & = 4 \int_{\theta_0}^{\theta_1} \frac{\probmax_{(\tiles-1)/2, (\tiles-1)/2}(\len, \theta)}{2\pi} \diff \theta \\
-& = \frac 2 \pi \int_{\arcsin\frac{\tiles-3}{2\len}}^{\arccos\frac{\tiles-3}{2\len}} \left( \len\sin\theta - \frac{\tiles-3} 2 \right)  \left( \len\cos\theta - \frac{\tiles-3} 2 \right) \diff \theta \\
-& = g\left( \len, \tiles-3, \tiles-3 \right ),
-\end{split}
-\end{equation}
-in accordance with the first expression in \eqref{eq: probmax}.
-
-For the second case in \eqref{eq: probmax}, $\probmax(\len)$ is obtained as above plus an additional term corresponding to the tile with lower-left corner $((\tiles-3)/2, (\tiles+1)/2)$, multiplied by $2$ to account for its symmetrical tile about the line $y=x$, which has lower-left corner $((\tiles+1)/2, (\tiles-3)/2)$, and by $4$ to include all quadrants.
-Observing the shaded rectangle in Figure~\ref{fig: probmax_funt_odd_offdiag},
-\begin{equation}
-\label{eq: probmax: probmax theta odd2}
-\probmax_{(\tiles-3)/2, (\tiles+1)/2}(\len, \theta) = \left(\len\sin\theta - \frac{\tiles-5}2\right) \left(\len\cos\theta - \frac{\tiles-1}2\right).
-\end{equation}
-The integration interval $(\theta_0, \theta_1)$ is in this case
-\begin{align}
-\theta_0 & = \arcsin\frac{\tiles-5}{2\len}, \\
-\theta_1 & = \arccos\frac{\tiles-1}{2\len}.
-\end{align}
-The additional contribution to $\probmax(\len)$ is thus
-\begin{multline}
-8 \int_{\theta_0}^{\theta_1} \frac{\probmax_{(\tiles-3)/2, (\tiles+1)/2}(\len, \theta)}{2\pi} \diff \theta \\
-= \frac 4 \pi \int_{\arcsin\frac{\tiles-5}{2\len}}^{\arccos\frac{\tiles-1}{2\len}} \left( \len\sin\theta - \frac{\tiles-5} 2 \right)  \left( \len\cos\theta - \frac{\tiles-1} 2 \right) \diff \theta,
-\end{multline}
-which is $2 g(\len, \tiles-5, \tiles-1)$. This proves the second equality in \eqref{eq: probmax}.
-
-The expressions for the third and fourth cases in \eqref{eq: probmax}, corresponding to $\tiles$ even, are established analogously.
 \end{proof}
 
-The technique used in the proof of Theorem~\ref{theo: probmax, sq} could in principle be used for obtaining the probability that the number of tiles visited by the segment equals or exceeds any given value $\tiles < \funt(\len)$.
-The resulting expressions, however, are cumbersome, because more cases (more subintervals for the length) need to be distinguished depending on how many tiles with lower-left corner $(i-1,j-1)$ such that $i+j-1=\tiles$ can be reached from the reference tile.
+The technique used in the proof Theorem~\ref{theo: probmax, sq} could in principle be used for obtaining the probability that the number of tiles visited by the segment equals or exceeds any given value $\tiles < \funt(\len)$.
+However, the process is more cumbersome.
+Specifically, with reference to Figure~\ref{fig: probmax_funt_odd}, the arc in this more general setting can enclose any number of corners of the considered tile, not necessarily $1$, which makes the characterization of the shaded rectangles more complicated.
+% because several cases need to be distinguished
 
@@
 
-Figure~\ref{fig: probmax_len} shows the probability $\probmax(\len)$ of visiting the maximum number of tiles on a unit square grid, as given by \eqref{eq: probmax}.
+The probability $\probmax(\len)$ of visiting the maximum number of tiles on a unit square grid, computed from Theorem~\ref{theo: probmax, sq}, is shown in Figure~\ref{fig: probmax_len} shows.
 As $\len$ grows, $\probmax(\len)$ has a jump when $\funt(\len)$ increases by $1$ (and there $\probmax(\len)$ takes the value approached from the left).
@@
 \begin{equation}
+\label{eq: len tiles}
 \len_\tiles = \begin{cases}
 \displaystyle
-\frac{\sqrt{(\tiles-3)^2 + (\tiles-1)^2}} {2} & \text{ for $\tiles$ odd, $\tiles \geq 3$} \\
+\frac{\sqrt{(\tiles-3)^2 + (\tiles-1)^2}} {2} = \sqrt{\frac{(\tiles-2)^2+1}{2}} & \text{ for $\tiles$ odd, $\tiles \geq 3$} \\
 \displaystyle
-\frac{\tiles-2}{\sqrt{2}} & \text{ for $\tiles$ odd, $\tiles \geq 4$},
+\frac{\tiles-2}{\sqrt{2}} & \text{ for $\tiles$ even, $\tiles \geq 4$},
 \end{cases}
@@
 
-The first continuous section seen in Figure~\ref{fig: probmax_len} corresponds to a maximum number of visited tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, where $\len_3= 1$.
+The first continuous section plotted in Figure~\ref{fig: probmax_len} corresponds to a maximum number of visited tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, where $\len_3= 1$.
 The second corresponds to $\tiles=4$, for lengths in $(\len_3, \len_4]$, where $\len_4= \sqrt{2}$.
 Within each continuous section the probability monotonically increases from $0$ to a maximum value.
-The heights of the maxima follow different trends for odd and even $\tiles$ respectively, as can be seen in the figure; but in each case they are asymptotically proportional to $1/\len_\tiles$.
-This is established by the next result.
+The heights of the maxima are asymptotically proportional to $1/\sqrt{\len_\tiles}$, as established by the next result.
 
@@
 \liminf_{\len \rightarrow \infty}\, \probmax(\len) & = 0, \\ 
-\label{eq: probmax lim odd}
-\lim_{\substack{\tiles \rightarrow \infty \\
-\tiles \ \mathrm{odd}}}\, \tiles \probmax(\len_\tiles) & = \frac 2 \pi, \\
-\label{eq: probmax lim even}
-\lim_{\substack{\tiles \rightarrow \infty \\ \tiles \ \mathrm{even}}}\, \tiles \probmax(\len_\tiles) & = \frac 8 {3\pi}, \\
-\label{eq: probmax limsup}
-\limsup_{\len \rightarrow \infty}\, \len \probmax(\len) & = \frac {4\sqrt{2}}{3\pi}.
+\label{eq: probmax lim sqrt tiles}
+\lim_{\tiles \rightarrow \infty}\, \sqrt{\tiles} \probmax(\len_\tiles) & = \frac {32\sqrt{2}}{105\pi}, \\
+\label{eq: probmax limsup sqrt len}
+\limsup_{\len \rightarrow \infty}\, \sqrt{\len} \probmax(\len) & = \frac {32\sqrt[4]{2}}{105\pi}.
 \end{align}
@@
 \begin{proof}
-The probability $\probmax(\len)$ becomes $0$ whenever the length causes the maximum number of visited tiles to increase by $1$. This proves \eqref{eq: probmax liminf}.
-
-For \eqref{eq: probmax lim odd} it suffices to note that
-\begin{equation}
-\begin{split}
-\label{eq: tedious limit, odd}
-&\lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2, \tiles-3, \tiles-3 \right) \\
-&\quad = \lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\sqrt{(\tiles-3)^2+(\tiles-1)^2}} 2, \tiles-5, \tiles-1 \right) = \frac 2 {3\pi},
-\end{split}
-\end{equation}
-as can be (somewhat tediously) checked using L'H\^opital's rule \cite[section~5.3]{Abbott15}.
-% It's very tedious by hand.
-I used Maple to compute the limit, and then checked it numerically.
-Maybe one or two intermediate steps from applying L'H\^opital's rule should be included.
-Perhaps those can be obtained with Maple too, and this part can be rewritten.
-Combining these expressions with the second equality in \eqref{eq: probmax} yields \eqref{eq: probmax lim odd}.
-
-Similarly,
-\begin{equation}
-\label{eq: tedious limit, even}
-\lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\tiles-2}{\sqrt 2}, \tiles-4, \tiles-2 \right) =
-\lim_{\tiles \rightarrow \infty}\, \tiles g\left(\frac{\tiles-2}{\sqrt 2}, \tiles-6, \tiles \right)
-= \frac 2 {3\pi},
-\end{equation}
-and combining with the fourth equality in \eqref{eq: probmax} gives \eqref{eq: probmax lim even}.
-
-For large $\tiles$, $\len_\tiles$ asymptotically approaches $\tiles/\sqrt{2}$.
-Using this into \eqref{eq: probmax lim even}, which is greater than \eqref{eq: probmax lim odd}, yields \eqref{eq: probmax limsup}.
+The probability $\probmax(\len)$ can be made arbitrarily close to $0$ by choosing $\len$ close enough to the left endpoint of either \eqref{eq: square, odd max tiles, lengths} or \eqref{eq: square, even max tiles, lengths}.
+This proves \eqref{eq: probmax liminf}.
+
+For the purpose of establishing \eqref{eq: probmax lim sqrt tiles}, consider the graph in Figure~\ref{fig: probmax_nterms} with $\len= \len_\tiles$.
+It is convenient to rotate this graph by $45^\circ$ and shift it so that the original $x+y = \tiles-1$ line (shown dashed in Figure~\ref{fig: probmax_nterms}) becomes the new horizontal axis.
+Figure~\ref{fig: probmax_asint_turned} shows the result, using $\tiles=9$ as an example.
+The circular markers in this figure correspond, in the original graph before rotating, to the lower-left corners $(i-1,j-1)$ of the tiles with $i+j-1=\tiles$, $(i-2)^2+(j-1)^2<\len_\tiles$.
+Each of these tiles can be (partially) seen above the corresponding marker in Figure~\ref{fig: probmax_asint_turned}, its sides forming $45^\circ$ degrees with respect to the horizontal axis.
+Let $x_{\tiles, n}$ denote the coordinates of these points along the horizontal axis of Figure~\ref{fig: probmax_asint_turned}.
+
+Using the coordinate axes in Figure~\ref{fig: probmax_asint_turned}, the circular arc satisfies the equation $y = \funcirc(x)$ with
+\begin{equation}
+\label{eq: funcirc}
+\funcirc(x) = h_\tiles - \len_\tiles + \sqrt{\len_\tiles^2-x^2}.
+\end{equation}
+The $y$ coordinate of the arc for $x=x_{\tiles,n}$ will be denoted as $y_{\tiles,n}$.
+The half-width of the arc along the horizontal axis, $w_\tiles$, is computed in the proof of Theorem~\ref{theo: probmax, sq} (or from the right triangle in Figure~\ref{fig: probmax_asint_turned}), and using \eqref{eq: len tiles} it can be expressed as
+\begin{equation}
+\label{eq: w}
+w_\tiles = \sqrt{\len_\tiles^2-\frac{(\tiles-3)^2}{2}} =
+\begin{cases}
+\displaystyle
+\sqrt{\tiles-2} & \text{ for $\tiles$ odd} \\
+\displaystyle
+\sqrt{\tiles-5/2} & \text{ for $\tiles$ even}.
+\end{cases}
+\end{equation}
+The inequality
+\begin{equation}
+\label{eq: sqrt asympt}
+\sqrt{1+\genvar} \leq 1 + \frac \genvar 2
+\end{equation}
+implies that
+\begin{equation}
+\label{eq: asympt w}
+w_\tiles \leq \sqrt{\tiles-2} < \sqrt{\tiles} - 1/\sqrt\tiles.
+\end{equation}
+Similarly, \eqref{eq: len tiles} and \eqref{eq: sqrt asympt} yield, considering that $\tiles \geq 3$,
+\begin{equation}
+\label{eq: asympt len tiles}
+\frac{\tiles-2}{\sqrt{2}} \leq \len_\tiles \leq \frac{\tiles-2}{\sqrt{2}} + \frac{1}{2\sqrt{2}(\tiles-2)}
+\leq \frac{\tiles-3/2}{\sqrt{2}}.
+\end{equation}
+The arc height,
+\begin{equation}
+\label{eq: h tiles}
+h_\tiles = \len_\tiles - \frac{\tiles-3}{\sqrt{2}},
+\end{equation}
+can be bounded from \eqref{eq: asympt len tiles} as
+\begin{equation}
+\label{eq: h bound}
+\frac 1 {\sqrt{2}} \leq h_\tiles \leq \frac 1 {\sqrt{2}} + \frac{1}{2\sqrt{2}(\tiles-2)}.
+\end{equation}
+
+\begin{figure}%
+\centering
+\includegraphics[width=.72\textwidth]{probmax_asint_turned}%
+\caption{Reachable tiles that maximize the number of visited tiles on a unit square grid; rotated view. Example for $\tiles=9$}
+\label{fig: probmax_asint_turned}%
+\end{figure}%
+
+As argued in the proof of Theorem~\ref{theo: probmax, sq}, $\probmax(\len_\tiles)$ is the sum of several terms $\probmax_{i,j}(\len_\tiles)$, $i+j-1=\tiles$.
+Each term is associated with one of the tiles shown in Figure~\ref{fig: probmax_asint_turned} (or in Figure~\ref{fig: probmax_nterms}), and includes a factor of $4$ to account for the symmetrical tiles in the other quadrants.
+For this proof it is more convenient to compute $\tiles \probmax_{i,j}(\len_\tiles)$, and to change the indexing as follows:
+\begin{equation}
+P_{\tiles,n} = \tiles \probmax_{i,j}(\len_\tiles),
+\end{equation}
+where $n$ equals $1$ for the smallest $i$ in \eqref{eq: probmax impl}, increases with $i$, and has a maximum value $N$ equal to the number of summands in that expression. Thus
+\begin{equation}
+\label{eq: sum p t n}
+\tiles \probmax(\len_\tiles) = \sum_{n=1}^N P_{\tiles,n}.
+\end{equation}
+
+Consider the tile depicted with thick line in Figure~\ref{fig: probmax_asint_turned}, for a generic value of the index $n$ in Figure~\ref{fig: probmax_asint_turned}, or of the indices $i,j$ in Figure~\ref{fig: probmax_nterms}.
+For this tile, $P_{\tiles,n}/t = \probmax_{i,j}(\len_\tiles)$ is obtained as the area of the shaded rectangle averaged over $\theta$ uniformly distributed on $[0,2\pi)$, and then multiplied by $4$, in the same way as in the proof of Theorem~\ref{theo: probmax, sq}.
+Equivalently, the average can be restricted to the range of values of $\theta$ for which the shaded area is non-zero, corresponding to the arc with thick line in the figure, of length $c_{\tiles,n}$, and the result multiplied by $c_{\tiles,n}/(2\pi\len_\tiles)$.
+Denoting the average over this restricted range by $A_{\tiles,n}$,
+\begin{equation}
+\label{eq: P i j asympt 1}
+P_{\tiles,n} = \frac{2 \tiles A_{\tiles,n} c_{\tiles,n}}{\pi\len_\tiles}.
+\end{equation}
+Taking into account that $A_{\tiles,n}$ and $c_{\tiles,n}$ are upper-bounded by the area of a tile and one quarter of the length of a circle of radius $1$, respectively,
+\begin{align}
+\label{eq: A tiles n bound}
+0 &< A_{\tiles,n} < 1, \\
+\label{eq: c tiles n bound}
+0 &< c_{\tiles,n} \leq \frac \pi 2,
+\end{align}
+and using \eqref{eq: asympt len tiles},
+\begin{align}
+\label{eq: P i j asympt 2 1}
+P_{\tiles,n}
+&\leq \left(1 + \frac 2 {\tiles-2} \right) \frac{2 \sqrt{2} A_{\tiles,n} c_{\tiles,n}}{\pi} = \frac{2\sqrt{2} A_{\tiles,n} c_{\tiles,n}}{\pi} + \frac{2\sqrt{2}}{\tiles-2}, \\
+\label{eq: P i j asympt 2 2}
+P_{\tiles,n}
+&\geq \left(1 + \frac {3/2} {\tiles-3/2} \right) \frac{2 \sqrt{2} A_{\tiles,n} c_{\tiles,n}}{\pi} = \frac{2 \sqrt{2} A_{\tiles,n} c_{\tiles,n}}{\pi} + \frac{3/\sqrt{2}}{\tiles-3/2}.
+\end{align}
+
+To calculate $A_{\tiles,n}$ and $c_{\tiles,n}$ it is easier to replace the arc by its chord, as depicted in Figure~\ref{fig: probmax_asint_detail}.
+The approximation error for either $A_{\tiles,n}$ or $c_{\tiles,n}$ tends to $0$ as $\tiles \rightarrow \infty$.
+Specifically, regarding $A_{\tiles,n}$, the difference in $y$ between the arc and the chord is upper-bounded by the length of the chord multiplied by the maximum of the absolute value of the arc slope.
+The former is at most $\sqrt{2}$, and the latter is bounded using \eqref{eq: funcirc}, \eqref{eq: w} and \eqref{eq: asympt len tiles} as
+\begin{equation}
+\label{eq: asympt slope bound}
+\left| \frac{\diff \funcirc(x)}{\diff x} \right| = \frac{x}{\sqrt{\len_\tiles^2-x^2}} \leq \frac{w_\tiles}{\len_\tiles} \leq \sqrt{\frac{2}{\tiles-2}}.
+\end{equation}
+Thus the difference in $y$ is at most $2/\sqrt{\tiles-2}$.
+This translates into a difference of at most  $\sqrt{2/(\tiles-2)}$ in each dimension of the shaded rectangle; or, since the maximum dimensions are $1$, with at least one of them being strictly less, a difference in area smaller than twice that value.
+Therefore the approximation error in $A_{\tiles,n}$ is less than $2\sqrt{2}/\sqrt{\tiles-2}$.
+
+As for approximating $c_{\tiles,n}$ by the chord length, the latter is $2\len_\tiles \sin(c_{\tiles,n}/(2\len_\tiles))$. Taking into account that
+\begin{equation}
+\label{eq: sin asympt pos}
+\sin\genvar > \genvar - \frac{\genvar^3} 6 \quad \text{for } s > 0
+\end{equation}
+and making use of \eqref{eq: c tiles n bound} and \eqref{eq: asympt len tiles}, this is bounded as
+\begin{equation}
+2\len_\tiles \sin \frac{c_{\tiles,n}}{2\len_\tiles} > c_{\tiles,n} -\frac{c_{\tiles,n}^2}{24\len_\tiles^2} \geq c_{\tiles,n} -\frac{\pi^2}{96\len_\tiles^2} \geq c_{\tiles,n} -\frac {\pi^2} {48(\tiles-2)^2},
+\end{equation}
+and thus the length difference is at most $\pi^2/(48(\tiles-2)^2)$.
+
+\begin{figure}
+\centering
+\includegraphics[width=.7\textwidth]{probmax_asint_detail}%
+\caption{Computation of $P_{\tiles,n}$}%
+\label{fig: probmax_asint_detail}%
+\end{figure}%
+
+Let $\tilde A_{\tiles,n}$ and $\tilde c_{\tiles,n}$ denote the approximations of $A_{\tiles,n}$ and $c_{\tiles,n}$ obtained using the chord instead of the arc. From the preceding two paragraphs,
+\begin{equation}
+\label{eq: P i j asympt 3}
+\tilde A_{\tiles,n} \tilde c_{\tiles,n} < A_{\tiles,n} c_{\tiles,n} <
+\tilde A_{\tiles,n} \tilde c_{\tiles,n} + \frac{\pi^2 A_{\tiles,n}}{48(\tiles-2^2)} + \frac{2\sqrt{2} c_{\tiles,n}}{\sqrt{\tiles-2}}.
+\end{equation}
+Using \eqref{eq: A tiles n bound} and \eqref{eq: c tiles n bound} into \eqref{eq: P i j asympt 3} and combining with \eqref{eq: P i j asympt 2 1} and \eqref{eq: P i j asympt 2 2},
+\begin{equation}
+\label{eq: P i j asympt 3 bis}
+P_{\tiles,n} = \frac{2\sqrt{2} \tilde A_{\tiles,n} \tilde c_{\tiles,n}}{\pi} + E_{\tiles, n}
+\end{equation}
+with $|E_{\tiles, n}| < e(\tiles)$, where $e$ is a function that depends only on $\tiles$ and tends to $0$ as $\tiles \rightarrow 0$.
+
+Denoting the slope of the chord by $m_{\tiles,n}$, it is clear from Figure~\ref{fig: probmax_asint_detail} that
+\begin{equation}
+\frac{y_{\tiles,n}-a_{\tiles,n}/\sqrt{2}}{a_{\tiles,n}/\sqrt{2}} = m_{\tiles,n},
+\end{equation}
+and therefore $a_{\tiles,n} = \sqrt{2}y_{\tiles,n}/(1+m_{\tiles,n})$.
+Similarly, $b_{\tiles,n} = \sqrt{2}y_{\tiles,n}/(1-m_{\tiles,n})$.
+Both $\tilde A_{\tiles,n}$ and $\tilde c_{\tiles,n}$ can be expressed in terms of $a_{\tiles,n}$ and $b_{\tiles,n}$, and thus of $y_{\tiles,n}$ and $m_{\tiles,n}$:
+\begin{align}
+\tilde A_{\tiles,n} &= a_{\tiles,n}b_{\tiles,n} \int_0^1 \eta(1-\eta)\, \diff \eta = \frac{a_{\tiles,n}b_{\tiles,n}} 6 = \frac{y_{\tiles,n}^2}{3(1-m_{\tiles,n}^2)}, \\
+\tilde c_{\tiles,n} &= \sqrt{a_{\tiles,n}^2+b_{\tiles,n}^2} = \frac{2 y_{\tiles,n} \sqrt{1+m_{\tiles,n}^2}}{1-m_{\tiles,n}^2}.
+\end{align}
+Substituting into \eqref{eq: P i j asympt 3 bis},
+\begin{equation}
+\label{eq: P i j asympt 4}
+P_{\tiles,n}
+= \frac{4\sqrt{2} \sqrt{1+m_{\tiles,n}^2}} {3\pi(1-m_{\tiles,n}^2)^2} y_{\tiles,n}^3 + E_{\tiles, n}.
+\end{equation}
+The slope $m_{\tiles,n}$ is bounded by \eqref{eq: asympt slope bound}, which implies that $m_{\tiles,n}^2 \leq 2/(\tiles-2)$.
+Using this inequality together with \eqref{eq: sqrt asympt}, the term depending on $m_{\tiles,n}^2$ in \eqref{eq: P i j asympt 4} can be bounded as
+\begin{equation}
+\label{eq: P i j asympt 4 bis}
+\frac{\sqrt{1+m_{\tiles, n}^2}} {(1-m_{\tiles, n}^2)^2} \leq \frac{(\tiles-2)^2}{(\tiles-4)^2} \left(1+\frac 1 {\tiles-2}\right) = 1 + \frac{5\tiles - 14}{(\tiles-4)^2}.
+\end{equation}
+On the other hand, $y_{\tiles,n}$ can be bounded from \eqref{eq: h bound} as $y_{\tiles,n} \leq h_\tiles \leq 1$.
+Using this inequality as well as \eqref{eq: P i j asympt 4 bis} into \eqref{eq: P i j asympt 4},
+\begin{equation}
+\label{eq: P i j asympt 5}
+P_{\tiles,n} = \frac{4\sqrt{2}} {3\pi} y_{\tiles,n}^3 + E_{\tiles,n} + E'_{\tiles,n}
+\end{equation}
+with $|E'_{\tiles,n}| < e'(t)$, where $e'(t)$ tends to $0$ as $t \rightarrow \infty$.
+
+Let $\funcircesc_\tiles$ be defined as a scaled version of $\funcirc$, where the variable $x \in [-w_\tiles, w_\tiles]$ is replaced by $\zeta = x/w_\tiles \in [-1,1]$:
+\begin{equation}
+\label{eq: funcircest tiles 1}
+\funcircesc_\tiles(\zeta) = \funcirc\left(w_\tiles \zeta\right) = h_\tiles - \len_\tiles + \sqrt{\len_\tiles^2-w^2_\tiles\zeta^2}.
+% = h_\tiles - \len_\tiles + \len_\tiles \sqrt{1-w^2_\tiles\zeta^2/\len_\tiles^2}.
+\end{equation}
+As $\tiles \rightarrow \infty$, the sequence of functions $\funcircesc_\tiles$ converges pointwise to
+\begin{equation}
+\label{eq: funcircesc}
+\funcircesc(\zeta) = \lim_{\tiles \rightarrow \infty} \funcircesc_\tiles(\zeta) = \frac{1 -\zeta^2}{\sqrt{2}}.
+% \frac{\sqrt{2}}{2}(1 -\zeta^2)
+\end{equation}
+Furthermore, the convergence is uniform.
+To show this, it is easier to consider the subsequences of $\tiles$ odd and even separately.
+For $\tiles$ odd, substituting $\len_\tiles$, $w_\tiles$ and $h_\tiles$ given by \eqref{eq: len tiles}, \eqref{eq: w} and \eqref{eq: h tiles} into \eqref{eq: funcircest tiles 1},
+\begin{equation}
+\label{eq: funcircesc, odd 1}
+\funcircesc_\tiles(\zeta) =
+% -\frac{\tiles-3}{\sqrt{2}} + \sqrt{\frac{(\tiles-2)^2+1}{2} - (\tiles-2)\zeta^2}
+-\frac{\tiles-3}{\sqrt{2}} + \sqrt{\frac{(\tiles-2)^2 - 2(\tiles-2)\zeta^2 +1 }{2}}.
+\end{equation}
+Computing $\partial \funcircesc_\tiles(\zeta)/\partial\tiles$ as if $\tiles$ were a continuous variable, and taking into account that $\zeta \in [-1,1]$,
+\begin{equation}
+\label{eq: der funcircesc, odd 1}
+\frac{\partial \funcircesc_\tiles(\zeta)}{\partial \tiles} = \frac 1 {\sqrt{2}} \left(
+-1 + \frac{\tiles-2-\zeta^2}{\sqrt{(\tiles-2)^2 - 2(\tiles-2)\zeta^2 + 1}} \right) \leq 0.
+\end{equation}
+This implies that the subsequence $\funcircesc_\tiles$ with $\tiles$ odd is monotone non-increasing.
+Since the limit function $\funcircesc$ is continuous, by Dini's theorem \cite[section~1.1]{Hirsch99} the convergence of this subsequence is uniform.
+On the other hand, for $\tiles$ even
+\begin{equation}
+\label{eq: funcircesc, even 1}
+\funcircesc_\tiles(\zeta) =
+% -\frac{\tiles-3}{\sqrt{2}} + \sqrt{\frac{(\tiles-2)^2}{2} - \left(\tiles-\frac 5 2\right)\zeta^2}
+-\frac{\tiles-3}{\sqrt{2}} + \sqrt{\frac{(\tiles-2)^2 - 2(\tiles-2)\zeta^2 + \zeta^2}{2}},
+\end{equation}
+and, for $\zeta \in [-1,1]$,
+\begin{equation}
+\label{eq: der funcircesc, even 1}
+\frac{\partial \funcircesc_\tiles(\zeta)}{\partial \tiles} = \frac 1 {\sqrt{2}} \left(
+-1 + \frac{\tiles-2-\zeta^2}{\sqrt{(\tiles-2)^2 - 2(\tiles-2)\zeta^2 + \zeta^2}} \right) \leq 0.
+\end{equation}
+Thus the subsequence $\funcircesc_\tiles$ with $\tiles$ even is also monotone non-increasing, and therefore converges uniformly.
+Consequently, the full sequence $\funcircesc_\tiles$ converges to $\funcircesc$ uniformly on $[-1,1]$.
+This means that
+\begin{equation}
+\label{eq: funcircest tiles 4}
+y_{\tiles,n} = \funcirc(x_{\tiles,n}) = \funcircesc_\tiles\left( \frac{x_{\tiles,n}} {w_\tiles} \right) = \funcircesc\left( \frac{x_{\tiles,n}} {w_\tiles} \right) + E''_{\tiles,n}
+\end{equation}
+with $|E''_{\tiles,n}| < e''(t)$, where $e''(\tiles)$ tends to $0$ as $\tiles \rightarrow \infty$.
+
+Combining \eqref{eq: sum p t n}, \eqref{eq: P i j asympt 5} and \eqref{eq: funcircest tiles 4},
+\begin{equation}
+\label{eq: sum p t n aprox 1}
+\tiles \probmax(\len_\tiles) = \sum_{n=1}^N P_{\tiles,n} = \frac{4 \sqrt{2}} {3\pi} \sum_{n=1}^N \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right) + \sum_{n=1}^N \left(E_{\tiles,n} + E'_{\tiles,n} + E''_{\tiles,n} \right),
+\end{equation}
+which can be rewritten as
+\begin{equation}
+\label{eq: sum p t n aprox 2}
+\sqrt{\tiles} \probmax(\len_\tiles) = \frac{4 w} {3\pi\sqrt{\tiles}} \frac {\sqrt{2}} {w} \sum_{n=1}^N \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right) +
+\frac{N} {\sqrt{t}} \frac 1 N \sum_{n=1}^N (E_{\tiles,n} + E'_{\tiles,n} + E''_{\tiles,n}).
+\end{equation}
+Noting that the values $\zeta_{\tiles,n}$ are equispaced from $-1$ to $1$ with step $\sqrt{2}/w$, the term $\sqrt{2}/w \sum_{n=1}^N \funcircesc^3 ( x_{\tiles,n}/w_\tiles)$ in \eqref{eq: sum p t n aprox 2} can be interpreted as a Riemmann sum that approximates the integral of $\funcircesc^3(\zeta)$ over $[-1,1]$.
+Since this function is continuous the sum indeed converges to the integral \cite[section~7.2]{Abbott15}.
+On the other hand,
+\begin{equation}
+\left| \frac 1 N \sum_{n=1}^N \left( E_{\tiles,n} + E'_{\tiles,n} + E''_{\tiles,n} \right) \right| < e(\tiles)+e'(\tiles)+e''(\tiles)
+\end{equation}
+and the right-hand side tends to $0$ as $\tiles \rightarrow \infty$.
+The terms $w/\sqrt{\tiles}$ and $N/\sqrt{\tiles}$ in \eqref{eq: sum p t n aprox 2} tend to $1$ and $\sqrt{2}$ respectively as $\tiles \rightarrow \infty$.
+Therefore, taking limits and substituting \eqref{eq: funcircesc},
+\begin{equation}
+\label{eq: sum p t n lim}
+\lim_{\tiles \rightarrow \infty} \sqrt{\tiles} \probmax(\len_\tiles) =
+\frac{4} {3\pi} \int_{-1}^1 \funcircesc^3(\zeta) \,\diff \zeta = \frac{2\sqrt{2}} {3\pi} \int_{0}^1 (1-\zeta^2)^3 \,\diff \zeta = \frac{32\sqrt{2}} {105\pi},
+\end{equation}
+which establishes \eqref{eq: probmax lim sqrt tiles}.
+
+Lastly, \eqref{eq: probmax limsup sqrt len} is obtained from \eqref{eq: probmax lim sqrt tiles} by observing that $\limsup_{\len \rightarrow \infty} \sqrt{\len} \probmax(\len) = \lim_{\tiles \rightarrow \infty} \sqrt{\len_\tiles} \probmax(\len_\tiles)$ and that $ \lim_{\tiles \rightarrow \infty} \len_\tiles/\tiles = 1 / \sqrt{2}$.
 \end{proof}

### 2023-04-10 figures/figures.m

@@ -1627,6 +1627,7 @@ for k = [0 1+3j 2+2j 3+1j 1-3j 2-2j 3-1j -3+1j -2+2j -1+3j -3-1j -2-2j -1-3j]
         plot(real(k)+[0 1 1 0 0], imag(k)+[0 0 1 1 0], 'k--', 'linewidth', 1.4)
     end
 end
+set(gcf, 'Position', [680 558 560 400])
 
 
 %% #63. Figure for proof of probmax: funt odd, diagonal
@@ -2398,7 +2399,7 @@ plot(num_k2, '--')
 all(num_k==num_k2) % checked
 
 
-%% #84. Computing probability non-asymptotically using the number of terms from #83
+%% #85. Computing probability non-asymptotically using the number of terms from #84
 
 clear, close all
 L_fine = linspace(10,1000,1e5);
@@ -2533,7 +2534,7 @@ plot(L_fine.*ind_even./ind_even, result_computed4.*ind_even./ind_even, 'k--', 'l
 g = @(r,u,v) (2*(acos(u./r)-asin(v./r)).*u.*v + r.^2 + u.^2 + v.^2 - 2*u.*sqrt(r.^2-v.^2) - 2*v.*sqrt(r.^2-u.^2))/pi;
 ind_odd = mod(funt_fine,2)==1;
 ind_even = ~ind_odd;
-result_computed3 = NaN(size(funt_fine));
+result_computed5 = NaN(size(funt_fine));
 for ind_fine = 1:numel(funt_fine)
     t = funt_fine(ind_fine);
     L = L_fine(ind_fine);
@@ -2548,8 +2549,521 @@ for ind_fine = 1:numel(funt_fine)
         jj = actual_j(ind_ij);
         res = res + g(L, ii-1, jj-1);
     end
-    result_computed3(ind_fine) = res;
+    result_computed5(ind_fine) = res;
 end
-plot(L_fine.*ind_odd./ind_odd, result_computed4.*ind_odd./ind_odd, 'k--', 'linewidth', .7), hold on
-plot(L_fine.*ind_even./ind_even, result_computed4.*ind_even./ind_even, 'k--', 'linewidth', .7)
+plot(L_fine.*ind_odd./ind_odd, result_computed5.*ind_odd./ind_odd, 'k--', 'linewidth', .7), hold on
[…]

### 2023-04-17 paper.tex
Proofs (hopefully) finished

@@
 \newcommand{\mss}{M}
-\newcommand{\nextrat}{K}
+%\newcommand{\nextrat}{K}
 \newcommand{\funcirc}{\Omega}
 \newcommand{\funcircesc}{\bar \Omega}
+\newcommand{\fdp}{p}
 
@@
 \displaystyle
-0 & \text{for } \tiles =1, 2 \\[1.4mm]
+0 & \text{for } \tiles =1, 2 \\[1.4mm] % *!*
 \displaystyle
-\frac{\tiles-3}{\sqrt{2}} & \text{for } \tiles \text{ odd, } \tiles \geq 3 \\[4.5mm]
+\frac{\tiles-3}{\sqrt{2}} & \text{for } \tiles \text{ odd, } \tiles \geq 3 \\[4.5mm] % *!*
 \displaystyle
@@
 \displaystyle
-0 & \text{for } \tiles =1, 2 \\[1.4mm]
+0 & \text{for } \tiles =1, 2 \\[1.4mm]  % *!*
 \displaystyle
@@
 \displaystyle
-1 & \text{for } \tiles \leq 3 \\[1.3mm]
+1 & \text{for } \tiles \leq 3 \\[1.3mm] % *!*
 \displaystyle
@@
 \displaystyle
-1 &\text{if\ \ } \displaystyle n =1 \\[1 mm]
+1 &\text{if\ \ } \displaystyle n =1 \\[1 mm] % *!*
 \displaystyle
-\frac{2\len}{\pi a} \left(f\left(\frac{a(n-1)}{\len}\right)-f\left(\frac{a(n-2)}{\len} \right)\right) &\text{if\ \ } \displaystyle 2 \leq n < \frac \len a + 1 \\[4 mm]
+\frac{2\len}{\pi a} \left(f\left(\frac{a(n-1)}{\len}\right)-f\left(\frac{a(n-2)}{\len} \right)\right) &\text{if\ \ } \displaystyle 2 \leq n < \frac \len a + 1 \\[4 mm] % *!*
 \displaystyle
-\frac{2\len}{\pi a}\,\, \left(1 - f\left(\frac{a(n-2)}{\len}\right)\right) &\text{if\ \ } \displaystyle \frac\len a + 1 \leq n < \frac\len a+2 \\[3 mm]
+\frac{2\len}{\pi a}\,\, \left(1 - f\left(\frac{a(n-2)}{\len}\right)\right) &\text{if\ \ } \displaystyle \frac\len a + 1 \leq n < \frac\len a+2 \\[3 mm] % *!*
 \displaystyle
-0 &\text{if\ \ } \displaystyle \frac \len a + 2 \leq  n; \\
+0 &\text{if\ \ } \displaystyle \frac \len a + 2 \leq  n;
 \end{cases}
@@
 
-The probability of visiting the maximum number of tiles for a rectangular grid is difficult to compute, due to the irregularity of the optimal $(i,j)$ pairs (which was discussed in \S\ref{part: fund results}; in particular see Figure~\ref{fig: pairs_bounds}).
-In a square grid, however, the simplicity of the set of optimal pairs makes the problem more tractable.
+The probability of visiting the maximum number of tiles for a rectangular grid is difficult to compute.
+The reason is analogous to the irregularity of the optimal $(i,j)$ pairs which was discussed in \S\ref{part: fund results}.
+In a square grid, however, the problem is more tractable.
 
@@
 Given $\len \in \mathbb R^+$, consider a unit square grid and a uniformly random segment of length $\len$, as previously defined.
-The probability $\probmax(\len)$ that the segment visits the maximum number of tiles is
+The probability $\probmax(\len)$ that the segment visits the maximum number of tiles $\tiles= \funt(\len)$ is
 \begin{equation}
@@
 \end{equation}
-or equivalently, defining $\nextrat = \sqrt{\len^2/2-(\tiles-3)^2/4}$,
+or equivalently, defining $w = \sqrt{\len^2-(\tiles-3)^2/2}$,
 \begin{equation}
@@
 g\left(\len, \frac{\tiles-3} 2, \frac{\tiles-3} 2 \right) +
-2 \sum_{k=1}^{\lceil\nextrat-1 \rceil} g\left(\len, \frac{\tiles-3} 2 - k, \frac{\tiles-3} 2 + k \right) &\text{for $\tiles$ odd} \\[6mm] % *!*
+2 \sum_{k=1}^{\left\lceil \frac w {\sqrt{2}} - 1 \right\rceil} g\left(\len, \frac{\tiles-3} 2 - k, \frac{\tiles-3} 2 + k \right) &\text{for $\tiles$ odd} \\[6mm] % *!*
 \displaystyle
 2 g\left(\len, \frac{\tiles-4} 2, \frac{\tiles-2} 2 \right) +
-2 \sum_{k=1}^{\lceil\nextrat-3/2 \rceil} g\left(\len, \frac{\tiles-4} 2 - k, \frac{\tiles-2} 2 + k \right) &\text{for $\tiles$ even}.
+2 \sum_{k=1}^{\left\lceil \frac w {\sqrt{2}} - \frac 3 2 \right\rceil} g\left(\len, \frac{\tiles-4} 2 - k, \frac{\tiles-2} 2 + k \right) &\text{for $\tiles$ even}.
 \end{cases}
@@
 Consider $\len \in \mathbb R^+$, and let $\tiles = \funt(\len)$.
-By Proposition~\ref{prop: i+j-1}, the segment visits $\tiles$ tiles almost surely if the dimensions $i$, $j$ of its discrete bounding rectangle satisfy $i+j-1=\tiles$.
-In addition, a segment that visits the maximum number of tiles always has $i, j \geq 2$, because if $i=1$ the segment can always be shifted and possibly tilted to cross a vertical grid line, thus increasing $i$ to $2$, without reducing $j$.
+By Proposition~\ref{prop: i+j-1}, the segment visits $\tiles$ tiles almost surely if the dimensions $i$, $j$ of its discrete bounding rectangle are such that $i+j-1=\tiles$.
+In addition, a segment that visits the maximum number of tiles always has $i, j \geq 2$.
+Namely, if $i=1$ the segment can always be shifted and possibly tilted to cross a vertical grid line, thus increasing $i$ to $2$, without reducing $j$.
 The reasoning for $j=1$ is analogous.
@@
 An example is shown in Figure~\ref{fig: probmax_symmetric_tiles}.
-Consequently, $\probmax_{i,j}(\len)$ can be obtained by computing the probability that the second endpoint is in the tile with lower-left corner $(i-1,j-1)$ and then multiplying by $4$.
+Consequently, $\probmax_{i,j}(\len)$ can be obtained by computing the probability that the second endpoint is in the tile with lower-left corner $(i-1,j-1)$ and multiplying by $4$.
 
@@
 \centering%
-\includegraphics[width=.5\textwidth]{probmax_symmetric_tiles_}%
-\caption{Symmetric tiles for the computation of $\probmax_{i,j}(\len)$ on a unit square grid. Example for $i=3$, $j=4$%
+\includegraphics[width=.53\textwidth]{probmax_symmetric_tiles}%
+\caption{Symmetrical tiles for the computation of $\probmax_{i,j}(\len)$ on a unit square grid. Example for $i=3$, $j=4$%
 }%
@@
 
-For a fixed orientation $\theta$, the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ if and only if shifting the segment up and to the right so that its first endpoint is at $(1,1)$ results in the second endpoint still being in the tile with lower-left corner $(i-1,j-1)$, as shown in Figure~\ref{fig: probmax_funt_odd}.
+For a fixed orientation $\theta$, the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ if and only if shifting the segment up and to the right so that its first endpoint is at $(1,1)$ results in the second endpoint still being in the tile with lower-left corner $(i-1,j-1)$.
+This is illustrated in Figure~\ref{fig: probmax_funt_odd}.
 It cannot be the case that the segment ``overshoots'' past this tile, because that would yield a number of visited tiles exceeding $\tiles$, which is the maximum.
@@
 \includegraphics[width=.56\textwidth]{probmax_funt_odd}%
-\caption{Reachability from the reference tile for a given $\tiles = \funt(\len)$ on a unit square grid. Example for $\tiles = 7$
+\caption{Segment orientations and positions for a given $\tiles = \funt(\len)$ on a unit square grid. Example for $\tiles = 7$
 }%
@@
 There is a range of values of $\theta$ for which the above condition is satisfied.
-This range corresponds to the part of the arc with solid line in Figure~\ref{fig: probmax_funt_odd}.
-For each $\theta$ in the allowed range, the valid positions for the second endpoint of the segment are in a rectangle contained in the tile with lower-left corner $(i-1,j-1)$.
+This corresponds to the part of the arc with solid line in Figure~\ref{fig: probmax_funt_odd}.
+For each $\theta$ in this range, the valid positions for the second endpoint of the segment are in a rectangle contained in the tile with lower-left corner $(i-1,j-1)$.
 These are the positions that cause the first endpoint to be in the reference tile.
@@
 The above arguments show that $\probmax_{i,j}(\len)$ is given by \eqref{eq: probmax: cond on theta} if the condition \eqref{eq: i-2 j-2 len} holds, and equals $0$ otherwise.
-Using this into \eqref{eq: probmax len probmax i,j len} yields \eqref{eq: probmax impl} .
+Using this into \eqref{eq: probmax len probmax i,j len} yields \eqref{eq: probmax impl}.
 
 Given $\tiles$, the range of lengths $\len$ such that $\tiles = \funt(\len)$ is either \eqref{eq: square, odd max tiles, lengths} or \eqref{eq: square, even max tiles, lengths}.
-For $\len$ in this range, the number of tiles with lower-left corners $(i-1,j-1)$, $i,j \geq 2$, $i+j-1=\tiles$ that can be reached, i.e.~the number of terms in \eqref{eq: probmax impl}, depends on $\len$.
-If $\tiles$ is odd the minimum number is $1$, corresponding to a tile with lower-left corner on the line $x=y$, which gives the first term in the first expression in \eqref{eq: probmax expl}.
-For $\tiles$ even the minimum number is $2$, corresponding to two symmetrical tiles whose $x$ and $y$ coordinates differ by $1$, which give the first term in the second expression.
+For $\len$ in this range, the number of reachable tiles with lower-left corners $(i-1,j-1)$, $i,j \geq 2$, $i+j-1=\tiles$, i.e.~the number of terms in \eqref{eq: probmax impl}, depends on $\len$.
+If $\tiles$ is odd the minimum number is $1$, corresponding to a tile with its lower-left corner on the line $x=y$, which gives the first term in the first expression in \eqref{eq: probmax expl}.
+For $\tiles$ even the minimum number is $2$, corresponding to two symmetrical tiles whose lower-left $x$ and $y$ coordinates differ by $1$, which give the first term in the second expression.
 Either for $\tiles$ odd or even, additional tiles may become reachable as $\len$ grows (within the range allowed by $\tiles$), and these always occur in symmetrical pairs.
@@
 As can be seen in Figure~\ref{fig: probmax_terms}, for each $\tiles$ there is a maximum number of tiles that can be reached, beyond which incrementing $\len$ only results in $\tiles$ increasing by $1$.
-For a given $\len$, the number of reachable tiles can be determined by counting how many markers are covered by the length $w$ defined in Figure~\ref{fig: probmax_nterms}.
-This length satisfies the relationship $w^2 = \len^2 - (\tiles-3)^2/2$, and thus $K$ as defined in the theorem equals $w/\sqrt{2}$.
-From the figure, the number of terms in the sum over $k$ (empty circular markers) is $\lceil w/\sqrt 2 - 1 \rceil$ or $\lceil w/\sqrt 2 - 3/2 \rceil$ for $\tiles$ odd or even, i.e.~$\lceil \nextrat - 1 \rceil$ or $\lceil \nextrat - 3/2 \rceil$ respectively.
+For a given $\len$, the number of reachable tiles can be determined by counting how many markers are covered by the stretch $w$ defined in Figure~\ref{fig: probmax_nterms}.
+This satisfies the relationship $w^2 = \len^2 - (\tiles-3)^2/2$.
+From the figure, the number of terms in the sum over $k$ (empty circular markers) is $\lceil w/\sqrt 2 - 1 \rceil$ for $\tiles$ odd and $\lceil w/\sqrt 2 - 3/2 \rceil$ for $\tiles$ even.
 This establishes \eqref{eq: probmax expl}.
+
 \begin{figure}
@@
 \end{figure}%
-\end{proof}
-
-The technique used in the proof Theorem~\ref{theo: probmax, sq} could in principle be used for obtaining the probability that the number of tiles visited by the segment equals or exceeds any given value $\tiles < \funt(\len)$.
+
+It should be noted that, in principle, the technique used in this proof could be employed for obtaining the probability that the number of tiles visited by the segment equals or exceeds any given value $\tiles < \funt(\len)$.
 However, the process is more cumbersome.
 Specifically, with reference to Figure~\ref{fig: probmax_funt_odd}, the arc in this more general setting can enclose any number of corners of the considered tile, not necessarily $1$, which makes the characterization of the shaded rectangles more complicated.
-% because several cases need to be distinguished
-
-\begin{figure}%
-\centering%
-\includegraphics[width=.75\textwidth]{probmax_len}%
-\caption{Probability $\probmax(\len)$ that a random segment of length $\len$ visits the maximum number of tiles on a unit square grid%
-}%
-\label{fig: probmax_len}%
-\end{figure}%
-
-The probability $\probmax(\len)$ of visiting the maximum number of tiles on a unit square grid, computed from Theorem~\ref{theo: probmax, sq}, is shown in Figure~\ref{fig: probmax_len} shows.
-As $\len$ grows, $\probmax(\len)$ has a jump when $\funt(\len)$ increases by $1$ (and there $\probmax(\len)$ takes the value approached from the left).
+\end{proof}
+
+The probability $\probmax(\len)$ of visiting the maximum number of tiles on a unit square grid, computed from Theorem~\ref{theo: probmax, sq}, is shown in Figure~\ref{fig: probmax_len}.
+As $\len$ grows, $\probmax(\len)$ has a jump when $\funt(\len)$ increases by $1$.
 This happens when $\len$ equals the right endpoint of the interval \eqref{eq: square, odd max tiles, lengths} for $\tiles$ odd or of the interval \eqref{eq: square, even max tiles, lengths} for $\tiles$ even.
@@
 \displaystyle
-\frac{\sqrt{(\tiles-3)^2 + (\tiles-1)^2}} {2} = \sqrt{\frac{(\tiles-2)^2+1}{2}} & \text{ for $\tiles$ odd, $\tiles \geq 3$} \\
+\frac{\sqrt{(\tiles-3)^2 + (\tiles-1)^2}} {2} = \sqrt{\frac{(\tiles-2)^2+1}{2}} & \text{ for $\tiles$ odd, $\tiles \geq 3$} \\[2 mm] % *!*
 \displaystyle
-\frac{\tiles-2}{\sqrt{2}} & \text{ for $\tiles$ even, $\tiles \geq 4$},
+\frac{\tiles-2}{\sqrt{2}} & \text{ for $\tiles$ even, $\tiles \geq 4$}.
 \end{cases}
 \end{equation}
-and satisfy $\funt(\len_\tiles) = \tiles$.
-
-The first continuous section plotted in Figure~\ref{fig: probmax_len} corresponds to a maximum number of visited tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, where $\len_3= 1$.
+Clearly, $\funt(\len_\tiles) = \tiles$ and $\lim_{\len \rightarrow \len_\tiles^- } \probmax(\len) = \probmax(\len_\tiles)$.
+
+\begin{figure}%
+\centering%
+\includegraphics[width=.74\textwidth]{probmax_len}%
+\caption{Probability $\probmax(\len)$ that a random segment of length $\len$ visits the maximum number of tiles on a unit square grid%
+}%
+\label{fig: probmax_len}%
+\end{figure}%
+
+The first continuous section seen in Figure~\ref{fig: probmax_len} corresponds to a maximum number of visited tiles $\tiles=3$, for lengths in the interval $(0, \len_3]$, where $\len_3= 1$.
 The second corresponds to $\tiles=4$, for lengths in $(\len_3, \len_4]$, where $\len_4= \sqrt{2}$.
@@
 
+As argued in the proof of Theorem~\ref{theo: probmax, sq}, $\probmax(\len_\tiles)$ is the sum of several terms $\probmax_{i,j}(\len_\tiles) = g(\len_\tiles, i-2, j-2)$, where the sum runs over $i,j$ as given by \eqref{eq: probmax impl}.
+Each term is associated with one of the tiles shown in Figure~\ref{fig: probmax_nterms}, and includes a factor of $4$ to account for the symmetrical tiles in the other quadrants.
 For the purpose of establishing \eqref{eq: probmax lim sqrt tiles}, consider the graph in Figure~\ref{fig: probmax_nterms} with $\len= \len_\tiles$.
-It is convenient to rotate this graph by $45^\circ$ and shift it so that the original $x+y = \tiles-1$ line (shown dashed in Figure~\ref{fig: probmax_nterms}) becomes the new horizontal axis.
+It is convenient to rotate this graph by $45^\circ$ and shift it so that the original $x+y+1 = \tiles$ line (shown dashed in Figure~\ref{fig: probmax_nterms}) becomes coincident with the $x$ axis, and the arc center is at $(0,-(\tiles-3)/\sqrt{2})$.
 Figure~\ref{fig: probmax_asint_turned} shows the result, using $\tiles=9$ as an example.
-The circular markers in this figure correspond, in the original graph before rotating, to the lower-left corners $(i-1,j-1)$ of the tiles with $i+j-1=\tiles$, $(i-2)^2+(j-1)^2<\len_\tiles$.
-Each of these tiles can be (partially) seen above the corresponding marker in Figure~\ref{fig: probmax_asint_turned}, its sides forming $45^\circ$ degrees with respect to the horizontal axis.
-Let $x_{\tiles, n}$ denote the coordinates of these points along the horizontal axis of Figure~\ref{fig: probmax_asint_turned}.
-
-Using the coordinate axes in Figure~\ref{fig: probmax_asint_turned}, the circular arc satisfies the equation $y = \funcirc(x)$ with
+The circular markers in this figure correspond, in the graph before rotating, to the lower-left corners $(i-1,j-1)$ of the tiles with $i+j-1=\tiles$, $(i-2)^2+(j-1)^2<\len_\tiles$.
+Each such tile can be (partially) seen above its marker in Figure~\ref{fig: probmax_asint_turned}, its sides forming $45^\circ$ degrees with respect to the horizontal direction.
+
+Given $\tiles$, let $N_\tiles$ be defined as the number of summands in \eqref{eq: probmax impl} for $\len=\len_\tiles$, and $i_\tiles^- = \min\{i \geq 2 \st (i-2)^2 + (\tiles-i-1)^2 < \len_\tiles^2\}$.
+For $n=1,\ldots,N_\tiles$, let
+\begin{align}
+\label{eq: i equiv n}
+i(\tiles,n) &= i_\tiles^- + n -1, \\
+\label{eq: j equiv n}
+j(\tiles,n) &= \tiles+1-i(\tiles,n) = \tiles + 2 - i_\tiles^- - n.
+\end{align}
+Then the indices $i,j$ in \eqref{eq: probmax impl} for $\len=\len_\tiles$ can be replaced by the single index $n$:
+\begin{equation}
+\label{eq: probmax impl n}
+\probmax(\len_\tiles) = \sum_{n=1}^{N_\tiles} \probmax_{i(\tiles,n), j(\tiles,n)}(\len_\tiles) = \sum_{n=1}^{N_\tiles} g\left(\len_\tiles, i(\tiles,n)-2, j(\tiles,n)-2 \right).
+\end{equation}
+The index $n = 1,\ldots,N_\tiles$  increases from left to right in Figure~\ref{fig: probmax_asint_turned}.
+Let $x_{\tiles, n}$ denote the $x$ coordinate of the $n$-th tile corner (circular marker) in this figure, and define
+\begin{equation}
+P_{\tiles,n} = \tiles \probmax_{i(\tiles,n),j(\tiles,n)}(\len_\tiles).
+\end{equation}
+Then, from \eqref{eq: probmax impl n},
+\begin{equation}
+\label{eq: sum p t n}
+\tiles \probmax(\len_\tiles) = \sum_{n=1}^{N_\tiles} P_{\tiles,n}.
+\end{equation}
+
+\begin{figure}%
+\centering
+\includegraphics[width=.73\textwidth]{probmax_asint_turned}%
+\caption{Reachable tiles that maximize the number of visited tiles on a unit square grid; rotated view. Example for $\tiles=9$}
+\label{fig: probmax_asint_turned}%
+\end{figure}%
+
+Using the coordinate axes defined in Figure~\ref{fig: probmax_asint_turned}, the circular arc satisfies the equation $y = \funcirc(x)$ with
 \begin{equation}
 \label{eq: funcirc}
-\funcirc(x) = h_\tiles - \len_\tiles + \sqrt{\len_\tiles^2-x^2}.
-\end{equation}
-The $y$ coordinate of the arc for $x=x_{\tiles,n}$ will be denoted as $y_{\tiles,n}$.
-The half-width of the arc along the horizontal axis, $w_\tiles$, is computed in the proof of Theorem~\ref{theo: probmax, sq} (or from the right triangle in Figure~\ref{fig: probmax_asint_turned}), and using \eqref{eq: len tiles} it can be expressed as
+\funcirc(x) = h_\tiles - \len_\tiles + \sqrt{\len_\tiles^2-x^2},
+\end{equation}
+where the arc height $h_\tiles$ is 
+\begin{equation}
+\label{eq: h tiles}
+h_\tiles = \len_\tiles - \frac{\tiles-3}{\sqrt{2}}.
+\end{equation}
+The $y$ coordinate of the arc for $x=x_{\tiles,n}$ will be denoted as $y_{\tiles,n} = \funcirc(x_{\tiles,n})$.
+The half-width of the arc, $w_\tiles$, is computed in the proof of Theorem~\ref{theo: probmax, sq} (or from the right triangle in Figure~\ref{fig: probmax_asint_turned}), and using \eqref{eq: len tiles} it can be expressed as
 \begin{equation}
@@
 \label{eq: sqrt asympt}
-\sqrt{1+\genvar} \leq 1 + \frac \genvar 2
+\sqrt{1+\genvar} < 1 + \frac \genvar 2 \quad\text{for } s \in \mathbb R, s \neq 0
 \end{equation}
@@
 \label{eq: asympt len tiles}
-\frac{\tiles-2}{\sqrt{2}} \leq \len_\tiles \leq \frac{\tiles-2}{\sqrt{2}} + \frac{1}{2\sqrt{2}(\tiles-2)}
+\frac{\tiles-2}{\sqrt{2}} \leq \len_\tiles < \frac{\tiles-2}{\sqrt{2}} + \frac{1}{2\sqrt{2}(\tiles-2)}
 \leq \frac{\tiles-3/2}{\sqrt{2}}.
 \end{equation}
-The arc height,
-\begin{equation}
-\label{eq: h tiles}
-h_\tiles = \len_\tiles - \frac{\tiles-3}{\sqrt{2}},
-\end{equation}
-can be bounded from \eqref{eq: asympt len tiles} as
+The arc height $h_\tiles$ is bounded from \eqref{eq: h tiles} and \eqref{eq: asympt len tiles} as
 \begin{equation}
 \label{eq: h bound}
-\frac 1 {\sqrt{2}} \leq h_\tiles \leq \frac 1 {\sqrt{2}} + \frac{1}{2\sqrt{2}(\tiles-2)}.
-\end{equation}
-
-\begin{figure}%
-\centering
-\includegraphics[width=.72\textwidth]{probmax_asint_turned}%
-\caption{Reachable tiles that maximize the number of visited tiles on a unit square grid; rotated view. Example for $\tiles=9$}
-\label{fig: probmax_asint_turned}%
-\end{figure}%
-
-As argued in the proof of Theorem~\ref{theo: probmax, sq}, $\probmax(\len_\tiles)$ is the sum of several terms $\probmax_{i,j}(\len_\tiles)$, $i+j-1=\tiles$.
-Each term is associated with one of the tiles shown in Figure~\ref{fig: probmax_asint_turned} (or in Figure~\ref{fig: probmax_nterms}), and includes a factor of $4$ to account for the symmetrical tiles in the other quadrants.
-For this proof it is more convenient to compute $\tiles \probmax_{i,j}(\len_\tiles)$, and to change the indexing as follows:
-\begin{equation}
-P_{\tiles,n} = \tiles \probmax_{i,j}(\len_\tiles),
-\end{equation}
-where $n$ equals $1$ for the smallest $i$ in \eqref{eq: probmax impl}, increases with $i$, and has a maximum value $N$ equal to the number of summands in that expression. Thus
-\begin{equation}
-\label{eq: sum p t n}
-\tiles \probmax(\len_\tiles) = \sum_{n=1}^N P_{\tiles,n}.
-\end{equation}
-
-Consider the tile depicted with thick line in Figure~\ref{fig: probmax_asint_turned}, for a generic value of the index $n$ in Figure~\ref{fig: probmax_asint_turned}, or of the indices $i,j$ in Figure~\ref{fig: probmax_nterms}.
-For this tile, $P_{\tiles,n}/t = \probmax_{i,j}(\len_\tiles)$ is obtained as the area of the shaded rectangle averaged over $\theta$ uniformly distributed on $[0,2\pi)$, and then multiplied by $4$, in the same way as in the proof of Theorem~\ref{theo: probmax, sq}.
+\frac {\sqrt{2}} 2 \leq h_\tiles <
+% \frac 1 {\sqrt{2}} + \frac{1}{2\sqrt{2}(\tiles-2)} \leq
+\frac{3\sqrt{2}}{4}.
+\end{equation}
+The number $N_\tiles$ of reachable tiles shown in the figure (circular markers) is easily obtained from $w_\tiles$ as
+\begin{equation}
+\label{eq: N tiles}
+N_\tiles = 
+\begin{cases}
+\displaystyle
+2\left\lceil \frac{w_\tiles}{\sqrt{2}} \right\rceil - 1 = 2\left\lceil \sqrt{\frac{\tiles-2}{2}} \right\rceil - 1 & \text{ for $\tiles$ odd} \\[5mm] % *!*
+\displaystyle
+2\left\lceil \frac{w_\tiles}{\sqrt{2}} - \frac 1 2 \right\rceil = 2\left\lceil \sqrt{\frac{\tiles-5/2}{2}} - \frac 1 2 \right\rceil & \text{ for $\tiles$ even}.
+\end{cases}
+\end{equation}
+
+The asymptotic analysis to be presented is based on the following observations.
+The half-width of the arc $w_\tiles$ grows asymptotically as $\sqrt{\tiles}$ for $\tiles \rightarrow \infty$, whereas the radius $\len_\tiles$ is asymptotically proportional to  $\tiles$, and the height $h_\tiles$ converges to $\sqrt{2}/2$.
+As a result, the slope $\diff \funcirc(x)/\diff x$ tends to $0$.
+
+Consider the tile depicted with thick line in Figure~\ref{fig: probmax_asint_turned}, for a generic value of the index $n$.
+For this tile, $P_{\tiles,n}/t = \probmax_{i(\tiles,n),j(\tiles,n)}(\len_\tiles)$ can be obtained as the area of the shaded rectangle in the figure averaged over $\theta$ uniformly distributed on $[0,2\pi)$, and then multiplied by $4$, in the same way as in the proof of Theorem~\ref{theo: probmax, sq}.
 Equivalently, the average can be restricted to the range of values of $\theta$ for which the shaded area is non-zero, corresponding to the arc with thick line in the figure, of length $c_{\tiles,n}$, and the result multiplied by $c_{\tiles,n}/(2\pi\len_\tiles)$.
@@
 P_{\tiles,n}
-&\leq \left(1 + \frac 2 {\tiles-2} \right) \frac{2 \sqrt{2} A_{\tiles,n} c_{\tiles,n}}{\pi} = \frac{2\sqrt{2} A_{\tiles,n} c_{\tiles,n}}{\pi} + \frac{2\sqrt{2}}{\tiles-2}, \\
+&\leq \left(1 + \frac 2 {\tiles-2} \right) \frac{2 \sqrt{2} A_{\tiles,n} c_{\tiles,n}}{\pi} < \frac{2\sqrt{2} A_{\tiles,n} c_{\tiles,n}}{\pi} + \frac{2\sqrt{2}}{\tiles-2}, \\
 \label{eq: P i j asympt 2 2}
 P_{\tiles,n}
-&\geq \left(1 + \frac {3/2} {\tiles-3/2} \right) \frac{2 \sqrt{2} A_{\tiles,n} c_{\tiles,n}}{\pi} = \frac{2 \sqrt{2} A_{\tiles,n} c_{\tiles,n}}{\pi} + \frac{3/\sqrt{2}}{\tiles-3/2}.
+&\geq \left(1 + \frac {3/2} {\tiles-3/2} \right) \frac{2 \sqrt{2} A_{\tiles,n} c_{\tiles,n}}{\pi} > \frac{2 \sqrt{2} A_{\tiles,n} c_{\tiles,n}}{\pi}.
 \end{align}
 
-To calculate $A_{\tiles,n}$ and $c_{\tiles,n}$ it is easier to replace the arc by its chord, as depicted in Figure~\ref{fig: probmax_asint_detail}.
-The approximation error for either $A_{\tiles,n}$ or $c_{\tiles,n}$ tends to $0$ as $\tiles \rightarrow \infty$.
-Specifically, regarding $A_{\tiles,n}$, the difference in $y$ between the arc and the chord is upper-bounded by the length of the chord multiplied by the maximum of the absolute value of the arc slope.
-The former is at most $\sqrt{2}$, and the latter is bounded using \eqref{eq: funcirc}, \eqref{eq: w} and \eqref{eq: asympt len tiles} as
-\begin{equation}
-\label{eq: asympt slope bound}
-\left| \frac{\diff \funcirc(x)}{\diff x} \right| = \frac{x}{\sqrt{\len_\tiles^2-x^2}} \leq \frac{w_\tiles}{\len_\tiles} \leq \sqrt{\frac{2}{\tiles-2}}.
-\end{equation}
-Thus the difference in $y$ is at most $2/\sqrt{\tiles-2}$.
-This translates into a difference of at most  $\sqrt{2/(\tiles-2)}$ in each dimension of the shaded rectangle; or, since the maximum dimensions are $1$, with at least one of them being strictly less, a difference in area smaller than twice that value.
-Therefore the approximation error in $A_{\tiles,n}$ is less than $2\sqrt{2}/\sqrt{\tiles-2}$.
-
-As for approximating $c_{\tiles,n}$ by the chord length, the latter is $2\len_\tiles \sin(c_{\tiles,n}/(2\len_\tiles))$. Taking into account that
-\begin{equation}
-\label{eq: sin asympt pos}
-\sin\genvar > \genvar - \frac{\genvar^3} 6 \quad \text{for } s > 0
-\end{equation}
-and making use of \eqref{eq: c tiles n bound} and \eqref{eq: asympt len tiles}, this is bounded as
-\begin{equation}
-2\len_\tiles \sin \frac{c_{\tiles,n}}{2\len_\tiles} > c_{\tiles,n} -\frac{c_{\tiles,n}^2}{24\len_\tiles^2} \geq c_{\tiles,n} -\frac{\pi^2}{96\len_\tiles^2} \geq c_{\tiles,n} -\frac {\pi^2} {48(\tiles-2)^2},
-\end{equation}
-and thus the length difference is at most $\pi^2/(48(\tiles-2)^2)$.
+Instead of calculating  $A_{\tiles,n}$ and $c_{\tiles,n}$ directly, it is easier to compute approximate versions of these, based on replacing the arc by its chord, as depicted in Figure~\ref{fig: probmax_asint_detail}.
+The ratio between the approximate and exact values in each case will be bounded from below and from above by functions of $\tiles$ that approach $1$ as $\tiles \rightarrow \infty$.
+Regarding $c_{\tiles,n}$, the chord length, denoted as $\tilde c_{\tiles,n}$, is obtained from Figure~\ref{fig: probmax_asint_turned} as
+\begin{equation}
+\tilde c_{\tiles,n} =  2\len_\tiles \sin\frac{c_{\tiles,n}}{2\len_\tiles}. 
+\end{equation}
+Taking into account that $\sin\genvar > \genvar - {\genvar^3}/6$ for $s > 0$ and using of \eqref{eq: asympt len tiles} and \eqref{eq: c tiles n bound},
+\begin{equation}
+\tilde c_{\tiles,n}> c_{\tiles,n} -\frac{c_{\tiles,n}^3}{24\len_\tiles^2} \geq c_{\tiles,n} -\frac{\pi^3}{192\len_\tiles^2} \geq c_{\tiles,n} - \frac {\pi^3} {96(\tiles-2)^2},
+\end{equation}
+and thus
+\begin{equation}
+\label{eq: tilde c bound}
+1 - \frac {\pi^3} {96(\tiles-2)^2} < \frac{\tilde c_{\tiles,n}}{c_{\tiles,n}} < 1.
+\end{equation}
 
@@
 \centering
-\includegraphics[width=.7\textwidth]{probmax_asint_detail}%
+\includegraphics[width=.82\textwidth]{probmax_asint_detail}%
 \caption{Computation of $P_{\tiles,n}$}%
@@
 
-Let $\tilde A_{\tiles,n}$ and $\tilde c_{\tiles,n}$ denote the approximations of $A_{\tiles,n}$ and $c_{\tiles,n}$ obtained using the chord instead of the arc. From the preceding two paragraphs,
-\begin{equation}
-\label{eq: P i j asympt 3}
-\tilde A_{\tiles,n} \tilde c_{\tiles,n} < A_{\tiles,n} c_{\tiles,n} <
-\tilde A_{\tiles,n} \tilde c_{\tiles,n} + \frac{\pi^2 A_{\tiles,n}}{48(\tiles-2^2)} + \frac{2\sqrt{2} c_{\tiles,n}}{\sqrt{\tiles-2}}.
-\end{equation}
-Using \eqref{eq: A tiles n bound} and \eqref{eq: c tiles n bound} into \eqref{eq: P i j asympt 3} and combining with \eqref{eq: P i j asympt 2 1} and \eqref{eq: P i j asympt 2 2},
+As for $A_{\tiles,n}$, let $\tilde A_{\tiles,n}$ denote its approximate version where the upper corner of the shaded rectangle in Figure~\ref{fig: probmax_asint_detail} is uniformly distributed on the chord rather than on the arc.
+This approximation entails two types of errors: the distribution of the angle $\theta$ is no longer uniform; and the shaded rectangle becomes smaller because its corner is on the chord, not on the arc.
+To analyze the first type of error, let $\fdp(\theta)$ denote the probability density function for the original, uniform distribution of the angle $\theta$ on the arc depicted with thick line in Figure~\ref{fig: probmax_asint_turned}.
+In the approximate version, the uniform distribution on the chord translates into a non-uniform distribution of $\theta$, with probability density function $\tilde\fdp(\theta)$.
+In the following it will be assumed that $\tiles \geq 4$.
+It stems from Figure~\ref{fig: probmax_asint_turned} that any point $(x,y)$ on the circular arc defined by \eqref{eq: funcirc} satisfies the equality $\cot \theta = \sqrt{2} x /(\tiles-3)$.
+Therefore
+\begin{equation}
+\label{eq: d theta d x}
+\frac{\diff \theta}{\diff x} = -\frac{\sqrt{2}}{\tiles-3} \sin^2 \theta.
+\end{equation}
+Along this arc, with $x \in [-w_\tiles, w_\tiles]$, the minimum of $|\sin \theta|$ is attained when $x$ equals $w_\tiles$ or $-w_\tiles$. Thus, using \eqref{eq: asympt len tiles},
+\begin{equation}
+\label{eq: sin theta bound}
+1 \geq |\sin \theta| \geq \frac{\tiles-3}{\sqrt{2}\len_\tiles} > \frac{\tiles-3}{\tiles-3/2}.
+\end{equation}
+The uniform distribution of the corner of the shaded rectangle on the chord is equivalent to a uniform distribution of its $x$ coordinate on the projection of the chord onto the horizontal axis.
+Denoting the slope of the chord by $m_{\tiles,n}$, the projection has length $\tilde c_{\tiles,n} / |m_{\tiles,n}|$, and $\tilde\fdp(\theta)$ is obtained, using the transformation theorem for continuous random variables \cite[section~5.2]{Papoulis91}, as
+\begin{equation}
+\label{eq: tilde fdp theta}
+\tilde \fdp(\theta) = \frac {|m_{\tiles,n}|} {\tilde c_{\tiles,n} |\diff \theta / \diff x|}
+\end{equation}
+for the range of $\theta$ corresponding to the arc with thick line in Figure~\ref{fig: probmax_asint_turned}. It follows from \eqref{eq: d theta d x}--\eqref{eq: tilde fdp theta} that
+% *!*
+\begin{equation}
+\label{eq: max min tilde pdf}
+\frac {\max_\theta \tilde \fdp(\theta)} {\min_\theta \tilde \fdp(\theta)} < \left(\frac{\tiles-3/2}{\tiles-3}\right)^2.
+\end{equation}
+Since $\fdp(\theta)$ and $\tilde\fdp(\theta)$ are probability density functions, it cannot be the case that one of them is greater than the other for all values of $\theta$.
+In addition, both functions have the same support, on which $\fdp(\theta)$ is constant.
+Taking into account that $\tiles \geq 4$, this implies that
+\begin{align}
+\label{eq: tilde pdf pdf bound >}
+\frac {\tilde \fdp(\theta)} {\fdp(\theta)} &> \left(\frac{\tiles-3}{\tiles-3/2}\right)^2 = \left(1-\frac{3/2}{\tiles-3/2}\right)^2 > 1 - \frac 3 {\tiles-3/2}, \\
+\label{eq: tilde pdf pdf bound <}
+\frac {\tilde \fdp(\theta)} {\fdp(\theta)} &< \left(\frac{\tiles-3/2}{\tiles-3}\right)^2 = \left(1+\frac{3/2}{\tiles-3}\right)^2 < 1 + \frac {21/4} {\tiles-3}.
+\end{align}
+Thus, changing the distribution of $\theta$ from $\fdp(\theta)$ to $\tilde \fdp(\theta)$ results in $A_{\tiles,n}$ being multiplied by a factor greater than the rightmost side of \eqref{eq: tilde pdf pdf bound >} and smaller than that of \eqref{eq: tilde pdf pdf bound <}.
+
+The second type of error in the approximation of $A_{\tiles,n}$ arises because, given the $x$ coordinate of the upper corner of the shaded rectangle in Figure~\ref{fig: probmax_asint_detail}, its $y$ coordinate is reduced from that of the arc to that of the chord.
+Considering the tangent line displayed in the figure (dashed line), it is easy to see that the difference between the vertical coordinate of the arc, $y$, and that of the chord, $\tilde y$, is less than the chord length, $\tilde c_{\tiles,n}$, multiplied by the maximum of the absolute value of the arc slope.
+The former is at most $\sqrt{2}$.
+The latter is upper-bounded by the value of $|\diff \funcirc(x)/\diff x|$ at $x=w_\tiles$ or $x=-w_\tiles$.
+From \eqref{eq: funcirc}, \eqref{eq: w} and \eqref{eq: sqrt asympt}, making use of the assumption that $\tiles \geq 4$,
+\begin{equation}
+\label{eq: asympt slope bound}
+\begin{split}
+\left| \frac{\diff \funcirc(x)}{\diff x} \right| &= \frac{|x|}{\sqrt{\len_\tiles^2-x^2}} \leq \frac{w_\tiles}{\sqrt{\len_\tiles^2-w_\tiles^2}} \leq \frac{\sqrt{2(\tiles-2)}}{\tiles-3} \\
+& < \sqrt{\frac 2 {\tiles-3}} \left(1 + \frac 1 {2(\tiles-3)} \right) \leq \frac 3 2 \sqrt{\frac 2 {\tiles-3}},
+\end{split}
+\end{equation}
+and thus
+\begin{equation}
+\label{eq: y: arc - chord}
+0 < y - \tilde y < \frac 3 {\sqrt{\tiles-3}}.
+\end{equation}
+This reduction from $y$ to $\tilde y$ translates into a difference smaller than $3 / \sqrt{2(\tiles-3)}$ in each dimension of the shaded rectangle; or, since those dimensions are at most $1$, a difference in area smaller than twice that value.
+Considering \eqref{eq: A tiles n bound}, this introduces a (second) error in $A_{\tiles,n}$ by a factor between $1-3\sqrt{2}/\sqrt{\tiles-3}$ and $1$.
+Combining with the factor from the previous paragraph, $\tilde A_{\tiles,n} / A_{\tiles,n}$ is bounded for $\tiles \geq 4$ as
+\begin{equation}
+\label{eq: tilde A bound}
+1 - \frac {3} {\tiles-3/2} - \frac{3\sqrt{2}} {\sqrt{\tiles-3}} < \frac{\tilde A_{\tiles,n}}{A_{\tiles,n}} < 1+\frac {21/4} {\tiles-3}.
+\end{equation}
+
+Multiplying \eqref{eq: tilde c bound} and \eqref{eq: tilde A bound} yields, for $\tiles \geq 4$,
+\begin{equation}
+\label{eq: tilde A c bound rel}
+1 - \frac {3} {\tiles-3/2} - \frac{3\sqrt{2}}  {\sqrt{\tiles-3}} - \frac {\pi^3} {96(\tiles-2)^2} < \frac{\tilde A_{\tiles,n} \tilde c_{\tiles,n}}{A_{\tiles,n} c_{\tiles,n}} < 1+\frac {21/4} {\tiles-3},
+\end{equation}
+which, using \eqref{eq: A tiles n bound} and \eqref{eq: c tiles n bound}, is transformed into 
+\begin{equation}
+\label{eq: tilde A c bound diff}
+- \frac {3\pi/2} {\tiles-3/2} - \frac{3\pi/\sqrt{2}}  {\sqrt{\tiles-3}} - \frac {\pi^4} {192(\tiles-2)^2} < \tilde A_{\tiles,n} \tilde c_{\tiles,n}- A_{\tiles,n} c_{\tiles,n} < \frac {21\pi/2} {\tiles-3}.
+\end{equation}
+Combining \eqref{eq: tilde A c bound diff} with \eqref{eq: P i j asympt 2 1} and \eqref{eq: P i j asympt 2 2} gives
 \begin{equation}
@@
 \end{equation}
-with $|E_{\tiles, n}| < e(\tiles)$, where $e$ is a function that depends only on $\tiles$ and tends to $0$ as $\tiles \rightarrow 0$.
-
-Denoting the slope of the chord by $m_{\tiles,n}$, it is clear from Figure~\ref{fig: probmax_asint_detail} that
-\begin{equation}
-\frac{y_{\tiles,n}-a_{\tiles,n}/\sqrt{2}}{a_{\tiles,n}/\sqrt{2}} = m_{\tiles,n},
-\end{equation}
-and therefore $a_{\tiles,n} = \sqrt{2}y_{\tiles,n}/(1+m_{\tiles,n})$.
-Similarly, $b_{\tiles,n} = \sqrt{2}y_{\tiles,n}/(1-m_{\tiles,n})$.
-Both $\tilde A_{\tiles,n}$ and $\tilde c_{\tiles,n}$ can be expressed in terms of $a_{\tiles,n}$ and $b_{\tiles,n}$, and thus of $y_{\tiles,n}$ and $m_{\tiles,n}$:
-\begin{align}
-\tilde A_{\tiles,n} &= a_{\tiles,n}b_{\tiles,n} \int_0^1 \eta(1-\eta)\, \diff \eta = \frac{a_{\tiles,n}b_{\tiles,n}} 6 = \frac{y_{\tiles,n}^2}{3(1-m_{\tiles,n}^2)}, \\
-\tilde c_{\tiles,n} &= \sqrt{a_{\tiles,n}^2+b_{\tiles,n}^2} = \frac{2 y_{\tiles,n} \sqrt{1+m_{\tiles,n}^2}}{1-m_{\tiles,n}^2}.
-\end{align}
-Substituting into \eqref{eq: P i j asympt 3 bis},
+with $|E_{\tiles, n}| < e(\tiles)$, where the function $e$ depends only on $\tiles$ and tends to $0$ as $\tiles \rightarrow 0$.
+
+Let $\tilde y_{\tiles,n}$ denote the $y$ coordinate of the chord for $x = x_{\tiles,n}$, as shown in Figure~\ref{fig: probmax_asint_detail}.
+By construction $|m_{\tiles,n}|<1$.
+Consider $a_{\tiles,n}$ and $b_{\tiles,n}$ as defined in the figure.
+It can be seen that
+\begin{equation}
+\frac{\tilde y_{\tiles,n}-a_{\tiles,n}/\sqrt{2}}{a_{\tiles,n}/\sqrt{2}} = m_{\tiles,n},
+\end{equation}
+which gives $a_{\tiles,n} = \sqrt{2}\tilde y_{\tiles,n}/(1+m_{\tiles,n})$.
+Likewise, $b_{\tiles,n} = \sqrt{2}\tilde y_{\tiles,n}/(1-m_{\tiles,n})$.
+With the upper corner of the shaded rectangle uniformly distributed along the chord, its average area $\tilde A_{\tiles,n}$ is easily computed in terms of $a_{\tiles,n}$ and $b_{\tiles,n}$, and therefore of $\tilde y_{\tiles,n}$ and $m_{\tiles,n}$:
+\begin{equation}
+\label{eq: tilde A y}
+\tilde A_{\tiles,n} = a_{\tiles,n}b_{\tiles,n} \int_0^1 \eta(1-\eta)\, \diff \eta = \frac{a_{\tiles,n}b_{\tiles,n}} 6 = \frac{\tilde y_{\tiles,n}^2}{3(1-m_{\tiles,n}^2)}.
+\end{equation}
+Similarly,
+\begin{equation}
+\label{eq: tilde c y}
+\tilde c_{\tiles,n} = \sqrt{a_{\tiles,n}^2+b_{\tiles,n}^2} = \frac {2 \tilde y_{\tiles,n} \sqrt{1+m_{\tiles,n}^2}} {1-m_{\tiles,n}^2}.
+\end{equation}
+Substituting \eqref{eq: tilde A y} and \eqref{eq: tilde c y} into \eqref{eq: P i j asympt 3 bis},
 \begin{equation}
@@
 P_{\tiles,n}
-= \frac{4\sqrt{2} \sqrt{1+m_{\tiles,n}^2}} {3\pi(1-m_{\tiles,n}^2)^2} y_{\tiles,n}^3 + E_{\tiles, n}.
-\end{equation}
-The slope $m_{\tiles,n}$ is bounded by \eqref{eq: asympt slope bound}, which implies that $m_{\tiles,n}^2 \leq 2/(\tiles-2)$.
-Using this inequality together with \eqref{eq: sqrt asympt}, the term depending on $m_{\tiles,n}^2$ in \eqref{eq: P i j asympt 4} can be bounded as
+= \frac{4\sqrt{2} \sqrt{1+m_{\tiles,n}^2}} {3\pi(1-m_{\tiles,n}^2)^2}\, \tilde y_{\tiles,n}^3 + E_{\tiles, n}.
+\end{equation}
+The inequality \eqref{eq: asympt slope bound} applies in particular to $|m_{\tiles,n}|$, and gives $m_{\tiles,n}^2 < 9/(2(\tiles-3))$.
+For simplicity, assume $\tiles \geq 8$ so that this bound is less than $1$.
+Then, making use of \eqref{eq: sqrt asympt}, the term depending on $m_{\tiles,n}^2$ in \eqref{eq: P i j asympt 4} satisfies
 \begin{equation}
 \label{eq: P i j asympt 4 bis}
-\frac{\sqrt{1+m_{\tiles, n}^2}} {(1-m_{\tiles, n}^2)^2} \leq \frac{(\tiles-2)^2}{(\tiles-4)^2} \left(1+\frac 1 {\tiles-2}\right) = 1 + \frac{5\tiles - 14}{(\tiles-4)^2}.
-\end{equation}
-On the other hand, $y_{\tiles,n}$ can be bounded from \eqref{eq: h bound} as $y_{\tiles,n} \leq h_\tiles \leq 1$.
-Using this inequality as well as \eqref{eq: P i j asympt 4 bis} into \eqref{eq: P i j asympt 4},
+\frac{\sqrt{1+m_{\tiles, n}^2}} {(1-m_{\tiles, n}^2)^2} \leq \frac{(\tiles-3)^2}{(\tiles-15/2)^2} \left(1+\frac {9/4} {\tiles-3}\right) = 1 + \frac{9(5\tiles - 24)}{(2\tiles-15)^2}.
+\end{equation}
+From \eqref{eq: h bound},
+\begin{equation}
+\label{eq: P i j asympt 4 ter}
+\tilde y_{\tiles,n} < y_{\tiles,n} \leq h_\tiles \leq \frac{3\sqrt{2}} 4.
+\end{equation}
+Substituting \eqref{eq: P i j asympt 4 bis} into \eqref{eq: P i j asympt 4} and taking into account \eqref{eq: P i j asympt 4 ter},
 \begin{equation}
 \label{eq: P i j asympt 5}
-P_{\tiles,n} = \frac{4\sqrt{2}} {3\pi} y_{\tiles,n}^3 + E_{\tiles,n} + E'_{\tiles,n}
+P_{\tiles,n} = \frac{4\sqrt{2}} {3\pi} \tilde y_{\tiles,n}^3 + E_{\tiles,n} + E'_{\tiles,n}
 \end{equation}
@@
 
-Let $\funcircesc_\tiles$ be defined as a scaled version of $\funcirc$, where the variable $x \in [-w_\tiles, w_\tiles]$ is replaced by $\zeta = x/w_\tiles \in [-1,1]$:
+Let $\funcircesc_\tiles$ be defined as a horizontally scaled version of $\funcirc$, where the variable $x \in [-w_\tiles, w_\tiles]$ is replaced by $\zeta = x/w_\tiles \in [-1,1]$:
 \begin{equation}
@@
 \end{equation}
-As $\tiles \rightarrow \infty$, the sequence of functions $\funcircesc_\tiles$ converges pointwise to
+As $\tiles \rightarrow \infty$, the sequence $\funcircesc_\tiles$ converges pointwise to the function
 \begin{equation}
@@
 Furthermore, the convergence is uniform.
-To show this, it is easier to consider the subsequences of $\tiles$ odd and even separately.
-For $\tiles$ odd, substituting $\len_\tiles$, $w_\tiles$ and $h_\tiles$ given by \eqref{eq: len tiles}, \eqref{eq: w} and \eqref{eq: h tiles} into \eqref{eq: funcircest tiles 1},
+To show this, it is easier to analyze the subsequences with $\tiles$ odd and even separately.
+For $\tiles$ odd, substituting $\len_\tiles$, $h_\tiles$  and $w_\tiles$ given by \eqref{eq: len tiles}, \eqref{eq: h tiles} and \eqref{eq: w} into \eqref{eq: funcircest tiles 1},
 \begin{equation}
@@
 Since the limit function $\funcircesc$ is continuous, by Dini's theorem \cite[section~1.1]{Hirsch99} the convergence of this subsequence is uniform.
-On the other hand, for $\tiles$ even
-\begin{equation}
-\label{eq: funcircesc, even 1}
-\funcircesc_\tiles(\zeta) =
-% -\frac{\tiles-3}{\sqrt{2}} + \sqrt{\frac{(\tiles-2)^2}{2} - \left(\tiles-\frac 5 2\right)\zeta^2}
--\frac{\tiles-3}{\sqrt{2}} + \sqrt{\frac{(\tiles-2)^2 - 2(\tiles-2)\zeta^2 + \zeta^2}{2}},
-\end{equation}
-and, for $\zeta \in [-1,1]$,
-\begin{equation}
-\label{eq: der funcircesc, even 1}
-\frac{\partial \funcircesc_\tiles(\zeta)}{\partial \tiles} = \frac 1 {\sqrt{2}} \left(
--1 + \frac{\tiles-2-\zeta^2}{\sqrt{(\tiles-2)^2 - 2(\tiles-2)\zeta^2 + \zeta^2}} \right) \leq 0.
-\end{equation}
-Thus the subsequence $\funcircesc_\tiles$ with $\tiles$ even is also monotone non-increasing, and therefore converges uniformly.
-Consequently, the full sequence $\funcircesc_\tiles$ converges to $\funcircesc$ uniformly on $[-1,1]$.
-This means that
-\begin{equation}
-\label{eq: funcircest tiles 4}
-y_{\tiles,n} = \funcirc(x_{\tiles,n}) = \funcircesc_\tiles\left( \frac{x_{\tiles,n}} {w_\tiles} \right) = \funcircesc\left( \frac{x_{\tiles,n}} {w_\tiles} \right) + E''_{\tiles,n}
-\end{equation}
-with $|E''_{\tiles,n}| < e''(t)$, where $e''(\tiles)$ tends to $0$ as $\tiles \rightarrow \infty$.
-
-Combining \eqref{eq: sum p t n}, \eqref{eq: P i j asympt 5} and \eqref{eq: funcircest tiles 4},
+An analogous argument establishes the uniform convergence of the subsequence for $\tiles$ even.
+%\begin{equation}
+%\label{eq: funcircesc, even 1}
+%\funcircesc_\tiles(\zeta) =
+%-\frac{\tiles-3}{\sqrt{2}} + \sqrt{\frac{(\tiles-2)^2 - 2(\tiles-2)\zeta^2 + \zeta^2}{2}},
+%\end{equation}
+%\begin{equation}
+%\label{eq: der funcircesc, even 1}
+%\frac{\partial \funcircesc_\tiles(\zeta)}{\partial \tiles} = \frac 1 {\sqrt{2}} \left( -1 + \frac{\tiles-2-\zeta^2}{\sqrt{(\tiles-2)^2 - 2(\tiles-2)\zeta^2 + \zeta^2}} \right) \leq 0.
+%\end{equation}
+Therefore, the full sequence $\funcircesc_\tiles$ converges to $\funcircesc$ uniformly on $[-1,1]$.
+
+%***Releer
+
+The inequality \eqref{eq: y: arc - chord} applies, for $\tiles \geq 4$, to $y_{\tiles,n} = \funcircesc_\tiles(x_{\tiles,n}/w_\tiles)$ and $\tilde y_{\tiles,n}$.
+Together with \eqref{eq: P i j asympt 4 ter}, and making use of the uniform convergence of $\funcircesc_\tiles$ to $\funcircesc$,
+this implies that
+\begin{equation}
+\label{eq: funcircest tiles 5}
+\tilde y_{\tiles,n}^3 = \funcircesc^3\left( \frac{x_{\tiles,n}} {w_\tiles} \right) + E''_{\tiles,n}
+\end{equation}
+where $e''(\tiles)$ tends to $0$ as $\tiles \rightarrow \infty$.
+
+From \eqref{eq: sum p t n}, \eqref{eq: P i j asympt 5} and \eqref{eq: funcircest tiles 5},
 \begin{equation}
 \label{eq: sum p t n aprox 1}
-\tiles \probmax(\len_\tiles) = \sum_{n=1}^N P_{\tiles,n} = \frac{4 \sqrt{2}} {3\pi} \sum_{n=1}^N \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right) + \sum_{n=1}^N \left(E_{\tiles,n} + E'_{\tiles,n} + E''_{\tiles,n} \right),
+\tiles \probmax(\len_\tiles) = \frac{4 \sqrt{2}} {3\pi} \sum_{n=1}^{N_\tiles} \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right) + \sum_{n=1}^{N_\tiles} \left(E_{\tiles,n} + E'_{\tiles,n} + \frac{4 \sqrt{2}} {3\pi} E''_{\tiles,n} \right),
 \end{equation}
@@
 \label{eq: sum p t n aprox 2}
-\sqrt{\tiles} \probmax(\len_\tiles) = \frac{4 w} {3\pi\sqrt{\tiles}} \frac {\sqrt{2}} {w} \sum_{n=1}^N \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right) +
-\frac{N} {\sqrt{t}} \frac 1 N \sum_{n=1}^N (E_{\tiles,n} + E'_{\tiles,n} + E''_{\tiles,n}).
-\end{equation}
-Noting that the values $\zeta_{\tiles,n}$ are equispaced from $-1$ to $1$ with step $\sqrt{2}/w$, the term $\sqrt{2}/w \sum_{n=1}^N \funcircesc^3 ( x_{\tiles,n}/w_\tiles)$ in \eqref{eq: sum p t n aprox 2} can be interpreted as a Riemmann sum that approximates the integral of $\funcircesc^3(\zeta)$ over $[-1,1]$.
+\sqrt{\tiles} \probmax(\len_\tiles) = \frac{4 w_\tiles} {3\pi\sqrt{\tiles}} \frac {\sqrt{2}} {w_\tiles} \sum_{n=1}^{N_\tiles} \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right)
++ \frac{{N_\tiles}} {\sqrt{t}} \frac 1 {N_\tiles} \sum_{n=1}^{N_\tiles} \left(E_{\tiles,n} + E'_{\tiles,n} + \frac{4 \sqrt{2}} {3\pi} E''_{\tiles,n} \right).
+\end{equation}
+It is clear from Figure~\ref{fig: probmax_asint_turned} that
+\begin{alignat}{2}
+-w_\tiles &< x_{\tiles,1} && \leq -w_\tiles + \sqrt{2}, \\
+w_\tiles - \sqrt{2} &\leq x_{\tiles,N_\tiles} && < w_\tiles.
+\end{alignat}
+Thus the values $x_{\tiles,n}/w_\tiles$, $n=1,\ldots,N_\tiles$ are equispaced with step $\sqrt{2}/w_\tiles$.
+In addition, $x_{\tiles,1}$ and $x_{\tiles,N_\tiles}$ tend to $-1$ and $1$  respectively as $\tiles \rightarrow \infty$.
+In consequence, the term
+\[
+\frac{\sqrt{2}}{w_\tiles} \sum_{n=1}^{N_\tiles} \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right)
+\]
+in \eqref{eq: sum p t n aprox 2} can be interpreted as a Riemmann sum that approximates the integral of $\funcircesc^3(\zeta)$ over $[-1,1]$.
 Since this function is continuous the sum indeed converges to the integral \cite[section~7.2]{Abbott15}.
@@
 \begin{equation}
-\left| \frac 1 N \sum_{n=1}^N \left( E_{\tiles,n} + E'_{\tiles,n} + E''_{\tiles,n} \right) \right| < e(\tiles)+e'(\tiles)+e''(\tiles)
+\left| \frac 1 {N_\tiles} \sum_{n=1}^{N_\tiles} \left(E_{\tiles,n} + E'_{\tiles,n} + \frac{4 \sqrt{2}} {3\pi} E''_{\tiles,n} \right) \right| < e(\tiles)+e'(\tiles) + \frac{4 \sqrt{2}} {3\pi} e''(\tiles)
 \end{equation}
 and the right-hand side tends to $0$ as $\tiles \rightarrow \infty$.
-The terms $w/\sqrt{\tiles}$ and $N/\sqrt{\tiles}$ in \eqref{eq: sum p t n aprox 2} tend to $1$ and $\sqrt{2}$ respectively as $\tiles \rightarrow \infty$.
-Therefore, taking limits and substituting \eqref{eq: funcircesc},
+The terms $w_\tiles/\sqrt{\tiles}$ and $N_\tiles/\sqrt{\tiles}$ in \eqref{eq: sum p t n aprox 2} tend to $1$ and $\sqrt{2}$ respectively, according to \eqref{eq: w} and \eqref{eq: N tiles}.
+Thus, taking limits and substituting \eqref{eq: funcircesc},
 \begin{equation}

### 2023-04-17 figures/figures.m

@@ -2197,9 +2197,8 @@ set(gca, 'XScale', 'log', 'YScale', 'log') % a gap is visible
 % are far from convergence
 %   Yes. The sum of terms divided by the estimated number of terms sqrt(t/2) gives around 0.4571,
 % which multiplied by 2 is 0.914. There was a factor of 2 or almost 2 missing, because the terms for
-% k>0 should appear twice in the sum; and that for k=0 too for t even. So the estimated number of
+% k>0 should appear twice in the sum, as should that for k=0 too for t even. So the estimated number of
 % terms should be sqrt(2*t), not sqrt(t/2).
-%   So the problem is that sum(g_actual_k/sqrt(2*t))
 
 clear, close all
 g = @(r,u,v) ((acos(u/2./r)-asin(v/2./r)).*u.*v + 2*r.^2 + u.^2/2 + v.^2/2 - u.*sqrt(4*r.^2-v.^2) - v.*sqrt(4*r.^2-u.^2))/2/pi;
@@ -2671,8 +2670,12 @@ for L = 8.32-2*sqrt(2) %[7.9 8.24 8.59 8.85 9.12]-sqrt(2)
     jj = jj(ind);
     ind_extra = abs(ii-jj)>1;
     for n = 1:numel(ii)
-        plot(ii(n)-1 + [0 1], jj(n)-1 + [0 0], 'k-', 'linewidth', 1) % horizontal sides of tiles
-        plot(ii(n)-1 + [0 0], jj(n)-1 + [0 1], 'k-', 'linewidth', 1) % vertical sides of tiles
+        patch(ii(n)-1+[0 1 1 0 0], jj(n)-1+[0 0 1 1 0], .88*[1 1 1], 'edgecolor', 'none')
+        lw = .5;
+        plot(ii(n)-1 + [0 1], jj(n)-1 + [0 0], 'k-', 'linewidth', lw) % horizontal sides of tiles
+        plot(ii(n)-1 + [0 0], jj(n)-1 + [0 1], 'k-', 'linewidth', lw) % vertical sides of tiles
[…]

### 2023-04-17 paper.tex
Correction of errata found by the reviewer

@@
 For each $\tiles \geq 3$, the pair $(i_\tiles,j_\tiles)$ should be chosen as that on the line $i+j-1=\tiles$ which minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$.
-This allows maximizing the sum $i+j-1$, and thus the number of visited tiles, for a given length restriction; or visiting a specified number of files with lengths as small as possible.
+This allows maximizing the sum $i+j-1$, and thus the number of visited tiles, for a given length restriction; or visiting a specified number of tiles with lengths as small as possible.
 
@@
 
-For $\tiles \geq 3$, the term $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$ can be interpreted geometrically as the distance between $(i_\tiles, j_\tiles)$ and $(2a, 2b)$.
-As can be seen with the help of Figure~\ref{fig: min suff set, form}, this distance is the hypotenuse of a right triangle whose other two sides extend from $(2a, 2b)$ to $(\isolr a, \jsolr b)$ and from $(\isolr a, \jsolr b)$ to $(i_\tiles, j_\tiles)$ respectively.
+For $\tiles \geq 3$, the term $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$ can be interpreted geometrically as the distance between $(i_\tiles a, j_\tiles a)$ and $(2a, 2b)$.
+As can be seen with the help of Figure~\ref{fig: min suff set, form}, this distance is the hypotenuse of a right triangle whose other two sides extend from $(2a, 2b)$ to $(\isolr a, \jsolr b)$ and from $(\isolr a, \jsolr b)$ to $(i_\tiles a, j_\tiles a)$ respectively.
 Therefore,
@@
 \end{equation}
-and its second derivative is easily seen to be negative. Therefore $\genfun'(\genvar)$ can be bounded for $\genvar \geq 3$ as
+and its second derivative is easily seen to be negative. Therefore $\genfun'(\genvar)$ can be bounded for $\genvar > 3$ as
 \begin{equation}
@@
 \end{equation}
-For $\leni \geq 2$, by the mean value theorem \cite[section~5.3]{Abbott15}, when $\leni$ is increased to $\leni+1$ the term $\sqrt{2\leni^2-2}$ in \eqref{eq: funti leni} has an increment that equals $\genfun'(\genvar)$ for some $\leni < \genvar < \leni+1$.
+For $\leni \geq 3$,
+% I think the above this could be done for $\leni > 2$, replacing $3/2$ in \eqref{eq: der bounds} by $s\sqrt{2/3}$, which is also less than $2$.
+Then, only the case  $\funti(2)-\funti(1) = 2$ would have to be checked separately.
+by the mean value theorem \cite[section~5.3]{Abbott15}, when $\leni$ is increased to $\leni+1$ the term $\sqrt{2\leni^2-2}$ in \eqref{eq: funti leni} has an increment that equals $\genfun'(\genvar)$ for some $\leni < \genvar < \leni+1$.
 Therefore
@@
 Firstly, since the number $\pi$ is involved in \eqref{eq: funta, form}, it is possible to design a simple probabilistic experiment to estimate the value of $\pi$, as in Buffon's original problem.
-For example, choosing $a=b=\len=1$ gives $\funta(\len) = 1/\pi+1$.
+For example, choosing $a=b=\len=1$ gives $\funta(\len) = 4/\pi+1$.
 

### 2023-04-17 figures/figures.m
Modified range for figure

@@ -2793,7 +2793,7 @@ set(gcf, 'Position', [680 558 560 380])
 %% #91. sqrt(t)*probmax(l_t)
 
 clear, clf
-t_all = 3:200;
+t_all = 10:1000;
 ind = mod(t_all,2)==1;
 L_all(ind) = hypot(t_all(ind)-3,t_all(ind)-1)/2;
 ind = mod(t_all,2)==0;
@@ -2824,7 +2824,7 @@ for ind_fine = 1:numel(funt_fine)
     end
 end
 
-marker = '.'; ms = 6;
+marker = '.'; ms = 5;
 hold on, grid on, box on, set(gca, 'colororderindex', 2)
 plot(t_all.*ind_odd./ind_odd, result_computed.*sqrt(t_all).*ind_odd./ind_odd, marker, 'markersize', ms)
 set(gca, 'colororderindex', 1)

### 2023-04-18 paper.tex
Wording
Minor adjustments

@@
 \newcommand{\fdp}{p}
+\newcommand{\lra}{\alpha}
+\newcommand{\lrb}{\beta}
 
@@
 The function $\funt$ is piecewise constant and left-continuous, with unit-height jumps.
-A jump occurs at $\len$ if and only if $\len = \funl(\tiles)$ for some $\tiles \in \mathbb N$, $\tiles \geq 4$; and then $\funt(\len) = \tiles-1$, $\lim_{h \rightarrow 0+} \funt(\len+h) = \tiles$.
+A jump occurs at $\len$ if and only if $\len = \funl(\tiles)$ for some $\tiles \in \mathbb N$, $\tiles \geq 4$; and then $\funt(\len) = \tiles-1$, $\lim_{\delta \rightarrow 0+} \funt(\len+\delta) = \tiles$.
 \end{theorem}
@@
 \end{equation}
-and there exists $\epsilon > 0$ such that $\funt(\len+h) = \tiles$ for $0 < h < \epsilon$. This implies that
+and there exists $\epsilon > 0$ such that $\funt(\len+\delta) = \tiles$ for $0 < \delta < \epsilon$. This implies that
 \begin{equation}
 \label{eq: theo funt, form: prop 3}
-\lim_{h \rightarrow 0+} \funt(\len+h) = \tiles,
+\lim_{\delta \rightarrow 0+} \funt(\len+\delta) = \tiles,
 \end{equation}
@@
 \label{eq: f}
-f(\genvar) = \int_0^\genvar \arccos z \, \diff z = \genvar \arccos \genvar - \sqrt{1-\genvar^2} + 1.
+f(\genvar) = \int_0^\genvar \arccos x \, \diff x = \genvar \arccos \genvar - \sqrt{1-\genvar^2} + 1.
 \end{equation}
@@
 The probability of visiting the maximum number of tiles for a rectangular grid is difficult to compute.
-The reason is analogous to the irregularity of the optimal $(i,j)$ pairs which was discussed in \S\ref{part: fund results}.
+The reason is the irregularity of the relevant $(i,j)$ pairs for general $a$, $b$, analogous to that observed in \S\ref{part: fund results} for the pairs $(i,j)$ that maximize the number of visited tiles.
 In a square grid, however, the problem is more tractable.
@@
 \end{equation}
-Computing this integral and using the identities $\cos^2 \alpha = 1-\sin^2 \alpha$ and $\cos(2\alpha) = 2\cos^2\alpha-1 = 1 - 2\sin^2\alpha$,
+Computing this integral and using the identities $\cos^2 z = 1-\sin^2 z$ and $\cos(2z) = 2\cos^2z-1 = 1 - 2\sin^2z$,
 \begin{equation}
@@
 By construction $|m_{\tiles,n}|<1$.
-Consider $a_{\tiles,n}$ and $b_{\tiles,n}$ as defined in the figure.
-It can be seen that
-\begin{equation}
-\frac{\tilde y_{\tiles,n}-a_{\tiles,n}/\sqrt{2}}{a_{\tiles,n}/\sqrt{2}} = m_{\tiles,n},
-\end{equation}
-which gives $a_{\tiles,n} = \sqrt{2}\tilde y_{\tiles,n}/(1+m_{\tiles,n})$.
-Likewise, $b_{\tiles,n} = \sqrt{2}\tilde y_{\tiles,n}/(1-m_{\tiles,n})$.
-With the upper corner of the shaded rectangle uniformly distributed along the chord, its average area $\tilde A_{\tiles,n}$ is easily computed in terms of $a_{\tiles,n}$ and $b_{\tiles,n}$, and therefore of $\tilde y_{\tiles,n}$ and $m_{\tiles,n}$:
+Consider the lengths $\lra_{\tiles,n}$ and $\lrb_{\tiles,n}$ defined in the figure.
+Then
+\begin{equation}
+\frac{\tilde y_{\tiles,n}-\lra_{\tiles,n}/\sqrt{2}}{\lra_{\tiles,n}/\sqrt{2}} = m_{\tiles,n},
+\end{equation}
+which gives $\lra_{\tiles,n} = \sqrt{2}\tilde y_{\tiles,n}/(1+m_{\tiles,n})$.
+Likewise, $\lrb_{\tiles,n} = \sqrt{2}\tilde y_{\tiles,n}/(1-m_{\tiles,n})$.
+With the upper corner of the shaded rectangle uniformly distributed along the chord, its average area $\tilde A_{\tiles,n}$ is easily computed in terms of $\lra_{\tiles,n}$ and $\lrb_{\tiles,n}$, and therefore of $\tilde y_{\tiles,n}$ and $m_{\tiles,n}$:
 \begin{equation}
 \label{eq: tilde A y}
-\tilde A_{\tiles,n} = a_{\tiles,n}b_{\tiles,n} \int_0^1 \eta(1-\eta)\, \diff \eta = \frac{a_{\tiles,n}b_{\tiles,n}} 6 = \frac{\tilde y_{\tiles,n}^2}{3(1-m_{\tiles,n}^2)}.
+\tilde A_{\tiles,n} = \lra_{\tiles,n}\lrb_{\tiles,n} \int_0^1 \eta(1-\eta)\, \diff \eta = \frac{\lra_{\tiles,n}\lrb_{\tiles,n}} 6 = \frac{\tilde y_{\tiles,n}^2}{3(1-m_{\tiles,n}^2)}.
 \end{equation}
@@
 \label{eq: tilde c y}
-\tilde c_{\tiles,n} = \sqrt{a_{\tiles,n}^2+b_{\tiles,n}^2} = \frac {2 \tilde y_{\tiles,n} \sqrt{1+m_{\tiles,n}^2}} {1-m_{\tiles,n}^2}.
+\tilde c_{\tiles,n} = \sqrt{\lra_{\tiles,n}^2+\lrb_{\tiles,n}^2} = \frac {2 \tilde y_{\tiles,n} \sqrt{1+m_{\tiles,n}^2}} {1-m_{\tiles,n}^2}.
 \end{equation}

### 2023-04-18 figures/figures.m

@@ -2904,9 +2904,9 @@ for ind = 1:3
     plot(x_dots(ind)+[sqrt(.5) 0], [sqrt(.5) 0], 'k-', 'linewidth', .5)
 end
 set(gca, 'colororderindex', 1)
-plot([-1j*y_aux z0], 'k-', 'linewidth', .5)
+plot([-1j*y_aux z0], '-', 'linewidth', .5)
 theta = linspace(0,angle(z0+1j*y_aux), 80);
-plot(-1j*y_aux + .35*exp(1j*theta), 'k-')
+plot(-1j*y_aux + .35*exp(1j*theta), '-')
 text(-.35, -3.88, char(hex2dec('03B8')))
 text(-1.86, .77, 'c_{t,n}')
 plot(-1j*y_aux+[0 .55], 'k:')
@@ -3129,15 +3129,17 @@ yticklabels({'0' '' '' [hex2dec('221A') '2/2']})
 xlabel x, ylabel y
 %set(gca, 'XColor', 'none', 'Ycolor', 'none')
 %figure(gcf)
-text(-1.79,.254,'a_{t,n}')
-text(-.945,.41,'b_{t,n}')
+%text(-1.79,.254,'a_{t,n}')
+text(-1.787,.254,[hex2dec('03B1') '_{t,n}'])
+%text(-.945,.41,'b_{t,n}')
+text(-.945,.41,[hex2dec('03B2') '_{t,n}'])
 text(-1.16,.705,'c_{t,n}')
 %text(-1.533,.437,'y_{t,n}')
 %text(-1.533,.436,[hex2dec('1EF9') '_{t,n}']) % "y" with tilde
-x = -2.41; y = .495; text(x,y,['y_{t,n}']), text(x+.002,y+.0322,'~','fontsize',8)
+x = -2.41; y = .495; text(x,y,['y_{t,n}']), text(x+.001,y+.0322,'~','fontsize',8)
 x = -2.41; y = .595; text(x,y,['y_{t,n}'])
-set(gcf, 'Position', [680 558 550 330]) % smaller figure, so font is larger in document
-x = -1.053; y = .5658; text(x,y,'c_{t,n}'), text(x+.002,y+.0322,'~','fontsize',8) % there is no "c" with tilde
+set(gcf, 'Position', [280 258 550 330]) % smaller figure, so font is larger in document
[…]

### 2023-05-22 paper.tex
A few changes before re-submitting to Mathematika

@@
 % Authors with different affiliations: from https://tex.stackexchange.com/a/370813/90222
-\author[1]{Luis Mendo}
-\author[2]{Alex Arkhipov}
-
-\affil[1]{\small Universidad Polit\'ecnica de Madrid. E-mail:  \texttt{luis.mendo@upm.es}}
+\author[1]{Alex Arkhipov}
+\author[2]{Luis Mendo}
+
+\affil[1]{\small E-mail: \texttt{arkhipov@alum.mit.edu}}
+\affil[2]{\small Universidad Polit\'ecnica de Madrid. E-mail:  \texttt{luis.mendo@upm.es}}
 %Information Processing and Telecommunications Center, Universidad Polit\'ecnica de Madrid\\
 %Avenida Complutense, 30. 28040 Madrid, Spain.\\
-\affil[2]{\small E-mail: \texttt{arkhipov@alum.mit.edu}}
 
@@
 \end{align}
-For $\theta$ outside of this range the conditional probability $\probmax_{i, j}(\len, \theta)$ is $0$.
-Averaging $\probmax_{i, j}(\len, \theta)$ over $\theta$ gives the probability that the segment visits this tile, and then multiplying by $4$ yields $\probmax_{i.j}(\len)$:
+For $\theta$ outside of this range, the conditional probability $\probmax_{i, j}(\len, \theta)$ is $0$.
+Averaging $\probmax_{i, j}(\len, \theta)$ over $\theta$ gives the probability that the segment visits this tile, and then multiplying by $4$ yields $\probmax_{i,j}(\len)$:
 \begin{equation}
@@
 For $\len$ in this range, the number of reachable tiles with lower-left corners $(i-1,j-1)$, $i,j \geq 2$, $i+j-1=\tiles$, i.e.~the number of terms in \eqref{eq: probmax impl}, depends on $\len$.
-If $\tiles$ is odd the minimum number is $1$, corresponding to a tile with its lower-left corner on the line $x=y$, which gives the first term in the first expression in \eqref{eq: probmax expl}.
-For $\tiles$ even the minimum number is $2$, corresponding to two symmetrical tiles whose lower-left $x$ and $y$ coordinates differ by $1$, which give the first term in the second expression.
+If $\tiles$ is odd, the minimum number is $1$, corresponding to a tile with its lower-left corner on the line $x=y$, which gives the first term in the first expression in \eqref{eq: probmax expl}.
+For $\tiles$ even, the minimum number is $2$, corresponding to two symmetrical tiles whose lower-left $x$ and $y$ coordinates differ by $1$, which give the first term in the second expression.
 Either for $\tiles$ odd or even, additional tiles may become reachable as $\len$ grows (within the range allowed by $\tiles$), and these always occur in symmetrical pairs.
@@
 \begin{proposition}
-The function $\probmax$ has the following asymptotic behaviour:
+The probability $\probmax(\len)$ of visiting the maximum number of tiles on a unit square grid has the following asymptotic behaviour:
 \begin{align}

### 2023-05-22 figures/figures.m

@@ -2751,7 +2751,7 @@ set(gcf, 'Position', [680 558 560 400])
 
 
 %% #90. Figure: probmax, incorporating the correction from #85
-% I usee #85 computed4, additionally changing the order of second and third inputs of g
+% I use #85 computed4, additionally changing the order of second and third inputs of g
 
 clear, clc
 L_fine = .001:.001:16; % spacing should be small because curves are almost vertical at some points
@@ -3255,3 +3255,114 @@ for c = 1:numel(t_all)
 end
 isequal(num_terms, num_terms2) % check
 
+
+%% #98. Based on #85: figure for comparing computed and simulated probability
+
+clear, close all
+L_fine = 15:.001:20;
+
+% From #77
+g = @(r,u,v) ((acos(u/2./r)-asin(v/2./r)).*u.*v + 2*r.^2 + u.^2/2 + v.^2/2 - u.*sqrt(4*r.^2-v.^2) - v.*sqrt(4*r.^2-u.^2))/2/pi;
+ii = ceil(L_fine/sqrt(2)) + 1;
+jj = ceil(sqrt(L_fine.^2 - (ii-2).^2 )) + 1;
+funt_fine = ii+jj-1; % formula valid for real-valued lengths
+result_computed = NaN(size(funt_fine));
+num_k =  NaN(size(funt_fine));
+for ind_fine = 1:numel(funt_fine)
+    t = funt_fine(ind_fine);
+    L = L_fine(ind_fine);
+    if mod(t,2) % odd
+        res = g(L, t-3, t-3); % this term is always present
+        potential_k = 1:(t-3)/2;
+        %%potential_k = 1; % old (wrong), that is, #65
+        ind_actual_k = ((t-3)/2-potential_k).^2 + ((t-3)/2+potential_k).^2 < L^2;
+        actual_k = potential_k(ind_actual_k);
+        for k = actual_k
+            res = res + 2*g(L, t-3-2*k, t-3+2*k);
+        end
+    else % even
+        res = 2*g(L, t-4, t-2); % this term is always present
+        potential_k = 1:(t-4)/2;
+        %%potential_k = 1; % old (wrong), that is, #65
+        ind_actual_k = ((t-4)/2-potential_k).^2 + ((t-2)/2+potential_k).^2 < L^2;
[…]

### 2023-07-01 paper.tex
Errata corrected, for second revision

@@
 
-For any $i, j \geq 2$ with $i+j-1=\tiles$, the segment has a discrete bounding rectangle with dimensions $i, j$ if and only if its second endpoint is in the tile with lower-left corner $(i-1,j-1)$ or in one of the other three symmetrical tiles with respect to the lines $x=1/2$ and $y=1/2$, i.e.~those with lower-left corners $(i-1,1-j)$, $(1-i,1-j)$ and $(1-i,1-j)$.
+For any $i, j \geq 2$ with $i+j-1=\tiles$, the segment has a discrete bounding rectangle with dimensions $i, j$ if and only if its second endpoint is in the tile with lower-left corner $(i-1,j-1)$ or in one of the other three symmetrical tiles with respect to the lines $x=1/2$ and $y=1/2$, i.e.~those with lower-left corners $(i-1,1-j)$, $(1-i,j-1)$ and $(1-i,1-j)$.
 An example is shown in Figure~\ref{fig: probmax_symmetric_tiles}.
@@
 \label{eq: i-2 j-2 len}
-(i-2)^2+(j-2)<\len^2.
+(i-2)^2+(j-2)^2<\len^2.
 \end{equation}
@@
 \label{eq: sqrt asympt}
-\sqrt{1+\genvar} < 1 + \frac \genvar 2 \quad\text{for } s \in \mathbb R, s \neq 0
+\sqrt{1+\genvar} < 1 + \frac \genvar 2 \quad\text{for } s \geq -1, \ s \neq 0
 \end{equation}
@@
 The uniform distribution of the corner of the shaded rectangle on the chord is equivalent to a uniform distribution of its $x$ coordinate on the projection of the chord onto the horizontal axis.
-Denoting the slope of the chord by $m_{\tiles,n}$, the projection has length $\tilde c_{\tiles,n} / |m_{\tiles,n}|$, and $\tilde\fdp(\theta)$ is obtained, using the transformation theorem for continuous random variables \cite[section~5.2]{Papoulis91}, as
+Denoting the slope of the chord by $m_{\tiles,n}$, the projection has length
+\[
+\frac{\tilde c_{\tiles,n}} {\sqrt{1 + m_{\tiles,n}^2}},
+\]
+and $\tilde\fdp(\theta)$ is obtained, using the transformation theorem for continuous random variables \cite[section~5.2]{Papoulis91}, as
 \begin{equation}
 \label{eq: tilde fdp theta}
-\tilde \fdp(\theta) = \frac {|m_{\tiles,n}|} {\tilde c_{\tiles,n} |\diff \theta / \diff x|}
+\tilde \fdp(\theta) = \frac {\sqrt{1 + m_{\tiles,n}^2}} {\tilde c_{\tiles,n} |\diff \theta / \diff x|}
 \end{equation}
@@
 \label{eq: tilde pdf pdf bound <}
-\frac {\tilde \fdp(\theta)} {\fdp(\theta)} &< \left(\frac{\tiles-3/2}{\tiles-3}\right)^2 = \left(1+\frac{3/2}{\tiles-3}\right)^2 < 1 + \frac {21/4} {\tiles-3}.
+\frac {\tilde \fdp(\theta)} {\fdp(\theta)} &< \left(\frac{\tiles-3/2}{\tiles-3}\right)^2 = \left(1+\frac{3/2}{\tiles-3}\right)^2 \leq 1 + \frac {21/4} {\tiles-3}.
 \end{align}
@@
 \label{eq: y: arc - chord}
-0 < y - \tilde y < \frac 3 {\sqrt{\tiles-3}}.
+0 \leq y - \tilde y < \frac 3 {\sqrt{\tiles-3}}.
 \end{equation}
@@
 
-%***Releer
-
 The inequality \eqref{eq: y: arc - chord} applies, for $\tiles \geq 4$, to $y_{\tiles,n} = \funcircesc_\tiles(x_{\tiles,n}/w_\tiles)$ and $\tilde y_{\tiles,n}$.
@@
 \end{equation}
-where $e''(\tiles)$ tends to $0$ as $\tiles \rightarrow \infty$.
+with $|E''_{\tiles,n}| < e''(t)$, where $e''(\tiles)$ tends to $0$ as $\tiles \rightarrow \infty$.
 
@@
 Thus the values $x_{\tiles,n}/w_\tiles$, $n=1,\ldots,N_\tiles$ are equispaced with step $\sqrt{2}/w_\tiles$.
-In addition, $x_{\tiles,1}$ and $x_{\tiles,N_\tiles}$ tend to $-1$ and $1$  respectively as $\tiles \rightarrow \infty$.
+In addition, $x_{\tiles,1} / w_\tiles$ and $x_{\tiles,N_\tiles} / w_\tiles$ tend to $-1$ and $1$  respectively as $\tiles \rightarrow \infty$.
 In consequence, the term
@@
 \]
-in \eqref{eq: sum p t n aprox 2} can be interpreted as a Riemmann sum that approximates the integral of $\funcircesc^3(\zeta)$ over $[-1,1]$.
+in \eqref{eq: sum p t n aprox 2} can be interpreted as a Riemann sum that approximates the integral of $\funcircesc^3(\zeta)$ over $[-1,1]$.
 Since this function is continuous the sum indeed converges to the integral \cite[section~7.2]{Abbott15}.

### 2023-07-10 paper.tex
Second revision, submitted

@@
 \usepackage{authblk}
+\usepackage{bigints}
 
@@
 \newcommand{\lrb}{\beta}
+\newcommand{\intervx}{I}
 
@@
 Instead of calculating  $A_{\tiles,n}$ and $c_{\tiles,n}$ directly, it is easier to compute approximate versions of these, based on replacing the arc by its chord, as depicted in Figure~\ref{fig: probmax_asint_detail}.
-The ratio between the approximate and exact values in each case will be bounded from below and from above by functions of $\tiles$ that approach $1$ as $\tiles \rightarrow \infty$.
+The difference between the approximate and exact values for each variable will be bounded from below and from above by functions of $\tiles$ that tend to $0$ as $\tiles \rightarrow \infty$.
 Regarding $c_{\tiles,n}$, the chord length, denoted as $\tilde c_{\tiles,n}$, is obtained from Figure~\ref{fig: probmax_asint_turned} as
@@
 \end{equation}
-Taking into account that $\sin\genvar > \genvar - {\genvar^3}/6$ for $s > 0$ and using of \eqref{eq: asympt len tiles} and \eqref{eq: c tiles n bound},
-\begin{equation}
-\tilde c_{\tiles,n}> c_{\tiles,n} -\frac{c_{\tiles,n}^3}{24\len_\tiles^2} \geq c_{\tiles,n} -\frac{\pi^3}{192\len_\tiles^2} \geq c_{\tiles,n} - \frac {\pi^3} {96(\tiles-2)^2},
-\end{equation}
-and thus
-\begin{equation}
-\label{eq: tilde c bound}
-1 - \frac {\pi^3} {96(\tiles-2)^2} < \frac{\tilde c_{\tiles,n}}{c_{\tiles,n}} < 1.
+Taking into account that $\sin\genvar > \genvar - {\genvar^3}/6$ for $s > 0$ and using \eqref{eq: asympt len tiles},
+\begin{equation}
+\label{eq: tilde c bound diff}
+0 > \tilde c_{\tiles,n} - c_{\tiles,n} > -\frac{c_{\tiles,n}^3}{24\len_\tiles^2} \geq -\frac{\pi^3}{192\len_\tiles^2} \geq -\frac {\pi^3} {96(\tiles-2)^2}.
 \end{equation}
@@
 This approximation entails two types of errors: the distribution of the angle $\theta$ is no longer uniform; and the shaded rectangle becomes smaller because its corner is on the chord, not on the arc.
-To analyze the first type of error, let $\fdp(\theta)$ denote the probability density function for the original, uniform distribution of the angle $\theta$ on the arc depicted with thick line in Figure~\ref{fig: probmax_asint_turned}.
-In the approximate version, the uniform distribution on the chord translates into a non-uniform distribution of $\theta$, with probability density function $\tilde\fdp(\theta)$.
+It is convenient to describe the two sources of error in terms of, respectively, the horizontal and vertical coordinates of the upper corner of the shaded rectangle.
+More specifically, let $x$ and $y$ denote the coordinates of the corner along the arc, and let $\intervx_{\tiles,n}$ denote the interval of values of $x$ corresponding to the projection of the arc (or of the chord) onto the horizontal axis.
+This interval is shown in Figure~\ref{fig: probmax_asint_detail} with dashed, thick line.
+Firstly, in the exact case the distribution of $x$ is non-uniform on $\intervx_{\tiles,n}$ (corresponding to a uniform distribution of $\theta$), whereas the approximation implies that $x$ is uniformly distributed on $\intervx_{\tiles,n}$.
+The corresponding probability density functions will be respectively denoted as $\fdp_{\tiles,n}(x)$ and $\tilde \fdp_{\tiles,n}(x)$.
+Secondly, for a given $x$, the vertical coordinate in the exact case, $y$ (determined by the arc), is replaced by a smaller value $\tilde y$ in the approximation (chord).
+Let $R_{\tiles,n}(x)$ and $\tilde R_{\tiles,n}(x)$ denote the area of the shaded rectangle that has its upper corner on the arc and on the chord respectively, as a function of $x$.
+Then
+\begin{align}
+\label{eq: A int R fdp x}
+A_{\tiles,n} &= \int_{\intervx_{\tiles,n}} R_{\tiles,n}(x) \fdp_{\tiles,n}(x) \, \diff x, \\
+\label{eq: tilde A int R fdp x}
+\tilde A_{\tiles,n} &= \int_{\intervx_{\tiles,n}} \tilde R_{\tiles,n}(x) \tilde \fdp_{\tiles,n}(x) \, \diff x.
+\end{align}
+
+Consider the first type of error in the approximation of $A_{\tiles,n}$, which arises from using $\tilde \fdp_{\tiles,n}(x)$ instead of $\fdp_{\tiles,n}(x)$.
 In the following it will be assumed that $\tiles \geq 4$.
-It stems from Figure~\ref{fig: probmax_asint_turned} that any point $(x,y)$ on the circular arc defined by \eqref{eq: funcirc} satisfies the equality $\cot \theta = \sqrt{2} x /(\tiles-3)$.
+It stems from Figure~\ref{fig: probmax_asint_turned} that any point $(x,y)$ on the full arc defined by \eqref{eq: funcirc} satisfies the equality $\cot \theta = \sqrt{2} x /(\tiles-3)$.
 Therefore
 \begin{equation}
-\label{eq: d theta d x}
-\frac{\diff \theta}{\diff x} = -\frac{\sqrt{2}}{\tiles-3} \sin^2 \theta.
+\label{eq: d x d theta}
+\frac{\diff x}{\diff \theta} = -\frac{\tiles-3}{\sqrt{2} \sin^2 \theta}.
 \end{equation}
@@
 \end{equation}
-The uniform distribution of the corner of the shaded rectangle on the chord is equivalent to a uniform distribution of its $x$ coordinate on the projection of the chord onto the horizontal axis.
-Denoting the slope of the chord by $m_{\tiles,n}$, the projection has length
-\[
-\frac{\tilde c_{\tiles,n}} {\sqrt{1 + m_{\tiles,n}^2}},
-\]
-and $\tilde\fdp(\theta)$ is obtained, using the transformation theorem for continuous random variables \cite[section~5.2]{Papoulis91}, as
-\begin{equation}
-\label{eq: tilde fdp theta}
-\tilde \fdp(\theta) = \frac {\sqrt{1 + m_{\tiles,n}^2}} {\tilde c_{\tiles,n} |\diff \theta / \diff x|}
-\end{equation}
-for the range of $\theta$ corresponding to the arc with thick line in Figure~\ref{fig: probmax_asint_turned}. It follows from \eqref{eq: d theta d x}--\eqref{eq: tilde fdp theta} that
+In the exact case $\theta$ is uniformly distributed along the arc shown with thick line in Figure~\ref{fig: probmax_asint_turned}, and the probability density function of $x$ can be obtained using the transformation theorem for continuous random variables \cite[section~5.2]{Papoulis91}:
+\begin{equation}
+\label{eq: fdp x}
+\fdp_{\tiles,n}(x) = \frac {\len_\tiles} {c_{\tiles,n} |\diff x / \diff \theta|} \quad\text{for } x \in \intervx_{\tiles,n}.
+\end{equation}
+It follows from \eqref{eq: d x d theta}--\eqref{eq: fdp x} that
 % *!*
@@
 \label{eq: max min tilde pdf}
-\frac {\max_\theta \tilde \fdp(\theta)} {\min_\theta \tilde \fdp(\theta)} < \left(\frac{\tiles-3/2}{\tiles-3}\right)^2.
-\end{equation}
-Since $\fdp(\theta)$ and $\tilde\fdp(\theta)$ are probability density functions, it cannot be the case that one of them is greater than the other for all values of $\theta$.
-In addition, both functions have the same support, on which $\fdp(\theta)$ is constant.
-Taking into account that $\tiles \geq 4$, this implies that
+\frac {\max_{x \in \intervx_{\tiles,n}} \fdp_{\tiles,n}(x)} {\min_{x \in \intervx_{\tiles,n}} \fdp_{\tiles,n}(x)} < \left(\frac{\tiles-3/2}{\tiles-3}\right)^2.
+\end{equation}
+Since $\fdp_{\tiles,n}(x)$ and $\tilde\fdp_{\tiles,n}(x)$ are probability density functions, it cannot be the case that one of them is greater than the other for all values of $x$.
+In addition, both functions have the same support, namely $\intervx_{\tiles,n}$, on which $\tilde \fdp_{\tiles,n}(x)$ is constant.
+This implies that, for $x \in \intervx_{\tiles,n}$ and $\tiles \geq 4$,
 \begin{align}
 \label{eq: tilde pdf pdf bound >}
-\frac {\tilde \fdp(\theta)} {\fdp(\theta)} &> \left(\frac{\tiles-3}{\tiles-3/2}\right)^2 = \left(1-\frac{3/2}{\tiles-3/2}\right)^2 > 1 - \frac 3 {\tiles-3/2}, \\
+\frac {\tilde \fdp_{\tiles,n}(x)} {\fdp_{\tiles,n}(x)} &> \left(\frac{\tiles-3}{\tiles-3/2}\right)^2 = \left(1-\frac{3/2}{\tiles-3/2}\right)^2 > 1 - \frac 3 {\tiles-3/2}, \\
 \label{eq: tilde pdf pdf bound <}
-\frac {\tilde \fdp(\theta)} {\fdp(\theta)} &< \left(\frac{\tiles-3/2}{\tiles-3}\right)^2 = \left(1+\frac{3/2}{\tiles-3}\right)^2 \leq 1 + \frac {21/4} {\tiles-3}.
-\end{align}
-Thus, changing the distribution of $\theta$ from $\fdp(\theta)$ to $\tilde \fdp(\theta)$ results in $A_{\tiles,n}$ being multiplied by a factor greater than the rightmost side of \eqref{eq: tilde pdf pdf bound >} and smaller than that of \eqref{eq: tilde pdf pdf bound <}.
-
-The second type of error in the approximation of $A_{\tiles,n}$ arises because, given the $x$ coordinate of the upper corner of the shaded rectangle in Figure~\ref{fig: probmax_asint_detail}, its $y$ coordinate is reduced from that of the arc to that of the chord.
-Considering the tangent line displayed in the figure (dashed line), it is easy to see that the difference between the vertical coordinate of the arc, $y$, and that of the chord, $\tilde y$, is less than the chord length, $\tilde c_{\tiles,n}$, multiplied by the maximum of the absolute value of the arc slope.
+\frac {\tilde \fdp_{\tiles,n}(x)} {\fdp_{\tiles,n}(x)} &< \left(\frac{\tiles-3/2}{\tiles-3}\right)^2 = \left(1+\frac{3/2}{\tiles-3}\right)^2 \leq 1 + \frac {21/4} {\tiles-3}.
+\end{align}
+
+To analyze the second type of error in the approximation of $A_{\tiles,n}$, caused by using $\tilde R_{\tiles,n}(x)$ instead of $R_{\tiles,n}(x)$, consider the tangent line displayed in Figure~\ref{fig: probmax_asint_detail} (dashed line).
+This makes it easy to see that for a given $x$ the difference between the vertical coordinate of the arc, $y$, and that of the chord, $\tilde y$, is less than the chord length, $\tilde c_{\tiles,n}$, multiplied by the maximum of the absolute value of the arc slope.
 The former is at most $\sqrt{2}$.
@@
 \end{equation}
-This reduction from $y$ to $\tilde y$ translates into a difference smaller than $3 / \sqrt{2(\tiles-3)}$ in each dimension of the shaded rectangle; or, since those dimensions are at most $1$, a difference in area smaller than twice that value.
-Considering \eqref{eq: A tiles n bound}, this introduces a (second) error in $A_{\tiles,n}$ by a factor between $1-3\sqrt{2}/\sqrt{\tiles-3}$ and $1$.
-Combining with the factor from the previous paragraph, $\tilde A_{\tiles,n} / A_{\tiles,n}$ is bounded for $\tiles \geq 4$ as
-\begin{equation}
-\label{eq: tilde A bound}
-1 - \frac {3} {\tiles-3/2} - \frac{3\sqrt{2}} {\sqrt{\tiles-3}} < \frac{\tilde A_{\tiles,n}}{A_{\tiles,n}} < 1+\frac {21/4} {\tiles-3}.
-\end{equation}
-
-Multiplying \eqref{eq: tilde c bound} and \eqref{eq: tilde A bound} yields, for $\tiles \geq 4$,
-\begin{equation}
-\label{eq: tilde A c bound rel}
-1 - \frac {3} {\tiles-3/2} - \frac{3\sqrt{2}}  {\sqrt{\tiles-3}} - \frac {\pi^3} {96(\tiles-2)^2} < \frac{\tilde A_{\tiles,n} \tilde c_{\tiles,n}}{A_{\tiles,n} c_{\tiles,n}} < 1+\frac {21/4} {\tiles-3},
-\end{equation}
-which, using \eqref{eq: A tiles n bound} and \eqref{eq: c tiles n bound}, is transformed into 
-\begin{equation}
-\label{eq: tilde A c bound diff}
-- \frac {3\pi/2} {\tiles-3/2} - \frac{3\pi/\sqrt{2}}  {\sqrt{\tiles-3}} - \frac {\pi^4} {192(\tiles-2)^2} < \tilde A_{\tiles,n} \tilde c_{\tiles,n}- A_{\tiles,n} c_{\tiles,n} < \frac {21\pi/2} {\tiles-3}.
-\end{equation}
-Combining \eqref{eq: tilde A c bound diff} with \eqref{eq: P i j asympt 2 1} and \eqref{eq: P i j asympt 2 2} gives
+This reduction from $y$ to $\tilde y$ translates into a difference smaller than $3 / \sqrt{2(\tiles-3)}$ in each dimension of the shaded rectangle; or, since those dimensions are at most $1$, a difference in area smaller than twice that value:.
+\begin{equation}
+\label{eq: R: arc - chord}
+0 \leq R_{\tiles,n}(x) - \tilde R_{\tiles,n}(x) < \frac{3\sqrt{2}} {\sqrt{\tiles-3}}.
+\end{equation}
+
+Combining \eqref{eq: tilde pdf pdf bound >} and \eqref{eq: R: arc - chord} with \eqref{eq: tilde A int R fdp x}, and using \eqref{eq: A tiles n bound} and \eqref{eq: A int R fdp x}, $\tilde A_{\tiles,n} - A_{\tiles,n}$ can be bounded for $\tiles \geq 4$ as
+\begin{equation}
+\label{eq: tilde A bound diff 1}
+\begin{split}
+\tilde A_{\tiles,n} - A_{\tiles,n} &= \int_{\intervx_{\tiles,n}} \tilde R_{\tiles,n}(x) \tilde \fdp_{\tiles,n}(x) \, \diff x - A_{\tiles,n} \\
+&> \left( 1 - \frac 3 {\tiles-3/2} \right) \bigintssss_{\intervx_{\tiles,n}} \biggl( R_{\tiles,n}(x) - \frac{3\sqrt{2}} {\sqrt{\tiles-3}} \biggr) \fdp_{\tiles,n}(x) \, \diff x - A_{\tiles,n} \\
+&> \left( 1 - \frac 3 {\tiles-3/2} \right) \biggl( A_{\tiles,n} - \frac{3\sqrt{2}}{\sqrt{\tiles-3}} \biggr) - A_{\tiles,n} > -\frac 3 {\tiles-3/2} - \frac{3\sqrt{2}}{\sqrt{\tiles-3}}.
+\end{split}
+\end{equation}
+Likewise, with \eqref{eq: tilde pdf pdf bound <} instead of \eqref{eq: tilde pdf pdf bound >},
+\begin{equation}
+\label{eq: tilde A bound diff 2}
+\tilde A_{\tiles,n} - A_{\tiles,n} \leq \left( 1 + \frac {21/4} {\tiles-3} \right) \int_{\intervx_{\tiles,n}} R_{\tiles,n}(x) \fdp_{\tiles,n}(x) \, \diff x - A_{\tiles,n} = \frac {21/4} {\tiles-3}\, A_{\tiles,n} < \frac {21/4} {\tiles-3}.
+\end{equation}
+
+The difference between $\tilde A_{\tiles,n} \tilde c_{\tiles,n}$ and $A_{\tiles,n} c_{\tiles,n}$ is bounded as follows:
+\begin{equation}
+\begin{split}
+\label{eq: tilde A c bound diff 1}
+|\tilde A_{\tiles,n} \tilde c_{\tiles,n} - A_{\tiles,n} c_{\tiles,n}|
+&= |A_{\tiles,n} (\tilde c_{\tiles,n} - c_{\tiles,n}) + (\tilde A_{\tiles,n} - A_{\tiles,n}) \tilde c_{\tiles,n})| \\
+&\leq A_{\tiles,n} \left|\tilde c_{\tiles,n} - c_{\tiles,n} \right| + \tilde c_{\tiles,n} |\tilde A_{\tiles,n} - A_{\tiles,n})|.
+\end{split}
+\end{equation}
+Using \eqref{eq: A tiles n bound}, \eqref{eq: c tiles n bound}, \eqref{eq: tilde c bound diff}, \eqref{eq: tilde A bound diff 1} and \eqref{eq: tilde A bound diff 2} into \eqref{eq: tilde A c bound diff 1},
+\begin{equation}
+\begin{split}
+\label{eq: tilde A c bound diff 2}
+|\tilde A_{\tiles,n} \tilde c_{\tiles,n} - A_{\tiles,n} c_{\tiles,n}|
+&< \frac {\pi^3} {96(\tiles-2)^2} + \frac{\pi}{2} \max\left\{
+\frac 3 {\tiles-3/2} + \frac{3\sqrt{2}}{\sqrt{\tiles-3}}, \frac {21/4} {\tiles-3}
+\right\}.
+\end{split}
+\end{equation}
+From \eqref{eq: P i j asympt 2 1}, \eqref{eq: P i j asympt 2 2} and \eqref{eq: tilde A c bound diff 2} it stems that
 \begin{equation}
@@
 \end{equation}
-Substituting \eqref{eq: P i j asympt 4 bis} into \eqref{eq: P i j asympt 4} and taking into account \eqref{eq: P i j asympt 4 ter},
+Using \eqref{eq: P i j asympt 4 bis} into \eqref{eq: P i j asympt 4} and taking into account \eqref{eq: P i j asympt 4 ter},
 \begin{equation}

### 2023-07-10 figures/figures.m

@@ -3116,6 +3116,8 @@ s1 = .5825; plot(x_dots(ind_dot)+[-sqrt(.5) 0]*s1, [sqrt(.5) 0]*s1, 'k-', 'linew
 s2 = 1; plot(x_dots(ind_dot)+[sqrt(.5) 0]*s2, [sqrt(.5) 0]*s2, 'k-', 'linewidth', 1) % s manually
 set(gca, 'colororderindex', 1)
 plot([x_dots(ind_dot)-sqrt(.5)*s1 x_dots(ind_dot)+sqrt(.5)*s2], [sqrt(.5)*s1 sqrt(.5)*s2], 'linewidth', 1)
+plot([x_dots(ind_dot)-sqrt(.5)*s1 x_dots(ind_dot)+sqrt(.5)*s2], [0 0], 'k--', 'linewidth', 1.2) % projection
+text(-1.205, .068 , 'I_{t,n}') % its text
 set(gca, 'colororderindex', 1)
 plot([x_dots(ind_dot) x_dots(ind_dot)+1j*.520557], 'k-', 'linewidth', .5) % manually
 plot([z0 z1], 'k:')
